### 2020-03-10 egpaper_final.tex
initial commit

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[breaklinks=true,bookmarks=false]{hyperref}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the EV Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{4321}
\begin{document}

%%%%%%%%% TITLE
\title{\LaTeX\ Author Guidelines for EV Proceedings}

\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
   The ABSTRACT is to be in fully-justified italicized text, at the top
   of the left-hand column, below the author and affiliation
   information. Use the word ``Abstract'' as the title, in 12-point
   Times, boldface type, centered relative to the column, initially
   capitalized. The abstract is to be in 10-point, single-spaced type.
   Leave two blank lines after the Abstract, then begin the main text.
   Look at previous EV abstracts to get a feel for style and length.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}

Please follow the steps outlined below when submitting your manuscript to
the IEEE Computer Society Press.  This style guide now has several
important modifications (for example, you are no longer warned against the
use of sticky tape to attach your artwork to the paper), so all authors
should read this new version.

%-------------------------------------------------------------------------
\subsection{Language}

All manuscripts must be in English.

\subsection{Dual submission}

Please refer to the author guidelines on the CVPR 2020 web page for a
discussion of the policy on dual submissions.

\subsection{Paper length}
Papers, excluding the references section,
must be no longer than eight pages in length. The references section
will not be included in the page count, and there is no limit on the
length of the references section. For example, a paper of eight pages
with two pages of references would have a total length of 10 pages.
{\bf There will be no extra page charges for CVPR 2020.}

Overlength papers will simply not be reviewed.  This includes papers
where the margins and formatting are deemed to have been significantly
altered from those laid down by this style guide.  Note that this
\LaTeX\ guide already sets figure captions and references in a smaller font.
The reason such papers will not be reviewed is that there is no provision for
supervised revisions of manuscripts.  The reviewing process cannot determine
the suitability of the paper for presentation in eight pages if it is
reviewed in eleven.

%-------------------------------------------------------------------------
\subsection{The ruler}
The \LaTeX\ style defines a printed ruler which should be present in the
version submitted for review.  The ruler is provided in order that
reviewers may comment on particular lines in the paper without
circumlocution.  If you are preparing a document using a non-\LaTeX\
document preparation system, please arrange for an equivalent ruler to
appear on the final output pages.  The presence or absence of the ruler
should not change the appearance of any other content on the page.  The
camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
the \verb'\cvprfinalcopy' command in the document preamble.)  Reviewers:
note that the ruler measurements do not align well with lines in the paper
--- this turns out to be very difficult to do well when the paper contains
many figures and equations, and, when done, looks ugly.  Just use fractional
references (e.g.\ this line is $095.5$), although in most cases one would
expect that the approximate location will be adequate.

\subsection{Mathematics}

Please number all of your sections and displayed equations.  It is
important for readers to be able to refer to any particular equation.  Just
because you didn't refer to it in the text doesn't mean some future reader
might not need to refer to it.  It is cumbersome to have to use
circumlocutions like ``the equation second from the top of page 3 column
1''.  (Note that the ruler will not be present in the final copy, so is not
an alternative to equation numbers).  All authors will benefit from reading
Mermin's description of how to write mathematics:
\url{http://www.pamitc.org/documents/mermin.pdf}.

\subsection{Blind review}

Many authors misunderstand the concept of anonymizing for blind
review.  Blind review does not mean that one must remove
citations to one's own work---in fact it is often impossible to
review a paper unless the previous citations are known and
available.

Blind review means that you do not use the words ``my'' or ``our''
when citing previous work.  That is all.  (But see below for
techreports.)

Saying ``this builds on the work of Lucy Smith [1]'' does not say
that you are Lucy Smith; it says that you are building on her
work.  If you are Smith and Jones, do not say ``as we show in
[7]'', say ``as Smith and Jones show in [7]'' and at the end of the
paper, include reference 7 as you would any other cited work.

An example of a bad paper just asking to be rejected:
\begin{quote}
\begin{center}
    An analysis of the frobnicatable foo filter.
\end{center}

   In this paper we present a performance analysis of our
   previous paper [1], and show it to be inferior to all
   previously known methods.  Why the previous paper was
   accepted without this analysis is beyond me.

   [1] Removed for blind review
\end{quote}

An example of an acceptable paper:

\begin{quote}
\begin{center}
     An analysis of the frobnicatable foo filter.
\end{center}

   In this paper we present a performance analysis of the
   paper of Smith \etal [1], and show it to be inferior to
   all previously known methods.  Why the previous paper
   was accepted without this analysis is beyond me.

   [1] Smith, L and Jones, C. ``The frobnicatable foo
   filter, a fundamental contribution to human knowledge''.
   Nature 381(12), 1-213.
\end{quote}

If you are making a submission to another conference at the same time,
which covers similar or overlapping material, you may need to refer to that
submission in order to explain the differences, just as you would if you
had previously published related work.  In such cases, include the
anonymized parallel submission~\cite{Authors14} as additional material and
cite it as
\begin{quote}
[1] Authors. ``The frobnicatable foo filter'', F\&G 2014 Submission ID 324,
Supplied as additional material {\tt fg324.pdf}.
\end{quote}

Finally, you may feel you need to tell the reader that more details can be
found elsewhere, and refer them to a technical report.  For conference
submissions, the paper must stand on its own, and not {\em require} the
reviewer to go to a techreport for further details.  Thus, you may say in
the body of the paper ``further details may be found
in~\cite{Authors14b}''.  Then submit the techreport as additional material.
Again, you may not assume the reviewers will read this material.

Sometimes your paper is about a problem which you tested using a tool which
is widely known to be restricted to a single institution.  For example,
let's say it's 1969, you have solved a key problem on the Apollo lander,
and you believe that the CVPR70 audience would like to hear about your
solution.  The work is a development of your celebrated 1968 paper entitled
``Zero-g frobnication: How being the only people in the world with access to
the Apollo lander source code makes us a wow at parties'', by Zeus \etal.

You can handle this paper like any other.  Don't write ``We show how to
improve our previous work [Anonymous, 1968].  This time we tested the
algorithm on a lunar lander [name of lander removed for blind review]''.
That would be silly, and would immediately identify the authors. Instead
write the following:
\begin{quotation}
\noindent
   We describe a system for zero-g frobnication.  This
   system is new because it handles the following cases:
   A, B.  Previous systems [Zeus et al. 1968] didn't
   handle case B properly.  Ours handles it by including
   a foo term in the bar integral.

   ...

   The proposed system was integrated with the Apollo
   lunar lander, and went all the way to the moon, don't
   you know.  It displayed the following behaviours
   which show how well we solved cases A and B: ...
\end{quotation}
As you can see, the above text follows standard scientific convention,
reads better than the first version, and does not explicitly name you as
the authors.  A reviewer might think it likely that the new paper was
written by Zeus \etal, but cannot make any decision based on that guess.
He or she would have to be sure that no other authors could have been
contracted to solve problem B.
\medskip

\noindent
FAQ\medskip\\
{\bf Q:} Are acknowledgements OK?\\
{\bf A:} No.  Leave them for the final copy.\medskip\\
{\bf Q:} How do I cite my results reported in open challenges?
{\bf A:} To conform with the double blind review policy, you can report results of other challenge participants together with your results in your paper. For your results, however, you should not identify yourself and should not mention your participation in the challenge. Instead present your results referring to the method proposed in your paper and draw conclusions based on the experimental comparison to other results.\medskip\\

\begin{figure}[t]
\begin{center}
\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   %\includegraphics[width=0.8\linewidth]{egfigure.eps}
\end{center}
   \caption{Example of caption.  It is set in Roman so that mathematics
   (always set in Roman: $B \sin A = A \sin B$) may be included without an
   ugly clash.}
\label{fig:long}
\label{fig:onecol}
\end{figure}

\subsection{Miscellaneous}

\noindent
Compare the following:\\
\begin{tabular}{ll}
 \verb'$conf_a$' &  $conf_a$ \\
 \verb'$\mathit{conf}_a$' & $\mathit{conf}_a$
\end{tabular}\\
See The \TeX book, p165.

The space after \eg, meaning ``for example'', should not be a
sentence-ending space. So \eg is correct, {\em e.g.} is not.  The provided
\verb'\eg' macro takes care of this.

When citing a multi-author paper, you may save space by using ``et alia'',
shortened to ``\etal'' (not ``{\em et.\ al.}'' as ``{\em et}'' is a complete word.)
However, use it only when there are three or more authors.  Thus, the
following is correct: ``
   Frobnication has been trendy lately.
   It was introduced by Alpher~\cite{Alpher02}, and subsequently developed by
   Alpher and Fotheringham-Smythe~\cite{Alpher03}, and Alpher \etal~\cite{Alpher04}.''

This is incorrect: ``... subsequently developed by Alpher \etal~\cite{Alpher03} ...''
because reference~\cite{Alpher03} has just two authors.  If you use the
\verb'\etal' macro provided, then you need not worry about double periods
when used at the end of a sentence as in Alpher \etal.

For this citation style, keep multiple citations in numerical (not
chronological) order, so prefer \cite{Alpher03,Alpher02,Authors14} to
\cite{Alpher02,Alpher03,Authors14}.

\begin{figure*}
\begin{center}
\fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
\end{center}
   \caption{Example of a short caption, which should be centered.}
\label{fig:short}
\end{figure*}

%------------------------------------------------------------------------
\section{Formatting your paper}

All text must be in a two-column format. The total allowable width of the
text area is $6\frac78$ inches (17.5 cm) wide by $8\frac78$ inches (22.54
cm) high. Columns are to be $3\frac14$ inches (8.25 cm) wide, with a
$\frac{5}{16}$ inch (0.8 cm) space between them. The main title (on the
first page) should begin 1.0 inch (2.54 cm) from the top edge of the
page. The second and following pages should begin 1.0 inch (2.54 cm) from
the top edge. On all pages, the bottom margin should be 1-1/8 inches (2.86
cm) from the bottom edge of the page for $8.5 \times 11$-inch paper; for A4
paper, approximately 1-5/8 inches (4.13 cm) from the bottom edge of the
page.

%-------------------------------------------------------------------------
\subsection{Margins and page numbering}

All printed material, including text, illustrations, and charts, must be kept
within a print area 6-7/8 inches (17.5 cm) wide by 8-7/8 inches (22.54 cm)
high.
Page numbers should be in footer with page numbers, centered and .75
inches from the bottom of the page and make it start at the correct page
number rather than the 4321 in the example.  To do this fine the line (around
line 23)
\begin{verbatim}
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{4321}
\end{verbatim}
where the number 4321 is your assigned starting page.

Make sure the first page is numbered by commenting out the first page being
empty on line 46
\begin{verbatim}
%\thispagestyle{empty}
\end{verbatim}

%-------------------------------------------------------------------------
\subsection{Type-style and fonts}

Wherever Times is specified, Times Roman may also be used. If neither is
available on your word processor, please use the font closest in
appearance to Times to which you have access.

MAIN TITLE. Center the title 1-3/8 inches (3.49 cm) from the top edge of
the first page. The title should be in Times 14-point, boldface type.
Capitalize the first letter of nouns, pronouns, verbs, adjectives, and
adverbs; do not capitalize articles, coordinate conjunctions, or
prepositions (unless the title begins with such a word). Leave two blank
lines after the title.

AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title
and printed in Times 12-point, non-boldface type. This information is to
be followed by two blank lines.

The ABSTRACT and MAIN TEXT are to be in a two-column format.

MAIN TEXT. Type main text in 10-point Times, single-spaced. Do NOT use
double-spacing. All paragraphs should be indented 1 pica (approx. 1/6
inch or 0.422 cm). Make sure your text is fully justified---that is,
flush left and flush right. Please do not place any additional blank
lines between paragraphs.

Figure and table captions should be 9-point Roman type as in
Figures~\ref{fig:onecol} and~\ref{fig:short}.  Short captions should be centred.

\noindent Callouts should be 9-point Helvetica, non-boldface type.
Initially capitalize only the first word of section titles and first-,
second-, and third-order headings.

FIRST-ORDER HEADINGS. (For example, {\large \bf 1. Introduction})
should be Times 12-point boldface, initially capitalized, flush left,
with one blank line before, and one blank line after.

SECOND-ORDER HEADINGS. (For example, { \bf 1.1. Database elements})
should be Times 11-point boldface, initially capitalized, flush left,
with one blank line before, and one after. If you require a third-order
heading (we discourage it), use 10-point Times, boldface, initially
capitalized, flush left, preceded by one blank line, followed by a period
and your text on the same line.

%-------------------------------------------------------------------------
\subsection{Footnotes}

Please use footnotes\footnote {This is what a footnote looks like.  It
often distracts the reader from the main flow of the argument.} sparingly.
Indeed, try to avoid footnotes altogether and include necessary peripheral
observations in
the text (within parentheses, if you prefer, as in this sentence).  If you
wish to use a footnote, place it at the bottom of the column on the page on
which it is referenced. Use Times 8-point type, single-spaced.

%-------------------------------------------------------------------------
\subsection{References}

List and number all bibliographical references in 9-point Times,
single-spaced, at the end of your paper. When referenced in the text,
enclose the citation number in square brackets, for
example~\cite{Authors14}.  Where appropriate, include the name(s) of
editors of referenced books.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|}
\hline
Method & Frobnability \\
\hline\hline
Theirs & Frumpy \\
Yours & Frobbly \\
Ours & Makes one's heart Frob\\
\hline
\end{tabular}
\end{center}
\caption{Results.   Ours is better.}
\end{table}

%-------------------------------------------------------------------------
\subsection{Illustrations, graphs, and photographs}

All graphics should be centered.  Please ensure that any point you wish to
make is resolvable in a printed copy of the paper.  Resize fonts in figures
to match the font in the body text, and choose line widths which render
effectively in print.  Many readers (and reviewers), even of an electronic
copy, will choose to print your paper in order to read it.  You cannot
insist that they do otherwise, and therefore must not assume that they can
zoom in to see tiny details on a graphic.

When placing figures in \LaTeX, it's almost always best to use
\verb+\includegraphics+, and to specify the  figure width as a multiple of
the line width as in the example below
{\small\begin{verbatim}
   \usepackage[dvips]{graphicx} ...
   \includegraphics[width=0.8\linewidth]
                   {myfile.eps}
\end{verbatim}
}

%-------------------------------------------------------------------------
\subsection{Color}

Please refer to the author guidelines on the CVPR 2020 web page for a discussion
of the use of color in your document.

%------------------------------------------------------------------------
\section{Final copy}

You must include your signed IEEE copyright release form when you submit
your finished paper. We MUST have this form before your paper can be
published in the proceedings.

Please direct any questions to the production editor in charge of these 
proceedings at the IEEE Computer Society Press: 
\url{https://www.computer.org/about/contact}.

{\small
\bibliographystyle{ieee_fullname}
\bibliography{egbib}
}

\end{document}

### 2020-03-12 egpaper_for_review.tex
Updates from Overleaf

--- 2020-03-10 egpaper_final.tex
@@
 
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\A}{\mathcal{A}}
+\newcommand{\Z}{\mathbb{Z}}
+\newcommand{\m}[1]{\pmb{\mathrm{#1}}}
+
 % If you comment hyperref and then uncomment it, you should delete
@@
 % run, let it finish, and you should be clear).
-\usepackage[breaklinks=true,bookmarks=false]{hyperref}
+\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
 
-\cvprfinalcopy % *** Uncomment this line for the final submission
+% \cvprfinalcopy % *** Uncomment this line for the final submission
 
@@
 % Pages are numbered in submission mode, and unnumbered in camera-ready
-%\ifcvprfinal\pagestyle{empty}\fi
-\setcounter{page}{4321}
+\ifcvprfinal\pagestyle{empty}\fi
 \begin{document}
@@
 %%%%%%%%% TITLE
-\title{\LaTeX\ Author Guidelines for EV Proceedings}
-
-\author{First Author\\
-Institution1\\
+\title{Learning filters from markers}
+\author{Italos Estilon de Souza\\
+Institute of Computing - University of Campinas (UNICAMP)\\
 Institution1 address\\
-{\tt\small firstauthor@i1.org}
+{\tt\small italos.souza@ic.unicamp.br}
 % For a paper whose authors are all at the same institution,
@@
 \and
-Second Author\\
-Institution2\\
-First line of institution2 address\\
-{\tt\small secondauthor@i2.org}
+Alexandre Falcão\\
+{\tt\small afalcao@ic.unicamp.br}
 }
@@
 %%%%%%%%% ABSTRACT
-\begin{abstract}
-   The ABSTRACT is to be in fully-justified italicized text, at the top
-   of the left-hand column, below the author and affiliation
-   information. Use the word ``Abstract'' as the title, in 12-point
-   Times, boldface type, centered relative to the column, initially
-   capitalized. The abstract is to be in 10-point, single-spaced type.
-   Leave two blank lines after the Abstract, then begin the main text.
-   Look at previous EV abstracts to get a feel for style and length.
-\end{abstract}
 
@@
 
-Please follow the steps outlined below when submitting your manuscript to
-the IEEE Computer Society Press.  This style guide now has several
-important modifications (for example, you are no longer warned against the
-use of sticky tape to attach your artwork to the paper), so all authors
-should read this new version.
+\section{Preliminaries}
 
-%-------------------------------------------------------------------------
-\subsection{Language}
+Let $D \subset \Z^d$ for $d > 1$ and let $\m{I}: D \to \R^b$, with $b \ge 1$, be a function that assings to each $p \in D$ a vector $\m{I}(p) = (I_1(p), I_2(p), \ldots, I_b(p))$.
+A multi-band and multi-dimensional image is a pair ${\hat{I} = (D, \m{I})}$.
+We say that $D$ is the \textit{image domain}, $\m{I}(p)$ is the \textit{local feature vector} of a space element (\textit{spel}) $p \in D$, $b$ is the number of bands, and $d$ is the number of dimensions.
+In this work, we focus on images with $d=2$.
 
-All manuscripts must be in English.
-
-\subsection{Dual submission}
-
-Please refer to the author guidelines on the CVPR 2020 web page for a
-discussion of the policy on dual submissions.
-
-\subsection{Paper length}
-Papers, excluding the references section,
-must be no longer than eight pages in length. The references section
-will not be included in the page count, and there is no limit on the
-length of the references section. For example, a paper of eight pages
-with two pages of references would have a total length of 10 pages.
-{\bf There will be no extra page charges for CVPR 2020.}
-
-Overlength papers will simply not be reviewed.  This includes papers
-where the margins and formatting are deemed to have been significantly
-altered from those laid down by this style guide.  Note that this
-\LaTeX\ guide already sets figure captions and references in a smaller font.
-The reason such papers will not be reviewed is that there is no provision for
-supervised revisions of manuscripts.  The reviewing process cannot determine
-the suitability of the paper for presentation in eight pages if it is
-reviewed in eleven.
-
-%-------------------------------------------------------------------------
-\subsection{The ruler}
-The \LaTeX\ style defines a printed ruler which should be present in the
-version submitted for review.  The ruler is provided in order that
-reviewers may comment on particular lines in the paper without
-circumlocution.  If you are preparing a document using a non-\LaTeX\
-document preparation system, please arrange for an equivalent ruler to
-appear on the final output pages.  The presence or absence of the ruler
-should not change the appearance of any other content on the page.  The
-camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
-the \verb'\cvprfinalcopy' command in the document preamble.)  Reviewers:
-note that the ruler measurements do not align well with lines in the paper
---- this turns out to be very difficult to do well when the paper contains
-many figures and equations, and, when done, looks ugly.  Just use fractional
-references (e.g.\ this line is $095.5$), although in most cases one would
-expect that the approximate location will be adequate.
-
-\subsection{Mathematics}
-
-Please number all of your sections and displayed equations.  It is
-important for readers to be able to refer to any particular equation.  Just
-because you didn't refer to it in the text doesn't mean some future reader
-might not need to refer to it.  It is cumbersome to have to use
-circumlocutions like ``the equation second from the top of page 3 column
-1''.  (Note that the ruler will not be present in the final copy, so is not
-an alternative to equation numbers).  All authors will benefit from reading
-Mermin's description of how to write mathematics:
-\url{http://www.pamitc.org/documents/mermin.pdf}.
-
-\subsection{Blind review}
-
-Many authors misunderstand the concept of anonymizing for blind
-review.  Blind review does not mean that one must remove
-citations to one's own work---in fact it is often impossible to
-review a paper unless the previous citations are known and
-available.
-
-Blind review means that you do not use the words ``my'' or ``our''
-when citing previous work.  That is all.  (But see below for
-techreports.)
-
-Saying ``this builds on the work of Lucy Smith [1]'' does not say
-that you are Lucy Smith; it says that you are building on her
-work.  If you are Smith and Jones, do not say ``as we show in
-[7]'', say ``as Smith and Jones show in [7]'' and at the end of the
-paper, include reference 7 as you would any other cited work.
-
-An example of a bad paper just asking to be rejected:
-\begin{quote}
-\begin{center}
-    An analysis of the frobnicatable foo filter.
-\end{center}
-
-   In this paper we present a performance analysis of our
-   previous paper [1], and show it to be inferior to all
-   previously known methods.  Why the previous paper was
-   accepted without this analysis is beyond me.
-
-   [1] Removed for blind review
-\end{quote}
-
-An example of an acceptable paper:
-
-\begin{quote}
-\begin{center}
-     An analysis of the frobnicatable foo filter.
-\end{center}
-
-   In this paper we present a performance analysis of the
-   paper of Smith \etal [1], and show it to be inferior to
-   all previously known methods.  Why the previous paper
-   was accepted without this analysis is beyond me.
-
-   [1] Smith, L and Jones, C. ``The frobnicatable foo
-   filter, a fundamental contribution to human knowledge''.
-   Nature 381(12), 1-213.
-\end{quote}
-
-If you are making a submission to another conference at the same time,
-which covers similar or overlapping material, you may need to refer to that
-submission in order to explain the differences, just as you would if you
-had previously published related work.  In such cases, include the
-anonymized parallel submission~\cite{Authors14} as additional material and
-cite it as
-\begin{quote}
-[1] Authors. ``The frobnicatable foo filter'', F\&G 2014 Submission ID 324,
-Supplied as additional material {\tt fg324.pdf}.
-\end{quote}
-
-Finally, you may feel you need to tell the reader that more details can be
-found elsewhere, and refer them to a technical report.  For conference
-submissions, the paper must stand on its own, and not {\em require} the
-reviewer to go to a techreport for further details.  Thus, you may say in
-the body of the paper ``further details may be found
-in~\cite{Authors14b}''.  Then submit the techreport as additional material.
-Again, you may not assume the reviewers will read this material.
-
-Sometimes your paper is about a problem which you tested using a tool which
-is widely known to be restricted to a single institution.  For example,
-let's say it's 1969, you have solved a key problem on the Apollo lander,
-and you believe that the CVPR70 audience would like to hear about your
-solution.  The work is a development of your celebrated 1968 paper entitled
-``Zero-g frobnication: How being the only people in the world with access to
-the Apollo lander source code makes us a wow at parties'', by Zeus \etal.
-
-You can handle this paper like any other.  Don't write ``We show how to
-improve our previous work [Anonymous, 1968].  This time we tested the
-algorithm on a lunar lander [name of lander removed for blind review]''.
-That would be silly, and would immediately identify the authors. Instead
-write the following:
-\begin{quotation}
-\noindent
-   We describe a system for zero-g frobnication.  This
-   system is new because it handles the following cases:
-   A, B.  Previous systems [Zeus et al. 1968] didn't
-   handle case B properly.  Ours handles it by including
-   a foo term in the bar integral.
-
-   ...
-
-   The proposed system was integrated with the Apollo
-   lunar lander, and went all the way to the moon, don't
-   you know.  It displayed the following behaviours
-   which show how well we solved cases A and B: ...
-\end{quotation}
-As you can see, the above text follows standard scientific convention,
-reads better than the first version, and does not explicitly name you as
-the authors.  A reviewer might think it likely that the new paper was
-written by Zeus \etal, but cannot make any decision based on that guess.
-He or she would have to be sure that no other authors could have been
-contracted to solve problem B.
-\medskip
-
-\noindent
-FAQ\medskip\\
-{\bf Q:} Are acknowledgements OK?\\
-{\bf A:} No.  Leave them for the final copy.\medskip\\
-{\bf Q:} How do I cite my results reported in open challenges?
-{\bf A:} To conform with the double blind review policy, you can report results of other challenge participants together with your results in your paper.
-For your results, however, you should not identify yourself and should not mention your participation in the challenge.
-Instead present your results referring to the method proposed in your paper and draw conclusions based on the experimental comparison to other results.\medskip\\
-
-\begin{figure}[t]
-\begin{center}
-\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
-   %\includegraphics[width=0.8\linewidth]{egfigure.eps}
-\end{center}
-   \caption{Example of caption.  It is set in Roman so that mathematics
-   (always set in Roman: $B \sin A = A \sin B$) may be included without an
-   ugly clash.}
-\label{fig:long}
-\label{fig:onecol}
-\end{figure}
-
-\subsection{Miscellaneous}
-
-\noindent
-Compare the following:\\
-\begin{tabular}{ll}
- \verb'$conf_a$' &  $conf_a$ \\
- \verb'$\mathit{conf}_a$' & $\mathit{conf}_a$
-\end{tabular}\\
-See The \TeX book, p165.
-
-The space after \eg, meaning ``for example'', should not be a
-sentence-ending space. So \eg is correct, {\em e.g.} is not.  The provided
-\verb'\eg' macro takes care of this.
-
-When citing a multi-author paper, you may save space by using ``et alia'',
-shortened to ``\etal'' (not ``{\em et.\ al.}'' as ``{\em et}'' is a complete word.)
-However, use it only when there are three or more authors.  Thus, the
-following is correct: ``
-   Frobnication has been trendy lately.
-   It was introduced by Alpher~\cite{Alpher02}, and subsequently developed by
-   Alpher and Fotheringham-Smythe~\cite{Alpher03}, and Alpher \etal~\cite{Alpher04}.''
-
-This is incorrect: ``... subsequently developed by Alpher \etal~\cite{Alpher03} ...''
-because reference~\cite{Alpher03} has just two authors.  If you use the
-\verb'\etal' macro provided, then you need not worry about double periods
-when used at the end of a sentence as in Alpher \etal.
-
-For this citation style, keep multiple citations in numerical (not
-chronological) order, so prefer \cite{Alpher03,Alpher02,Authors14} to
-\cite{Alpher02,Alpher03,Authors14}.
-
-\begin{figure*}
-\begin{center}
-\fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
-\end{center}
-   \caption{Example of a short caption, which should be centered.}
-\label{fig:short}
-\end{figure*}
-
-%------------------------------------------------------------------------
-\section{Formatting your paper}
-
-All text must be in a two-column format. The total allowable width of the
-text area is $6\frac78$ inches (17.5 cm) wide by $8\frac78$ inches (22.54
-cm) high. Columns are to be $3\frac14$ inches (8.25 cm) wide, with a
-$\frac{5}{16}$ inch (0.8 cm) space between them. The main title (on the
-first page) should begin 1.0 inch (2.54 cm) from the top edge of the
-page. The second and following pages should begin 1.0 inch (2.54 cm) from
-the top edge. On all pages, the bottom margin should be 1-1/8 inches (2.86
-cm) from the bottom edge of the page for $8.5 \times 11$-inch paper; for A4
-paper, approximately 1-5/8 inches (4.13 cm) from the bottom edge of the
-page.
-
-%-------------------------------------------------------------------------
-\subsection{Margins and page numbering}
-
-All printed material, including text, illustrations, and charts, must be kept
-within a print area 6-7/8 inches (17.5 cm) wide by 8-7/8 inches (22.54 cm)
-high.
-Page numbers should be in footer with page numbers, centered and .75
-inches from the bottom of the page and make it start at the correct page
-number rather than the 4321 in the example.  To do this fine the line (around
-line 23)
-\begin{verbatim}
-%\ifcvprfinal\pagestyle{empty}\fi
-\setcounter{page}{4321}
-\end{verbatim}
-where the number 4321 is your assigned starting page.
-
-Make sure the first page is numbered by commenting out the first page being
-empty on line 46
-\begin{verbatim}
-%\thispagestyle{empty}
-\end{verbatim}
-
-%-------------------------------------------------------------------------
-\subsection{Type-style and fonts}
-
-Wherever Times is specified, Times Roman may also be used. If neither is
-available on your word processor, please use the font closest in
-appearance to Times to which you have access.
-
-MAIN TITLE. Center the title 1-3/8 inches (3.49 cm) from the top edge of
-the first page. The title should be in Times 14-point, boldface type.
-Capitalize the first letter of nouns, pronouns, verbs, adjectives, and
-adverbs; do not capitalize articles, coordinate conjunctions, or
-prepositions (unless the title begins with such a word). Leave two blank
-lines after the title.
-
-AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title
-and printed in Times 12-point, non-boldface type. This information is to
-be followed by two blank lines.
-
-The ABSTRACT and MAIN TEXT are to be in a two-column format.
-
-MAIN TEXT. Type main text in 10-point Times, single-spaced. Do NOT use
-double-spacing. All paragraphs should be indented 1 pica (approx. 1/6
-inch or 0.422 cm). Make sure your text is fully justified---that is,
-flush left and flush right. Please do not place any additional blank
-lines between paragraphs.
-
-Figure and table captions should be 9-point Roman type as in
-Figures~\ref{fig:onecol} and~\ref{fig:short}.  Short captions should be centred.
-
-\noindent Callouts should be 9-point Helvetica, non-boldface type.
-Initially capitalize only the first word of section titles and first-,
-second-, and third-order headings.
-
-FIRST-ORDER HEADINGS. (For example, {\large \bf 1. Introduction})
-should be Times 12-point boldface, initially capitalized, flush left,
-with one blank line before, and one blank line after.
-
-SECOND-ORDER HEADINGS. (For example, { \bf 1.1. Database elements})
-should be Times 11-point boldface, initially capitalized, flush left,
-with one blank line before, and one after. If you require a third-order
-heading (we discourage it), use 10-point Times, boldface, initially
-capitalized, flush left, preceded by one blank line, followed by a period
-and your text on the same line.
-
-%-------------------------------------------------------------------------
-\subsection{Footnotes}
-
-Please use footnotes\footnote {This is what a footnote looks like.  It
-often distracts the reader from the main flow of the argument.} sparingly.
-Indeed, try to avoid footnotes altogether and include necessary peripheral
-observations in
-the text (within parentheses, if you prefer, as in this sentence).  If you
-wish to use a footnote, place it at the bottom of the column on the page on
-which it is referenced. Use Times 8-point type, single-spaced.
-
-%-------------------------------------------------------------------------
-\subsection{References}
-
-List and number all bibliographical references in 9-point Times,
-single-spaced, at the end of your paper. When referenced in the text,
-enclose the citation number in square brackets, for
-example~\cite{Authors14}.  Where appropriate, include the name(s) of
-editors of referenced books.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{|l|c|}
-\hline
-Method & Frobnability \\
-\hline\hline
-Theirs & Frumpy \\
-Yours & Frobbly \\
-Ours & Makes one's heart Frob\\
-\hline
-\end{tabular}
-\end{center}
-\caption{Results.   Ours is better.}
-\end{table}
-
-%-------------------------------------------------------------------------
-\subsection{Illustrations, graphs, and photographs}
-
-All graphics should be centered.  Please ensure that any point you wish to
-make is resolvable in a printed copy of the paper.  Resize fonts in figures
-to match the font in the body text, and choose line widths which render
-effectively in print.  Many readers (and reviewers), even of an electronic
-copy, will choose to print your paper in order to read it.  You cannot
-insist that they do otherwise, and therefore must not assume that they can
-zoom in to see tiny details on a graphic.
-
-When placing figures in \LaTeX, it's almost always best to use
-\verb+\includegraphics+, and to specify the  figure width as a multiple of
-the line width as in the example below
-{\small\begin{verbatim}
-   \usepackage[dvips]{graphicx} ...
-   \includegraphics[width=0.8\linewidth]
-                   {myfile.eps}
-\end{verbatim}
-}
-
-%-------------------------------------------------------------------------
-\subsection{Color}
-
-Please refer to the author guidelines on the CVPR 2020 web page for a discussion
-of the use of color in your document.
-
-%------------------------------------------------------------------------
-\section{Final copy}
-
-You must include your signed IEEE copyright release form when you submit
-your finished paper. We MUST have this form before your paper can be
-published in the proceedings.
-
-Please direct any questions to the production editor in charge of these 
-proceedings at the IEEE Computer Society Press: 
-\url{https://www.computer.org/about/contact}.
-
+An adjacency relation $\A \subseteq D \times D$ is a binary relation between spels $p, q \in D$.
+We focus on reflexive and translation-invariant relations.
+More precisely, we focus on adjacency relations given by $\A_{i,j} = \{ (p, q) \in D : \|x_p - x_q\| \le i, \|y_p - y_q\| \le j \}$
 {\small

### 2020-03-12 egpaper_for_review.tex

@@
 
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\A}{\mathcal{A}}
-\newcommand{\Z}{\mathbb{Z}}
-\newcommand{\m}[1]{\pmb{\mathrm{#1}}}
+\usepackage{subcaption}
+\usepackage{siunitx}
 
@@
 
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\A}{\mathcal{A}}
+\newcommand{\D}{\mathcal{D}}
+\newcommand{\I}{\hat{I}}
+\newcommand{\Z}{\mathbb{Z}}
+\newcommand{\m}[1]{{\mathrm{\bf #1}}}
+\newcommand{\E}{\tilde{\m{I}}}
+\newcommand{\F}{\hat{F}}
+\newcommand{\lI}{\m{I}}
+\newcommand{\mF}{\m{F}}
+\newcommand{\bF}{\mathcal{F}}
+\newcommand{\J}{\hat{J}}
+\newcommand{\lJ}{\m{J}}
+\newcommand{\pD}{D^\prime}
+\newcommand{\eF}{\hat{\mF}}
+\newcommand{\gN}{\m{N}}
+\newcommand{\W}{\m{W}}
+\newcommand{\M}{\mathcal{M}}
+\newcommand{\Pa}{\mathcal{P}}
+\newcommand{\pDD}{\D^\prime}
 % \cvprfinalcopy % *** Uncomment this line for the final submission
@@
 
-Let $D \subset \Z^d$ for $d > 1$ and let $\m{I}: D \to \R^b$, with $b \ge 1$, be a function that assings to each $p \in D$ a vector $\m{I}(p) = (I_1(p), I_2(p), \ldots, I_b(p))$.
+Let $D \subset \Z^d$ for $d > 1$ and let $\m{I}: D \to \R^b$, with $b \ge 1$, be a function that assings to each $p = (x_p, y_q) \in D$ a vector $\m{I}(p) = (I_1(p), I_2(p), \ldots, I_b(p))$.
 A multi-band and multi-dimensional image is a pair ${\hat{I} = (D, \m{I})}$.
@@
 In this work, we focus on images with $d=2$.
+The set of local feature vectors is called \textit{local feature space}.
 
@@
 We focus on reflexive and translation-invariant relations.
-More precisely, we focus on adjacency relations given by $\A_{i,j} = \{ (p, q) \in D : \|x_p - x_q\| \le i, \|y_p - y_q\| \le j \}$
+More precisely, we focus on adjacency relations given by \[\A_{i,j} = \{ (p, q) \in D : \|x_p - x_q\| \le i, \|y_p - y_q\| \le j \}.\] We say that $\A(p) = \{q \in D : (p, q) \in \A\}$ is the set of spels that are adjacent to $p$ with respect to $\A$.
+
+Let $\D$ be a dataset of images that have the same image domain, and let $\lambda(\I) = \{\omega_1, \omega_2, \ldots, \omega_c\}$ be the true label of an image $\I \in \D$ among $c$ classes.
+A dataset $\D$ is \textit{supervised} if all image $\I \in \D$ is inspected by an expert and she agrees with $\lambda(\I)$.
+A dataset $\D$ is \textit{unsupervised} if $\lambda(\I) = \emptyset$ for all $\I \in \D$, that is all true labels are unkown.
+If there exist images with known and unkown true labels, we say that dataset $\D$ is \textit{semi-supervised}.
+
+For a spel $p \in D$ and an adjacency relation $\A$, a \textit{locally expanded feature vector} is given by \[\E(p) = \lI(q_1) \oplus \lI(q_2) \oplus \ldots \oplus \lI(q_{|\A|})\] such as $q_i \in \A(p)$, for $i = 1, \ldots, |\A|$.
+Observe that for two different spels $p, q \in D$, it might be the case that $|\A(p)| \neq |\A(q)|$.
+
+%resolver a definição de vetor extendido localmente para assegurar que todos os vetores tenham a mesma quantidade de dimensões
+
+A multidimensioal and multi-band filter $\F$ is a pair $(\A, \mF)$ such that $\A \in D \times D$ is a adjacency relation and for all $(p, q) \in \A$, $\mF(p, q)$ assings a vector $(F_1(p, q), F_2(p, q), \ldots, F_b(p, q)) \in \R^b$ with $b \ge 1$ that satisfies the constraint that $\mF(p, q) = \mF(p, q - p)$.
+We can interpret $\F$ as a moving window that can be placed at any position $p \in D$.
+We say that $\bF = \{\F_1, \F_2, \ldots, \F_l\}$ is a \textit{filter bank} if for any $\F_i = (\A_i, \mF_i) \in \bF$ and $\F_j = (\A_j, \mF_j) \in \bF$, $\A_i = \A_j$, that is, all filters from $\bF$ has the same adjacency relation.
+
+Now, we define the convolution operation.
+The convolution between an image $\I = (D, \lI)$ and a filter bank $\bF$ with $m$ filters produces a image $\J = (\pD, \lJ)$, in which, for $p \in \pD$, $\lJ(p) = (J_1(p), J_2(p), \ldots, J_m(p))$, and \[J_i(p) = \sum_{q \in \A(p)}\lI(q) \cdot {\F_i(q - p)},\] for $m = 1, 2,  \ldots, m$.
+
+In convolution neural networks, the filters has the purpose of detecting relevant features.
+This happens by producing an positive value for spels which have those feaures and negative values for the ones which does not have them.
+This information is handle by the activation function.
+An activation function is a pair $\alpha$ that takes an image $\I = (D, \lI)$ as input and produce another image $\hat{A} = (D, \m{A})$ as output.
+In this work, we use the \textit{rectified linear unit}, which can be defined as $A_l(p) = \max(0, I_j(p))$, as activation function.
+
+Pooling has the purpose of aggregating nearby information and make convolutinal neural networks less sensible to small variations in data.
+In this work, we use only max pooling which is a operation that takes an image $\I = (D, \lI)$ with $b$ bands as input and produces an image $\hat{P} = (\pD, \m(P))$ where, for $p \in \pD$, ${\m{P}} (p) = (P_1(p), P_2(p), \ldots, P_b(p))$ such that \[P_i(p) = \max_{q \in \A(p)}{I_i(q)},\] for $i = 1, 2, \ldots, b$, and an adjacency relation $\A$.
+
+When computing the convolution between each image of a dataset $\D$ and each filter of a filter bank, followed by ReLU activation,
+at each position $p\in D$, points $\E(p)$ from the
+images $\I \in \D$ may be detected at the positive (or
+negative) side of a hyperplane with normal vector $\eF_l$ at the origin of $\R^{b \times |\A|}$. The distance
+between $\E(p)$ and that hyperplane is proportional to $J_l(p)
+=\I(p) \cdot \eF_l$, for $l \in
+\{1,2,\ldots,m\}$.
+
+One can then imagine that whenever local patterns $\E(p)$ related to images from different categories appear as
+distinct groups of points in $\R^{b\times |\A|}$, they should
+be at different sides of the hyperplanes of the filter bank, so
+different filters will be able to capture relevant details of distinct
+classes. This is precisely the role of the bias in convolutional
+neural networks. However, if the center of the points $\E(p)$ is at the origin of $\R^{b\times |\A|}$, the bias
+computation becomes unnecessary.
+
+Batch normalization is the operation that places the origin of
+$\R^{b\times |\A|}$ at the center of the cloud of points
+$\E(p)$, for $p\in D$, from all images $\hat{I}\in
+\D$, and then makes those clouds of points for all $p\in D$ so
+hyperspherical as possible. This can be done independently, per
+position and band, by transforming each image $\I=(D_I,\lI)$
+into an image $\hat{N}=(D_I,{\bf N})$, ${\bf
+  N}(p)=(N_1(p),N_2(p),\ldots,N_b(p))\in \R^b$, where
+\begin{eqnarray}
+  N_j(p) & = & \frac{I_j(p) - \mu_j(p)}{\sigma_j(p)} \\
+  \mu_j(p) & = & \frac{1}{|\D|} \sum_{\forall \hat{I}\in \D} I_{j}(p) \\
+  \sigma^2_j(p) & = & \frac{1}{|\D|} \sum_{\forall \hat{I}\in \D} \left(I_{j}(p)-\mu_j(p)\right)^2
+\end{eqnarray}
+
+A convolutional neural network may consist of a sequence of
+convolutional layers followed by a pattern classifier. The
+convolutional layers extract features from each input image $\I$
+and create a global feature vector $\gN$ at the output of the
+extra batch normalization after the last convolutional layer. The
+classifier may be a linear Support Vector Machine (SVM) or a
+Multi-Layer Perceptron (MLP). When using a MLP, the feature space of $\gN$ has usually
+its dimensions reduced by a sequence of fully connected layers before
+the decision layer.
+
+For instance, assume that ${\gN}={\gN}^{(0)}\in \R^{m\times |D|}$
+and let ${\W}^{(1)}_i \in \R^{m\times |D|}$ be the synaptic weight
+vector of each neuron $i=1,2,\ldots,k_1$ in a first fully connected
+layer with $k_1$ neurons. The result ${\gN}^{(1)}=(N^{(1)}_1,N^{(1)}_2,\ldots,N^{(1)}_{k_1})\in \R^{k_1}$, as
+$k_1 = m\times |D|$, is obtained by $k_1$ doct products with ${\W}^{(1)}_i$, $i=1,2,\ldots,k_1$, followed by activation values, such that
+\begin{eqnarray}
+  N^{(1)}_i & = & \max\{0,\langle {\gN}^{(0)}, {\W}^{(1)}_i\rangle\}.
+\end{eqnarray}
+
+For a sequence of $l$ fully connected layers with $k_1,  k_2, \ldots, k_l$ neurons per layer, the above equation becomes recursive 
+\begin{eqnarray}
+  N^{(l)}_i & = & \max\{0,\langle {\gN}^{(l-1)}, {\W}^{(l)}_i\rangle\}, 
+\end{eqnarray}
+$i=1,2,\ldots,k_l$, creating a final feature vector ${\gN}^{(l)}\in
+\R^{k_l}$ for classification by a decision layer.
+
+For $c$ categories $\omega_1, \omega_2, \ldots, \omega_c$, the
+decision layer consists of $c$ neurons with synaptic weights ${\W}_i\in \R^{k_l}$, $i=1,2,\ldots,c$, such that the input image
+$\I$ is assigned to class $\omega_j$, $j\in \{1,2,\ldots,c\}$,
+whenever
+\begin{eqnarray}
+  a_i & = & \max\{0,\langle {\gN}^{(L)}, {\W}_i\rangle\},\\ 
+  a_j & = & \max_{i=1,2,\ldots,c}\{ a_i \}.
+\end{eqnarray}
+
+Given an image $\I = (D, \lI)$ of a dataset $\D$ such that each image has some category $\omega \in \{1, 2, \ldots, c\}$, a \textit{marker} is a pair $(p, \omega)$ where $p \in D$.
+We denote by $M(\I)$ a set of markers for image $\I$.
+For a marker $m = (p, \omega) \in M(\I)$, a \textit{patch} is the locally expanded feauture vector of $p$ with respect to some adjacency relation $\A$.
+We denote by $P(\I, M, \A)$ the set of patches of an image $\I$ with respect to a set of markers and an adjacency relation.
+
+\section{Literature review}
+
+\section{The method}
+Our method is based on the idea that the network designer is an expert in the application domain or knows well enough to indicate which are the relevant characteristics of the objects of interest.
+The designer highlights these regions by placing markers.
+For each marker, the extended local feature vector is taken.
+These vectors are clustered and the centers of the clusterizes are used as convolutional layer filters.
+
+Let $\D$ be a dataset of images.
+The designer puts markers on only a subset of the images, let's say $\pDD \subseteq \D$.
+That is, she produces for each $\I \in \pDD$ a set of markers $\M(\I)$.
+This subset of images, using the information from the markers, is used to create a set of patches \[\Pa(\pDD) = \bigcup_{\I \in \pDD}{P(\I, M, \A)},\] with respect to an adjacency relationship also defined by the designer.
+In this work, we use only circular adjacency relationships.
+
+Given the set of patches $\Pa(\pDD)$ of the subset of images $\pDD$, we use a clustering algorithm to discover groups of $\Pa(\pDD)$.
+Each group represents a pattern (or set of patterns) highlighted by the designer through the markers.
+Since the center of each cluster must be a good representative of the characteristics of the group, we can use the centroid as filters to recognize these patterns in the images of $\D$.
+
+However, for it to work properly, patches and images need to be centered at the origin.
+For this, we use the information from the patches to calculate the mean and standard deviation per band.
+If the patches are representative of images of $\pDD$ the mean and standard deviation (per band) of $\Pa$ are very close to the mean and standard deviation (per band) of $\pDD$.
+
+Observe que a convolução em um spel $p \in D$ corresponde ao produto interno do vetor de carterísticas locais extendido e o vetor que representa o filtro.
+Dessa forma, seja $H$ um híperplano ortogonal ao vetor de uma filtro $\mF$.
+Esse hyperplano divide o espaço em três partes: os pontos de um lado do hyperplano, os pontos do outro lado e os pontos no hyperplano.
+Vetores parecidos com o vetor do filtro $\mF$ devem cair do mesmo lado do hiperplano que $\mF$.
+Ou seja, se o vetor $\E(p)$ está do mesmo lado de $H$ que $\mF$, entao o produto interno entre $\E(p)$ e $\mF$ é positivo, e é negativo caso contrário.
+
+\begin{figure}[t]
+  \begin{center}
+     \includegraphics[width=0.8\linewidth]{figures/tile0705-markers.png}
+  \end{center}
+     \caption{Example of markers.}
+  \label{fig:ex-markers}
+\end{figure}
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.37\textwidth}
+      \includegraphics[width=\textwidth]{figures/probabilidade-fratura.png}
+      \caption{Probability map for fractures.}
+      \label{fig:prob-fractures}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{0.37\textwidth}
+      \includegraphics[width=\textwidth]{figures/probabilidade-rocha.png}
+      \caption{Probability map for rocks.}
+    \label{fig:prob-rocks}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{0.37\textwidth}
+      \includegraphics[width=\textwidth]{figures/probabilidade-vegetacao.png}
+      \caption{Probability map for vegetation.}
+  \label{fig:prob-vegetation}
+  \end{subfigure}
+
+  \caption{Probability maps for the classes.}
+  \label{fig:prob_map}
+\end{figure}
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.37\textwidth}
+      \includegraphics[width=\textwidth]{figures/fratura.png}
+      \caption{Segmentation of fractures.}
+      \label{fig:prob-fractures}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{0.37\textwidth}
+      \includegraphics[width=\textwidth]{figures/rocha.png}
+      \caption{Segmentation of rocks.}
+    \label{fig:prob-rocks}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{0.37\textwidth}
+      \includegraphics[width=\textwidth]{figures/vegetacao.png}
+      \caption{Segmentation of vegetation vegetation.}
+  \label{fig:prob-vegetation}
+  \end{subfigure}
+
+  \caption{Segemetation of classes.}
+  \label{fig:prob_map}
+\end{figure}
+
+\section{Experiments}
+
+\section{Conclusions and future work}
+
 {\small

### 2020-03-12 egpaper_for_review.tex
describe method

@@
 
-Observe que a convolução em um spel $p \in D$ corresponde ao produto interno do vetor de carterísticas locais extendido e o vetor que representa o filtro.
-Dessa forma, seja $H$ um híperplano ortogonal ao vetor de uma filtro $\mF$.
-Esse hyperplano divide o espaço em três partes: os pontos de um lado do hyperplano, os pontos do outro lado e os pontos no hyperplano.
-Vetores parecidos com o vetor do filtro $\mF$ devem cair do mesmo lado do hiperplano que $\mF$.
-Ou seja, se o vetor $\E(p)$ está do mesmo lado de $H$ que $\mF$, entao o produto interno entre $\E(p)$ e $\mF$ é positivo, e é negativo caso contrário.
+Note that the convolution in a spel $p \in D$ corresponds to the doct product of the locally extended feature vecture and the vector that represents the filter.
+Thus, let $H$ be a hyperplane orthogonal to the vector of a $\mF$ filter.
+Thus, let $ H $ be a hyperplane orthogonal to the vector of a $ \mF$ filter.
+Vectors similar to the vector of filter $\mF$ must fall on the same side of the hyperplane as $\mF$.
+That is, if the vector $\E(p)$ is on the same side of $H$ as $\mF$, then the doct product between $\E (p)$ and $\mF$ is positive, and is negative otherwise.
+Thus, the doct product between the center of a cluster and each of the vectors of that cluster must be positive, which implies that the center is a good filter to identify the pattern described by that cluster.
+In the Figure (fazer a figura) we can see an example.
+
+Since all the patches used to produce the filters are centered around the origin, as well as the images of $\D$, we have that the hyperplanes ortogonal to the filters contain the origin, therefor there is no need for bias.
 

### 2020-03-13 texto.tex
Updates from Overleaf

--- 2020-03-12 egpaper_for_review.tex
@@
 \D$, and then makes those clouds of points for all $p\in D$ so
-hyperspherical as possible. This can be done independently, per
+hyperspherical as possible \cite{ioffe2015batch}. This can be done independently, per
 position and band, by transforming each image $\I=(D_I,\lI)$
@@
 
-\section{Literature review}
+\section{Related work}
 
@@
 
+Each convolutional layer is trained individually, one layer at a time.
+After the convolution operation, we apply the ReLU function to eliminate negative activations and the max pooling operation to aggregate local information.
+We preserve the output dimensions of convolutional layers so that we can use the same markers used in the previous layer.
+The new patches are generated from the convolutional layer output and the whole process is repeated to find the filters of the next layer.
+
 \begin{figure}[t]
@@
 
-\begin{figure}[t]
+To exemplify how our method works, we apply it to the problem of segmenting rock images into vegetation, rock and fractures.
+In Figure \ref{fig:ex-markers} we can see an example of the markers added by the network designer in regions that she considers relevant to characterize each of the classes.
+
+\begin{figure*}[t]
   \centering
@@
   \label{fig:prob_map}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[t]
+We designed a neural network with two convolutional layers and used the features extracted by it to produce the probability maps that we can see in Figure \ref{fig:prob_map}, one for each class.
+
+\begin{figure*}[t]
   \centering
@@
     \caption{Segmentation of fractures.}
-      \label{fig:prob-fractures}
+    \label{fig:segm-fractures}
   \end{subfigure}
@@
     \caption{Segmentation of rocks.}
-    \label{fig:prob-rocks}
+    \label{fig:segm-rocks}
   \end{subfigure}
@@
     \caption{Segmentation of vegetation vegetation.}
-  \label{fig:prob-vegetation}
+  \label{fig:segm-vegetation}
 \end{subfigure}
@@
 \caption{Segemetation of classes.}
-  \label{fig:prob_map}
-\end{figure}
+\label{fig:segm}
+\end{figure*}
+
+The probability maps were used to produce the segmentations that we can see in Figure \ ref {fig: segm}. For simplify viewing, each image shows the segmentation of one class.
 
 \section{Experiments}
+To evaluate our method in a quantitative way, we need a dataset with a ground truth and unfortunately the rock image dataset was not properly labeled.
+
+So, we chose to use a dataset formed from patches of coconut images.
+The dataset consists of images of dimensions $90 \times 90$ of coconut and non-coconut.
+The latter class having a large variability.
+The problem of interest is to classify the images in one of the two classes.
+
+As in reality, labeling images is a costly and time-consuming process, we assume that we only knew the true labels of a small set of samples (one hundred images for each class, totaling two hundred images).
+That way, we would have only those two hundred images to train any models that we would use to try to solve the problem.
+
+To reduce the impact of chance on the results, we randomly selected three sets of 200 images for training and three sets of 11387 images for testing.
+We wanted to represent a scenario where the network designer did not have much effort to annotate all these images.
+
+From each of these sets of training images, we selected 4 images using the 2D space projection of the 200 images using the tSNE.
+We tried to select images that were part of more cohesive groups.
+Thus, these images are more likely to have characteristics that represent their group.
+
+We placed markers on these images in regions that we consider important to identify the coconut and non-coconut (more diverse) classes.
+An example of a coconut tree image with markers can be seen in the Figure.
+
+With the markers, we created a convolutional layer with filters of dimensions $7 \times 7$.
+We did a padding to preserve the dimensions of the images.
+The network architecture can be seen in the Figure.
+We used a classifier similar to that of VGG.
+
+For clustering to find the convolutional layer filters, we used K-means to cluster the coconut tree patches in 35 clusters and the non-coconut tree patches in also 35 clusters.
+
+As the experiments were repeated three times, we reported the mean and standard deviation for each of the metrics in the test set. The results can be seen in the Table \ref{tab:results}.
+
+\begin{table*}[h]
+  \begin{center}
+  \begin{tabular}{|l|c|c|c|}
+  \hline
+   Method & Precision & Recall & F-score \\
+  \hline\hline
+  Ours & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.85 \pm 0.02$\\
+  Ours (back) & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.84 \pm 0.02$\\
+  VGG & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
+  VGG (fine tuned) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
+  \hline
+  \end{tabular}
+  \end{center}
+  \caption{Results.}
+  \label{tab:results}
+  \end{table*}
 

### 2020-03-13 egpaper_for_review.tex
add new figures

@@
 
+\begin{figure}[t]
+  \begin{center}
+     \includegraphics[width=0.8\linewidth]{figures/filter.png}
+  \end{center}
+     \caption{Filter learned from clustering.}
+  \label{fig:filter}
+\end{figure}
+
 Note that the convolution in a spel $p \in D$ corresponds to the doct product of the locally extended feature vecture and the vector that represents the filter.
@@
 Thus, the doct product between the center of a cluster and each of the vectors of that cluster must be positive, which implies that the center is a good filter to identify the pattern described by that cluster.
-In the Figure (fazer a figura) we can see an example.
+In the Figure \ref{fig:filter} we can see an example.
 
 Since all the patches used to produce the filters are centered around the origin, as well as the images of $\D$, we have that the hyperplanes ortogonal to the filters contain the origin, therefor there is no need for bias.
+
+Each convolutional layer is trained individually, one layer at a time.
+After the convolution operation, we apply the ReLU function to eliminate negative activations and the max pooling operation to aggregate local information.
+We preserve the output dimensions of convolutional layers so that we can use the same markers used in the previous layer.
+The new patches are generated from the convolutional layer output and the whole process is repeated to find the filters of the next layer.
 
@@
 
-\begin{figure}[t]
+To exemplify how our method works, we apply it to the problem of segmenting rock images into vegetation, rock and fractures.
+In Figure \ref{fig:ex-markers} we can see an example of the markers added by the network designer in regions that she considers relevant to characterize each of the classes.
+
+\begin{figure*}[t]
   \centering
-  \begin{subfigure}[b]{0.37\textwidth}
+  \begin{subfigure}[b]{0.33\textwidth}
       \includegraphics[width=\textwidth]{figures/probabilidade-fratura.png}
@@
   ~
-  \begin{subfigure}[b]{0.37\textwidth}
+  \begin{subfigure}[b]{0.33\textwidth}
       \includegraphics[width=\textwidth]{figures/probabilidade-rocha.png}
@@
   ~
-  \begin{subfigure}[b]{0.37\textwidth}
+  \begin{subfigure}[b]{0.33\textwidth}
       \includegraphics[width=\textwidth]{figures/probabilidade-vegetacao.png}
@@
   \label{fig:prob_map}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[t]
+We designed a neural network with two convolutional layers and used the features extracted by it to produce the probability maps that we can see in Figure \ref{fig:prob_map}, one for each class.
+
+\begin{figure*}[t]
   \centering
-  \begin{subfigure}[b]{0.37\textwidth}
+  \begin{subfigure}[b]{0.33\textwidth}
     \includegraphics[width=\textwidth]{figures/fratura.png}
     \caption{Segmentation of fractures.}
-      \label{fig:prob-fractures}
+    \label{fig:segm-fractures}
   \end{subfigure}
   ~
-  \begin{subfigure}[b]{0.37\textwidth}
+  \begin{subfigure}[b]{0.33\textwidth}
     \includegraphics[width=\textwidth]{figures/rocha.png}
     \caption{Segmentation of rocks.}
-    \label{fig:prob-rocks}
+    \label{fig:segm-rocks}
   \end{subfigure}
   ~
-  \begin{subfigure}[b]{0.37\textwidth}
+  \begin{subfigure}[b]{0.33\textwidth}
     \includegraphics[width=\textwidth]{figures/vegetacao.png}
     \caption{Segmentation of vegetation vegetation.}
-  \label{fig:prob-vegetation}
+  \label{fig:segm-vegetation}
 \end{subfigure}
@@
 \caption{Segemetation of classes.}
-  \label{fig:prob_map}
+\label{fig:segm}
+\end{figure*}
+
+The probability maps were used to produce the segmentations that we can see in Figure \ ref {fig: segm}. For simplify viewing, each image shows the segmentation of one class.
+
+\section{Experiments}
+To evaluate our method in a quantitative way, we need a dataset with a ground truth and unfortunately the rock image dataset was not properly labeled.
+
+So, we chose to use a dataset formed from patches of coconut images.
+The dataset consists of images of dimensions $90 \times 90$ of coconut and non-coconut.
+The latter class having a large variability.
+The problem of interest is to classify the images in one of the two classes.
+
+As in reality, labeling images is a costly and time-consuming process, we assume that we only knew the true labels of a small set of samples (one hundred images for each class, totaling two hundred images).
+That way, we would have only those two hundred images to train any models that we would use to try to solve the problem.
+
+To reduce the impact of chance on the results, we randomly selected three sets of 200 images for training and three sets of 11387 images for testing.
+We wanted to represent a scenario where the network designer did not have much effort to annotate all these images.
+
+From each of these sets of training images, we selected 4 images using the 2D space projection of the 200 images using the tSNE.
+We tried to select images that were part of more cohesive groups.
+Thus, these images are more likely to have characteristics that represent their group.
+
+We placed markers on these images in regions that we consider important to identify the coconut and non-coconut (more diverse) classes.
+An example of a coconut tree image with markers can be seen in Figure \ref{fig:coco-markers} and non-coconut tree image with markers in Figure \ref{fig:non-coco-markers}.
+
+\begin{figure}[h!]
+  \begin{center}
+     \includegraphics[width=0.4\linewidth]{figures/coco_markers.png}
+  \end{center}
+     \caption{Markers on an image which contains a coconut tree.}
+  \label{fig:coco-markers}
 \end{figure}
 
-\section{Experiments}
+\begin{figure}[h!]
+  \begin{center}
+     \includegraphics[width=0.4\linewidth]{figures/non_coco-markers.png}
+  \end{center}
+     \caption{Markers on an image which does not contain a coconut tree.}
+  \label{fig:non-coco-markers}
+\end{figure}
+
+With the markers, we created a convolutional layer with filters of dimensions $7 \times 7$.
+We did a padding to preserve the dimensions of the images.
+The network architecture can be seen in the Figure.
+We used a classifier similar to that of VGG.
+
+For clustering to find the convolutional layer filters, we used K-means to cluster the coconut tree patches in 35 clusters and the non-coconut tree patches in also 35 clusters.
+
+As the experiments were repeated three times, we reported the mean and standard deviation for each of the metrics in the test set. The results can be seen in the Table \ref{tab:results}.
+
+\begin{table*}[h]
+  \begin{center}
+  \begin{tabular}{|l|c|c|c|}
+  \hline
+   Method & Precision & Recall & F-score \\
+  \hline\hline
+  Ours & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.85 \pm 0.02$\\
+  Ours (back) & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.84 \pm 0.02$\\
+  VGG & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
+  VGG (fine tuned) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
+  \hline
+  \end{tabular}
+  \end{center}
+  \caption{Results.}
+  \label{tab:results}
+  \end{table*}
 

### 2020-03-14 texto.tex
Updates from Overleaf

@@
 
+\section{Related Work}
+
 \section{Preliminaries}
@@
 In this work, we focus on images with $d=2$.
-The set of local feature vectors is called \textit{local feature space}.
+The local feature vectors $\m{I}(p)$, of all $p\in D$, may be thought of as points of a \textit{local feature space} $\Re^b$.
 
@@
 
-A multidimensioal and multi-band filter $\F$ is a pair $(\A, \mF)$ such that $\A \in D \times D$ is a adjacency relation and for all $(p, q) \in \A$, $\mF(p, q)$ assings a vector $(F_1(p, q), F_2(p, q), \ldots, F_b(p, q)) \in \R^b$ with $b \ge 1$ that satisfies the constraint that $\mF(p, q) = \mF(p, q - p)$.
+A multidimensioal and multi-band filter $\F$ is a pair $(\A, \mF)$ such that $\A \in D \times D$ is a adjacency relation and for all $(p, q) \in \A$, $\mF(p, q)$ assings a vector $(F_1(p, q), F_2(p, q), \ldots, F_b(p, q)) \in \R^b$ with $b \ge 1$ that satisfies the constraint that $\mF(p, q) = \mF(q - p)$.
 We can interpret $\F$ as a moving window that can be placed at any position $p \in D$.
-We say that $\bF = \{\F_1, \F_2, \ldots, \F_l\}$ is a \textit{filter bank} if for any $\F_i = (\A_i, \mF_i) \in \bF$ and $\F_j = (\A_j, \mF_j) \in \bF$, $\A_i = \A_j$, that is, all filters from $\bF$ has the same adjacency relation.
+We say that $\bF = \{\F_1, \F_2, \ldots, \F_l\}$ is a \textit{filter bank} if for any $\F = (\A, \mF) \in \bF$ and $\F' = (\A', \mF') \in \bF$, $\A = \A'$, that is, all filters from $\bF$ has the same adjacency relation.
 
 Now, we define the convolution operation.
-The convolution between an image $\I = (D, \lI)$ and a filter bank $\bF$ with $m$ filters produces a image $\J = (\pD, \lJ)$, in which, for $p \in \pD$, $\lJ(p) = (J_1(p), J_2(p), \ldots, J_m(p))$, and \[J_i(p) = \sum_{q \in \A(p)}\lI(q) \cdot {\F_i(q - p)},\] for $m = 1, 2,  \ldots, m$.
+The convolution between an image $\I = (D, \lI)$ and a filter bank $\bF$ with $m$ filters produces a image $\J = (\pD, \lJ)$, in which, for $p \in \pD\subseteq D$, $\lJ(p) = (J_1(p), J_2(p), \ldots, J_m(p))$, and \[J_i(p) = \sum_{q \in \A(p)}\lI(q) \cdot {{\bf F}_i(q - p)},\] for $i = 1, 2,  \ldots, m$.
+For the sake of simplicity, we will assume that $\pD=D$.
 
 In convolution neural networks, the filters has the purpose of detecting relevant features.
-This happens by producing an positive value for spels which have those feaures and negative values for the ones which does not have them.
-This information is handle by the activation function.
-An activation function is a pair $\alpha$ that takes an image $\I = (D, \lI)$ as input and produce another image $\hat{A} = (D, \m{A})$ as output.
-In this work, we use the \textit{rectified linear unit}, which can be defined as $A_l(p) = \max(0, I_j(p))$, as activation function.
+This happens by producing an positive value for spels which have those features and negative values for the ones which does not have them.
+This information is handled by the activation function, which takes an image $\J = (D, \lJ)$ as input and produces another image $\hat{A} = (D, \m{A})$ as output.
+In this work, we use the \textit{rectified linear unit}, which the elements of $\m{A}$ are defined as $A_i(p) = \max(0, J_i(p))$, $i=1,2,\ldots,m$, as activation function.
 
-Pooling has the purpose of aggregating nearby information and make convolutinal neural networks less sensible to small variations in data.
-In this work, we use only max pooling which is a operation that takes an image $\I = (D, \lI)$ with $b$ bands as input and produces an image $\hat{P} = (\pD, \m(P))$ where, for $p \in \pD$, ${\m{P}} (p) = (P_1(p), P_2(p), \ldots, P_b(p))$ such that \[P_i(p) = \max_{q \in \A(p)}{I_i(q)},\] for $i = 1, 2, \ldots, b$, and an adjacency relation $\A$.
+Pooling has the purpose of aggregating nearby information and makes convolutional neural networks less sensible to small variations in data.
+In this work, we use only max pooling which is an operation that takes an image $\J = (D, \lJ)$ with $m$ bands as input and produces an image $\hat{P} = (\pD, \m{P})$ where, for $p \in \pD\subseteq D$, ${\m{P}} (p) = (P_1(p), P_2(p), \ldots, P_m(p))$ such that \[P_i(p) = \max_{q \in \A(p)}\{I_i(q)\},\] for $i = 1, 2, \ldots, m$, and an adjacency relation $\A$.
+For the sake of simplicity, we will assume that $\pD=D$.
 
@@
 images $\I \in \D$ may be detected at the positive (or
-negative) side of a hyperplane with normal vector $\eF_l$ at the origin of $\R^{b \times |\A|}$. The distance
-between $\E(p)$ and that hyperplane is proportional to $J_l(p)
-=\I(p) \cdot \eF_l$, for $l \in
+negative) side of a hyperplane with normal vector $\mF_i$ at the origin of $\R^{b \times |\A|}$. The distance
+between $\E(p)$ and that hyperplane is proportional to $J_i(p)
+=\E(p) \cdot \eF_i$, for $i \in
 \{1,2,\ldots,m\}$.
@@
 $\R^{b\times |\A|}$ at the center of the cloud of points
-$\E(p)$, for $p\in D$, from all images $\hat{I}\in
+$\E(p)$, for a given $p\in D$, from all images $\hat{I}\in
 \D$, and then makes those clouds of points for all $p\in D$ so
 hyperspherical as possible \cite{ioffe2015batch}. This can be done independently, per
-position and band, by transforming each image $\I=(D_I,\lI)$
-into an image $\hat{N}=(D_I,{\bf N})$, ${\bf
-  N}(p)=(N_1(p),N_2(p),\ldots,N_b(p))\in \R^b$, where
+position and band, by transforming each image $\I=(D,\lI)$
+into an image $\hat{N}=(D,{\bf N})$, ${\bf
+  N}(p)=(N_1(p),N_2(p),\ldots,N_b(p))\in \R^b$, where, for $i\in \{1,2,\ldots,b\}$,
 \begin{eqnarray}
-  N_j(p) & = & \frac{I_j(p) - \mu_j(p)}{\sigma_j(p)} \\
-  \mu_j(p) & = & \frac{1}{|\D|} \sum_{\forall \hat{I}\in \D} I_{j}(p) \\
-  \sigma^2_j(p) & = & \frac{1}{|\D|} \sum_{\forall \hat{I}\in \D} \left(I_{j}(p)-\mu_j(p)\right)^2
+  N_i(p) & = & \frac{I_i(p) - \mu_i(p)}{\sigma_i(p)} \\
+  \mu_i(p) & = & \frac{1}{|\D|} \sum_{\forall \hat{I}\in \D} I_{i}(p) \\
+  \sigma^2_i(p) & = & \frac{1}{|\D|} \sum_{\forall \hat{I}\in \D} \left(I_{i}(p)-\mu_i(p)\right)^2
 \end{eqnarray}
@@
 convolutional layers extract features from each input image $\I$
-and create a global feature vector $\gN$ at the output of the
+and create a global feature vector $\gN$ at the output of an
 extra batch normalization after the last convolutional layer. The
@@
 layer with $k_1$ neurons. The result ${\gN}^{(1)}=(N^{(1)}_1,N^{(1)}_2,\ldots,N^{(1)}_{k_1})\in \R^{k_1}$, as
-$k_1 = m\times |D|$, is obtained by $k_1$ doct products with ${\W}^{(1)}_i$, $i=1,2,\ldots,k_1$, followed by activation values, such that
+$k_1 = m\times |D|$, is obtained by $k_1$ inner products with ${\W}^{(1)}_i$, $i=1,2,\ldots,k_1$, followed by activation values, such that
 \begin{eqnarray}
@@
 Given an image $\I = (D, \lI)$ of a dataset $\D$ such that each image has some category $\omega \in \{1, 2, \ldots, c\}$, a \textit{marker} is a pair $(p, \omega)$ where $p \in D$.
-We denote by $M(\I)$ a set of markers for image $\I$.
-For a marker $m = (p, \omega) \in M(\I)$, a \textit{patch} is the locally expanded feauture vector of $p$ with respect to some adjacency relation $\A$.
-We denote by $P(\I, M, \A)$ the set of patches of an image $\I$ with respect to a set of markers and an adjacency relation.
+We denote by $\M(\I)$ a set of markers for image $\I$.
+For a marker $m = (p, \omega) \in \M(\I)$, a \textit{patch} is the locally expanded feature vector $\E(p)$ of $p$ with respect to some adjacency relation $\A$.
+We denote by $P(\I, \M, \A)$ the set of patches of an image $\I$ with respect to a set of markers and an adjacency relation.
 
-\section{Related work}
-
-\section{The method}
+\section{Learning filters from markers}
 Our method is based on the idea that the network designer is an expert in the application domain or knows well enough to indicate which are the relevant characteristics of the objects of interest.
 The designer highlights these regions by placing markers.
-For each marker, the extended local feature vector is taken.
-These vectors are clustered and the centers of the clusterizes are used as convolutional layer filters.
+For each marker, the extended local feature vectors of its spels are taken.
+These vectors are clustered and the centers of those groups are used as filters for the current convolutional layer.
+Since each convolutional layer transforms the input feaure space of the images into a new one, the same procedure can be repeated to create the filters of the subsequents convolutional layers.
+This way, it is enough to specify the number of desired convolutional layers and their hyperparameters.
 
 Let $\D$ be a dataset of images.
-The designer puts markers on only a subset of the images, let's say $\pDD \subseteq \D$.
+The designer draws markers on only a subset of the images, let's say $\pDD \subseteq \D$.
 That is, she produces for each $\I \in \pDD$ a set of markers $\M(\I)$.
-This subset of images, using the information from the markers, is used to create a set of patches \[\Pa(\pDD) = \bigcup_{\I \in \pDD}{P(\I, M, \A)},\] with respect to an adjacency relationship also defined by the designer.
-In this work, we use only circular adjacency relationships.
+This subset of images, using the information from the markers, is used to create a set of patches \[\Pa(\pDD) = \bigcup_{\I \in \pDD}{P(\I, \M, \A)},\] with respect to an adjacency relation also defined by the designer.
+In this work, we use only circular adjacency relations.
 
-Given the set of patches $\Pa(\pDD)$ of the subset of images $\pDD$, we use a clustering algorithm to discover groups of $\Pa(\pDD)$.
-Each group represents a pattern (or set of patterns) highlighted by the designer through the markers.
-Since the center of each cluster must be a good representative of the characteristics of the group, we can use the centroid as filters to recognize these patterns in the images of $\D$.
+Given the set of patches $\Pa(\pDD)$ of the subset of images $\pDD$, we use a clustering algorithm to discover groups in $\Pa(\pDD)$.
+Each group represents local patterns from one or more classes, highlighted by the designer through the markers.
+Since the center of each cluster must be a good representative of the characteristics of the group, we can use the group centroids as filters to enhance those patterns in the images of $\D$.
 
@@
 For this, we use the information from the patches to calculate the mean and standard deviation per band.
-If the patches are representative of images of $\pDD$ the mean and standard deviation (per band) of $\Pa$ are very close to the mean and standard deviation (per band) of $\pDD$.
+If the patches are representative of images in $\D$, the mean and standard deviation (per band) of $\Pa$ are very close to the mean and standard deviation (per band) in $\D$.
+
+\begin{figure}[t]
+  \begin{center}
+     \includegraphics[width=0.8\linewidth]{figures/filter.png}
+  \end{center}
+     \caption{Filter learned from clustering.}
+  \label{fig:filter}
+\end{figure}
 
@@
 Thus, the doct product between the center of a cluster and each of the vectors of that cluster must be positive, which implies that the center is a good filter to identify the pattern described by that cluster.
-In the Figure (fazer a figura) we can see an example.
+In the Figure \ref{fig:filter} we can see an example.
 
@@
   \centering
-  \begin{subfigure}[b]{0.37\textwidth}
+  \begin{subfigure}[b]{0.33\textwidth}
       \includegraphics[width=\textwidth]{figures/probabilidade-fratura.png}
@@
   \centering
-  \begin{subfigure}[b]{0.37\textwidth}
+  \begin{subfigure}[b]{0.33\textwidth}
     \includegraphics[width=\textwidth]{figures/fratura.png}
@@
   ~
-  \begin{subfigure}[b]{0.37\textwidth}
+  \begin{subfigure}[b]{0.33\textwidth}
     \includegraphics[width=\textwidth]{figures/rocha.png}
@@
   ~
-  \begin{subfigure}[b]{0.37\textwidth}
+  \begin{subfigure}[b]{0.33\textwidth}
     \includegraphics[width=\textwidth]{figures/vegetacao.png}
@@
 We placed markers on these images in regions that we consider important to identify the coconut and non-coconut (more diverse) classes.
-An example of a coconut tree image with markers can be seen in the Figure.
+An example of a coconut tree image with markers can be seen in Figure \ref{fig:coco-markers} and non-coconut tree image with markers in Figure \ref{fig:non-coco-markers}.
+
+\begin{figure}[h!]
+  \begin{center}
+     \includegraphics[width=0.4\linewidth]{figures/coco_markers.png}
+  \end{center}
+     \caption{Markers on an image which contains a coconut tree.}
+  \label{fig:coco-markers}
+\end{figure}
+
+\begin{figure}[h!]
+  \begin{center}
+     \includegraphics[width=0.4\linewidth]{figures/non_coco-markers.png}
+  \end{center}
+     \caption{Markers on an image which does not contain a coconut tree.}
+  \label{fig:non-coco-markers}
+\end{figure}
 

### 2020-05-28 texto.tex
Updates from Overleaf

@@
-\documentclass[10pt,twocolumn,letterpaper]{article}
+\documentclass[a4paper,conference]{IEEEtran}
 
-\usepackage{cvpr}
 \usepackage{times}
@@
 The network architecture can be seen in the Figure.
-We used a classifier similar to that of VGG.
+We used a classifier similar to that of VGG-16.
 
@@
   Ours & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.85 \pm 0.02$\\
-  Ours (back) & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.84 \pm 0.02$\\
-  VGG & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
-  VGG (fine tuned) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
+  Ours (fine tuned) & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.84 \pm 0.02$\\
+  VGG-16 & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
+  VGG-16 (fine tuned) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
   \hline

### 2020-05-28 texto.tex
Updates from Overleaf

@@
 \usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
+
+\usepackage{comment}
 
@@
 \newcommand{\pDD}{\D^\prime}
-% \cvprfinalcopy % *** Uncomment this line for the final submission
 
-\def\cvprPaperID{****} % *** Enter the EV Paper ID here
-\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
-
-% Pages are numbered in submission mode, and unnumbered in camera-ready
-\ifcvprfinal\pagestyle{empty}\fi
 \begin{document}
@@
 \author{Italos Estilon de Souza\\
-Institute of Computing - University of Campinas (UNICAMP)\\
-Institution1 address\\
 {\tt\small italos.souza@ic.unicamp.br}
@@
 \section{Preliminaries}
-
+\begin{comment}
 Let $D \subset \Z^d$ for $d > 1$ and let $\m{I}: D \to \R^b$, with $b \ge 1$, be a function that assings to each $p = (x_p, y_q) \in D$ a vector $\m{I}(p) = (I_1(p), I_2(p), \ldots, I_b(p))$.
@@
 We denote by $P(\I, \M, \A)$ the set of patches of an image $\I$ with respect to a set of markers and an adjacency relation.
+\end{comment}
 

### 2020-06-01 main.tex
add introduction

--- 2020-05-28 texto.tex
@@
+%% bare_conf.tex
+%% V1.4b
+%% 2015/08/26
+%% by Michael Shell
+%% See:
+%% http://www.michaelshell.org/
+%% for current contact information.
+%%
+%% This is a skeleton file demonstrating the use of IEEEtran.cls
+%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
+%% conference paper.
+%%
+%% Support sites:
+%% http://www.michaelshell.org/tex/ieeetran/
+%% http://www.ctan.org/pkg/ieeetran
+%% and
+%% http://www.ieee.org/
+%%*************************************************************************
+%% Legal Notice:
+%% This code is offered as-is without any warranty either expressed or
+%% implied; without even the implied warranty of MERCHANTABILITY or
+%% FITNESS FOR A PARTICULAR PURPOSE!
+%% User assumes all risk.
+%% In no event shall the IEEE or any contributor to this code be liable for
+%% any damages or losses, including, but not limited to, incidental,
+%% consequential, or any other damages, resulting from the use or misuse
+%% of any information contained here.
+%%
+%% All comments are the opinions of their respective authors and are not
+%% necessarily endorsed by the IEEE.
+%%
+%% This work is distributed under the LaTeX Project Public License (LPPL)
+%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
+%% distributed and modified. A copy of the LPPL, version 1.3, is included
+%% in the base LaTeX documentation of all distributions of LaTeX released
+%% 2003/12/01 or later.
+%% Retain all contribution notices and credits.
+%% ** Modified files should be clearly indicated as such, including  **
+%% ** renaming them and changing author support contact information. **
+%%*************************************************************************
+
 \documentclass[a4paper,conference]{IEEEtran}
 
-\usepackage{times}
-\usepackage{epsfig}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{amssymb}
+\def\citepunt{,}
 
-% Include other packages here, before hyperref.
+\usepackage[pdftex]{graphicx}
+\ifCLASSOPTIONcompsoc
+  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
+\else
+  \usepackage[caption=false,font=footnotesize]{subfig}
+\fi
+\usepackage{comment}
+\usepackage{float}
+\usepackage{mathtools}
+\usepackage{amsfonts}
 
-\usepackage{subcaption}
-\usepackage{siunitx}
-
-% If you comment hyperref and then uncomment it, you should delete
-% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
-% run, let it finish, and you should be clear).
-\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
-
-\usepackage{comment}
+\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
+\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
 
@@
 
+% correct bad hyphenation here
+\hyphenation{op-tical net-works semi-conduc-tor}
+
 \begin{document}
 
-%%%%%%%%% TITLE
-\title{Learning filters from markers}
-\author{Italos Estilon de Souza\\
-{\tt\small italos.souza@ic.unicamp.br}
-% For a paper whose authors are all at the same institution,
-% omit the following lines up until the closing ``}''.
-% Additional authors and addresses can be added with ``\and'',
-% just like the second author.
-% To save space, use either the email address or home page, not both
-\and
-Alexandre Falcão\\
-{\tt\small afalcao@ic.unicamp.br}
-}
+\title{Learning filters from user provided image makers}
 
+% make the title area
 \maketitle
-%\thispagestyle{empty}
 
-%%%%%%%%% ABSTRACT
+\begin{abstract}
 
-%%%%%%%%% BODY TEXT
+\end{abstract}
+
 \section{Introduction}
+Deep learning has proven to be applicable to different types of tasks, from image classification to the creation of synthetic data \cite{goodfellow2016deep}.
+Given the success of the application of deep learning in various areas of knowledge, this technique has also been applied to several tasks of remote sensing.
+Examples of this technique exist in various problems such as segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic, hamaguchi2018effective}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}.
 
-\section{Related Work}
+The detection of trees is important to produce knowledge about the land use of a region, from the assessment of crop diversity or the performance of plantations in a given season to the assessment of damages after natural disasters such as drilling and flooding.
+As these plantations can span very large areas, manual detection can be costly and susceptible to errors.
 
-\section{Preliminaries}
-\begin{comment}
-Let $D \subset \Z^d$ for $d > 1$ and let $\m{I}: D \to \R^b$, with $b \ge 1$, be a function that assings to each $p = (x_p, y_q) \in D$ a vector $\m{I}(p) = (I_1(p), I_2(p), \ldots, I_b(p))$.
-A multi-band and multi-dimensional image is a pair ${\hat{I} = (D, \m{I})}$.
-We say that $D$ is the \textit{image domain}, $\m{I}(p)$ is the \textit{local feature vector} of a space element (\textit{spel}) $p \in D$, $b$ is the number of bands, and $d$ is the number of dimensions.
-In this work, we focus on images with $d=2$.
-The local feature vectors $\m{I}(p)$, of all $p\in D$, may be thought of as points of a \textit{local feature space} $\Re^b$.
+In \cite{fassnacht2016review}, we can see a review of the problem of classification of tree species of aerial images.
+\cite{puttemans2018comparing} apply deep learning to the problem of detecting coconut trees in aerial images.
+Aparta et al.
+\cite{aparna2018cnn} apply convolutional neural networks to count coconut trees on plantations in India.
+Vargas-Muñoz et al.
+\cite{8899005} propose an active learning approach to detect coconut trees that seeks to reduce the effort of annotating images.
+The \cite{8899005} approach is divided into a stage of producing candidate patches for having a coconut tree and a stage where a CNN determines whether a patch has a coconut tree.
+We propose a different way to produce this classifier.
 
-An adjacency relation $\A \subseteq D \times D$ is a binary relation between spels $p, q \in D$.
-We focus on reflexive and translation-invariant relations.
-More precisely, we focus on adjacency relations given by \[\A_{i,j} = \{ (p, q) \in D : \|x_p - x_q\| \le i, \|y_p - y_q\| \le j \}.\] We say that $\A(p) = \{q \in D : (p, q) \in \A\}$ is the set of spels that are adjacent to $p$ with respect to $\A$.
+Despite this success, methods based on deep learning require a lot of annotated images for training, and annotating this volume of data is quite costly.
+Even interactive methods require large training sets.
+If these methods depend on huge training sets, their application may not be feasible in areas where it is impossible, or the cost of gathering and annotating these data sets is prohibitive.
+Much of the user's interaction with methods based on deep learning occurs through the annotation of the training set, and the user does not have much influence on the training process.
 
-Let $\D$ be a dataset of images that have the same image domain, and let $\lambda(\I) = \{\omega_1, \omega_2, \ldots, \omega_c\}$ be the true label of an image $\I \in \D$ among $c$ classes.
-A dataset $\D$ is \textit{supervised} if all image $\I \in \D$ is inspected by an expert and she agrees with $\lambda(\I)$.
-A dataset $\D$ is \textit{unsupervised} if $\lambda(\I) = \emptyset$ for all $\I \in \D$, that is all true labels are unkown.
-If there exist images with known and unkown true labels, we say that dataset $\D$ is \textit{semi-supervised}.
+Despite the advances in deep learning over the last decades, especially with most recent deep neural networks, some important questions related to human interaction remain unanswered: (1) How to choose the most reasonable model, in the sense that it provides acceptable effectiveness, being as simple and efficient as needed, for a given classification problem?
+(2) How to train that model from a reduced number of samples?
+(3) Can the user explain the decisions of that model?
+(4) Can that model improve from human corrections?
+The first question requires to explore human knowledge about the classification technique and the problem of interest.
+The second one requires to reduce as much as possible the human effort to train a model.
+The third issue is related to human understanding, and it may explore techniques (such as data visualization) to explain the decisions of the model and perhaps to guide human decisions concerning the design of the model.
+The fourth question is also important during training, and it is related to human control over the process.
+They all lead to the importance of involving human experts during the machine learning process.
 
-For a spel $p \in D$ and an adjacency relation $\A$, a \textit{locally expanded feature vector} is given by \[\E(p) = \lI(q_1) \oplus \lI(q_2) \oplus \ldots \oplus \lI(q_{|\A|})\] such as $q_i \in \A(p)$, for $i = 1, \ldots, |\A|$.
-Observe that for two different spels $p, q \in D$, it might be the case that $|\A(p)| \neq |\A(q)|$.
+In this paper, we present some advances towards the answers to the questions (1) and (2).
+We also discuss future research related to the questions (3) and (4).
+We present a method to the design of light-weighted convolutional neural networks from markers drawn by an expert on a few images to indicate representative regions of each object of interest for classification.
+From those markers, our approach can learn the filters of each layer, and the convolutional neural network can be designed in a layer-by-layer fashion, with no need for backpropagation in order to build a representative feature space for the classification problem.
+Backpropagation is only used to train a bank of multi-layer perceptrons for local pattern classification.
 
-%resolver a definição de vetor extendido localmente para assegurar que todos os vetores tenham a mesma quantidade de dimensões
-
-A multidimensioal and multi-band filter $\F$ is a pair $(\A, \mF)$ such that $\A \in D \times D$ is a adjacency relation and for all $(p, q) \in \A$, $\mF(p, q)$ assings a vector $(F_1(p, q), F_2(p, q), \ldots, F_b(p, q)) \in \R^b$ with $b \ge 1$ that satisfies the constraint that $\mF(p, q) = \mF(q - p)$.
-We can interpret $\F$ as a moving window that can be placed at any position $p \in D$.
-We say that $\bF = \{\F_1, \F_2, \ldots, \F_l\}$ is a \textit{filter bank} if for any $\F = (\A, \mF) \in \bF$ and $\F' = (\A', \mF') \in \bF$, $\A = \A'$, that is, all filters from $\bF$ has the same adjacency relation.
-
-Now, we define the convolution operation.
-The convolution between an image $\I = (D, \lI)$ and a filter bank $\bF$ with $m$ filters produces a image $\J = (\pD, \lJ)$, in which, for $p \in \pD\subseteq D$, $\lJ(p) = (J_1(p), J_2(p), \ldots, J_m(p))$, and \[J_i(p) = \sum_{q \in \A(p)}\lI(q) \cdot {{\bf F}_i(q - p)},\] for $i = 1, 2,  \ldots, m$.
-For the sake of simplicity, we will assume that $\pD=D$.
-
-In convolution neural networks, the filters has the purpose of detecting relevant features.
-This happens by producing an positive value for spels which have those features and negative values for the ones which does not have them.
-This information is handled by the activation function, which takes an image $\J = (D, \lJ)$ as input and produces another image $\hat{A} = (D, \m{A})$ as output.
-In this work, we use the \textit{rectified linear unit}, which the elements of $\m{A}$ are defined as $A_i(p) = \max(0, J_i(p))$, $i=1,2,\ldots,m$, as activation function.
-
-Pooling has the purpose of aggregating nearby information and makes convolutional neural networks less sensible to small variations in data.
-In this work, we use only max pooling which is an operation that takes an image $\J = (D, \lJ)$ with $m$ bands as input and produces an image $\hat{P} = (\pD, \m{P})$ where, for $p \in \pD\subseteq D$, ${\m{P}} (p) = (P_1(p), P_2(p), \ldots, P_m(p))$ such that \[P_i(p) = \max_{q \in \A(p)}\{I_i(q)\},\] for $i = 1, 2, \ldots, m$, and an adjacency relation $\A$.
-For the sake of simplicity, we will assume that $\pD=D$.
-
-When computing the convolution between each image of a dataset $\D$ and each filter of a filter bank, followed by ReLU activation,
-at each position $p\in D$, points $\E(p)$ from the
-images $\I \in \D$ may be detected at the positive (or
-negative) side of a hyperplane with normal vector $\mF_i$ at the origin of $\R^{b \times |\A|}$. The distance
-between $\E(p)$ and that hyperplane is proportional to $J_i(p)
-=\E(p) \cdot \eF_i$, for $i \in
-\{1,2,\ldots,m\}$.
-
-One can then imagine that whenever local patterns $\E(p)$ related to images from different categories appear as
-distinct groups of points in $\R^{b\times |\A|}$, they should
-be at different sides of the hyperplanes of the filter bank, so
-different filters will be able to capture relevant details of distinct
-classes. This is precisely the role of the bias in convolutional
-neural networks. However, if the center of the points $\E(p)$ is at the origin of $\R^{b\times |\A|}$, the bias
-computation becomes unnecessary.
-
-Batch normalization is the operation that places the origin of
-$\R^{b\times |\A|}$ at the center of the cloud of points
-$\E(p)$, for a given $p\in D$, from all images $\hat{I}\in
-\D$, and then makes those clouds of points for all $p\in D$ so
-hyperspherical as possible \cite{ioffe2015batch}. This can be done independently, per
-position and band, by transforming each image $\I=(D,\lI)$
-into an image $\hat{N}=(D,{\bf N})$, ${\bf
-  N}(p)=(N_1(p),N_2(p),\ldots,N_b(p))\in \R^b$, where, for $i\in \{1,2,\ldots,b\}$,
-\begin{eqnarray}
-  N_i(p) & = & \frac{I_i(p) - \mu_i(p)}{\sigma_i(p)} \\
-  \mu_i(p) & = & \frac{1}{|\D|} \sum_{\forall \hat{I}\in \D} I_{i}(p) \\
-  \sigma^2_i(p) & = & \frac{1}{|\D|} \sum_{\forall \hat{I}\in \D} \left(I_{i}(p)-\mu_i(p)\right)^2
-\end{eqnarray}
-
-A convolutional neural network may consist of a sequence of
-convolutional layers followed by a pattern classifier. The
-convolutional layers extract features from each input image $\I$
-and create a global feature vector $\gN$ at the output of an
-extra batch normalization after the last convolutional layer. The
-classifier may be a linear Support Vector Machine (SVM) or a
-Multi-Layer Perceptron (MLP). When using a MLP, the feature space of $\gN$ has usually
-its dimensions reduced by a sequence of fully connected layers before
-the decision layer.
-
-For instance, assume that ${\gN}={\gN}^{(0)}\in \R^{m\times |D|}$
-and let ${\W}^{(1)}_i \in \R^{m\times |D|}$ be the synaptic weight
-vector of each neuron $i=1,2,\ldots,k_1$ in a first fully connected
-layer with $k_1$ neurons. The result ${\gN}^{(1)}=(N^{(1)}_1,N^{(1)}_2,\ldots,N^{(1)}_{k_1})\in \R^{k_1}$, as
-$k_1 = m\times |D|$, is obtained by $k_1$ inner products with ${\W}^{(1)}_i$, $i=1,2,\ldots,k_1$, followed by activation values, such that
-\begin{eqnarray}
-  N^{(1)}_i & = & \max\{0,\langle {\gN}^{(0)}, {\W}^{(1)}_i\rangle\}.
-\end{eqnarray}
-
-For a sequence of $l$ fully connected layers with $k_1,  k_2, \ldots, k_l$ neurons per layer, the above equation becomes recursive 
-\begin{eqnarray}
-  N^{(l)}_i & = & \max\{0,\langle {\gN}^{(l-1)}, {\W}^{(l)}_i\rangle\}, 
-\end{eqnarray}
-$i=1,2,\ldots,k_l$, creating a final feature vector ${\gN}^{(l)}\in
-\R^{k_l}$ for classification by a decision layer.
-
-For $c$ categories $\omega_1, \omega_2, \ldots, \omega_c$, the
-decision layer consists of $c$ neurons with synaptic weights ${\W}_i\in \R^{k_l}$, $i=1,2,\ldots,c$, such that the input image
-$\I$ is assigned to class $\omega_j$, $j\in \{1,2,\ldots,c\}$,
-whenever
-\begin{eqnarray}
-  a_i & = & \max\{0,\langle {\gN}^{(L)}, {\W}_i\rangle\},\\ 
-  a_j & = & \max_{i=1,2,\ldots,c}\{ a_i \}.
-\end{eqnarray}
-
-Given an image $\I = (D, \lI)$ of a dataset $\D$ such that each image has some category $\omega \in \{1, 2, \ldots, c\}$, a \textit{marker} is a pair $(p, \omega)$ where $p \in D$.
-We denote by $\M(\I)$ a set of markers for image $\I$.
-For a marker $m = (p, \omega) \in \M(\I)$, a \textit{patch} is the locally expanded feature vector $\E(p)$ of $p$ with respect to some adjacency relation $\A$.
-We denote by $P(\I, \M, \A)$ the set of patches of an image $\I$ with respect to a set of markers and an adjacency relation.
-\end{comment}
+We propose a new way of learning the filters of the convolutional layers of deep neural networks that do not depend on the optimization of an error function and, therefore, do not need a large number of training samples.
+The training process is done with user intervention, which adds markers to the training images to indicate relevant features to characterize each class.
+To learn the filters, we get a patch around each marker of a class and cluster them.
+The centroid of each cluster is a filter that can be used to identify features of this cluster.
+The set of centroids of clusters of all clusters is the convolutional layer filters.
+For the next layer, the patches are generated on the output of the previous one.
+To evaluate our method, we use a dataset of aerial images to detect coconut trees.
+We identified that, with few layers and few training samples, we managed to get competitive results with state-of-the-art models.
 
@@
 
-\begin{figure}[t]
-  \begin{center}
-     \includegraphics[width=0.8\linewidth]{figures/tile0705-markers.png}
-  \end{center}
-     \caption{Example of markers.}
-  \label{fig:ex-markers}
-\end{figure}
-
 To exemplify how our method works, we apply it to the problem of segmenting rock images into vegetation, rock and fractures.
-In Figure \ref{fig:ex-markers} we can see an example of the markers added by the network designer in regions that she considers relevant to characterize each of the classes.
-
-\begin{figure*}[t]
-  \centering
-  \begin{subfigure}[b]{0.33\textwidth}
-      \includegraphics[width=\textwidth]{figures/probabilidade-fratura.png}
-      \caption{Probability map for fractures.}
-      \label{fig:prob-fractures}
-  \end{subfigure}
-  ~
-  \begin{subfigure}[b]{0.33\textwidth}
-      \includegraphics[width=\textwidth]{figures/probabilidade-rocha.png}
-      \caption{Probability map for rocks.}
-    \label{fig:prob-rocks}
-  \end{subfigure}
-  ~
-  \begin{subfigure}[b]{0.33\textwidth}
-      \includegraphics[width=\textwidth]{figures/probabilidade-vegetacao.png}
-      \caption{Probability map for vegetation.}
-  \label{fig:prob-vegetation}
-  \end{subfigure}
-
-  \caption{Probability maps for the classes.}
-  \label{fig:prob_map}
-\end{figure*}
-
-We designed a neural network with two convolutional layers and used the features extracted by it to produce the probability maps that we can see in Figure \ref{fig:prob_map}, one for each class.
-
-\begin{figure*}[t]
-  \centering
-  \begin{subfigure}[b]{0.33\textwidth}
-    \includegraphics[width=\textwidth]{figures/fratura.png}
-    \caption{Segmentation of fractures.}
-    \label{fig:segm-fractures}
-  \end{subfigure}
-  ~
-  \begin{subfigure}[b]{0.33\textwidth}
-    \includegraphics[width=\textwidth]{figures/rocha.png}
-    \caption{Segmentation of rocks.}
-    \label{fig:segm-rocks}
-  \end{subfigure}
-  ~
-  \begin{subfigure}[b]{0.33\textwidth}
-    \includegraphics[width=\textwidth]{figures/vegetacao.png}
-    \caption{Segmentation of vegetation vegetation.}
-  \label{fig:segm-vegetation}
-\end{subfigure}
-
-\caption{Segemetation of classes.}
-\label{fig:segm}
-\end{figure*}
-
-The probability maps were used to produce the segmentations that we can see in Figure \ ref {fig: segm}. For simplify viewing, each image shows the segmentation of one class.
+In Figure we can see an example of the markers added by the network designer in regions that she considers relevant to characterize each of the classes.
 
@@
 
-We placed markers on these images in regions that we consider important to identify the coconut and non-coconut (more diverse) classes.
-An example of a coconut tree image with markers can be seen in Figure \ref{fig:coco-markers} and non-coconut tree image with markers in Figure \ref{fig:non-coco-markers}.
-
-\begin{figure}[h!]
+\begin{figure}[t]
     \begin{center}
@@
 
-\begin{figure}[h!]
+  \begin{figure}[t]
     \begin{center}
@@
   \end{figure}
+
+We placed markers on these images in regions that we consider important to identify the coconut and non-coconut (more diverse) classes.
+An example of a coconut tree image with markers can be seen in Figure \ref{fig:coco-markers} and non-coconut tree image with markers in Figure \ref{fig:non-coco-markers}.
 
@@
 
-\begin{table*}[h]
+\begin{table}[!t]
   \begin{center}
@@
   \label{tab:results}
-  \end{table*}
+\end{table}
 
-\section{Conclusions and future work}
-
-{\small
-\bibliographystyle{ieee_fullname}
-\bibliography{egbib}
-}
+\bibliographystyle{IEEEtran}
+\bibliography{bibliography}
 

### 2020-06-01 main.tex
add results with SVM

@@
   Ours (fine tuned) & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.84 \pm 0.02$\\
+  Ours with SVM & $0.81 \pm 0.02$ & $0.77 \pm 0.03$ & $0.78 \pm 0.02$\\
   VGG-16 & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\

### 2020-06-03 main.tex
new figures

--- 2020-06-01 main.tex
@@
 \usepackage{amsfonts}
+\usepackage{bm}
 
@@
 
+\begin{comment}
 \begin{figure}[t]
@@
   \caption{Filter learned from clustering.}
+  \label{fig:filter}
+\end{figure}
+\end{comment}
+
+\begin{figure}[!t]
+  \centering
+  \subfloat[Groups of three classes centered on the origin.]{\includegraphics[width=.7\linewidth]{figures/example_groups/groups_centered}}
+  \\
+  \subfloat[Groups after applying convolution and activation operations.]{\includegraphics[width=.7\linewidth]{figures/example_groups/groups3d_centered}
+  \label{fig_second_case}}
+  \caption{Exemple of application of convolution with groups centers as filters' weightes.}
   \label{fig:filter}
@@
 
-\begin{figure}[t]
-    \begin{center}
-       \includegraphics[width=0.4\linewidth]{figures/coco_markers.png}
-    \end{center}
-       \caption{Markers on an image which contains a coconut tree.}
-    \label{fig:coco-markers}
-  \end{figure}
-
-  \begin{figure}[t]
-    \begin{center}
-       \includegraphics[width=0.4\linewidth]{figures/non_coco-markers.png}
-    \end{center}
-       \caption{Markers on an image which does not contain a coconut tree.}
-    \label{fig:non-coco-markers}
+\begin{figure}[!t]
+    \centering
+    \subfloat[Markers on an image which contains a coconut tree.]{\includegraphics[width=.45\linewidth]{figures/markers1}
+    \label{fig:markers1}}
+    ~
+    \subfloat[Markers on an image which does not contain a coconut tree.]{\includegraphics[width=.45\linewidth]{figures/markers2}
+    \label{fig:markers2}}
+    \\
+    \subfloat[Markers on an image which does not contain a coconut tree.]{\includegraphics[width=.45\linewidth]{figures/markers3}
+    \label{fig:markers3}}
+    ~
+    \subfloat[Markers on an image which does not contain a coconut tree.]{\includegraphics[width=.45\linewidth]{figures/markers4}
+    \label{fig:markers4}}
+    \caption{Markers used for training.}
 \end{figure}
@@
 We placed markers on these images in regions that we consider important to identify the coconut and non-coconut (more diverse) classes.
-An example of a coconut tree image with markers can be seen in Figure \ref{fig:coco-markers} and non-coconut tree image with markers in Figure \ref{fig:non-coco-markers}.
+An example of a coconut tree image with markers can be seen in Figure \ref{fig:markers1} and non-coconut tree image with markers in Figure \ref{fig:markers2}.
 
@@
 For clustering to find the convolutional layer filters, we used K-means to cluster the coconut tree patches in 35 clusters and the non-coconut tree patches in also 35 clusters.
+
+\begin{figure}
+  \centering
+  \subfloat[Visualization of original image feature space.]{\includegraphics[width=.45\linewidth]{figures/input}
+  \label{fig:vis-input}}
+  ~
+  \subfloat[Visualization of the feature extractor's outputs.]{\includegraphics[width=.45\linewidth]{figures/feature_extractor}
+  \label{fig:vis-feature-extractor}}
+  \\
+  \subfloat[Visualization of outputs from the last classifier hidden layer.]{\includegraphics[width=.45\linewidth]{figures/classifier}
+  \label{fig:vis-classifier}}
+  \caption{This image shows visualizations of the input images and the intermediate layer's outputs.
+Green points represent background images, and red points represent images that have a coconut tree.
+The projection from high dimensionality space to 2D was made using t-SNE.
+The original image shows much overlap between the classes; after applying the feature extractor with only one convolutional layer, we can see that the overlapping decreases considerably.
+Before the decision layer, the last layer put most of the images from one class on one side and the other on the other side.}
+\end{figure}
 
@@
   \hline\hline
-  Ours & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.85 \pm 0.02$\\
+    Ours & $\boldsymbol{0.86 \pm 0.01}$ & $0.84 \pm 0.02$ & $0.85 \pm 0.02$\\
   Ours (fine tuned) & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.84 \pm 0.02$\\
+  Ours with SVM & $0.81 \pm 0.02$ & $0.77 \pm 0.03$ & $0.78 \pm 0.02$\\
   VGG-16 & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
@@
 
+\section{Conclusions}
+
 \bibliographystyle{IEEEtran}

### 2020-06-03 main.tex
definition of adjacency

@@
 Our method is based on the idea that the network designer is an expert in the application domain or knows well enough to indicate which are the relevant characteristics of the objects of interest.
-The designer highlights these regions by placing markers.
-For each marker, the extended local feature vectors of its spels are taken.
+The designer highlights these regions by placing markers, which are associate to pixels.
+For each marker, we create a extended feature vector by taken the pixel's feature vector and of its neiborhs.
 These vectors are clustered and the centers of those groups are used as filters for the current convolutional layer.
@@
 Let $\D$ be a dataset of images.
+For an image $I \in \D$, the adjacency $\A$ of a pixel $p = (x, y)$ of $I$ is a set pixels $\A(p) = \{ p_1, p_2, \ldots, p_n \}$.
+In this work, we only consider a paritcular case of adjacency,
+\[\A_k(p) = \{(x^\prime, y^\prime) :\, \sqrt{(x - x^\prime)^2 + (y - y^\prime)^2} \le k\}.\]
 The designer draws markers on only a subset of the images, let's say $\pDD \subseteq \D$.
-That is, she produces for each $\I \in \pDD$ a set of markers $\M(\I)$.
+That is, she produces for each $I \in \pDD$ a set of markers $\M(I)$.
 This subset of images, using the information from the markers, is used to create a set of patches \[\Pa(\pDD) = \bigcup_{\I \in \pDD}{P(\I, \M, \A)},\] with respect to an adjacency relation also defined by the designer.

### 2020-06-05 main.tex
conclusion and abstract

--- 2020-06-03 main.tex
@@
 
+\DeclareMathOperator{\vect}{vec}
+
 \newcommand{\R}{\mathbb{R}}
@@
 \begin{abstract}
-
+  Identifying the species of trees in a region is essential to identify the type of land use that is taking place.
+It is also essential to assess and monitor plantation performance and analyze the impact of natural disasters.
+The manual identification of these trees is tedious, costly, and error-prone, so the application of automatic methods is necessary.
+Deep learning is quite effective in applications from different domains; however, it usually needs a considerable amount of annotated data so that the models' appropriate training can be done.
+We propose a training method for feature extractors of convolutional neural networks that need only a minimal set of images.
+The method uses markers that the user places in relevant regions of the training images to learn the convolutional layers' filters.
+The method was applied to the problem of determining whether area image patches contain coconut trees or not.
+The model trained with this method achieved competitive results with state of the art.
 \end{abstract}
@@
 \section{Learning filters from markers}
+\label{sec:method}
+
+Convolutional neural networks perform well in image classification tasks due to its capability of extracting features.
+That is, convolutional layers produce great descriptors that can be used by the classifier.
+Conventionally, during the training process, an algorithm tries to minimize a loss function that measures how good the classifier is performing with the features given by the feature extractor.
+The algorithm of optimization tries through backpropagation to find out which features to extract.
 Our method is based on the idea that the network designer is an expert in the application domain or knows well enough to indicate which are the relevant characteristics of the objects of interest.
-The designer highlights these regions by placing markers.
-For each marker, the extended local feature vectors of its spels are taken.
-These vectors are clustered and the centers of those groups are used as filters for the current convolutional layer.
-Since each convolutional layer transforms the input feaure space of the images into a new one, the same procedure can be repeated to create the filters of the subsequents convolutional layers.
-This way, it is enough to specify the number of desired convolutional layers and their hyperparameters.
+Hence, it does not rely on the optimization of some loss function to learn filters for those relevant characteristics.
+In our approach, the network designer highlights features that she believes are the ones that characterize each class by placing markers.
+Those markers are used to learn the filters to identify the features they represent.
 
-Let $\D$ be a dataset of images.
-The designer draws markers on only a subset of the images, let's say $\pDD \subseteq \D$.
-That is, she produces for each $\I \in \pDD$ a set of markers $\M(\I)$.
-This subset of images, using the information from the markers, is used to create a set of patches \[\Pa(\pDD) = \bigcup_{\I \in \pDD}{P(\I, \M, \A)},\] with respect to an adjacency relation also defined by the designer.
-In this work, we use only circular adjacency relations.
+Let $\D$ be a dataset of images, and let $I \in \D$ be an image with dimensions $X \times Y$.
+A marker $m$ is associated with a pixel $p(m) = (x, y)$ and has a label $L(m) = c$.
+Let $M(I)$ be the set of markers of image $I$.
+For each maker $m \in M(I)$, we create a patch $P$ with dimensions $k \times k$ from a window centered in the marker's pixel as $P(I, x ,y) = I(x+i, y+j)$ for $i, j = -\floor{k/2}, \ldots, \floor{k/2}$.We pad images with $0$ so that we can create patches from border pixels.
+For each class, we create a set of patches made using the markers of that class.
+Let \[\mathcal{P}_c = \bigcup_{ I \in \D, m \in M(I) :\, L(m) = c}{P(I, p(m))}\] be the set of all patches made from markers of class $c$.
 
-Given the set of patches $\Pa(\pDD)$ of the subset of images $\pDD$, we use a clustering algorithm to discover groups in $\Pa(\pDD)$.
-Each group represents local patterns from one or more classes, highlighted by the designer through the markers.
-Since the center of each cluster must be a good representative of the characteristics of the group, we can use the group centroids as filters to enhance those patterns in the images of $\D$.
+Given set of patches $\Pa_c$, we can use a clustering algorithm to discover groups in $\Pa_c$.
+Each group represents local patterns of class $c$, highlighted by the designer through the markers.
+Since the centroid of each cluster must be a good representative of the characteristics of the group, we can use them as filters to enhance those patterns in the images of $\D$.
+We must do this for each class.
 
 However, for it to work properly, patches and images need to be centered at the origin.
-For this, we use the information from the patches to calculate the mean and standard deviation per band.
-If the patches are representative of images in $\D$, the mean and standard deviation (per band) of $\Pa$ are very close to the mean and standard deviation (per band) in $\D$.
+For this, we use the information from the patches of all classes to calculate the mean and standard deviation per band.
+If the patches are representative of images in $\D$, the mean and standard deviation per band of $\Pa$ are close to the mean and standard deviation per band in $\D$.
 
@@
 
-\begin{figure}[!t]
-  \centering
-  \subfloat[Groups of three classes centered on the origin.]{\includegraphics[width=.7\linewidth]{figures/example_groups/groups_centered}}
-  \\
-  \subfloat[Groups after applying convolution and activation operations.]{\includegraphics[width=.7\linewidth]{figures/example_groups/groups3d_centered}
-  \label{fig_second_case}}
-  \caption{Exemple of application of convolution with groups centers as filters' weightes.}
-  \label{fig:filter}
-\end{figure}
-
-Note that the convolution in a spel $p \in D$ corresponds to the doct product of the locally extended feature vecture and the vector that represents the filter.
-Thus, let $H$ be a hyperplane orthogonal to the vector of a $\mF$ filter.
-Thus, let $ H $ be a hyperplane orthogonal to the vector of a $ \mF$ filter.
-Vectors similar to the vector of filter $\mF$ must fall on the same side of the hyperplane as $\mF$.
-That is, if the vector $\E(p)$ is on the same side of $H$ as $\mF$, then the doct product between $\E (p)$ and $\mF$ is positive, and is negative otherwise.
-Thus, the doct product between the center of a cluster and each of the vectors of that cluster must be positive, which implies that the center is a good filter to identify the pattern described by that cluster.
-In the Figure \ref{fig:filter} we can see an example.
+Let $F$ be a filter with dimension $k \times k$, and let $\pi$ be the patch around pixel $(x,y)$ also with dimension $k \times k$.
+By the definition of convolution in deep learning, we can interpret this operation for a pixel $(x, y)$ as the dot product between the kernel vector $\vect(F)$ and the vector of a patch $\vect(\pi)$ around this pixel, that is, the convolution in $(x, y)$ is the same as $\vect(F) \cdot \vect(\pi)$.
+Thus, let $H$ be a hyperplane orthogonal to vector $\vect(F)$.
+Vectors similar to $\vect(F)$ must fall on the same side of the hyperplane as $\vect(F)$.
+Therefore, if vector $\vect(\pi)$ is on the same side of $H$ as $\vect(F)$, then the doct product between $\vect(\pi)$ and $\vect(F)$ is positive, and it is negative otherwise.
+Thus, the doct product between the centroid of a cluster and each of the vectors of that cluster must be positive, which implies that the centroid is a good filter to identify the pattern described by that cluster.
+Hence, if we throw away the negative values, applying convolution with centroids as filters must highlight regions that has a certain pattern that the network designer believes is important.
 
@@
 We preserve the output dimensions of convolutional layers so that we can use the same markers used in the previous layer.
-The new patches are generated from the convolutional layer output and the whole process is repeated to find the filters of the next layer.
+The new patches are generated from the previeus convolutional layer output and the whole process is repeated to find the filters of the next one.
 
-To exemplify how our method works, we apply it to the problem of segmenting rock images into vegetation, rock and fractures.
-In Figure we can see an example of the markers added by the network designer in regions that she considers relevant to characterize each of the classes.
+\begin{figure}[!t]
+  \centering
+  \subfloat[Groups of three classes centered on the origin.\label{fig:ex-groups}]
+  {\includegraphics[width=.45\linewidth]{figures/example_groups/groups_centered}}
+  ~
+  \subfloat[Groups after applying convolution and activation operations. \label{fig:ex-groups-after}]{\includegraphics[width=.45\linewidth]{figures/example_groups/groups3d_centered}
+  }
+  \caption{Exemple of application of convolution with groups centers as filters' weightes.}
+  \label{fig:filter}
+\end{figure}
+
+To exemplify our method, we create a synthetic dataset of points of two classes.
+We can see this dataset in Figure \ref{fig:ex-groups}.
+The points are distributed in two regions, and it is not possible to separate the yellow class linearly from the violet class.
+To center the points,  we subtracted the mean and divided by the standard deviation.
+If we look only to violet points, it is clear that they form two clusters.
+In turn, the yellow dots form a single cluster.
+We then take the center of each of these three clusters and apply convolution to the set of points using these centers as filters.
+The result of the convolution after applying the ReLU function can be seen in Figure \ref{fig:ex-groups-after}.
+For most points after the transformation made by convolution and the ReLU function, only one coordinate was positive, this being the coordinate which value is the internal product with the center of the cluster of which these points are part.
+These points lie along the three axes of space $\mathbb{R}^3$, each cluster on an axis.
+The other points are farthest from their cluster center, and the internal product with the center of another cluster is also positive.
+Cluster centers functioned as good filters to detect similar points, even though they detected points from other clusters.
 
 \section{Experiments}
-To evaluate our method in a quantitative way, we need a dataset with a ground truth and unfortunately the rock image dataset was not properly labeled.
-
-So, we chose to use a dataset formed from patches of coconut images.
+We use a dataset formed from patches of coconut images.
 The dataset consists of images of dimensions $90 \times 90$ of coconut and non-coconut.
-The latter class having a large variability.
+The latter class has a large variability.
 The problem of interest is to classify the images in one of the two classes.
 
-As in reality, labeling images is a costly and time-consuming process, we assume that we only knew the true labels of a small set of samples (one hundred images for each class, totaling two hundred images).
+As in reality, labeling images is a costly and time-consuming process, and then we assume that we only knew the correct labels of a small set of samples (one hundred images for each class, totaling two hundred images).
 That way, we would have only those two hundred images to train any models that we would use to try to solve the problem.
@@
 We wanted to represent a scenario where the network designer did not have much effort to annotate all these images.
-
-From each of these sets of training images, we selected 4 images using the 2D space projection of the 200 images using the tSNE.
+From each of these sets of training images, we selected four images using the 2D space projection of the 200 images using the t-SNE.
 We tried to select images that were part of more cohesive groups.
@@
     \caption{Markers used for training.}
+    \label{fig:markers}
 \end{figure}
@@
 We placed markers on these images in regions that we consider important to identify the coconut and non-coconut (more diverse) classes.
-An example of a coconut tree image with markers can be seen in Figure \ref{fig:markers1} and non-coconut tree image with markers in Figure \ref{fig:markers2}.
-
-With the markers, we created a convolutional layer with filters of dimensions $7 \times 7$.
-We did a padding to preserve the dimensions of the images.
+In Figure \ref{fig:markers} we can the four images of one the splits.
+With the markers, we created a convolutional layer with filters of dimensions $7 \times 7$ using the method described in Section \ref{sec:method}.
 The network architecture can be seen in the Figure.
 We used a classifier similar to that of VGG-16.
+We used K-means to cluster the coconut tree patches in 30 clusters and the non-coconut tree patches in also 30 clusters to find the convolutional layer filter.
 
-For clustering to find the convolutional layer filters, we used K-means to cluster the coconut tree patches in 35 clusters and the non-coconut tree patches in also 35 clusters.
+\begin{table}[!t]
+  \begin{center}
+  \begin{tabular}{|l|c|c|c|}
+  \hline
+   Method & Precision & Recall & F-score \\
+  \hline\hline
+    I - Ours & $\boldsymbol{0.86 \pm 0.01}$ & $0.84 \pm 0.02$ & $0.85 \pm 0.02$\\
+    II - Ours (fine tuned) & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.84 \pm 0.02$\\
+    III - Ours with SVM & $0.81 \pm 0.02$ & $0.77 \pm 0.03$ & $0.78 \pm 0.02$\\
+    IV - VGG & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
+    V - VGG (fine tuned) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
+  \hline
+  \end{tabular}
+  \end{center}
+  \caption{Results.}
+  \label{tab:results}
+\end{table}
+
+We tested three scenarios: (I) using out method to learn the feature extractor weights and the classifier with backpropagation; (II) using our method as initialization and train the whole model with backpropagation; (III) using SVM Linear as classification.
+For VGG-16, we tested two secenarios: (IV) initializing with Xavier initialization and training from screath; (V) using ImageNet weights and training with the training set images.
+The results can be seen in Table \ref{tab:results}.
+As the experiments were repeated three times, we reported the mean and standard deviation of each the metrics in the test set.
+
+Our method achieved better precision than conventionally trained VGG, and the difference between precision and recall is much smaller.
+The Linear SVM did not prove to be a good classifier like the MLP, but, on average, it still got a little better than the VGG-16.
+Using our method as initialization and training the entire network with backpropagation did not improve the result; this implies that the filters we created are already good filters given the training set.
+Our feature extractor training method was able to learn good filters from a minimal set of images.
+VGG has many parameters to learn, so it is susceptible to overfitting with such a small training set.
+We do not use any technique to handle small data sets to train VGG because we intended to verify our approach viability, and these techniques could also be used with our method.
 
@@
 
-As the experiments were repeated three times, we reported the mean and standard deviation for each of the metrics in the test set. The results can be seen in the Table \ref{tab:results}.
-
-\begin{table}[!t]
-  \begin{center}
-  \begin{tabular}{|l|c|c|c|}
-  \hline
-   Method & Precision & Recall & F-score \\
-  \hline\hline
-    Ours & $\boldsymbol{0.86 \pm 0.01}$ & $0.84 \pm 0.02$ & $0.85 \pm 0.02$\\
-  Ours (fine tuned) & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.84 \pm 0.02$\\
-  Ours with SVM & $0.81 \pm 0.02$ & $0.77 \pm 0.03$ & $0.78 \pm 0.02$\\
-  VGG-16 & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
-  VGG-16 (fine tuned) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
-  \hline
-  \end{tabular}
-  \end{center}
-  \caption{Results.}
-  \label{tab:results}
-\end{table}
+To understand how the feature extractor trained with our method transforms the image spaces, we made projections of three stages of the network using t-SNE.
+In Figure \ref{fig:vis-input}, we can see the projection of the original images (in the LAB color space).
+Red points are images that contain coconut trees, and green points are images that do not contain coconut trees.
+In the projection in the image of Figure \ref{fig:vis-input}, we see that the images of coconut trees and non-coconut trees are almost entirely overlapping.
+Coconut tree images are more frequent in some regions but are very dispersed.
+In the image of Figure \ref{fig:vis-feature-extractor}, we see the projection of the feature extractor output.
+The coconut tree images are more concentrated, forming some clusters, and it is possible to notice that there is much less overlap.
+Finally, we see in Figure \ref{fig:vis-classifier} the output of the previous layer of the decision layer of the classifier.
+The classifier managed to organize conquest and non-coconut images in different regions of space, although there is overlap as the model failed to achieve 100\% accuracy.
 
 \section{Conclusions}
+
+We present a way to train a feature extractor without using backpropagation capable of learning good filters from a minimal set of images.
+To evaluate the method, we proposed experiments to classify images that contain achievements.
+Our approach has achieved competitive results with state of the art.
 

### 2020-06-07 main.tex
dataset description

@@
 \section{Experiments}
-We use a dataset formed from patches of coconut images.
-The dataset consists of images of dimensions $90 \times 90$ of coconut and non-coconut.
+We use a dataset formed from patches of aerial images made by \cite{8899005} from imagery provided by WeRobotics.
+The imagery with a resolution of 8cm was captured by the World Bank's UAVs for Disaster Resilience Program, collaborating with WeRobotics and OpenAerialMap in the Kindom of Tonga in October 2017.
+The Humanitarian OpenStreetMap community provided the annotation.
+The dataset consists of 13587 patches of dimensions $90 \times 90$ of coconut (10268) and non-coconut (3319).
 The latter class has a large variability.
@@
   \caption{This image shows visualizations of the input images and the intermediate layer's outputs.
-Green points represent background images, and red points represent images that have a coconut tree.
+Red points represent background images, and green points represent images that have a coconut tree.
 The projection from high dimensionality space to 2D was made using t-SNE.

### 2020-06-07 main.tex
testing new caption description

@@
   \centering
-  \subfloat[Groups of three classes centered on the origin.\label{fig:ex-groups}]
-  {\includegraphics[width=.45\linewidth]{figures/example_groups/groups_centered}}
+  \subfloat[\label{fig:ex-groups}]
+  {\includegraphics[width=.35\linewidth]{figures/example_groups/groups_centered}}
   ~
-  \subfloat[Groups after applying convolution and activation operations. \label{fig:ex-groups-after}]{\includegraphics[width=.45\linewidth]{figures/example_groups/groups3d_centered}
+  \subfloat[\label{fig:ex-groups-after}]{\includegraphics[width=.45\linewidth]{figures/example_groups/groups3d_centered}
   }
-  \caption{Exemple of application of convolution with groups centers as filters' weightes.}
+  \caption{Exemple of application of convolution with groups centers as filters' weightes: (a) Groups of three classes centered on the origin; (b) Groups after applying convolution and activation operations.
+}
   \label{fig:filter}

### 2020-06-08 main.tex
qualitative analysis

@@
 
-\title{Learning filters from user provided image makers}
+\title{Learning filters from user provided image makers for convolutional neural networks}
 
@@
 Deep learning is quite effective in applications from different domains; however, it usually needs a considerable amount of annotated data so that the models' appropriate training can be done.
-We propose a training method for feature extractors of convolutional neural networks that need only a minimal set of images.
-The method uses markers that the user places in relevant regions of the training images to learn the convolutional layers' filters.
+As deep as the model gets, it needs more data.
+We propose a training method for feature extractors of convolutional neural networks that need only a minimal set of images and that the CNN designer has more control over.
+The method uses markers that the user places in relevant regions of some training images to learn the convolutional layers' filters allowing the user to interfere with the training process directly.
+As this method does not rely on backpropagation or minimization of some cost function to learn the feature extractor filters, it does not require many training images.
 The method was applied to the problem of determining whether area image patches contain coconut trees or not.
@@
 The problem of interest is to classify the images in one of the two classes.
+We only report experiments with one dataset due to the space limitation.
 
@@
     \centering
-    \subfloat[Markers on an image which contains a coconut tree.]{\includegraphics[width=.45\linewidth]{figures/markers1}
-    \label{fig:markers1}}
+    \subfloat[\label{fig:markers1}]{\includegraphics[width=.3\linewidth]{figures/markers1}
+    }
     ~
-    \subfloat[Markers on an image which does not contain a coconut tree.]{\includegraphics[width=.45\linewidth]{figures/markers2}
-    \label{fig:markers2}}
+    \subfloat[\label{fig:markers2}]{\includegraphics[width=.3\linewidth]{figures/markers2}
+    }
     \\
-    \subfloat[Markers on an image which does not contain a coconut tree.]{\includegraphics[width=.45\linewidth]{figures/markers3}
-    \label{fig:markers3}}
+    \subfloat[\label{fig:markers3}]{\includegraphics[width=.3\linewidth]{figures/markers3}
+    }
     ~
-    \subfloat[Markers on an image which does not contain a coconut tree.]{\includegraphics[width=.45\linewidth]{figures/markers4}
-    \label{fig:markers4}}
-    \caption{Markers used for training.}
+    \subfloat[\label{fig:markers4}]{\includegraphics[width=.3\linewidth]{figures/markers4}
+    }
+    \caption{Markers used for training.
+(a) Markers on an image which contains a coconut tree.
+(b) Markers on an image which does not contain a coconut tree.
+(c) Markers on an image which does not contain a coconut tree.
+(d) Markers on an image which does not contain a coconut tree.}
     \label{fig:markers}
@@
   \hline\hline
-    I - Ours & $\boldsymbol{0.86 \pm 0.01}$ & $0.84 \pm 0.02$ & $0.85 \pm 0.02$\\
-    II - Ours (fine tuned) & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.84 \pm 0.02$\\
-    III - Ours with SVM & $0.81 \pm 0.02$ & $0.77 \pm 0.03$ & $0.78 \pm 0.02$\\
-    IV - VGG & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
-    V - VGG (fine tuned) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
+    I-Ours & $\boldsymbol{0.86 \pm 0.01}$ & $\boldsymbol{0.84 \pm 0.02}$ & $\boldsymbol{0.85 \pm 0.02}$\\
+    II-Ours (fine tuned) & $\boldsymbol{0.86 \pm 0.01}$ & $\boldsymbol{0.84 \pm 0.02}$ & $0.84 \pm 0.02$\\
+    III-Ours with SVM & $0.81 \pm 0.02$ & $0.77 \pm 0.03$ & $0.78 \pm 0.02$\\
+    IV-VGG & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
+    V-VGG (fine tuned) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
   \hline
@@
 
+\begin{figure}
+  \centering
+  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/ours/upper-coco-108.png}
+  }
+  ~
+  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/ours/upper-coco-2473.png}
+  }
+  ~
+  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/ours/upper-non_coco-506.png}
+  }
+  \\
+  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/vgg/lower-coco-1098.png}
+  }
+  ~
+  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/vgg/lower-coco-2400.png}
+  }
+  ~
+  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/vgg/lower-non_coco-105.png}
+  }
+  \label{fig:ex-classification}
+  \caption{Misclassified images.
+The first row has images misclassified by the network trained by our method, and the second row has images misclassified by VGG-16.
+The two first images of a row have coconut trees.}
+\end{figure}
+
 Our method achieved better precision than conventionally trained VGG, and the difference between precision and recall is much smaller.
@@
   \centering
-  \subfloat[Visualization of original image feature space.]{\includegraphics[width=.45\linewidth]{figures/input}
-  \label{fig:vis-input}}
+  \subfloat[\label{fig:vis-input}]{\includegraphics[width=.3\linewidth]{figures/input}
+  }
   ~
-  \subfloat[Visualization of the feature extractor's outputs.]{\includegraphics[width=.45\linewidth]{figures/feature_extractor}
-  \label{fig:vis-feature-extractor}}
-  \\
-  \subfloat[Visualization of outputs from the last classifier hidden layer.]{\includegraphics[width=.45\linewidth]{figures/classifier}
+  \subfloat[\label{fig:vis-feature-extractor}]{\includegraphics[width=.3\linewidth]{figures/feature_extractor}
+  }
+  ~
+  \subfloat[]{\includegraphics[width=.3\linewidth]{figures/classifier}
   \label{fig:vis-classifier}}
@@
 The original image shows much overlap between the classes; after applying the feature extractor with only one convolutional layer, we can see that the overlapping decreases considerably.
-Before the decision layer, the last layer put most of the images from one class on one side and the other on the other side.}
+Before the decision layer, the last layer put most of the images from one class on one side and the other on the other side.
+(a) Visualization of original image feature space.
+(b) Visualization of the feature extractor's outputs.
+(c) Visualization of outputs from the last classifier hidden layer.}
 \end{figure}
 
-To understand how the feature extractor trained with our method transforms the image spaces, we made projections of three stages of the network using t-SNE.
+As motivated by Rauber et al.
+\cite{rauber2016visualizing}, we made projections of three stages of the network using t-SNE \cite{maaten2008visualizing} to understand how the feature extractor trained with our method transforms the image spaces.
 In Figure \ref{fig:vis-input}, we can see the projection of the original images (in the LAB color space).

### 2020-06-08 main.tex
qualitative analysis text

@@
   }
-  \label{fig:ex-classification}
   \caption{Misclassified images.
@@
 The two first images of a row have coconut trees.}
+  \label{fig:ex-classification}
 \end{figure}
@@
 We do not use any technique to handle small data sets to train VGG because we intended to verify our approach viability, and these techniques could also be used with our method.
+In Figure \ref{fig:ex-classification} we can see examples of misclassified images.
+In these images, coconut trees appear in different angles, sizes, and shapes, or the boundaries between the coconut tree and the background are tenuous, making it more difficult to identify them.
+In turn, images that do not contain coconut trees, contain trees that resemble their shape.
 

### 2020-06-08 main.tex
Updates from Overleaf
update SVM results

@@
 \usepackage{bm}
+\usepackage{pgf,tikz,pgfplots}
+\usetikzlibrary{calc}
 
@@
 \newcommand{\pDD}{\D^\prime}
+
+\definecolor{navy}{RGB}{0,0,137}
+\definecolor{tealDeer}{RGB}{148,232,180}
+\definecolor{dodgerBlue}{RGB}{18,161,255}
+\definecolor{citrine}{RGB}{230, 194, 8}
+\definecolor{violet}{RGB}{112,5,164}
+\definecolor{navyPurple}{RGB}{172,86,253}
+\definecolor{heliotrope}{RGB}{236,93,253}
 
@@
 
+\begin{figure}
+  \begin{center}
+  \tikzset{
+    treenode/.style = {shape=rectangle, rounded corners,
+                       draw=black, align=center, line width=width},
+    specialedge/.style = {line width=1.5pt}
+  }
+
+  \begin{tikzpicture}
+    [
+      grow                    = down,
+      sibling distance        = 5em,
+      level distance          = 3em,
+      edge from parent/.style = {draw=black},
+      every node/.style       = {treenode},
+      sloped
+    ]
+    \node [treenode, fill=tealDeer, label = {[label distance = 0.5em]}] (v0) {Conv$(60 \times 7 \times 7)$}
+      child {
+        child {
+          child {
+                child {
+                  node [treenode, fill=navyPurple] {Classifier}
+                  edge from parent
+                }
+                node [treenode, fill=tealDeer] {BatchNorm}
+                edge from parent
+           }
+          node [treenode, fill=tealDeer] {MaxPool$(3\times3)$}
+          edge from parent
+        }
+        node [treenode, fill=dodgerBlue, label = {[label distance = 0.5em] }] {ReLU} 
+        edge from parent
+      };
+
+  \end{tikzpicture}
+
+  \end{center}
+  \caption[An output example of the algorithm to compute a core stable coalition structure.]{An output example of the algorithm to compute a core stable coalition structure for games represented by forests.
+Each color represents a coalition.}
+  \label{fig:tree}
+\end{figure}
+
 To exemplify our method, we create a synthetic dataset of points of two classes.
@@
     II-Ours (fine tuned) & $\boldsymbol{0.86 \pm 0.01}$ & $\boldsymbol{0.84 \pm 0.02}$ & $0.84 \pm 0.02$\\
-    III-Ours with SVM & $0.81 \pm 0.02$ & $0.77 \pm 0.03$ & $0.78 \pm 0.02$\\
+    III-Ours with SVM & $0.836 \pm 0.015 $ & $ 0.793 \pm 0.025$ & $ 0.804 \pm 0.022$\\
     IV-VGG & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
@@
 In turn, images that do not contain coconut trees, contain trees that resemble their shape.
+The nete
 

### 2020-06-08 main.tex
Updates from Overleaf

--- 2020-06-08 main.tex
@@
 
-\title{Learning filters from user provided image makers for convolutional neural networks}
+\title{Learning CNN filters from user-drawn image makers for coconut-tree image classification}
 
@@
 \begin{abstract}
-  Identifying the species of trees in a region is essential to identify the type of land use that is taking place.
-It is also essential to assess and monitor plantation performance and analyze the impact of natural disasters.
-The manual identification of these trees is tedious, costly, and error-prone, so the application of automatic methods is necessary.
-Deep learning is quite effective in applications from different domains; however, it usually needs a considerable amount of annotated data so that the models' appropriate training can be done.
-As deep as the model gets, it needs more data.
-We propose a training method for feature extractors of convolutional neural networks that need only a minimal set of images and that the CNN designer has more control over.
-The method uses markers that the user places in relevant regions of some training images to learn the convolutional layers' filters allowing the user to interfere with the training process directly.
-As this method does not rely on backpropagation or minimization of some cost function to learn the feature extractor filters, it does not require many training images.
-The method was applied to the problem of determining whether area image patches contain coconut trees or not.
-The model trained with this method achieved competitive results with state of the art.
+ Identifying species of trees in aerial images is essential for land-use classification, plantation monitoring, and impact assessment of natural disasters.
+The manual identification of trees in aerial images is tedious, costly, and error-prone, so automatic classification methods are necessary.
+Convolutional Neural Network (CNN) models have well succeeded in image classification applications from different domains.
+However, CNN models usually require intensive manual identification to create large training sets.
+We present a training method for CNN feature extraction that needs only a minimal set of images selected by the user.
+The method learns the filters of each convolutional layer from user-drawn markers in image regions that discriminate classes, allowing direct expert interference in the training process.
+It does not rely on optimization based on backpropagation, which considerably reduces the number of required images.
+We demonstrate its advantages on the binary classification of coconut-tree images against a popular CNN model.
 \end{abstract}
@@
 \section{Introduction}
-Deep learning has proven to be applicable to different types of tasks, from image classification to the creation of synthetic data \cite{goodfellow2016deep}.
-Given the success of the application of deep learning in various areas of knowledge, this technique has also been applied to several tasks of remote sensing.
-Examples of this technique exist in various problems such as segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic, hamaguchi2018effective}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}.
+Deep learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}.
+In remote sensing, the applications have involved segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic, hamaguchi2018effective}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}.
+In this work, we are interested in
+ identifying species of trees from aerial images.
+The topic is important for land-use classification, plantation monitoring during a season, and damage assessment of natural disasters, such as drilling and flooding.
+As these plantations can span very large areas, manual identification can be costly, tedious, and susceptible to errors, so automatic methods are necessary.
 
-The detection of trees is important to produce knowledge about the land use of a region, from the assessment of crop diversity or the performance of plantations in a given season to the assessment of damages after natural disasters such as drilling and flooding.
-As these plantations can span very large areas, manual detection can be costly and susceptible to errors.
-
-In \cite{fassnacht2016review}, we can see a review of the problem of classification of tree species of aerial images.
-\cite{puttemans2018comparing} apply deep learning to the problem of detecting coconut trees in aerial images.
+Classification of tree species in aerial images has been actively investigated~\cite{fassnacht2016review}.
+In \cite{puttemans2018comparing}, Puttemans et al.
+present a solution to detect coconut trees based on deep learning.
 Aparta et al.
@@
 In turn, images that do not contain coconut trees, contain trees that resemble their shape.
+The network designer could put markers in those images in other to improve the feature extractor.
 

### 2020-06-08 main.tex
Updates from Overleaf

@@
 Classification of tree species in aerial images has been actively investigated~\cite{fassnacht2016review}.
-In \cite{puttemans2018comparing}, Puttemans et al.
-present a solution to detect coconut trees based on deep learning.
-Aparta et al.
-\cite{aparna2018cnn} apply convolutional neural networks to count coconut trees on plantations in India.
-Vargas-Muñoz et al.
-\cite{8899005} propose an active learning approach to detect coconut trees that seeks to reduce the effort of annotating images.
-The \cite{8899005} approach is divided into a stage of producing candidate patches for having a coconut tree and a stage where a CNN determines whether a patch has a coconut tree.
+In \cite{puttemans2018comparing, aparna2018cnn}, the authors present solutions to detect coconut trees based on convolutional neural network (CNN) models.
+Vargas-Muñoz et al.~\cite{8899005} propose an active learning approach to reduce human effort in annotating images of coconut trees.
+They divide the problem in two stages: a stage that produces candidate image patches and a stage in which a CNN identifies the candidate patches with a coconut tree.
 We propose a different way to produce this classifier.
@@
   \end{tikzpicture}
-
   \end{center}
-  \caption[An output example of the algorithm to compute a core stable coalition structure.]{An output example of the algorithm to compute a core stable coalition structure for games represented by forests.
-Each color represents a coalition.}
   \label{fig:tree}

### 2020-06-08 main.tex
update ours results

--- 2020-06-08 main.tex
@@
 
-\title{Learning CNN filters from user-drawn image makers for coconut-tree image classification}
+\title{Learning filters from user provided image makers for convolutional neural networks}
 
@@
 \begin{abstract}
- Identifying species of trees in aerial images is essential for land-use classification, plantation monitoring, and impact assessment of natural disasters.
-The manual identification of trees in aerial images is tedious, costly, and error-prone, so automatic classification methods are necessary.
-Convolutional Neural Network (CNN) models have well succeeded in image classification applications from different domains.
-However, CNN models usually require intensive manual identification to create large training sets.
-We present a training method for CNN feature extraction that needs only a minimal set of images selected by the user.
-The method learns the filters of each convolutional layer from user-drawn markers in image regions that discriminate classes, allowing direct expert interference in the training process.
-It does not rely on optimization based on backpropagation, which considerably reduces the number of required images.
-We demonstrate its advantages on the binary classification of coconut-tree images against a popular CNN model.
+  Identifying the species of trees in a region is essential to identify the type of land use that is taking place.
+It is also essential to assess and monitor plantation performance and analyze the impact of natural disasters.
+The manual identification of these trees is tedious, costly, and error-prone, so the application of automatic methods is necessary.
+Deep learning is quite effective in applications from different domains; however, it usually needs a considerable amount of annotated data so that the models' appropriate training can be done.
+As deep as the model gets, it needs more data.
+We propose a training method for feature extractors of convolutional neural networks that need only a minimal set of images and that the CNN designer has more control over.
+The method uses markers that the user places in relevant regions of some training images to learn the convolutional layers' filters allowing the user to interfere with the training process directly.
+As this method does not rely on backpropagation or minimization of some cost function to learn the feature extractor filters, it does not require many training images.
+The method was applied to the problem of determining whether area image patches contain coconut trees or not.
+The model trained with this method achieved competitive results with state of the art.
 \end{abstract}
@@
 \section{Introduction}
-Deep learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}.
-In remote sensing, the applications have involved segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic, hamaguchi2018effective}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}.
-In this work, we are interested in
- identifying species of trees from aerial images.
-The topic is important for land-use classification, plantation monitoring during a season, and damage assessment of natural disasters, such as drilling and flooding.
-As these plantations can span very large areas, manual identification can be costly, tedious, and susceptible to errors, so automatic methods are necessary.
+Deep learning has proven to be applicable to different types of tasks, from image classification to the creation of synthetic data \cite{goodfellow2016deep}.
+Given the success of the application of deep learning in various areas of knowledge, this technique has also been applied to several tasks of remote sensing.
+Examples of this technique exist in various problems such as segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic, hamaguchi2018effective}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}.
 
-Classification of tree species in aerial images has been actively investigated~\cite{fassnacht2016review}.
-In \cite{puttemans2018comparing}, Puttemans et al.
-present a solution to detect coconut trees based on deep learning.
+The detection of trees is important to produce knowledge about the land use of a region, from the assessment of crop diversity or the performance of plantations in a given season to the assessment of damages after natural disasters such as drilling and flooding.
+As these plantations can span very large areas, manual detection can be costly and susceptible to errors.
+
+In \cite{fassnacht2016review}, we can see a review of the problem of classification of tree species of aerial images.
+\cite{puttemans2018comparing} apply deep learning to the problem of detecting coconut trees in aerial images.
 Aparta et al.
@@
   \hline\hline
-    I-Ours & $\boldsymbol{0.86 \pm 0.01}$ & $\boldsymbol{0.84 \pm 0.02}$ & $\boldsymbol{0.85 \pm 0.02}$\\
+    I-Ours & $\boldsymbol{0.856 \pm 0.014}$ & $\boldsymbol{0.836 \pm 0.020}$ & $\boldsymbol{0.842 \pm 0.018}$\\
     II-Ours (fine tuned) & $\boldsymbol{0.86 \pm 0.01}$ & $\boldsymbol{0.84 \pm 0.02}$ & $0.84 \pm 0.02$\\

### 2020-06-08 main.tex
Updates from Overleaf

--- 2020-06-08 main.tex
@@
 Classification of tree species in aerial images has been actively investigated~\cite{fassnacht2016review}.
-In \cite{puttemans2018comparing, aparna2018cnn}, the authors present solutions to detect coconut trees based on convolutional neural network (CNN) models.
+In \cite{puttemans2018comparing} and~\cite{aparna2018cnn}, the authors present automatic solutions to detect coconut trees based on convolutional neural network (CNN) models.
+Despite these recent advances, CNN models usually require considerable human effort in image annotation to create large training sets.
 Vargas-Muñoz et al.~\cite{8899005}  propose an active learning approach to reduce human effort in annotating images of coconut trees.
-They divide the problem in two stages: a stage that produces candidate image patches and a stage in which a CNN identifies the candidate patches with a coconut tree.
+They divide the problem in two stages: a stage that produces candidate image patches for having or not a coconut tree and a stage in which a CNN identifies the candidate patches with a coconut tree.
 We propose a different way to produce this classifier.
 
-Despite this success, methods based on deep learning require a lot of annotated images for training, and annotating this volume of data is quite costly.
+methods based on deep learning require a lot of annotated images for training, and annotating this volume of data is quite costly.
 Even interactive methods require large training sets.

### 2020-06-08 main.tex
Updates from Overleaf

--- 2020-06-08 main.tex
@@
 Convolutional Neural Network (CNN) models have well succeeded in image classification applications from different domains.
-However, CNN models usually require intensive manual identification to create large training sets.
-We present a training method for CNN feature extraction that needs only a minimal set of images selected by the user.
-The method learns the filters of each convolutional layer from user-drawn markers in image regions that discriminate classes, allowing direct expert interference in the training process.
+However, CNN models usually require intensive manual annotation to create large training sets.
+A CNN may be conceptually divided into convolutional layers for feature extraction and fully connected layers for classification.
+We present a method that needs a minimal set of user-selected images to train the CNN's feature extractor, considerably reducing the number of images to train its classifier.
+The method learns the filters of each convolutional layer from user-drawn markers in image regions that discriminate classes, allowing direct human expert interference in the training process.
 It does not rely on optimization based on backpropagation, which considerably reduces the number of required images.
-We demonstrate its advantages on the binary classification of coconut-tree images against a popular CNN model.
+We demonstrate its advantages on binary classification of coconut-tree images against a popular CNN model.
 \end{abstract}
@@
  identifying species of trees from aerial images.
-The topic is important for land-use classification, plantation monitoring during a season, and damage assessment of natural disasters, such as drilling and flooding.
-As these plantations can span very large areas, manual identification can be costly, tedious, and susceptible to errors, so automatic methods are necessary.
+The topic is important for land-use classification, plantation monitoring, and damage assessment of natural disasters.
+As the plantations can span very large areas, the manual identification of each tree is costly, tedious, and error-prone, and so automatic classification methods are necessary.
 
@@
 Despite these recent advances, CNN models usually require considerable human effort in image annotation to create large training sets.
-Vargas-Muñoz et al.~\cite{8899005}  propose an active learning approach to reduce human effort in annotating images of coconut trees.
-They divide the problem in two stages: a stage that produces candidate image patches for having or not a coconut tree and a stage in which a CNN identifies the candidate patches with a coconut tree.
-We propose a different way to produce this classifier.
+Vargas-Muñoz et al.~\cite{8899005} propose an approach to mitigate the problem based on active learning.
+The method explores data projection techniques to allow simultaneous annotation of multiple patches as having or not a coconut tree.
+It then uses a CNN to identify the candidate patches with a coconut tree.
+We adopt another alternative in this work -- the design of simplified CNN models from reduced annotated image sets and assistance of a human expert.
 
-methods based on deep learning require a lot of annotated images for training, and annotating this volume of data is quite costly.
-Even interactive methods require large training sets.
-If these methods depend on huge training sets, their application may not be feasible in areas where it is impossible, or the cost of gathering and annotating these data sets is prohibitive.
-Much of the user's interaction with methods based on deep learning occurs through the annotation of the training set, and the user does not have much influence on the training process.
-
-Despite the advances in deep learning over the last decades, especially with most recent deep neural networks, some important questions related to human interaction remain unanswered: (1) How to choose the most reasonable model, in the sense that it provides acceptable effectiveness, being as simple and efficient as needed, for a given classification problem?
+The design of CNN models without the expert as part of the training loop leaves several research questions unanswered: (1) How to choose the most reasonable model, in the sense that it provides acceptable effectiveness, being as simple and efficient as needed, for a given classification problem?
 (2) How to train that model from a reduced number of samples?
@@
 The second one requires to reduce as much as possible the human effort to train a model.
-The third issue is related to human understanding, and it may explore techniques (such as data visualization) to explain the decisions of the model and perhaps to guide human decisions concerning the design of the model.
+The third issue is related to human understanding, and it may explore data visualization techniques to explain the decisions of the model and perhaps to guide human decisions concerning the design of the model.
 The fourth question is also important during training, and it is related to human control over the process.
@@
 
-In this paper, we present some advances towards the answers to the questions (1) and (2).
+In this paper, we present some advances towards answering the questions (1) and (2).
 We also discuss future research related to the questions (3) and (4).
-We present a method to the design of light-weighted convolutional neural networks from markers drawn by an expert on a few images to indicate representative regions of each object of interest for classification.
-From those markers, our approach can learn the filters of each layer, and the convolutional neural network can be designed in a layer-by-layer fashion, with no need for backpropagation in order to build a representative feature space for the classification problem.
+We present a method to the design of simplified convolutional neural networks from markers drawn by a human expert on a few images to indicate the most discriminant regions among the classes.
+From those markers, our approach can learn the filters of each convolutional layer, such that the CNN can be created in a layer-by-layer fashion, with no need for backpropagation in order to build a representative feature space for the classification problem.
 Backpropagation is only used to train a bank of multi-layer perceptrons for local pattern classification.
@@
   }
-  \caption{Exemple of application of convolution with groups centers as filters' weightes: (a) Groups of three classes centered on the origin; (b) Groups after applying convolution and activation operations.
+  \caption{Example of application of convolution with groups centers as filters' weights: (a) Groups of three classes centered on the origin; (b) Groups after applying convolution and activation operations.
 }
@@
         }
-        node [treenode, fill=dodgerBlue, label = {[label distance = 0.5em] }] {ReLU} 
+        node [treenode, fill=tealDeer, label = {[label distance = 0.5em] }] {ReLU} 
         edge from parent
@@
   \hline\hline
-    I-Ours & $\boldsymbol{0.86 \pm 0.01}$ & $\boldsymbol{0.84 \pm 0.02}$ & $\boldsymbol{0.85 \pm 0.02}$\\
-    II-Ours (fine tuned) & $\boldsymbol{0.86 \pm 0.01}$ & $\boldsymbol{0.84 \pm 0.02}$ & $0.84 \pm 0.02$\\
+    I-Ours & $0.856 \pm 0.014$ & $0.836 \pm 0.020$ & $0.842 \pm 0.018$\\
+    II-Ours (fine tuned) & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.84 \pm 0.02$\\
     III-Ours with SVM & $0.836 \pm 0.015 $ & $ 0.793 \pm 0.025$ & $ 0.804 \pm 0.022$\\

### 2020-06-08 main.tex
Updates from Overleaf

@@
 
-The design of CNN models without the expert as part of the training loop leaves several research questions unanswered: (1) How to choose the most reasonable model, in the sense that it provides acceptable effectiveness, being as simple and efficient as needed, for a given classification problem?
-(2) How to train that model from a reduced number of samples?
+The design of CNN models without a human expert as part of the training loop leaves several questions unanswered: (1) How to choose the most reasonable model, with acceptable effectiveness, and as simple as needed for a given classification problem?
+(2) How to train that model from a reduced number of images?
 (3) Can the user explain the decisions of that model?
@@
 The first question requires to explore human knowledge about the classification technique and the problem of interest.
-The second one requires to reduce as much as possible the human effort to train a model.
-The third issue is related to human understanding, and it may explore data visualization techniques to explain the decisions of the model and perhaps to guide human decisions concerning the design of the model.
-The fourth question is also important during training, and it is related to human control over the process.
-They all lead to the importance of involving human experts during the machine learning process.
+The second one requires to reduce as much as possible the human effort to train a CNN model.
+The third issue is related to human understanding.
+It may explore data visualization techniques to explain the model's decisions and guide human decisions concerning the design of the model.
+The fourth question is also essential during training, and it is related to human control over the process.
+They all lead to the importance of involving human experts (users) during the machine learning process.
 
-In this paper, we present some advances towards answering the questions (1) and (2).
-We also discuss future research related to the questions (3) and (4).
-We present a method to the design of simplified convolutional neural networks from markers drawn by a human expert on a few images to indicate the most discriminant regions among the classes.
-From those markers, our approach can learn the filters of each convolutional layer, such that the CNN can be created in a layer-by-layer fashion, with no need for backpropagation in order to build a representative feature space for the classification problem.
-Backpropagation is only used to train a bank of multi-layer perceptrons for local pattern classification.
+In this paper, we present significant advances related to the questions (1) and (2).
+First, we conceptually divide a CNN into convolutional layers for feature extraction and fully connected layers for feature space reduction and classification.
+Each convolutional layer contains a filter bank, an activation function, and alternative operations (e.g., pooling, batch normalization).
+As the number of convolutional layers increases (deeper is the model), higher is the number of annotated samples required to train the model by backpropagation.
 
-We propose a new way of learning the filters of the convolutional layers of deep neural networks that do not depend on the optimization of an error function and, therefore, do not need a large number of training samples.
-The training process is done with user intervention, which adds markers to the training images to indicate relevant features to characterize each class.
-To learn the filters, we get a patch around each marker of a class and cluster them.
-The centroid of each cluster is a filter that can be used to identify features of this cluster.
-The set of centroids of clusters of all clusters is the convolutional layer filters.
-For the next layer, the patches are generated on the output of the previous one.
-To evaluate our method, we use a dataset of aerial images to detect coconut trees.
-We identified that, with few layers and few training samples, we managed to get competitive results with state-of-the-art models.
+We present a method that needs a minimal set of images to learn the filters of each convolutional layer.
+The user selects the number of convolutional layers, a few training images, and draws markers in regions that best discriminate the classes.
+We explain how the classes may appear in distinct clusters of a local feature space defined by patches as extracted from marker pixels at the input of each convolutional layer.
+Patch size (number of weights per filter) is also a user choice.
+The architecture of the network can be optimized by observing its results on a small validation set or based on data visualization techniques~\cite{rauber2016visualizing}.
+The method uses clustering and defines the weights of the filters from the centroids of the clusters.
+Such filters must activate the discriminant regions in the output of the given convolutional layer, and the process repeats in a layer-by-layer fashion with usually no need for new markers.
+
+We demonstrate the advantages of this new technique on binary image classification of coconut-tree images against a popular CNN model, VGG-16~\cite{simonyan2014very}.
+First, the user can directly interfere and verify the markers' effectiveness in the training process.
+The strategy allows user control and a better understanding of CNN models.
+Second, very few images per class (e.g., less than five) are usually enough.
+Third, by eliminating the need for optimization of convolutional layers by backpropagation, the method can considerably reduce the number of annotated images to train the fully connected layers.
+Fourth, the technique is application-independent, and so it might be useful in other image classification problems.
 
@@
 Vectors similar to $\vect(F)$ must fall on the same side of the hyperplane as $\vect(F)$.
-Therefore, if vector $\vect(\pi)$ is on the same side of $H$ as $\vect(F)$, then the doct product between $\vect(\pi)$ and $\vect(F)$ is positive, and it is negative otherwise.
-Thus, the doct product between the centroid of a cluster and each of the vectors of that cluster must be positive, which implies that the centroid is a good filter to identify the pattern described by that cluster.
-Hence, if we throw away the negative values, applying convolution with centroids as filters must highlight regions that has a certain pattern that the network designer believes is important.
+Therefore, if vector $\vect(\pi)$ is on the same side of $H$ as $\vect(F)$, then the dot product between $\vect(\pi)$ and $\vect(F)$ is positive, and it is negative otherwise.
+Thus, the dot product between the centroid of a cluster and each of the vectors of that cluster must be positive, which implies that the centroid is a good filter to identify the pattern described by that cluster.
+Hence, if we throw away the negative values, applying convolution with centroids as filters must highlight regions with a certain pattern that the network designer believes is important.
 
-Since all the patches used to produce the filters are centered around the origin, as well as the images of $\D$, we have that the hyperplanes ortogonal to the filters contain the origin, therefor there is no need for bias.
+Since all patches used to produce the filters are centered around the origin, as well as the images of $\D$, we have that the hyperplanes orthogonal to the filters contain the origin.
+Therefore, there is no need for bias.
 
@@
 After the convolution operation, we apply the ReLU function to eliminate negative activations and the max pooling operation to aggregate local information.
-We preserve the output dimensions of convolutional layers so that we can use the same markers used in the previous layer.
-The new patches are generated from the previeus convolutional layer output and the whole process is repeated to find the filters of the next one.
+We preserve the convolutional layers' output dimensions so that we can use the same markers used in the previous layer.
+The new patches are generated from the previous convolutional layer output, and the whole process is repeated to find the filters of the next one.
 
@@
 We use a dataset formed from patches of aerial images made by \cite{8899005} from imagery provided by WeRobotics.
-The imagery with a resolution of 8cm was captured by the World Bank's UAVs for Disaster Resilience Program, collaborating with WeRobotics and OpenAerialMap in the Kindom of Tonga in October 2017.
+The imagery with a resolution of 8cm was captured by the World Bank's UAVs for Disaster Resilience Program, collaborating with WeRobotics and OpenAerialMap in the Kingdom of Tonga in October 2017.
 The Humanitarian OpenStreetMap community provided the annotation.
@@
     I-Ours & $0.856 \pm 0.014$ & $0.836 \pm 0.020$ & $0.842 \pm 0.018$\\
-    II-Ours (fine tuned) & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.84 \pm 0.02$\\
-    III-Ours with SVM & $0.836 \pm 0.015 $ & $ 0.793 \pm 0.025$ & $ 0.804 \pm 0.022$\\
+    II-Ours (B) & $0.851 \pm 0.018$ & $0.834 \pm 0.023$ & $0.839 \pm 0.021$\\
+    III-Ours (SVM) & $0.836 \pm 0.015 $ & $ 0.793 \pm 0.025$ & $ 0.804 \pm 0.022$\\
     IV-VGG & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
-    V-VGG (fine tuned) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
+    V-VGG (PT) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
   \hline

### 2020-06-11 main.tex
Updates from Overleaf

@@
 However, CNN models usually require intensive manual annotation to create large training sets.
-A CNN may be conceptually divided into convolutional layers for feature extraction and fully connected layers for classification.
-We present a method that needs a minimal set of user-selected images to train the CNN's feature extractor, considerably reducing the number of images to train its classifier.
-The method learns the filters of each convolutional layer from user-drawn markers in image regions that discriminate classes, allowing direct human expert interference in the training process.
-It does not rely on optimization based on backpropagation, which considerably reduces the number of required images.
-We demonstrate its advantages on binary classification of coconut-tree images against a popular CNN model.
+One may conceptually divide a CNN into convolutional layers for feature extraction and fully connected layers for feature space reduction and classification.
+We present a method that needs a minimal set of user-selected images to train the CNN's feature extractor, reducing the number of required images to train the fully connected layers.
+The method learns the filters of each convolutional layer from user-drawn markers in image regions that discriminate classes, allowing better user control and understanding of the training process.
+It does not rely on optimization based on backpropagation, and we demonstrate its advantages on the binary classification of coconut-tree images against one of the most popular CNN models.
 \end{abstract}
@@
 Deep learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}.
-In remote sensing, the applications have involved segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic, hamaguchi2018effective}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}.
+In remote sensing, the applications may involve segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic, hamaguchi2018effective}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}, for instance.
 In this work, we are interested in
@@
 Despite these recent advances, CNN models usually require considerable human effort in image annotation to create large training sets.
-Vargas-Muñoz et al.~\cite{8899005} propose an approach to mitigate the problem based on active learning.
-The method explores data projection techniques to allow simultaneous annotation of multiple patches as having or not a coconut tree.
-It then uses a CNN to identify the candidate patches with a coconut tree.
-We adopt another alternative in this work -- the design of simplified CNN models from reduced annotated image sets and assistance of a human expert.
+Vargas-Muñoz et al.~\cite{8899005} propose an active learning approach to mitigate the problem.
+The method explores data projection techniques to allow simultaneous annotation of multiple patches as having or not one coconut tree.
+It then uses a CNN to identify the candidate patches with one coconut tree.
+We adopt another alternative -- the design of simplified CNN models from small training sets with user interaction, being the user an expert in the design of CNN models.
 
-The design of CNN models without a human expert as part of the training loop leaves several questions unanswered: (1) How to choose the most reasonable model, with acceptable effectiveness, and as simple as needed for a given classification problem?
-(2) How to train that model from a reduced number of images?
-(3) Can the user explain the decisions of that model?
-(4) Can that model improve from human corrections?
-The first question requires to explore human knowledge about the classification technique and the problem of interest.
-The second one requires to reduce as much as possible the human effort to train a CNN model.
+The design of CNN models without a human as part of the training loop leaves several questions unanswered: (1) How to find a useful and simplified model for a given classification problem?
+(2) How to train that model from a minimum number of annotated images?
+(3) Can the user explain the decisions of the model?
+(4) Can the model improve from label corrections?
+The first question requires human knowledge about CNN modeling and the problem of interest.
+The second one requires to reduce the human effort to train a CNN model.
 The third issue is related to human understanding.
-It may explore data visualization techniques to explain the model's decisions and guide human decisions concerning the design of the model.
-The fourth question is also essential during training, and it is related to human control over the process.
-They all lead to the importance of involving human experts (users) during the machine learning process.
-
-In this paper, we present significant advances related to the questions (1) and (2).
+It may explore data visualization techniques to explain the model's decisions and to guide the user in the design of the model.
+The fourth question is also essential during training, and it is related to user control over the process.
+They all lead to the importance of involving an expert user in the machine learning process.
+In this paper, we present significant advances related to questions (1) and (2).
 First, we conceptually divide a CNN into convolutional layers for feature extraction and fully connected layers for feature space reduction and classification.
 Each convolutional layer contains a filter bank, an activation function, and alternative operations (e.g., pooling, batch normalization).
-As the number of convolutional layers increases (deeper is the model), higher is the number of annotated samples required to train the model by backpropagation.
+As the number of convolutional layers increases (deeper is the model), and so higher is the number of annotated images required to train the model by backpropagation.
 
-We present a method that needs a minimal set of images to learn the filters of each convolutional layer.
-The user selects the number of convolutional layers, a few training images, and draws markers in regions that best discriminate the classes.
+In order to reduce the need for large annotated training sets by exploiting the user knowledge, we present a method, called FLIM (\emph{Feature Learning from Image Markers}), that needs a minimal set of images to learn the filters of each convolutional layer.
+The user selects the number of convolutional layers, their filter sizes, a few training images, and draws markers in regions of the input images that best discriminate the classes.
 We explain how the classes may appear in distinct clusters of a local feature space defined by patches as extracted from marker pixels at the input of each convolutional layer.
-Patch size (number of weights per filter) is also a user choice.
-The architecture of the network can be optimized by observing its results on a small validation set or based on data visualization techniques~\cite{rauber2016visualizing}.
-The method uses clustering and defines the weights of the filters from the centroids of the clusters.
-Such filters must activate the discriminant regions in the output of the given convolutional layer, and the process repeats in a layer-by-layer fashion with usually no need for new markers.
+FLIM defines the weights of the filters as the centroids of the clusters.
+Such filters are expected to activate discriminant regions in the output of the given convolutional layer, and the process repeats in a layer-by-layer fashion.
+The architecture of the CNN could also be optimized by observing its results on a small validation set or based on data visualization techniques~\cite{rauber2016visualizing}.
 
-We demonstrate the advantages of this new technique on binary image classification of coconut-tree images against a popular CNN model, VGG-16~\cite{simonyan2014very}.
-First, the user can directly interfere and verify the markers' effectiveness in the training process.
-The strategy allows user control and a better understanding of CNN models.
-Second, very few images per class (e.g., less than five) are usually enough.
-Third, by eliminating the need for optimization of convolutional layers by backpropagation, the method can considerably reduce the number of annotated images to train the fully connected layers.
-Fourth, the technique is application-independent, and so it might be useful in other image classification problems.
+We demonstrate the advantages of this interactive technique on binary image classification of coconut-tree images against  VGG-16~\cite{simonyan2014very}.
+First, the user can better understand and control the training process by observing the effectiveness of  marker selection.
+Second, very few images per class (e.g., less than five) seem to be  enough.
+Third, by eliminating the optimization of the convolutional layers by backpropagation, FLIM can reduce the number of annotated images to train the fully connected layers.
+Fourth, FLIM is application-independent, and so it might be useful in other image classification problems.
 
-\section{Learning filters from markers}
+\section{FLIM: Feature Learning  from Image Markers}
 \label{sec:method}
 
-Convolutional neural networks perform well in image classification tasks due to its capability of extracting features.
-That is, convolutional layers produce great descriptors that can be used by the classifier.
-Conventionally, during the training process, an algorithm tries to minimize a loss function that measures how good the classifier is performing with the features given by the feature extractor.
-The algorithm of optimization tries through backpropagation to find out which features to extract.
-Our method is based on the idea that the network designer is an expert in the application domain or knows well enough to indicate which are the relevant characteristics of the objects of interest.
-Hence, it does not rely on the optimization of some loss function to learn filters for those relevant characteristics.
-In our approach, the network designer highlights features that she believes are the ones that characterize each class by placing markers.
-Those markers are used to learn the filters to identify the features they represent.
+A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation in those regions by eliminating negative activations, and aggregate the resulting activations (pooling and flattening) into a highly dimensional and sparse feature space, in which even a linear classifier should separate the classes with reasonable success.
+Nevertheless, it is common to assume that the classes are at least piecewise linearly separable and use fully connected layers -- a Multi-Layer Perceptron (MLP) classifier -- for feature space reduction and classification.
 
-Let $\D$ be a dataset of images, and let $I \in \D$ be an image with dimensions $X \times Y$.
+The training of a CNN usually rely on  weight optimization by backpropagation, which requires a higher number of annotated images as deeper is the model.
+In FLIM, we eliminate the need for backpropagation by finding a set ${\cal F}$ of filters that can detect discriminant regions from all classes.
+
+The user selects a few training images $I$ to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes.
+We with to estimate the weights of filters $F\in {\cal F}$, such that the convolution between an image $I$ and $F$ will generate positive activations in regions whose pattern is represented by $F$.
+
+For a problem with $c$ classes, let $\lambda(p)=i\in \{1,2,\ldots,c\}$ be the label of a marker pixel $p$ from class $i$.
+Let ${\cal M}_I$ be a set of marker pixels drawn in image $I$ and $P_I(p)$ be a patch of size $k\times k$ pixels and $m$ image channels, around a marker pixel $p\in {\cal M}_I$.
+The patch values compose a \emph{local feature vector} $\vect\left[P_I(p)\right]$ with $k\times k\times m$ values.
+Let ${\cal P}$ be the set of all patches around marker pixels from all images $I\in {\cal D}$, with representative examples  $\vect\left[P_I(p)\right]$ from each of the $c$ classes.
+\begin{eqnarray}
+{\cal P} & = & \bigcup_{I\in {\cal D},p\in {\cal M}_I, \lambda(p)\in [1,c]} P_I(p)
+\label{eq.pset}
+\end{eqnarray}
+The patch set ${\cal P}$ defines samples from all classes in a local feature space $\Re^{k\times k\times m}$.
+
+The classes are usually not separated in this space but the convolution of an image $I$ with a set of filters ${\cal F}$ of sizes $k\times k\times m$ should facilitate their separation in the new space
+
+is equivalent to computing the inner product between $\vect\left[P_I(q)\right]$ and  $\vect\left[F\right]$ for all pixels $q$ in the image $I$.
+
  A marker $m$ is associated with a pixel $p(m) = (x, y)$ and has a label $L(m) = c$.
@@
   \hline\hline
-    I-Ours & $0.856 \pm 0.014$ & $0.836 \pm 0.020$ & $0.842 \pm 0.018$\\
-    II-Ours (B) & $0.851 \pm 0.018$ & $0.834 \pm 0.023$ & $0.839 \pm 0.021$\\
-    III-Ours (SVM) & $0.836 \pm 0.015 $ & $ 0.793 \pm 0.025$ & $ 0.804 \pm 0.022$\\
-    IV-VGG & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
-    V-VGG (PT) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
+      FLIM + SVM & $0.856 \pm 0.011 $ & $ 0.831 \pm 0.019$ & $ 0.838 \pm 0.017$\\ 
+       VGG-FE + SVM & $0.855 \pm 0.001$ & $0.816 \pm 0.007$ & $0.826 \pm 0.006 $ \\\hline
+
+      FLIM & $0.863 \pm 0.002$ & $0.849 \pm 0.005$ & $0.854 \pm 0.004$\\
+    FLIM-FT &  &  & \\
+
+    VGG &  &  & \\
+    VGG-FT & $0.872 \pm 0.007$ & $0.844 \pm 0.015$ & $0.851 \pm 0.014 $ \\
+
   \hline
@@
 
+\begin{table*}
+  \begin{center}
+  \begin{tabular}{|l|c|c|c|c|c|c| }
+  \hline
+   Method & Pr0 & Pr1 & Re0 & Re1 & Fs0 & Fs1 \\
+  \hline\hline
+      FLIM + SVM & $0.933 \pm 0.003 $ & $ 0.612 \pm 0.038$ & $ 0.838 \pm 0.024$ & $0.809 \pm 0.007 $ & $ 0.883 \pm 0.014$ & $0.697 \pm 0.025 $ \\ 
+       VGG-FE + SVM & $0.941 \pm 0.005$ & $0.580 \pm 0.013$ & $0.809 \pm 0.014 $ & $ 0.839 \pm 0.016$ & $ 0.870 \pm 0.006$ & $0.686 \pm 0.005 $ \\\hline
+
+      FLIM & $0.929 \pm 0.002$ & $0.654 \pm 0.016$ & $0.869 \pm 0.010$ & $0.789 \pm 0.009 $ & $0.898 \pm 0.004 $ & $ 0.715 \pm 0.006$ \\
+    FLIM-FT &  &  &  &  &  &  \\
+
+    VGG &  &  &  &  &  & \\
+    VGG-FT & $0.948 \pm 0.001$ & $0.629 \pm 0.032$ & $0.841 \pm 0.021 $ & $ 0.854 \pm 0.006$ & $ 0.891 \pm 0.012$ & $0.724 \pm 0.020 $  \\
+
+  \hline
+  \end{tabular}
+  \end{center}
+  \caption{Results per class.}
+  \label{tab:results_per_class}
+\end{table*}
+
 We tested three scenarios: (I) using out method to learn the feature extractor weights and the classifier with backpropagation; (II) using our method as initialization and train the whole model with backpropagation; (III) using SVM Linear as classification.
-For VGG-16, we tested two secenarios: (IV) initializing with Xavier initialization and training from screath; (V) using ImageNet weights and training with the training set images.
+For VGG-16, we tested two scenarios: (IV) initializing with Xavier initialization and training from scratch; (V) using ImageNet weights and training with the training set images.
 The results can be seen in Table \ref{tab:results}.

### 2020-06-11 main.tex
Updates from Overleaf

@@
 Nevertheless, it is common to assume that the classes are at least piecewise linearly separable and use fully connected layers -- a Multi-Layer Perceptron (MLP) classifier -- for feature space reduction and classification.
-
 The training of a CNN usually rely on  weight optimization by backpropagation, which requires a higher number of annotated images as deeper is the model.
@@
 
-The user selects a few training images $I$ to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes.
-We with to estimate the weights of filters $F\in {\cal F}$, such that the convolution between an image $I$ and $F$ will generate positive activations in regions whose pattern is represented by $F$.
+In FLIM, the user selects a few training images $I$ to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes.
+We wish to estimate the weights of filters $F\in {\cal F}$, such that the convolution between an image $I$ and $F$ will generate positive activations in regions whose local patterns are detected by $F$.
+In order to understand this process, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect\left[P_I(p)\right] \cdot  \vect\left[F\right]$, where $P_I(p)$ is a patch with $k\times k$ pixels and $m$ image channels around $p$, $F$ is a filter with $k\times k\times m$ weights, and $\vect\left[.\right]$ is the vectorization operation.
+A filter $F$ can discriminate a class $i\in \{1,2,\ldots,c\}$ among $c$ classes when it generates positive results for local patterns $\vect\left[P_I(p)\right]$ of class $i$ and negative results for patterns from other classes.
+That is, $\vect\left[F\right]$ plays the role of a normal vector of a hyperplane in $\Re^{k\times k\times m}$ and we expect that ${\cal F}$ will contain filters to detect local patterns from all classes in different positions $p$.
 
-For a problem with $c$ classes, let $\lambda(p)=i\in \{1,2,\ldots,c\}$ be the label of a marker pixel $p$ from class $i$.
-Let ${\cal M}_I$ be a set of marker pixels drawn in image $I$ and $P_I(p)$ be a patch of size $k\times k$ pixels and $m$ image channels, around a marker pixel $p\in {\cal M}_I$.
-The patch values compose a \emph{local feature vector} $\vect\left[P_I(p)\right]$ with $k\times k\times m$ values.
-Let ${\cal P}$ be the set of all patches around marker pixels from all images $I\in {\cal D}$, with representative examples  $\vect\left[P_I(p)\right]$ from each of the $c$ classes.
+For a problem with $c$ classes where $\lambda(p)=i\in \{1,2,\ldots,c\}$ is the label of a marker pixel $p$ from class $i$, let ${\cal M}_I$ be a set of marker pixels drawn in image $I\in {\cal D}$ and $P_I(p)$ be the respective patch around a pixel $p\in {\cal M}_I$.
+Let ${\cal P}$ be the set of all patches around marker pixels from all images $I\in {\cal D}$, with representative examples  $\vect\left[P_I(p)\right]$ from all $c$ classes.
 \begin{eqnarray}
@@
 \end{eqnarray}
-The patch set ${\cal P}$ defines samples from all classes in a local feature space $\Re^{k\times k\times m}$.
-
-The classes are usually not separated in this space but the convolution of an image $I$ with a set of filters ${\cal F}$ of sizes $k\times k\times m$ should facilitate their separation in the new space
-
-is equivalent to computing the inner product between $\vect\left[P_I(q)\right]$ and  $\vect\left[F\right]$ for all pixels $q$ in the image $I$.
-
- A marker $m$ is associated with a pixel $p(m) = (x, y)$ and has a label $L(m) = c$.
-Let $M(I)$ be the set of markers of image $I$.
-For each maker $m \in M(I)$, we create a patch $P$ with dimensions $k \times k$ from a window centered in the marker's pixel as $P(I, x ,y) = I(x+i, y+j)$ for $i, j = -\floor{k/2}, \ldots, \floor{k/2}$.We pad images with $0$ so that we can create patches from border pixels.
-For each class, we create a set of patches made using the markers of that class.
-Let \[\mathcal{P}_c = \bigcup_{ I \in \D, m \in M(I) :\, L(m) = c}{P(I, p(m))}\] be the set of all patches made from markers of class $c$.
-
-Given set of patches $\Pa_c$, we can use a clustering algorithm to discover groups in $\Pa_c$.
-Each group represents local patterns of class $c$, highlighted by the designer through the markers.
-Since the centroid of each cluster must be a good representative of the characteristics of the group, we can use them as filters to enhance those patterns in the images of $\D$.
-We must do this for each class.
-
-However, for it to work properly, patches and images need to be centered at the origin.
-For this, we use the information from the patches of all classes to calculate the mean and standard deviation per band.
-If the patches are representative of images in $\D$, the mean and standard deviation per band of $\Pa$ are close to the mean and standard deviation per band in $\D$.
+The patch set ${\cal P}$ defines samples from all classes in $\Re^{k\times k\times m}$ and we should expect that a clustering operation will create groups with similar local patterns.
+Ideally, each group should contain local patterns of a same class.
+However, especially in the first layers, some classes might share similar patterns and we should at least expect that they will be detected in different regions of the image.
+The groups must be shifted to the origin of $\Re^{k\times k\times m}$ by subtracting the mean vector of the samples in ${\cal P}$.
+Optionally, the standard deviation of each feature can be computed and used for feature standardization.
+We call this operation \emph{patch-based normalization}.
+The centroid of each cluster defines the weights of each filter $F\in {\cal F}$ and we also force norm $1$ to avoid preferences among them.
+This centralization for filter definition is paramount to eliminate activations in regions from other classes, whose patterns fall in the negative side of the hyperplane represented by $\vect\left[F\right]$.
+Figure~\ref{fig:filter} illustrates how it works.
+....
 
@@
 
-Let $F$ be a filter with dimension $k \times k$, and let $\pi$ be the patch around pixel $(x,y)$ also with dimension $k \times k$.
+%Let $F$ be a filter with dimension $k \times k$, and let $\pi$ be the patch around pixel $(x,y)$ also with dimension $k \times k$.
 By the definition of convolution in deep learning, we can interpret this operation for a pixel $(x, y)$ as the dot product between the kernel vector $\vect(F)$ and the vector of a patch $\vect(\pi)$ around this pixel, that is, the convolution in $(x, y)$ is the same as $\vect(F) \cdot \vect(\pi)$.
@@
 
-Since all patches used to produce the filters are centered around the origin, as well as the images of $\D$, we have that the hyperplanes orthogonal to the filters contain the origin.
+%Since all patches used to produce the filters are centered around the origin, as well as the images of $\D$, we have that the hyperplanes orthogonal to the filters contain the origin.
 Therefore, there is no need for bias.
@@
 
-\begin{table*}
-  \begin{center}
-  \begin{tabular}{|l|c|c|c|c|c|c| }
-  \hline
-   Method & Pr0 & Pr1 & Re0 & Re1 & Fs0 & Fs1 \\
-  \hline\hline
-      FLIM + SVM & $0.933 \pm 0.003 $ & $ 0.612 \pm 0.038$ & $ 0.838 \pm 0.024$ & $0.809 \pm 0.007 $ & $ 0.883 \pm 0.014$ & $0.697 \pm 0.025 $ \\ 
-       VGG-FE + SVM & $0.941 \pm 0.005$ & $0.580 \pm 0.013$ & $0.809 \pm 0.014 $ & $ 0.839 \pm 0.016$ & $ 0.870 \pm 0.006$ & $0.686 \pm 0.005 $ \\\hline
-
-      FLIM & $0.929 \pm 0.002$ & $0.654 \pm 0.016$ & $0.869 \pm 0.010$ & $0.789 \pm 0.009 $ & $0.898 \pm 0.004 $ & $ 0.715 \pm 0.006$ \\
-    FLIM-FT &  &  &  &  &  &  \\
-
-    VGG &  &  &  &  &  & \\
-    VGG-FT & $0.948 \pm 0.001$ & $0.629 \pm 0.032$ & $0.841 \pm 0.021 $ & $ 0.854 \pm 0.006$ & $ 0.891 \pm 0.012$ & $0.724 \pm 0.020 $  \\
-
-  \hline
-  \end{tabular}
-  \end{center}
-  \caption{Results per class.}
-  \label{tab:results_per_class}
-\end{table*}
-
 We tested three scenarios: (I) using out method to learn the feature extractor weights and the classifier with backpropagation; (II) using our method as initialization and train the whole model with backpropagation; (III) using SVM Linear as classification.

### 2020-06-12 main.tex
new resutls
Updates from Overleaf

@@
 
-A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation in those regions by eliminating negative activations, and aggregate the resulting activations (pooling and flattening) into a highly dimensional and sparse feature space, in which even a linear classifier should separate the classes with reasonable success.
+A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation by eliminating negative activations, and aggregate the resulting activations (pooling and flattening) into a highly dimensional and sparse feature space, in which even a linear classifier should separate the classes with reasonable success.
 Nevertheless, it is common to assume that the classes are at least piecewise linearly separable and use fully connected layers -- a Multi-Layer Perceptron (MLP) classifier -- for feature space reduction and classification.
 The training of a CNN usually rely on  weight optimization by backpropagation, which requires a higher number of annotated images as deeper is the model.
-In FLIM, we eliminate the need for backpropagation by finding a set ${\cal F}$ of filters that can detect discriminant regions from all classes.
+In FLIM, we eliminate the need of backpropagation to train convolutional layers by finding a set ${\cal F}$ of filters that can detect discriminant regions from all classes.
 
-In FLIM, the user selects a few training images $I$ to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes.
-We wish to estimate the weights of filters $F\in {\cal F}$, such that the convolution between an image $I$ and $F$ will generate positive activations in regions whose local patterns are detected by $F$.
-In order to understand this process, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect\left[P_I(p)\right] \cdot  \vect\left[F\right]$, where $P_I(p)$ is a patch with $k\times k$ pixels and $m$ image channels around $p$, $F$ is a filter with $k\times k\times m$ weights, and $\vect\left[.\right]$ is the vectorization operation.
+The user selects a few training images $I$ to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes.
+The convolution between an image $I$ and a filter $F\in {\cal F}$ generates positive activations in regions whose local patterns are detected by $F$.
+We wish then to estimate the weights of $F$ such that those local patterns are characteristic of one given class.
+For sake of clarity, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect\left[P_I(p)\right] \cdot  \vect\left[F\right]$, where $P_I(p)$ is a patch with $k\times k$ pixels and $m$ image channels around $p$, $F$ is a filter with $k\times k\times m$ weights, and $\vect\left[.\right]$ is the vectorization operation.
 A filter $F$ can discriminate a class $i\in \{1,2,\ldots,c\}$ among $c$ classes when it generates positive results for local patterns $\vect\left[P_I(p)\right]$ of class $i$ and negative results for patterns from other classes.
-That is, $\vect\left[F\right]$ plays the role of a normal vector of a hyperplane in $\Re^{k\times k\times m}$ and we expect that ${\cal F}$ will contain filters to detect local patterns from all classes in different positions $p$.
+That is, $\vect\left[F\right]$ is the normal vector of a hyperplane in $\Re^{k\times k\times m}$ which detects the patterns of class $i$ at its positive side.
+We wish to estimate ${\cal F}$ such that its filters will detect patterns from all classes in different positions $p$.
 
@@
 \end{eqnarray}
-The patch set ${\cal P}$ defines samples from all classes in $\Re^{k\times k\times m}$ and we should expect that a clustering operation will create groups with similar local patterns.
-Ideally, each group should contain local patterns of a same class.
-However, especially in the first layers, some classes might share similar patterns and we should at least expect that they will be detected in different regions of the image.
-The groups must be shifted to the origin of $\Re^{k\times k\times m}$ by subtracting the mean vector of the samples in ${\cal P}$.
+The patch set ${\cal P}$ defines samples from all classes in $\Re^{k\times k\times m}$ and a clustering operation on ${\cal P}$ creates groups with similar local patterns.
+Ideally, each group should contain patterns of a same class.
+However, some classes might share similar patterns.
+We should at least expect that their patterns be detected in different image regions.
+The groups must be shifted to the origin of the feature space by subtracting the mean vector of the samples in ${\cal P}$.
 Optionally, the standard deviation of each feature can be computed and used for feature standardization.
 We call this operation \emph{patch-based normalization}.
-The centroid of each cluster defines the weights of each filter $F\in {\cal F}$ and we also force norm $1$ to avoid preferences among them.
-This centralization for filter definition is paramount to eliminate activations in regions from other classes, whose patterns fall in the negative side of the hyperplane represented by $\vect\left[F\right]$.
-Figure~\ref{fig:filter} illustrates how it works.
-....
+The centroid of each cluster defines the weights of each filter $F\in {\cal F}$ and we force norm $\|\vect\left[F\right]\|=1$ to avoid preferences among them.
+The centralization for filter definition is paramount to eliminate activations in regions from other classes, whose patterns fall in the negative side of the hyperplane with normal $\vect\left[F\right]$.
+Figure~\ref{fig:filter}a illustrates three groups with color-coded samples from two classes in a hypothetical 2D feature space.
+The path-based normalization followed by convolution with filters obtained as centroids of the three groups creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters.
 
-\begin{comment}
-\begin{figure}[t]
-  \begin{center}
-     \includegraphics[width=0.8\linewidth]{figures/filter.png}
-  \end{center}
-  \caption{Filter learned from clustering.}
-  \label{fig:filter}
-\end{figure}
-\end{comment}
-
-%Let $F$ be a filter with dimension $k \times k$, and let $\pi$ be the patch around pixel $(x,y)$ also with dimension $k \times k$.
-By the definition of convolution in deep learning, we can interpret this operation for a pixel $(x, y)$ as the dot product between the kernel vector $\vect(F)$ and the vector of a patch $\vect(\pi)$ around this pixel, that is, the convolution in $(x, y)$ is the same as $\vect(F) \cdot \vect(\pi)$.
-Thus, let $H$ be a hyperplane orthogonal to vector $\vect(F)$.
-Vectors similar to $\vect(F)$ must fall on the same side of the hyperplane as $\vect(F)$.
-Therefore, if vector $\vect(\pi)$ is on the same side of $H$ as $\vect(F)$, then the dot product between $\vect(\pi)$ and $\vect(F)$ is positive, and it is negative otherwise.
-Thus, the dot product between the centroid of a cluster and each of the vectors of that cluster must be positive, which implies that the centroid is a good filter to identify the pattern described by that cluster.
-Hence, if we throw away the negative values, applying convolution with centroids as filters must highlight regions with a certain pattern that the network designer believes is important.
-
-%Since all patches used to produce the filters are centered around the origin, as well as the images of $\D$, we have that the hyperplanes orthogonal to the filters contain the origin.
-Therefore, there is no need for bias.
-
-Each convolutional layer is trained individually, one layer at a time.
-After the convolution operation, we apply the ReLU function to eliminate negative activations and the max pooling operation to aggregate local information.
-We preserve the convolutional layers' output dimensions so that we can use the same markers used in the previous layer.
-The new patches are generated from the previous convolutional layer output, and the whole process is repeated to find the filters of the next one.
-
-\begin{figure}[!t]
+\begin{figure}[!ht]
   \centering
@@
   }
-  \caption{Example of application of convolution with groups centers as filters' weights: (a) Groups of three classes centered on the origin; (b) Groups after applying convolution and activation operations.
-}
+  \caption{(a) A 2D feature space with color-coded samples from two classes distributed in three groups.
+(b) After patch-based normalization and convolution with filters obtained from each group, the classes can be more easily detected by two filters in a new 3D feature space.}
   \label{fig:filter}
 \end{figure}
+
+Each convolutional layer is trained individually, one layer at a time, to find its filter set ${\cal F}$.
+The number of filters per layer depends on the clustering technique.
+We use $K$-means, then the user must specify the number $K=|{\cal F}|$ of filters.
+After the convolution operation, we apply the ReLU function to eliminate negative activations and the max-pooling operation to aggregate local information.
+Note that, apart from the initial marker selection, the training process is automatic.
+We have preserved the convolutional layers' output dimensions, such that we can use the same markers used in the input layer to find the filters of layer $L+1$ from the resulting images obtained at the output of layer $L$.
+Figure~\ref{fig:tree} shows one example of a simple network projected by FLIM for the experiments of the next section.
 
@@
   \end{center}
+  \caption{Example of a network with a single convolutional layer followed by a classifier, as projected by FLIM and used for the experiments in this work.}
   \label{fig:tree}
 \end{figure}
-
-To exemplify our method, we create a synthetic dataset of points of two classes.
-We can see this dataset in Figure \ref{fig:ex-groups}.
-The points are distributed in two regions, and it is not possible to separate the yellow class linearly from the violet class.
-To center the points,  we subtracted the mean and divided by the standard deviation.
-If we look only to violet points, it is clear that they form two clusters.
-In turn, the yellow dots form a single cluster.
-We then take the center of each of these three clusters and apply convolution to the set of points using these centers as filters.
-The result of the convolution after applying the ReLU function can be seen in Figure \ref{fig:ex-groups-after}.
-For most points after the transformation made by convolution and the ReLU function, only one coordinate was positive, this being the coordinate which value is the internal product with the center of the cluster of which these points are part.
-These points lie along the three axes of space $\mathbb{R}^3$, each cluster on an axis.
-The other points are farthest from their cluster center, and the internal product with the center of another cluster is also positive.
-Cluster centers functioned as good filters to detect similar points, even though they detected points from other clusters.
 

### 2020-06-12 main.tex
Updates from Overleaf

@@
 In remote sensing, the applications may involve segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic, hamaguchi2018effective}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}, for instance.
-In this work, we are interested in
- identifying species of trees from aerial images.
+In this work, we are interested in identifying species of trees from aerial images.
 The topic is important for land-use classification, plantation monitoring, and damage assessment of natural disasters.
-As the plantations can span very large areas, the manual identification of each tree is costly, tedious, and error-prone, and so automatic classification methods are necessary.
+As the plantations can span vast areas, the manual identification of each tree is costly, tedious, and error-prone, and so automatic classification methods are necessary.
 
@@
 Vargas-Muñoz et al.~\cite{8899005} propose an active learning approach to mitigate the problem.
-The method explores data projection techniques to allow simultaneous annotation of multiple patches as having or not one coconut tree.
-It then uses a CNN to identify the candidate patches with one coconut tree.
+The method explores data projection techniques to allow simultaneous annotation of multiple patches as having or not coconut trees.
+It then uses a CNN to identify the candidate patches with coconut trees.
 We adopt another alternative -- the design of simplified CNN models from small training sets with user interaction, being the user an expert in the design of CNN models.
@@
 The user selects the number of convolutional layers, their filter sizes, a few training images, and draws markers in regions of the input images that best discriminate the classes.
-We explain how the classes may appear in distinct clusters of a local feature space defined by patches as extracted from marker pixels at the input of each convolutional layer.
-FLIM defines the weights of the filters as the centroids of the clusters.
-Such filters are expected to activate discriminant regions in the output of the given convolutional layer, and the process repeats in a layer-by-layer fashion.
+\textcolor{blue}{Classes may appear in distinct clusters of a local feature space defined by patches extracted from marker pixels at each convolutional layer's input.} FLIM defines the weights of the filters as the centroids of the clusters.
+Such filters are expected to activate discriminant regions in the given convolutional layer's output, and the process repeats in a layer-by-layer fashion.
 The architecture of the CNN could also be optimized by observing its results on a small validation set or based on data visualization techniques~\cite{rauber2016visualizing}.
@@
 Second, very few images per class (e.g., less than five) seem to be enough.
-Third, by eliminating the optimization of the convolutional layers by backpropagation, FLIM can reduce the number of annotated images to train the fully connected layers.
-Fourth, FLIM is application-independent, and so it might be useful in other image classification problems.
+\textcolor{red}{Third, by eliminating the optimization of the convolutional layers by backpropagation, FLIM can reduce the number of annotated images to train the fully connected layers.} Fourth, FLIM is application-independent, and so it might be useful in other image classification problems.
 
@@
 
-A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation by eliminating negative activations, and aggregate the resulting activations (pooling and flattening) into a highly dimensional and sparse feature space, in which even a linear classifier should separate the classes with reasonable success.
-Nevertheless, it is common to assume that the classes are at least piecewise linearly separable and use fully connected layers -- a Multi-Layer Perceptron (MLP) classifier -- for feature space reduction and classification.
-The training of a CNN usually rely on  weight optimization by backpropagation, which requires a higher number of annotated images as deeper is the model.
-In FLIM, we eliminate the need of backpropagation to train convolutional layers by finding a set ${\cal F}$ of filters that can detect discriminant regions from all classes.
+\textcolor{blue}{A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation by eliminating negative activations, and aggregate the resulting activations locally (pooling) and globally (flattening) into a highly dimensional and sparse feature space, in which even a linear classifier should separate the classes with reasonable success.} Nevertheless, it is common to assume that the classes are at least piecewise linearly separable and apply fully connected layers -- a Multi-Layer Perceptron (MLP) classifier -- for feature space reduction and classification.
+The training of a CNN usually relies on weight optimization using backpropagation, which requires a higher number of annotated images as deep as the model.
+In FLIM, we eliminate the need for backpropagation to train convolutional layers by finding a set ${\cal F}$ of filters that can detect discriminant regions from all classes.
 
-The user selects a few training images $I$ to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes.
+The user selects a few training images to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes.
 The convolution between an image $I$ and a filter $F\in {\cal F}$ generates positive activations in regions whose local patterns are detected by $F$.
 We wish then to estimate the weights of $F$ such that those local patterns are characteristic of one given class.
-For sake of clarity, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect\left[P_I(p)\right] \cdot  \vect\left[F\right]$, where $P_I(p)$ is a patch with $k\times k$ pixels and $m$ image channels around $p$, $F$ is a filter with $k\times k\times m$ weights, and $\vect\left[.\right]$ is the vectorization operation.
-A filter $F$ can discriminate a class $i\in \{1,2,\ldots,c\}$ among $c$ classes when it generates positive results for local patterns $\vect\left[P_I(p)\right]$ of class $i$ and negative results for patterns from other classes.
-That is, $\vect\left[F\right]$ is the normal vector of a hyperplane in $\Re^{k\times k\times m}$ which detects the patterns of class $i$ at its positive side.
+For sake of clarity, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect(P_I(p)) \cdot  \vect(F)$, where $P_I(p)$ is a patch with $k\times k$ pixels and $m$ image channels around $p$, $F$ is a filter with $k\times k\times m$ weights, and $\vect$ is the vectorization operation.
+A filter $F$ can discriminate a class $i\in \{1,2,\ldots,c\}$ among $c$ classes when it generates positive results for local patterns $\vect(P_I(p))$ of class $i$ and negative results for patterns from other classes.
+That is, $\vect(F)$ is the normal vector of a hyperplane in $\Re^{k\times k\times m}$ which detects the patterns of class $i$ at its positive side.
 We wish to estimate ${\cal F}$ such that its filters will detect patterns from all classes in different positions $p$.
@@
 For a problem with $c$ classes where $\lambda(p)=i\in \{1,2,\ldots,c\}$ is the label of a marker pixel $p$ from class $i$, let ${\cal M}_I$ be a set of marker pixels drawn in image $I\in {\cal D}$ and $P_I(p)$ be the respective patch around a pixel $p\in {\cal M}_I$.
-Let ${\cal P}$ be the set of all patches around marker pixels from all images $I\in {\cal D}$, with representative examples  $\vect\left[P_I(p)\right]$ from all $c$ classes.
+Let ${\cal P}$ be the set of all patches around marker pixels from all images $I\in {\cal D}$, with representative examples  $\vect(P_I(p))$ from all the $c$ classes.
 \begin{eqnarray}
@@
 We should at least expect that their patterns be detected in different image regions.
-The groups must be shifted to the origin of the feature space by subtracting the mean vector of the samples in ${\cal P}$.
-Optionally, the standard deviation of each feature can be computed and used for feature standardization.
-We call this operation \emph{patch-based normalization}.
-The centroid of each cluster defines the weights of each filter $F\in {\cal F}$ and we force norm $\|\vect\left[F\right]\|=1$ to avoid preferences among them.
-The centralization for filter definition is paramount to eliminate activations in regions from other classes, whose patterns fall in the negative side of the hyperplane with normal $\vect\left[F\right]$.
+The groups must be shifted to the origin of $\Re^{k\times k\times m}$ by subtracting the mean value of each feature in ${\cal P}$.
+Additionally, the standard deviation of each feature is computed and used for standardization.
+This operation allows batch normalization of image sets using the mean and standard deviation of the features from the image markers -- i.e.,  \emph{a marker-based batch normalization}.
+The centroid of each cluster defines the weights of each filter $F\in {\cal F}$ and we force norm $\|\vect(F)\|=1$ to avoid preferences among them.
+The centralization for filter definition is paramount to eliminate activations in regions from other classes, whose patterns fall in the negative side of the hyperplane with normal $\vect(F)$.
 Figure~\ref{fig:filter}a illustrates three groups with color-coded samples from two classes in a hypothetical 2D feature space.
-The path-based normalization followed by convolution with filters obtained as centroids of the three groups creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters.
+The marker-based batch normalization and convolution with filters obtained as centroids of the three groups creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters.
 
@@
   \caption{(a) A 2D feature space with color-coded samples from two classes distributed in three groups.
-(b) After patch-based normalization and convolution with filters obtained from each group, the classes can be more easily detected by two filters in a new 3D feature space.}
+(b) After marker-based batch normalization and convolution with filters obtained from each group, the classes can be more easily detected by two filters in a new 3D feature space.}
   \label{fig:filter}
@@
 Note that, apart from the initial marker selection, the training process is automatic.
-We have preserved the convolutional layers' output dimensions, such that we can use the same markers used in the input layer to find the filters of layer $L+1$ from the resulting images obtained at the output of layer $L$.
+We have preserved the convolutional layers' output dimensions and the marker pixels used to find the filters of layer $L+1$ from the output of layer $L$ are always the same for all convolutional layers.
 Figure~\ref{fig:tree} shows one example of a simple network projected by FLIM for the experiments of the next section.
@@
       level distance          = 3em,
-      edge from parent/.style = {draw=black},
+      edge from parent/.style = {draw=black, -latex},
       every node/.style       = {treenode},
@@
     ]
-    \node [treenode, fill=tealDeer, label = {[label distance = 0.5em]}] (v0) {Conv$(60 \times 7 \times 7)$}
+    \node [treenode, fill=tealDeer] (v0) {MarkerBased BatchNorm}
     child {
+        node [treenode, fill=tealDeer] {Conv$(60 \times 7 \times 7)$}
         child {
+          node [treenode, fill=tealDeer] {ReLU} 
           child {
+            node [treenode, fill=tealDeer] {MaxPool$(3\times3)$}
+            child {
+              node [treenode, fill=tealDeer] {BatchNorm}
               child {
@@
               }
-                node [treenode, fill=tealDeer] {BatchNorm}
-                edge from parent
             }
-          node [treenode, fill=tealDeer] {MaxPool$(3\times3)$}
-          edge from parent
           } 
-        node [treenode, fill=tealDeer, label = {[label distance = 0.5em] }] {ReLU} 
-        edge from parent
+        }
       };
@@
 \section{Experiments}
-We use a dataset formed from patches of aerial images made by \cite{8899005} from imagery provided by WeRobotics.
+
+We use a dataset from~\cite{8899005}, which contains aerial images with or without one coconut tree for counting. These images belong to a large imagery from
+
+selected aerial images from WeRobotics and used the same dataset of regions of interest in these images, as provided by\cite{8899005} .
 The imagery with a resolution of 8cm was captured by the World Bank's UAVs for Disaster Resilience Program, collaborating with WeRobotics and OpenAerialMap in the Kingdom of Tonga in October 2017.
@@
     FLIM & $0.863 \pm 0.002$ & $0.849 \pm 0.005$ & $0.854 \pm 0.004$\\
-    FLIM-FT &  &  & \\
+    FLIM-FT & $0.845 \pm 0.003$ & $0.825\pm 0.006$ & $0.832 \pm 0.005$ \\
 
-    VGG &  &  & \\
+    VGG & $0.827 \pm 0.003$ & $0.770 \pm 0.016$  & $ 0.784 \pm 0.014$\\
     VGG-FT & $0.872 \pm 0.007$ & $0.844 \pm 0.015$ & $0.851 \pm 0.014 $ \\

### 2020-06-12 main.tex
review definitions

@@
 We wish then to estimate the weights of $F$ such that those local patterns are characteristic of one given class.
-For sake of clarity, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect(P_I(p)) \cdot  \vect(F)$, where $P_I(p)$ is a patch with $k\times k$ pixels and $m$ image channels around $p$, $F$ is a filter with $k\times k\times m$ weights, and $\vect$ is the vectorization operation.
-A filter $F$ can discriminate a class $i\in \{1,2,\ldots,c\}$ among $c$ classes when it generates positive results for local patterns $\vect(P_I(p))$ of class $i$ and negative results for patterns from other classes.
-That is, $\vect(F)$ is the normal vector of a hyperplane in $\Re^{k\times k\times m}$ which detects the patterns of class $i$ at its positive side.
+For sake of clarity, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect(P_I(p)) \cdot \vect(F)$, where $P_I(p)$ is a $k\times k$ patch with $m$ channels around $p$, $F$ is a filter with shape $k\times k \times m$, and $\vect$ is the vectorization operation.
+A filter $F$ can discriminate a class $i \in \{1,2,\ldots,c\}$ among $c$ classes when it generates positive results for local patterns $\vect(P_I(p))$ of class $i$ and negative results for patterns from other classes.
+That is, $\vect(F)$ is the normal vector of a hyperplane in $\R^{k\times k\times m}$ which detects the patterns of class $i$ at its positive side.
 We wish to estimate ${\cal F}$ such that its filters will detect patterns from all classes in different positions $p$.
 
-For a problem with $c$ classes where $\lambda(p)=i\in \{1,2,\ldots,c\}$ is the label of a marker pixel $p$ from class $i$, let ${\cal M}_I$ be a set of marker pixels drawn in image $I\in {\cal D}$ and $P_I(p)$ be the respective patch around a pixel $p\in {\cal M}_I$.
+\textcolor{blue}{For a problem with $c$ classes where $\lambda(p)=i\in \{1,2,\ldots,c\}$ is the label of a marker pixel $p$ from class $i$, let ${\cal M}_I$ be a set of marker pixels drawn in image $I\in {\cal D}$ and $P_I(p)$ be the respective patch around a pixel $p\in {\cal M}_I$.
 Let ${\cal P}$ be the set of all patches around marker pixels from all images $I\in {\cal D}$, with representative examples  $\vect(P_I(p))$ from all the $c$ classes.
+We define the set of patches from class $i$ as
 \begin{eqnarray}
-{\cal P} & = & \bigcup_{I\in {\cal D},p\in {\cal M}_I, \lambda(p)\in [1,c]} P_I(p)
-\label{eq.pset}
+{\cal P}_i & = & \bigcup_{I\in {\cal D},p\in {\cal M}_I, \lambda(p) = i}{P_I(p)}.
+\label{eq:pset}
 \end{eqnarray}
-The patch set ${\cal P}$ defines samples from all classes in $\Re^{k\times k\times m}$ and a clustering operation on ${\cal P}$ creates groups with similar local patterns.
-Ideally, each group should contain patterns of a same class.
+A clustering operation on ${\cal P}$ creates groups with similar local patterns.
+Ideally, each group should contain patterns of class $c$.
 However, some classes might share similar patterns.
 We should at least expect that their patterns be detected in different image regions.
-The groups must be shifted to the origin of $\Re^{k\times k\times m}$ by subtracting the mean value of each feature in ${\cal P}$.
-Additionally, the standard deviation of each feature is computed and used for standardization.
-This operation allows batch normalization of image sets using the mean and standard deviation of the features from the image markers -- i.e.,  \emph{a marker-based batch normalization}.
+The groups must be shifted to the origin of $\R^{k\times k\times m}$ by subtracting the mean value of each patch in ${\cal P}$, where $\mathcal{P} = \bigcup_{i \in \{1, 2, \ldots, c\}}{\mathcal{P}_i}$.
+Additionally, the standard deviation of each patch is computed and used for standardization.
+This operation allows batch normalization of image sets using the mean and standard deviation of the patches from the image markers -- i.e.,  \emph{a marker-based batch normalization}.
 The centroid of each cluster defines the weights of each filter $F\in {\cal F}$ and we force norm $\|\vect(F)\|=1$ to avoid preferences among them.
 The centralization for filter definition is paramount to eliminate activations in regions from other classes, whose patterns fall in the negative side of the hyperplane with normal $\vect(F)$.
-Figure~\ref{fig:filter}a illustrates three groups with color-coded samples from two classes in a hypothetical 2D feature space.
-The marker-based batch normalization and convolution with filters obtained as centroids of the three groups creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters.
+Figure~\ref{fig:filter} illustrates three groups with color-coded samples from two classes in a hypothetical 2D feature space.
+The marker-based batch normalization and convolution with filters obtained as centroids of the three groups creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters.}
 

### 2020-06-12 main.tex
Updates from Overleaf

--- 2020-06-12 main.tex
@@
 
-We use a dataset from~\cite{8899005}, which contains aerial images with or without one coconut tree for counting. These images belong to a large imagery from
+We use a dataset from~\cite{8899005}, which contains regions of aerial images with and without coconut trees from the Kingdom of Tonga, as acquired in October 2017.
+Each region should be classified as containing one or none coconut tree, but part of coconut trees from adjacent regions might appears near the region's border.
+The Humanitarian OpenStreetMap community annotated the regions.
+The dataset consists of 13587 regions with $90 \times 90$ pixels of size $8\times 8$  $cm^2$, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
 
-selected aerial images from WeRobotics and used the same dataset of regions of interest in these images, as provided by\cite{8899005} .
-The imagery with a resolution of 8cm was captured by the World Bank's UAVs for Disaster Resilience Program, collaborating with WeRobotics and OpenAerialMap in the Kingdom of Tonga in October 2017.
-The Humanitarian OpenStreetMap community provided the annotation.
-The dataset consists of 13587 patches of dimensions $90 \times 90$ of coconut (10268) and non-coconut (3319).
-The latter class has a large variability.
-The problem of interest is to classify the images in one of the two classes.
-We only report experiments with one dataset due to the space limitation.
+In order to assess the impact of FLIM in reducing the need for annotated samples, we randomly select 100 regions from each class for training and left the remaining 11387 regions for testing.
+We have also repeated this procedure three times in order to compute mean and standard deviation of the results in precision, recall, and f-score.
+By vectorizing and projecting on 2D~\cite{maaten2008visualizing} the 200 training images of each split, the user identified cohesive groups of images from each class and selected two representative images from each class to draw markers.
+We have neither evaluated markers from a higher number of images nor multiple network architectures trained by FLIM.
+We just projected the simplest network with a single convolutional layer with weights obtained by FLIM from markers drawn in four images only.
+
+several architectures based on FLIM and
+
+analyzing the 2D t-SNE projection of the 200 training images, , the user selected
+
+For comparative analysis, we select the popular VGG-16~\cite{simonyan2014very} since it is a deep neural network that usually requires a much higher number  of annotated samples than 200 to be trained from the scratch.
+It is also a very competitive CNN when pretrained in the ImageNet dataset.
+Therefore, as baselines, we use three solutions based on VGG-16: (1) VGG is a network trained from the scratch, using the Xavier's weight-initialization approach, (2) VGG-FT is the network pretrained in the ImageNet dataset and fine-tuned in the coconut-tree training set, and (3) VGG-FE is the feature extractor obtained at the output of the last convolutional layer (layer 13th) of VGG-16 pretrained in the ImageNet dataset.
+VGG and VGG-FT use a MLP classifier as part of the network and we use the linear SVM-OVO classifier from sklearn~\cite{fan2008liblinear} with $C=0.01$ for VGG-FE.
+
+In order to evaluate the quality of our feature learning approach, we first compare the features produced by a single-layer network trained by FLIM with those of the last convolutional layer (layer 13th) of the VGG-16, pretrained on the ImageNet dataset.
+For this comparison, we use  (VGG-FE).
+Table~\ref{tab:results} shows the result of this comparison on the first two rows.
+
+We compare different solutions based on FLIM and In order to evaluate FLIM metho, we select VGG-16 as a baseline CNN. VGG-16
 

### 2020-06-12 main.tex
more review

--- 2020-06-12 main.tex
@@
 Note that, apart from the initial marker selection, the training process is automatic.
-We have preserved the convolutional layers' output dimensions and the marker pixels used to find the filters of layer $L+1$ from the output of layer $L$ are always the same for all convolutional layers.
+We have preserved the convolutional layers' output dimensions and the marker pixels so we can find the filters of layer $L+1$ from the output of layer $L$.
 Figure~\ref{fig:tree} shows one example of a simple network projected by FLIM for the experiments of the next section.

### 2020-06-12 main.tex
Updates from Overleaf

--- 2020-06-12 main.tex
@@
 
-In order to assess the impact of FLIM in reducing the need for annotated samples, we randomly select 100 regions from each class for training and left the remaining 11387 regions for testing.
-We have also repeated this procedure three times in order to compute mean and standard deviation of the results in precision, recall, and f-score.
-By vectorizing and projecting on 2D~\cite{maaten2008visualizing} the 200 training images of each split, the user identified cohesive groups of images from each class and selected two representative images from each class to draw markers.
-We have neither evaluated markers from a higher number of images nor multiple network architectures trained by FLIM.
-We just projected the simplest network with a single convolutional layer with weights obtained by FLIM from markers drawn in four images only.
-
-several architectures based on FLIM and
-
-analyzing the 2D t-SNE projection of the 200 training images, , the user selected
+In order to assess the impact of FLIM in reducing the need for annotated samples, we randomly selected 100 regions from each class for training and left the remaining 11387 regions for testing.
+We also repeated this procedure three times in order to compute mean and standard deviation of the results in precision, recall, and f-score.
+By vectorizing and projecting the 200 training images of each split on 2D~\cite{maaten2008visualizing}, the user identified cohesive groups of images from each class and selected two representative images from each class to draw markers.
+We neither evaluated markers from a higher number of images nor multiple network architectures trained by FLIM.
+We just trained by FLIM the simplest network with a single convolutional layer from markers drawn in four images only.
 

### 2020-06-13 main.tex
Updates from Overleaf

@@
 \usepackage{bm}
+\usepackage{siunitx}
 \usepackage{pgf,tikz,pgfplots}
@@
 The user selects the number of convolutional layers, their filter sizes, a few training images, and draws markers in regions of the input images that best discriminate the classes.
-\textcolor{blue}{Classes may appear in distinct clusters of a local feature space defined by patches extracted from marker pixels at each convolutional layer's input.} FLIM defines the weights of the filters as the centroids of the clusters.
+Classes may appear in distinct clusters of a local feature space defined by patches extracted from marker pixels at each convolutional layer's input.
+FLIM defines the weights of the filters as the centroids of the clusters.
 Such filters are expected to activate discriminant regions in the given convolutional layer's output, and the process repeats in a layer-by-layer fashion.
@@
 Second, very few images per class (e.g., less than five) seem to be enough.
-\textcolor{red}{Third, by eliminating the optimization of the convolutional layers by backpropagation, FLIM can reduce the number of annotated images to train the fully connected layers.} Fourth, FLIM is application-independent, and so it might be useful in other image classification problems.
+Third, by eliminating the optimization of the convolutional layers by backpropagation, FLIM can reduce the number of annotated images to train the fully connected layers.
+Fourth, FLIM is application-independent, and so it might be useful in other image classification problems.
 
@@
 
-\textcolor{blue}{A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation by eliminating negative activations, and aggregate the resulting activations locally (pooling) and globally (flattening) into a highly dimensional and sparse feature space, in which even a linear classifier should separate the classes with reasonable success.} Nevertheless, it is common to assume that the classes are at least piecewise linearly separable and apply fully connected layers -- a Multi-Layer Perceptron (MLP) classifier -- for feature space reduction and classification.
+A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation by eliminating negative activations, and aggregate the resulting activations locally (pooling) and globally (flattening) into a highly dimensional and sparse feature space, in which even a linear classifier should separate the classes with reasonable success.
+Nevertheless, it is common to assume that the classes are at least piecewise linearly separable and apply fully connected layers -- a Multi-Layer Perceptron (MLP) classifier -- for feature space reduction and classification.
 The training of a CNN usually relies on weight optimization using backpropagation, which requires a higher number of annotated images as deep as the model.
@@
 A filter $F$ can discriminate a class $i \in \{1,2,\ldots,c\}$ among $c$ classes when it generates positive results for local patterns $\vect(P_I(p))$ of class $i$ and negative results for patterns from other classes.
-That is, $\vect(F)$ is the normal vector of a hyperplane in $\R^{k\times k\times m}$ which detects the patterns of class $i$ at its positive side.
+That is, $\vect(F)$ is the normal vector of a hyperplane in $\R^{k\times k\times m}$, which detects the patterns of class $i$ at its positive side.
 We wish to estimate ${\cal F}$ such that its filters will detect patterns from all classes in different positions $p$.
 
-\textcolor{blue}{For a problem with $c$ classes where $\lambda(p)=i\in \{1,2,\ldots,c\}$ is the label of a marker pixel $p$ from class $i$, let ${\cal M}_I$ be a set of marker pixels drawn in image $I\in {\cal D}$ and $P_I(p)$ be the respective patch around a pixel $p\in {\cal M}_I$.
-Let ${\cal P}$ be the set of all patches around marker pixels from all images $I\in {\cal D}$, with representative examples  $\vect(P_I(p))$ from all the $c$ classes.
-We define the set of patches from class $i$ as
+For a problem with $c$ classes where $\lambda(p)=i\in \{1,2,\ldots,c\}$ is the label of a marker pixel $p$ from class $i$, let ${\cal M}_I$ be a set of marker pixels drawn in image $I\in {\cal D}$ and $P_I(p)$ be the respective patch around a pixel $p\in {\cal M}_I$.
+Let ${\cal P}_i$ be the set of all patches around marker pixels from all images $I\in {\cal D}$, with representative examples  $\vect(P_I(p))$ from class $i\in \{1,2,\ldots,c\}$.
 \begin{eqnarray}
@@
 \end{eqnarray}
-A clustering operation on ${\cal P}$ creates groups with similar local patterns.
-Ideally, each group should contain patterns of class $c$.
-However, some classes might share similar patterns.
-We should at least expect that their patterns be detected in different image regions.
-The groups must be shifted to the origin of $\R^{k\times k\times m}$ by subtracting the mean value of each patch in ${\cal P}$, where $\mathcal{P} = \bigcup_{i \in \{1, 2, \ldots, c\}}{\mathcal{P}_i}$.
-Additionally, the standard deviation of each patch is computed and used for standardization.
+A clustering operation on each set ${\cal P}_i$, $i\in [1,c]$, guarantees groups with similar local patterns for each class.
+The groups must be shifted to the origin of $\R^{k\times k\times m}$ by subtracting the mean value of all patches, set $\mathcal{P} = \bigcup_{i \in \{1, 2, \ldots, c\}}{\mathcal{P}_i}$.
+Additionally, the standard deviation of all patches is computed and used for standardization.
 This operation allows batch normalization of image sets using the mean and standard deviation of the patches from the image markers -- i.e.,  \emph{a marker-based batch normalization}.
-The centroid of each cluster defines the weights of each filter $F\in {\cal F}$ and we force norm $\|\vect(F)\|=1$ to avoid preferences among them.
+The centroid of each cluster defines the weights of each filter $F\in {\cal F}$, and we force norm $\|\vect(F)\|=1$ to avoid preferences among them.
 The centralization for filter definition is paramount to eliminate activations in regions from other classes, whose patterns fall in the negative side of the hyperplane with normal $\vect(F)$.
 Figure~\ref{fig:filter} illustrates three groups with color-coded samples from two classes in a hypothetical 2D feature space.
-The marker-based batch normalization and convolution with filters obtained as centroids of the three groups creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters.}
+The marker-based batch normalization and convolution with filters obtained as centroids of the three groups creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters.
 
@@
 The number of filters per layer depends on the clustering technique.
-We use $K$-means, then the user must specify the number $K=|{\cal F}|$ of filters.
+We use $K$-means, and then the user must specify the number $K=|{\cal F}|$ of filters.
 After the convolution operation, we apply the ReLU function to eliminate negative activations and the max-pooling operation to aggregate local information.
@@
 We have preserved the convolutional layers' output dimensions and the marker pixels so we can find the filters of layer $L+1$ from the output of layer $L$.
-Figure~\ref{fig:tree} shows one example of a simple network projected by FLIM for the experiments of the next section.
+Figure~\ref{fig:arch} shows one example of a simple network projected by FLIM for the experiments of the next section.
 
@@
       sibling distance        = 5em,
-      level distance          = 3em,
+      level distance          = 8mm,
       edge from parent/.style = {draw=black, -latex},
@@
   \caption{Example of a network with a single convolutional layer followed by a classifier, as projected by FLIM and used for the experiments in this work.}
-  \label{fig:tree}
+  \label{fig:arch}
 \end{figure}
@@
 
-We use a dataset from~\cite{8899005}, which contains regions of aerial images with and without coconut trees from the Kingdom of Tonga, as acquired in October 2017.
+We used a dataset from~\cite{8899005}, which contains regions of aerial images with and without coconut trees from the Kingdom of Tonga, as acquired in October 2017~\footnote{The images are available in \url{blog.werobotics.org/2018/01/11/open-ai-challenge-2}.}.
 Each region should be classified as containing one or none coconut tree, but part of coconut trees from adjacent regions might appears near the region's border.
 The Humanitarian OpenStreetMap community annotated the regions.
-The dataset consists of 13587 regions with $90 \times 90$ pixels of size $8\times 8$  $cm^2$, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
+The dataset consists of 13587 regions with $90 \times 90$ pixels with resolution under $10$\si{\square\centi\metre}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
+The region images are converted to the LAB color space, with each channel normalized within $[0,1]$.
 
-In order to assess the impact of FLIM in reducing the need for annotated samples, we randomly selected 100 regions from each class for training and left the remaining 11387 regions for testing.
-We also repeated this procedure three times in order to compute mean and standard deviation of the results in precision, recall, and f-score.
-By vectorizing and projecting the 200 training images of each split on 2D~\cite{maaten2008visualizing}, the user identified cohesive groups of images from each class and selected two representative images from each class to draw markers.
+To assess the impact of FLIM in reducing the need for annotated samples, we randomly selected 100 regions from each class for training and left the remaining 11387 regions for testing.
+We repeated this procedure three times to compute the mean and the standard deviation of the results in precision, recall, and f-score.
+By vectorizing and projecting the 200 training images of each split on 2D with t-SNE~\cite{maaten2008visualizing}, we identified cohesive groups of images from each class and selected two representative images from each class to draw markers.
+The selected images from one split can be seen in Figure \ref{fig:markers}.
+We placed markers on these images in regions that we thought to better distinguish between coconut and non-coconut classes.
+
 We neither evaluated markers from a higher number of images nor multiple network architectures trained by FLIM.
-We just trained by FLIM the simplest network with a single convolutional layer from markers drawn in four images only.
+We just trained the single-convolutional-layer network of Figure \ref{fig:arch} from markers drawn in two images per class.
+The convolutional layer has $60$ filters with dimension $7 \times 7$ and 3 color channels.
+It includes marker-based normalization before convolution, ReLu activation, and a max-pooling operation with a window of dimension $3 \times 3$.
+Since it is the single and last convolutional layer, we applied a stride of $4$ in max-pooling and batch normalization to create an input as close as possible to the one of the MLP classifier used in the popular VGG-16~\cite{simonyan2014very}.
 
-For comparative analysis, we select the popular VGG-16~\cite{simonyan2014very} since it is a deep neural network that usually requires a much higher number  of annotated samples than 200 to be trained from the scratch.
-It is also a very competitive CNN when pretrained in the ImageNet dataset.
-Therefore, as baselines, we use three solutions based on VGG-16: (1) VGG is a network trained from the scratch, using the Xavier's weight-initialization approach, (2) VGG-FT is the network pretrained in the ImageNet dataset and fine-tuned in the coconut-tree training set, and (3) VGG-FE is the feature extractor obtained at the output of the last convolutional layer (layer 13th) of VGG-16 pretrained in the ImageNet dataset.
-VGG and VGG-FT use a MLP classifier as part of the network and we use the linear SVM-OVO classifier from sklearn~\cite{fan2008liblinear} with $C=0.01$ for VGG-FE.
+We selected VGG-16 for comparative analysis since it usually requires many more samples than 200 to be trained from scratch.
+It is also a very competitive CNN when pretrained in the ImageNet dataset, working as an alternative solution for small training sets.
+Therefore, as baselines, we used three solutions based on VGG-16: (1) VGG is trained from scratch, (2) VGG-FT is pretrained in the ImageNet dataset and fine-tuned in the coconut-tree training set, and (3) VGG-FE is the feature extractor of VGG-16 (the output of the last convolutional layer) pretrained in the ImageNet dataset.
+In this case, we used for classification the  linear SVM classifier with $C=0.01$ from sklearn~\cite{fan2008liblinear, scikit-learn}.
 
-In order to evaluate the quality of our feature learning approach, we first compare the features produced by a single-layer network trained by FLIM with those of the last convolutional layer (layer 13th) of the VGG-16, pretrained on the ImageNet dataset.
-For this comparison, we use  (VGG-FE).
-Table~\ref{tab:results} shows the result of this comparison on the first two rows.
-
-We compare different solutions based on FLIM and In order to evaluate FLIM metho, we select VGG-16 as a baseline CNN. VGG-16
-
-As in reality, labeling images is a costly and time-consuming process, and then we assume that we only knew the correct labels of a small set of samples (one hundred images for each class, totaling two hundred images).
-That way, we would have only those two hundred images to train any models that we would use to try to solve the problem.
-
-To reduce the impact of chance on the results, we randomly selected three sets of 200 images for training and three sets of 11387 images for testing.
-We wanted to represent a scenario where the network designer did not have much effort to annotate all these images.
-From each of these sets of training images, we selected four images using the 2D space projection of the 200 images using the t-SNE.
-We tried to select images that were part of more cohesive groups.
-Thus, these images are more likely to have characteristics that represent their group.
+Table~\ref{tab:results} shows the results of three FLIM-based solutions using the SVM and MLP classifiers; "FLIM + SVM", "FLIM + MLP", and "FLIM-FT + MLP" (i.e., FLIM with fine-tuning); compared with the VGG-based solutions above.
+The first two rows indicate that FLIM  can provide a more effective feature extractor using a single convolutional layer than VGG-FE, which contains 13 convolutional layers pretrained on the ImageNet.
+The results of "FLIM + MLP" and "FLIM-FT + MLP" show that FLIM  may dismiss fine tuning by providing a more effective solution.
+By comparing "FLIM + MLP" with VGG, one can observe that FLIM reduces the number of required training images.
+Finally, the comparison between "FLIM + MLP" and VGG-FT demonstrates that FLIM can provide competitive results using a considerably more simplified architecture than  VGG-FT.
+The number of parameters of our feature extractor represents only $0.07\%$ of the number of parameter of VGG16's feature extractor.
 
@@
     }
-    \caption{Markers used for training.
-(a) Markers on an image which contains a coconut tree.
-(b) Markers on an image which does not contain a coconut tree.
-(c) Markers on an image which does not contain a coconut tree.
-(d) Markers on an image which does not contain a coconut tree.}
+    \caption{Markers used for training in image regions with (first column) and without (second column) a coconut tree.}
     \label{fig:markers}
 \end{figure}
-
-We placed markers on these images in regions that we consider important to identify the coconut and non-coconut (more diverse) classes.
-In Figure \ref{fig:markers} we can the four images of one the splits.
-With the markers, we created a convolutional layer with filters of dimensions $7 \times 7$ using the method described in Section \ref{sec:method}.
-The network architecture can be seen in the Figure.
-We used a classifier similar to that of VGG-16.
-We used K-means to cluster the coconut tree patches in 30 clusters and the non-coconut tree patches in also 30 clusters to find the convolutional layer filter.
 
@@
 
-    FLIM & $0.863 \pm 0.002$ & $0.849 \pm 0.005$ & $0.854 \pm 0.004$\\
-    FLIM-FT & $0.845 \pm 0.003$ & $0.825\pm 0.006$ & $0.832 \pm 0.005$ \\
+    FLIM + MLP & $0.863 \pm 0.002$ & $0.849 \pm 0.005$ & $0.854 \pm 0.004$\\
+    FLIM-FT + MLP & $0.845 \pm 0.003$ & $0.825\pm 0.006$ & $0.832 \pm 0.005$ \\
 
@@
   \end{center}
-  \caption{Results.}
+
+\caption{Mean and standard deviation of precision, recall, and f-score for each method: a feature extractor with a classifier.}
   \label{tab:results}
@@
 
-We tested three scenarios: (I) using out method to learn the feature extractor weights and the classifier with backpropagation; (II) using our method as initialization and train the whole model with backpropagation; (III) using SVM Linear as classification.
-For VGG-16, we tested two scenarios: (IV) initializing with Xavier initialization and training from scratch; (V) using ImageNet weights and training with the training set images.
-The results can be seen in Table \ref{tab:results}.
-As the experiments were repeated three times, we reported the mean and standard deviation of each the metrics in the test set.
-
-\begin{figure}
+\begin{figure*}
   \centering
-  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/ours/upper-coco-108.png}
+  \subfloat{\includegraphics[width=.13\linewidth]{figures/qualitative/ours/upper-coco-108.png}
   }
   ~
-  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/ours/upper-coco-2473.png}
+  \subfloat{\includegraphics[width=.13\linewidth]{figures/qualitative/ours/upper-coco-2473.png}
   }
   ~
-  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/ours/upper-non_coco-506.png}
-  }
-  \\
-  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/vgg/lower-coco-1098.png}
+  \subfloat{\includegraphics[width=.13\linewidth]{figures/qualitative/ours/upper-non_coco-506.png}
   }
   ~
-  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/vgg/lower-coco-2400.png}
+  \subfloat{\includegraphics[width=.13\linewidth]{figures/qualitative/vgg/lower-coco-1098.png}
   }
   ~
-  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/vgg/lower-non_coco-105.png}
+  \subfloat{\includegraphics[width=.13\linewidth]{figures/qualitative/vgg/lower-coco-2400.png}
+  }
+  ~
+  \subfloat{\includegraphics[width=.13\linewidth]{figures/qualitative/vgg/lower-non_coco-105.png}
   }
@@
   \label{fig:ex-classification}
-\end{figure}
+\end{figure*}
 
-Our method achieved better precision than conventionally trained VGG, and the difference between precision and recall is much smaller.
-The Linear SVM did not prove to be a good classifier like the MLP, but, on average, it still got a little better than the VGG-16.
-Using our method as initialization and training the entire network with backpropagation did not improve the result; this implies that the filters we created are already good filters given the training set.
-Our feature extractor training method was able to learn good filters from a minimal set of images.
-VGG has many parameters to learn, so it is susceptible to overfitting with such a small training set.
-We do not use any technique to handle small data sets to train VGG because we intended to verify our approach viability, and these techniques could also be used with our method.
-In Figure \ref{fig:ex-classification} we can see examples of misclassified images.
+ In Figure \ref{fig:ex-classification}, we can see examples of misclassified images.
 In these images, coconut trees appear in different angles, sizes, and shapes, or the boundaries between the coconut tree and the background are tenuous, making it more difficult to identify them.
@@
   \label{fig:vis-classifier}}
-  \caption{This image shows visualizations of the input images and the intermediate layer's outputs.
-Red points represent background images, and green points represent images that have a coconut tree.
-The projection from high dimensionality space to 2D was made using t-SNE.
-The original image shows much overlap between the classes; after applying the feature extractor with only one convolutional layer, we can see that the overlapping decreases considerably.
-Before the decision layer, the last layer put most of the images from one class on one side and the other on the other side.
-(a) Visualization of original image feature space.
-(b) Visualization of the feature extractor's outputs.
-(c) Visualization of outputs from the last classifier hidden layer.}
+  \caption{Projections by t-SNE of the images at (a) the input layer (original images), (b) the output of the convolutional layer of our network (Figure~\ref{fig:arch}), and (c) the output of the last hidden layer of the MLP classifier in "FLIM + MLP".}
 \end{figure}
@@
 As motivated by Rauber et al.
-\cite{rauber2016visualizing}, we made projections of three stages of the network using t-SNE \cite{maaten2008visualizing} to understand how the feature extractor trained with our method transforms the image spaces.
+\cite{rauber2016visualizing}, we made projections of three stages of the network "FLIM + MLP" using t-SNE \cite{maaten2008visualizing} to understand how our CNN transforms the image spaces.
 In Figure \ref{fig:vis-input}, we can see the projection of the original images (in the LAB color space).
-Red points are images that contain coconut trees, and green points are images that do not contain coconut trees.
-In the projection in the image of Figure \ref{fig:vis-input}, we see that the images of coconut trees and non-coconut trees are almost entirely overlapping.
-Coconut tree images are more frequent in some regions but are very dispersed.
-In the image of Figure \ref{fig:vis-feature-extractor}, we see the projection of the feature extractor output.
-The coconut tree images are more concentrated, forming some clusters, and it is possible to notice that there is much less overlap.
-Finally, we see in Figure \ref{fig:vis-classifier} the output of the previous layer of the decision layer of the classifier.
-The classifier managed to organize conquest and non-coconut images in different regions of space, although there is overlap as the model failed to achieve 100\% accuracy.
+Green points are images that contain coconut trees, and red points are images that do not contain coconut trees.
+The projection shows considerable overlapping between the two classes.
+Coconut-tree images are more frequent in some regions but they appear very dispersed.
+In Figure \ref{fig:vis-feature-extractor}, the projection of the output of our feature extractor shows some reduction in the dispersion of the coconut-tree images.
+The coconut tree images are more concentrated, forming some clusters, and it is possible to notice some overlapping reduction between classes.
+Finally, Figure \ref{fig:vis-classifier} shows the output of the last hidden layer of the MLP classifier.
+The overlapping between classes and the dispersion of coconut-tree images are considerably reduced.
 
-\section{Conclusions}
+\section{Conclusion}
 
-We present a way to train a feature extractor without using backpropagation capable of learning good filters from a minimal set of images.
-To evaluate the method, we proposed experiments to classify images that contain achievements.
-Our approach has achieved competitive results with state of the art.
+We introduced a first feature learning technique, named FLIM, that can estimate effective filter weights for the convolutional layers of a given network architecture from user-drawn markers in a few training images.
+The resulting feature extractor can be used with different classifiers and, when it is used with a MLP classifier, the number of required images to train the fully connected layers is reduced.
+We demonstrated the advantages of FLIM over solutions based on VGG-16 for binary classification of coconut-tree images.
+The experiments indicated that FLIM is an effective approach that can produce considerably simplified network architectures.
+By involving the user in the training process, FLIM improves the understanding about CNN models and the user control over their training process.
 
+We intend to further investigate feature learning from image markers for different applications, elaborate methodologies to optimize the network architecture, and extend FLIM to estimate weights in fully connected layers.
+
+\section*{Acknowledgements}
+
+The authors thank FAPESP (Proc. 2014/12236-1), CNPq (Proc. 303808/2018-7), and Unisinos (Viscarb and Visorg projects). 
 \bibliographystyle{IEEEtran}

### 2020-06-13 main.tex
Updates from Overleaf

@@
 \usetikzlibrary{calc}
+\usepackage{hyperref}
 
@@
 
-\begin{figure*}
+\begin{figure}
   \centering
-  \subfloat{\includegraphics[width=.13\linewidth]{figures/qualitative/ours/upper-coco-108.png}
+  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/ours/upper-coco-108.png}
   }
   ~
-  \subfloat{\includegraphics[width=.13\linewidth]{figures/qualitative/ours/upper-coco-2473.png}
+  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/ours/upper-coco-2473.png}
   }
   ~
-  \subfloat{\includegraphics[width=.13\linewidth]{figures/qualitative/ours/upper-non_coco-506.png}
+  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/ours/upper-non_coco-506.png}
+  }
+  \\
+  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/vgg/lower-coco-1098.png}
   }
   ~
-  \subfloat{\includegraphics[width=.13\linewidth]{figures/qualitative/vgg/lower-coco-1098.png}
+  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/vgg/lower-coco-2400.png}
   }
   ~
-  \subfloat{\includegraphics[width=.13\linewidth]{figures/qualitative/vgg/lower-coco-2400.png}
-  }
-  ~
-  \subfloat{\includegraphics[width=.13\linewidth]{figures/qualitative/vgg/lower-non_coco-105.png}
+  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/vgg/lower-non_coco-105.png}
   }
@@
   \label{fig:ex-classification}
-\end{figure*}
+\end{figure}
 

### 2020-06-13 main.tex
adjustment to fit in 4 pages

@@
 Deep learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}.
-In remote sensing, the applications may involve segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic, hamaguchi2018effective}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}, for instance.
+In remote sensing, the applications may involve segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}, for instance.
 In this work, we are interested in identifying species of trees from aerial images.
@@
 
-\begin{figure}
+\begin{figure*}
   \begin{center}
@@
     treenode/.style = {shape=rectangle, rounded corners,
-                       draw=black, align=center, line width=width},
+                       draw=black, align=center, line width=width, minimum width=2cm},
     specialedge/.style = {line width=1.5pt}
@@
     [
-      grow                    = down,
+      grow                    = right,
       sibling distance        = 5em,
-      level distance          = 8mm,
+      level distance          = 30mm,
       edge from parent/.style = {draw=black, -latex},
@@
     ]
-    \node [treenode, fill=tealDeer] (v0) {MarkerBased BatchNorm}
+    \node [treenode, fill=tealDeer] {MarkerBased \\ BatchNorm}
     child {
@@
   \label{fig:arch}
-\end{figure}
+\end{figure*}
 
@@
     \centering
-    \subfloat[\label{fig:markers1}]{\includegraphics[width=.3\linewidth]{figures/markers1}
+    \subfloat[\label{fig:markers1}]{\includegraphics[width=.25\linewidth]{figures/markers1}
     }
     ~
-    \subfloat[\label{fig:markers2}]{\includegraphics[width=.3\linewidth]{figures/markers2}
+    \subfloat[\label{fig:markers2}]{\includegraphics[width=.25\linewidth]{figures/markers2}
     }
     \\
-    \subfloat[\label{fig:markers3}]{\includegraphics[width=.3\linewidth]{figures/markers3}
+    \subfloat[\label{fig:markers3}]{\includegraphics[width=.25\linewidth]{figures/markers3}
     }
     ~
-    \subfloat[\label{fig:markers4}]{\includegraphics[width=.3\linewidth]{figures/markers4}
+    \subfloat[\label{fig:markers4}]{\includegraphics[width=.25\linewidth]{figures/markers4}
     }
@@
   \centering
-  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/ours/upper-coco-108.png}
+  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/ours/upper-coco-108.png}
   }
   ~
-  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/ours/upper-coco-2473.png}
+  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/ours/upper-coco-2473.png}
   }
   ~
-  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/ours/upper-non_coco-506.png}
+  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/ours/upper-non_coco-506.png}
   }
   \\
-  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/vgg/lower-coco-1098.png}
+  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/vgg/lower-coco-1098.png}
   }
   ~
-  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/vgg/lower-coco-2400.png}
+  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/vgg/lower-coco-2400.png}
   }
   ~
-  \subfloat{\includegraphics[width=.3\linewidth]{figures/qualitative/vgg/lower-non_coco-105.png}
+  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/vgg/lower-non_coco-105.png}
   }

### 2020-07-01 main.tex
Updates from Overleaf

@@
 
-\documentclass[a4paper,conference]{IEEEtran}
+\documentclass[journal, twoside, peerreview, draftclsnofoot]{IEEEtran}
 
@@
 \usepackage{comment}
-\usepackage{float}
 \usepackage{mathtools}
@@
 \usepackage{siunitx}
-\usepackage{pgf,tikz,pgfplots}
+\usepackage{tikz}
 \usetikzlibrary{calc}
@@
 
-\title{Learning CNN filters from user-drawn image makers for coconut-tree image classification}
+\title{Learning CNN filters from user-drawn image markers for coconut-tree image classification}
+
+\author{Italos~de~Souza,~
+        and~Alexandre~ Falc\~{a}o,~\IEEEmembership{Member,~IEEE}}
+
+\markboth{IEEE GEOSCIENCE AND REMOTE SENSING LETTERS,~Vol.~17, No.~11, November~2020}%
+{de Souza \MakeLowercase{\textit{et al.}}: Learning CNN filters from user-drawn image makers for coconut-tree image classification}
 
@@
 The method learns the filters of each convolutional layer from user-drawn markers in image regions that discriminate classes, allowing better user control and understanding of the training process.
-It does not rely on optimization based on backpropagation, and we demonstrate its advantages on the binary classification of coconut-tree images against one of the most popular CNN models.
+It does not rely on optimization based on backpropagation, and we demonstrate its advantages on the binary classification of coconut-tree aerial images against one of the most popular CNN models.
 \end{abstract}
 
+\begin{IEEEkeywords}
+Design of convolutional neural networks, remote sensing image analysis, interactive machine learning.
+\end{IEEEkeywords}
+
 \section{Introduction}
-Deep learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}.
+\IEEEPARstart{D}{epp} learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}.
 In remote sensing, the applications may involve segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}, for instance.
@@
 Vargas-Muñoz et al.~\cite{8899005} propose an active learning approach to mitigate the problem.
-The method explores data projection techniques to allow simultaneous annotation of multiple patches as having or not coconut trees.
-It then uses a CNN to identify the candidate patches with coconut trees.
-We adopt another alternative -- the design of simplified CNN models from small training sets with user interaction, being the user an expert in the design of CNN models.
+The method explores data projection techniques to allow simultaneous annotation of multiple regions as having or not coconut trees.
+It then uses a CNN to identify the candidate regions with coconut trees.
+We adopt another alternative -- the design of simplified CNN models from small training sets with user interaction, being the user knowledgeable about CNN modeling and the application domain.
 
@@
 The fourth question is also essential during training, and it is related to user control over the process.
-They all lead to the importance of involving an expert user in the machine learning process.
-In this paper, we present significant advances related to questions (1) and (2).
+They all lead to the importance of involving human experts during the machine learning process.
+In this paper, we present significant advances related to the above questions.
 First, we conceptually divide a CNN into convolutional layers for feature extraction and fully connected layers for feature space reduction and classification.
 Each convolutional layer contains a filter bank, an activation function, and alternative operations (e.g., pooling, batch normalization).
-As the number of convolutional layers increases (deeper is the model), and so higher is the number of annotated images required to train the model by backpropagation.
+As the number of convolutional layers increases (deeper is the model), and higher is the number of annotated images required to train the model by backpropagation.
 
@@
 
-We demonstrate the advantages of this interactive technique on binary image classification of coconut-tree images against  VGG-16~\cite{simonyan2014very}.
+We demonstrate the advantages of this interactive technique on the binary image classification of coconut-tree images against  VGG-16~\cite{simonyan2014very}.
 First, the user can better understand and control the training process by observing the effectiveness of marker selection.
@@
 
-A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation by eliminating negative activations, and aggregate the resulting activations locally (pooling) and globally (flattening) into a highly dimensional and sparse feature space, in which even a linear classifier should separate the classes with reasonable success.
-Nevertheless, it is common to assume that the classes are at least piecewise linearly separable and apply fully connected layers -- a Multi-Layer Perceptron (MLP) classifier -- for feature space reduction and classification.
-The training of a CNN usually relies on weight optimization using backpropagation, which requires a higher number of annotated images as deep as the model.
+A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation by eliminating negative activations, and aggregate the resulting activations locally (pooling) and globally (flattening) into a highly dimensional and sparse feature space, in which a linear classifier should separate the classes with reasonable success.
+Nevertheless, it is common to assume that the classes are piecewise linearly separable and apply fully connected layers -- a Multi-Layer Perceptron (MLP) classifier -- for feature space reduction and classification.
+The training of a CNN usually relies on weight optimization using backpropagation, which requires a higher number of annotated images as deep as the model is.
 In FLIM, we eliminate the need for backpropagation to train convolutional layers by finding a set ${\cal F}$ of filters that can detect discriminant regions from all classes.
@@
 Figure~\ref{fig:filter} illustrates three groups with color-coded samples from two classes in a hypothetical 2D feature space.
-The marker-based batch normalization and convolution with filters obtained as centroids of the three groups creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters.
+The marker-based batch normalization and convolution with filters obtained as centroids of the three groups in Figure~\ref{fig:filter}a creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters in distinct activation bands.
 
@@
 The dataset consists of 13587 regions with $90 \times 90$ pixels with resolution under $10$\si{\square\centi\metre}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
-The region images are converted to the LAB color space, with each channel normalized within $[0,1]$.
+We converted the region images to the LAB color space, with each channel normalized within $[0,1]$.
 
@@
 
-We neither evaluated markers from a higher number of images nor multiple network architectures trained by FLIM.
+We have not investigated the impact of placing markers on a higher number of images and network architecture optimization procedures yet.
 We just trained the single-convolutional-layer network of Figure \ref{fig:arch} from markers drawn in two images per class.
@@
 Since it is the single and last convolutional layer, we applied a stride of $4$ in max-pooling and batch normalization to create an input as close as possible to the one of the MLP classifier used in the popular VGG-16~\cite{simonyan2014very}.
+We use Pytorch to implement our feature extractor and the MLP classifier.
+We trained the MLP classifier for $40$ epochs and the whole network for $60$ epochs with a batch size of $64$.
+We use a learning rate of $0.0001$, a weight decay of $0.001$, and we update the learning rate by a factor of $0.1$ every $5$ epochs after epoch 30.
 
 We selected VGG-16 for comparative analysis since it usually requires many more samples than 200 to be trained from scratch.
-It is also a very competitive CNN when pretrained in the ImageNet dataset, working as an alternative solution for small training sets.
-Therefore, as baselines, we used three solutions based on VGG-16: (1) VGG is trained from scratch, (2) VGG-FT is pretrained in the ImageNet dataset and fine-tuned in the coconut-tree training set, and (3) VGG-FE is the feature extractor of VGG-16 (the output of the last convolutional layer) pretrained in the ImageNet dataset.
+It is also a very competitive CNN when pretrained in the ImageNet dataset, working as a counterpart solution for small training sets.
+Therefore, as baselines, we used three solutions based on VGG-16: (1) VGG is trained from scratch, (2) VGG-FT is pretrained in the ImageNet dataset and fine-tuned in the coconut-tree training set, and (3) VGG-FE is the feature extractor of VGG-16 (the output of its last convolutional layer) pretrained in the ImageNet dataset.
 In this case, we used for classification the  linear SVM classifier with $C=0.01$ from sklearn~\cite{fan2008liblinear, scikit-learn}.
+We use Pytorch's implementation of VGG.
+In (1) and (2), we trained for $95$ epochs with a batch size of $64$, learning rate of $0.001$, weight decay of $0.001$, and we update learning rate by a factor of $0.1$ every $5$ epochs after the model achieves an accuracy of $0.95$.
 
-Table~\ref{tab:results} shows the results of three FLIM-based solutions using the SVM and MLP classifiers; "FLIM + SVM", "FLIM + MLP", and "FLIM-FT + MLP" (i.e., FLIM with fine-tuning); compared with the VGG-based solutions above.
+Table~\ref{tab:results} shows the results of three FLIM-based solutions using the SVM and MLP classifiers; ``FLIM+SVM'', ``FLIM+MLP'', and ``FLIM-FT+MLP'' (i.e., FLIM with fine-tuning); as compared with the VGG-based solutions above.
 The first two rows indicate that FLIM  can provide a more effective feature extractor using a single convolutional layer than VGG-FE, which contains 13 convolutional layers pretrained on the ImageNet.
-The results of "FLIM + MLP" and "FLIM-FT + MLP" show that FLIM  may dismiss fine tuning by providing a more effective solution.
-By comparing "FLIM + MLP" with VGG, one can observe that FLIM reduces the number of required training images.
-Finally, the comparison between "FLIM + MLP" and VGG-FT demonstrates that FLIM can provide competitive results using a considerably more simplified architecture than  VGG-FT.
-The number of parameters of our feature extractor represents only $0.07\%$ of the number of parameter of VGG16's feature extractor.
+The results of ``FLIM+MLP'' and ``FLIM-FT+MLP'' show that FLIM  may dismiss fine tuning by directly providing a more effective solution.
+By comparing ``FLIM+MLP'' with VGG, one can observe that FLIM reduces the number of required training images.
+Finally, the comparison between ``FLIM+MLP'' and VGG-FT demonstrates that FLIM can provide competitive results using a considerably more simplified architecture than  VGG-FT.
+The number of parameters of our feature extractor represents only $0.07\%$ of the number of parameter of VGG-FE.
 
@@
   \hline\hline
-    FLIM + SVM & $0.856 \pm 0.011 $ & $ 0.831 \pm 0.019$ & $ 0.838 \pm 0.017$\\ 
-    VGG-FE + SVM & $0.855 \pm 0.001$ & $0.816 \pm 0.007$ & $0.826 \pm 0.006 $ \\\hline
+    FLIM+SVM & $0.856 \pm 0.011 $ & $ 0.831 \pm 0.019$ & $ 0.838 \pm 0.017$\\ 
+    VGG-FE+SVM & $0.855 \pm 0.001$ & $0.816 \pm 0.007$ & $0.826 \pm 0.006 $ \\\hline
 
-    FLIM + MLP & $0.863 \pm 0.002$ & $0.849 \pm 0.005$ & $0.854 \pm 0.004$\\
-    FLIM-FT + MLP & $0.845 \pm 0.003$ & $0.825\pm 0.006$ & $0.832 \pm 0.005$ \\
+    FLIM+MLP & $0.863 \pm 0.002$ & $0.849 \pm 0.005$ & $0.854 \pm 0.004$\\
+    FLIM-FT+MLP & $0.845 \pm 0.003$ & $0.825\pm 0.006$ & $0.832 \pm 0.005$ \\
 
@@
   ~
-  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/vgg/lower-coco-2400.png}
+  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/vgg/lower-coco-875.png}
   }
@@
   }
-  \caption{Misclassified images.
-The first row has images misclassified by the network trained by our method, and the second row has images misclassified by VGG-16.
-The two first images of a row have coconut trees.}
+  \caption{Misclassified images by ``FLIM + MLP'' (first row) and VGG-FT (second row), being the last column with images without coconut trees.}
   \label{fig:ex-classification}
@@
 In turn, images that do not contain coconut trees, contain trees that resemble their shape.
-The network designer could put markers in those images in other to improve the feature extractor.
+The network designer could add markers in those images to improve the feature extractor in an active learning loop.
 
@@
   \label{fig:vis-classifier}}
-  \caption{Projections by t-SNE of the images at (a) the input layer (original images), (b) the output of the convolutional layer of our network (Figure~\ref{fig:arch}), and (c) the output of the last hidden layer of the MLP classifier in "FLIM + MLP".}
+  \caption{Projections by t-SNE of the test images at (a) the input layer, (b) the output of the convolutional layer of  ``FLIM + MLP'' after batch normalization (Figure~\ref{fig:arch}), and (c) the output of the last hidden layer of the MLP classifier in  ``FLIM + MLP''.}
 \end{figure}
@@
 As motivated by Rauber et al.
-\cite{rauber2016visualizing}, we made projections of three stages of the network "FLIM + MLP" using t-SNE \cite{maaten2008visualizing} to understand how our CNN transforms the image spaces.
-In Figure \ref{fig:vis-input}, we can see the projection of the original images (in the LAB color space).
+\cite{rauber2016visualizing}, we created projections of three stages of the network ``FLIM + MLP" using t-SNE \cite{maaten2008visualizing} to understand how our CNN transforms image spaces along its layers.
+In Figure \ref{fig:vis-input}, we can see the projection of the test images in the LAB color space.
 Green points are images that contain coconut trees, and red points are images that do not contain coconut trees.
 The projection shows considerable overlapping between the two classes.
-Coconut-tree images are more frequent in some regions but they appear very dispersed.
-In Figure \ref{fig:vis-feature-extractor}, the projection of the output of our feature extractor shows some reduction in the dispersion of the coconut-tree images.
-The coconut tree images are more concentrated, forming some clusters, and it is possible to notice some overlapping reduction between classes.
-Finally, Figure \ref{fig:vis-classifier} shows the output of the last hidden layer of the MLP classifier.
-The overlapping between classes and the dispersion of coconut-tree images are considerably reduced.
+The samples of both classes are also dispersed in the projection.
+In Figure \ref{fig:vis-feature-extractor}, the projection of the output of our feature extractor shows some reduction in sample dispersion.
+The classes show more concentration in some parts of the projection and reduced overlapping between them.
+Finally, Figure \ref{fig:vis-classifier} shows the output of the last hidden layer of the MLP classifier in ``FLIM + MLP''.
+The overlapping between classes and sample dispersion are considerably reduced, explaining our results.
 
@@
 The resulting feature extractor can be used with different classifiers and, when it is used with a MLP classifier, the number of required images to train the fully connected layers is reduced.
-We demonstrated the advantages of FLIM over solutions based on VGG-16 for binary classification of coconut-tree images.
-The experiments indicated that FLIM is an effective approach that can produce considerably simplified network architectures.
-By involving the user in the training process, FLIM improves the understanding about CNN models and the user control over their training process.
+We demonstrated the advantages of FLIM over solutions based on VGG-16 for the binary classification of coconut-tree images.
+The experiments indicated that FLIM is an effective approach to produce considerably simplified network architectures.
+By involving the user in the training process, FLIM improves the understanding about CNN models and the user control over the training process.
 
-We intend to further investigate feature learning from image markers for different applications, elaborate methodologies to optimize the network architecture, and extend FLIM to estimate weights in fully connected layers.
+We intend to further investigate feature learning from image markers for different applications, elaborate methodologies to optimize network architecture, and extend FLIM to estimate weights in fully connected layers.
 
@@
 
-The authors thank FAPESP (Proc. 2014/12236-1), CNPq (Proc. 303808/2018-7), and Unisinos (Viscarb and Visorg projects). 
+This research was funded by FAPESP (2014/12236-1), CNPq (303808/2018-7), Petróleo Brasileiro S.A.
+(PETROBRAS) and Agência Nacional do Petróleo, Gás Natural e Biocombustíveis (ANP), Grant Numbers 4600556376 and 4600583791.
 \bibliographystyle{IEEEtran}

### 2020-07-21 main.tex
results with increased dataset

@@
 
-\documentclass[journal, twoside, peerreview, draftclsnofoot]{IEEEtran}
+\documentclass[journal, twoside]{IEEEtran}
 
@@
 \section{Introduction}
-\IEEEPARstart{D}{epp} learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}.
+\IEEEPARstart{D}{eep} learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}.
 In remote sensing, the applications may involve segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}, for instance.
@@
 
+\begin{table}[!t]
+  \begin{center}
+  \begin{tabular}{|l|c|c|c|}
+  \hline
+   Method & Precision & Recall & F-score \\
+  \hline\hline
+    FLIM+MLP & $0.894 \pm 0.006$ & $0.888 \pm 0.009$ & $0.890 \pm 0.008$\\
+    FLIM-FT+MLP & $0.888 \pm 0.001$ & $0.878\pm 0.001$ & $0.881 \pm 0.001$ \\
+    VGG-FT & $0.884 \pm 0.003$ & $0.877 \pm 0.006$ & $0.879 \pm 0.005 $ \\
+  \hline
+  \end{tabular}
+  \end{center}
+
+\caption{Mean and standard deviation of precision, recall, and f-score for each method: a feature extractor with a classifier with a larger train set.}
+  \label{tab:results-with-1200-images}
+\end{table}
+
 \section{Conclusion}

### 2020-07-21 main.tex
pipeline figure

@@
 \usepackage{siunitx}
+\usepackage{hyperref}
+
 \usepackage{tikz}
 \usetikzlibrary{calc}
-\usepackage{hyperref}
+\usetikzlibrary{shapes.geometric, arrows, fit}
+\usepackage{varwidth}
+\usetikzlibrary{positioning}
+\usetikzlibrary{backgrounds}
 
@@
 
+\begin{figure*}
+  \begin{center}
+
+    \tikzstyle{steps} = [rectangle, rounded corners, minimum width=1.3cm, minimum height=0.7cm,text centered, draw=black,fill=violet!30, execute at begin node={\begin{varwidth}{2.5cm}},
+    execute at end node={\end{varwidth}}]
+    \tikzstyle{image} = [minimum width=1.3cm, minimum height=0.7cm,text centered, draw=white,fill=white!30, inner sep=0pt]
+
+    \tikzstyle{myarrow} = [navy, thick]
+
+    \tikzstyle{arrow} = [thick,->,>=stealth]
+
+    \begin{tikzpicture}[node distance=1.5cm]
+      %images----------------------------
+     \node (img1) [image] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-coco-789.png}};
+     \node (img2) [image, below left=-5em of img1] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-coco-789.png}};
+     \node (img3) [image, below left=-5em of img2] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-coco-789.png}};
+
+     \node (imgs) [fit={($(img1.north west)+(-0.45,0.2)$) ($(img3.south east)+(0.45,-0.2)$)}, inner sep=0pt, draw=none, rounded corners, very thick, label = {[label distance = 0.3em]below:{Images}}] {};
+     %----------------------------------
+
+     \node (tsne) [steps, right=1em of imgs] {t-SNE};
+
+     \node (vis1) [image, right=1.5em of tsne, outer sep=-15pt] {\includegraphics[width=.15\textwidth]{figures/input.png}};
+
+     \node (select) [steps, right=1.35em of vis1] {Select Images};
+
+    %images----------------------------
+    \node (simg1) [image, above right=-2em and 3em of select] {\includegraphics[width=.08\textwidth]{figures/img1.png}};
+    \node (simg2) [image, below left=-5em of simg1] {\includegraphics[width=.08\textwidth]{figures/img2.png}};
+    \node (simg3) [image, below left=-5em of simg2] {\includegraphics[width=.08\textwidth]{figures/img3.png}};
+    \node (simg4) [image, below left=-5em of simg3] {\includegraphics[width=.08\textwidth]{figures/img4.png}};
+
+    \node (simgs) [fit={($(simg1.north west)+(-0.64,0.2)$) ($(simg4.south east)+(0.64,-0.2)$)}, inner sep=0pt, draw=none] {};
+    %----------------------------------
+
+     \node (put) [steps, right=1em of simgs] {Put Markers};
+
+     %images----------------------------
+    \node (mimg1) [image, above right=-2em and 3em of put] {\includegraphics[width=.08\textwidth]{figures/markers1.png}};
+    \node (mimg2) [image, below left=-5em of mimg1] {\includegraphics[width=.08\textwidth]{figures/markers2.png}};
+    \node (mimg3) [image, below left=-5em of mimg2] {\includegraphics[width=.08\textwidth]{figures/markers3.png}};
+    \node (mimg4) [image, below left=-5em of mimg3] {\includegraphics[width=.08\textwidth]{figures/markers4.png}};
+
+    \node (mimgs) [fit={($(mimg1.north west)+(-0.64,0.2)$) ($(mimg4.south east)+(0.64,-0.2)$)}, inner sep=0pt, draw=none] {};
+    %----------------------------------
+
+     \node (flim) [steps, below left=5em and -3em of mimgs] {FLIM};
+
+     \node (ft) [steps, left=2em of flim] {Feature Extractor};
+
+     \node (aft) [steps, left=2em of ft] {Apply Featute \\ Extractor};
+
+      %images----------------------------
+    \node (f1) [image, above left=-2.8em and 2em of aft] {\includegraphics[width=.07\textwidth]{figures/features/f1.png}};
+    \node (f2) [image, below left=-5em of f1] {\includegraphics[width=.07\textwidth]{figures/features/f2.png}};
+    \node (f3) [image, below left=-5em of f2] {\includegraphics[width=.07\textwidth]{figures/features/f3.png}};
+    \node (f4) [image, below left=-5em of f3] {\includegraphics[width=.07\textwidth]{figures/features/f4.png}};
+
+    \node (features) [fit={($(f1.north west)+(-0.17,0.2)$) ($(f4.south east)+(0.17,-0.2)$)}, draw=none, inner sep=0pt] {};
+    %----------------------------------
+
+     \node (clf) [steps, left=2em of features] {Train classifier};
+
+     %--------------------------------------------------------------arrows
+     \draw [arrow, myarrow] (imgs) -- (tsne);
+     \draw [arrow, myarrow] (tsne) -- (vis1);
+     \draw [arrow, myarrow] (vis1) -- (select);
+     \draw [arrow, myarrow] (select) -- (simgs);
+     \draw [arrow, myarrow] (simgs) -- (put);
+     \draw [arrow, myarrow] (put) -- (mimgs);
+     \draw [arrow, myarrow] (mimgs) -- (flim);
+     \draw [arrow, myarrow] (flim) -- (ft);
+     \draw [arrow, myarrow] (ft) -- (aft);
+     \draw [arrow, myarrow] (aft) -- (features);
+     \draw [arrow, myarrow] (features) -- (clf);
+     \draw [arrow, myarrow] (imgs) -- (aft);
+
+    \end{tikzpicture}
+  \end{center}
+  \caption{Pipeline.}
+  \label{fig:pipeline}
+\end{figure*}
+
 \section{Experiments}

### 2020-07-23 main.tex
plot of increasing train set size

@@
 
+\usepackage{pgfplots,filecontents}
+
 \DeclarePairedDelimiter\ceil{\lceil}{\rceil}
@@
 \definecolor{heliotrope}{RGB}{236,93,253}
+\definecolor{pink}{RGB}{250,38,160}
 
@@
 
+\begin{comment}
   \begin{table}[!t]
@@
   \end{center}
-
   \caption{Mean and standard deviation of precision, recall, and f-score for each method: a feature extractor with a classifier with a larger train set.}
@@
   \end{table}
+\end{comment}
+
+\begin{figure}
+\begin{tikzpicture}
+  \begin{axis}[
+    /pgf/number format/set thousands separator = {},
+    xlabel = Train set size,
+    ylabel = f-score,
+    xmin=300, xmax=1200,
+    grid=major,
+    grid style=dashed,
+    xtick={300,450, 600, 750, 900, 1050, 1200},
+    legend style={nodes={scale=0.65, transform shape}},
+    legend pos=south east
+    ]
+    \addplot table[x index=0,y index=1,header=false] {data/data.dat};
+    \addlegendentry{FLIM+MLP}
+
+    \addplot  table[x index=0,y index=2,header=false] {data/data.dat};
+    \addlegendentry{FLIM-FT+MLP}
+
+    \addplot  table[x index=0,y index=3,header=false] {data/data.dat};
+    \addlegendentry{VGG-FT}
+
+    \addplot table[x index=0,y index=4,header=false] {data/data.dat};
+    \addlegendentry{CL+MLP}
+
+    \legend{FLIM+MLP,FLIM-FT+MLP,VGG-FT,CL+MLP}
+
+  \end{axis}
+\end{tikzpicture}
+\caption{Graph showing the impact of training set size on quantitative results.
+FLIM+MLP feature extractor was trained only with four images, and the remaining training set was used to train the MLP classifier.}
+\end{figure}
 

### 2020-07-23 main.tex
professor's changes

@@
 
-A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation by eliminating negative activations, and aggregate the resulting activations locally (pooling) and globally (flattening) into a highly dimensional and sparse feature space, in which a linear classifier should separate the classes with reasonable success.
-Nevertheless, it is common to assume that the classes are piecewise linearly separable and apply fully connected layers -- a Multi-Layer Perceptron (MLP) classifier -- for feature space reduction and classification.
+\textcolor{red}{A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation by eliminating negative activations, and aggregate the resulting activations locally (pooling) and globally (flattening) into a highly dimensional and sparse feature space, in which a linear classifier should separate the classes with reasonable success.} Nevertheless, it is common to assume that the classes are piecewise linearly separable and apply fully connected layers -- a Multi-Layer Perceptron (MLP) classifier -- for feature space reduction and classification.
 The training of a CNN usually relies on weight optimization using backpropagation, which requires a higher number of annotated images as deep as the model is.
@@
 The Humanitarian OpenStreetMap community annotated the regions.
-The dataset consists of 13587 regions with $90 \times 90$ pixels with resolution under $10$\si{\square\centi\metre}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
-We converted the region images to the LAB color space, with each channel normalized within $[0,1]$.
+The dataset consists of 13587 regions with $90 \times 90$ pixels \textcolor{red}{with resolution under $10$\si{\square\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
+We converted the region images to the \textcolor{red}{LAB color space}, with each channel normalized within $[0,1]$.
 
-To assess the impact of FLIM in reducing the need for annotated samples, we randomly selected 100 regions from each class for training and left the remaining 11387 regions for testing.
-We repeated this procedure three times to compute the mean and the standard deviation of the results in precision, recall, and f-score.
-By vectorizing and projecting the 200 training images of each split on 2D with t-SNE~\cite{maaten2008visualizing}, we identified cohesive groups of images from each class and selected two representative images from each class to draw markers.
-The selected images from one split can be seen in Figure \ref{fig:markers}.
+To assess the impact of FLIM in reducing the need for annotated samples, we randomly selected 100 regions from each class for training and left \textcolor{red}{the remaining 11387 regions} for testing.
+\textcolor{red}{We repeated this procedure three times to compute the mean and the standard deviation of the results in precision, recall, and f-score.} By vectorizing and projecting the 200 training images of each split on 2D with t-SNE~\cite{maaten2008visualizing}, \textcolor{red}{we identified cohesive groups of images from each class and selected two representative images from each class to draw markers.}  The selected images from one split can be seen in Figure \ref{fig:markers}.
 We placed markers on these images in regions that we thought to better distinguish between coconut and non-coconut classes.

### 2020-07-23 IEEE-GRSL-Response-Letter.tex
Updates from Overleaf

%Fiquemos com Deus e Nossa Senhora!
\documentclass[a4paper,12pt]{article}

\usepackage[a4paper]{geometry}
\geometry{a4paper,left=20mm,right=20mm,top=35mm,bottom=35mm}

\usepackage[pdftex]{graphicx}
\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{color}
\usepackage{multicol}
\usepackage{mathtools}

\newtheorem{theo}{Theorem}
\newtheorem{defi}{Definition}%[section]
\newtheorem{prop}{Proposition}
\newtheorem{lemma}{Lemma}
\usepackage{url}
\setcounter{secnumdepth}{5}

\newcommand{\rwr}[1]{\par \medskip \noindent {\bf Reviewer #1: }}
\newcommand{\ans}{\smallskip\begin{quote}  \noindent }
\newcommand{\eans}{\end{quote}}

\newcommand{\edit}{\par \medskip\noindent {\bf Editor: }}

\title{Learning CNN filters from user-drawn image markers for coconut-tree image classification}

\author{Italos Estilon de Souza and Alexandre Xavier Falc\~{a}o}

\date{Manuscript GRSL-00821-2020}

\begin{document}
\maketitle

Dear Editor, please find in attach the revised version of our
manuscript GRSL-00821-2020. We would like to thank you and the
reviewers for the opportunity and comments to improve the
presentation of this work.

We have addressed all raised issues, highlighted in the text in red, and explained as follows.

\rwr{1} 
The manuscript entitled "Learning CNN ﬁlters from user-drawn image-makers for coconut-tree image classiﬁcation" presents a method that needs a minimal set of user-selected images to train the CNN’s feature extractor, reducing the number of required images to train the fully connected layers. The theme of this manuscript is interesting, and the text brings an important contribution to the remote sensing area using deep learning tools. However, to be considered adequated to publish in the journal "Geoscience and Remote Sensing Letters", it requires a revision on the text. The attached document presents all corrections and observations made for it. However, the inclusion of more experiments using more images to train the proposed model is a crucial issue in my point of view. This is important to prove the robustness of the proposed model.

\rwr{1} A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation by eliminating negative activations, and aggregate the resulting activations locally (pooling) and globally (flattening) into a highly dimensional and sparse feature space, in which a linear classifier should separate the classes with reasonable success.

\ans   \eans

\rwr{1} with resolution under 10 $cm^2$; LAB color space; 11387 regions; MLP classifier;

\ans CIE L*a*b* color space \eans

\rwr{1} We repeated this procedure three times to compute the mean and the standard deviation of the results in precision, recall, and f-score.

\ans \eans

\rwr{1} we identified cohesive groups of images from each class and selected two representative images from each class to draw markers.}

\ans \eans

\rwr{1} We have not investigated the impact of placing markers on a higher number of images and network architecture optimization procedures yet.

\ans \eans

\rwr{1} We use Pytorch to implement our feature extractor and the MLP classifier.

\ans \eans

\rwr{1} Caption of the previous Fig. 3 (Fig. 4 now): Markers used for training in image regions with (first column) and without (second column) a coconut tree.

\ans \eans

\rwr{1}{By comparing ``FLIM+MLP'' with VGG, one can observe that FLIM reduces the number of required training images. Finally, the comparison between ``FLIM+MLP'' and VGG-FT demonstrates that FLIM can provide competitive results using a considerably more simplified architecture than  VGG-FT.}

\ans \eans

\rwr{2} This paper presents a new feature learning approach that can be used to reduce the number of required images to train a deep neural network. It introduces a novel way to estimate filter weights for the convolutional layers using user-drawn markers in a few training images. The need for a large number of labeled input samples is a very well known and important problem for many CNN models and the proposed solution seems to be promising but the paper has some presentation and methodological problems that must be addressed.

\ans \eans

\rwr{2} Pg. 5 - Line 6: An error in the paper title is something that should be avoided at all costs as it is very confusing for the reader.  "...User-drawn Image MAKERS..." ??? The first word of the introduction is also wrong: "DePP learning"? Overall, there are not many language errors, but a proof-reading is highly recommended

\ans \eans

\rwr{2} Pg. 9 - Lines 14-18: Justify the choice of all hyperparameters (E.g. epochs, batch size, learning rate, etc). Have any previous experiments been made to tune these hyperparameters? Has the same dataset been used?

\ans \eans

\rwr{2} Pg. 10 - Table 1: The F-score difference from FLIM+MLP (0.854) and VGG+FT (0.851) is statistically significant considering the standard deviation?  Has a statistical hypothesis test been applied?

\ans \eans

\rwr{2} Pg. 11 - It would be worth investigating or discussing if the time taken to mark some images is always less than the time take to produce more labeled images (E.g.: how many new labeled images would take for VGG+FT beat FLIMP+MLP?). Is there a trade-off?

\ans \eans

\rwr{2} Pg. 12 - It seems that the literature review outdated (only one 2019/2020 work and from the same authors).

\ans

\eans

\end{document}

### 2020-07-23 main.tex
Updates from Overleaf

@@
 
-We have not investigated the impact of placing markers on a higher number of images and network architecture optimization procedures yet.
-We just trained the single-convolutional-layer network of Figure \ref{fig:arch} from markers drawn in two images per class.
+\textcolor{red}{We have not investigated the impact of placing markers on a higher number of images and network architecture optimization procedures yet.} We just trained the single-convolutional-layer network of Figure \ref{fig:arch} from markers drawn in two images per class.
 The convolutional layer has $60$ filters with dimension $7 \times 7$ and 3 color channels.
 It includes marker-based normalization before convolution, ReLu activation, and a max-pooling operation with a window of dimension $3 \times 3$.
-Since it is the single and last convolutional layer, we applied a stride of $4$ in max-pooling and batch normalization to create an input as close as possible to the one of the MLP classifier used in the popular VGG-16~\cite{simonyan2014very}.
-We use Pytorch to implement our feature extractor and the MLP classifier.
-We trained the MLP classifier for $40$ epochs and the whole network for $60$ epochs with a batch size of $64$.
+Since it is the single and last convolutional layer, we applied a stride of $4$ in max-pooling and batch normalization to create an input as close as possible to the one of the \textcolor{red}{MLP} classifier used in the popular VGG-16~\cite{simonyan2014very}.
+\textcolor{red}{We use Pytorch to implement our feature extractor and the MLP classifier.} We trained the MLP classifier for $40$ epochs and the whole network for $60$ epochs with a batch size of $64$.
 We use a learning rate of $0.0001$, a weight decay of $0.001$, and we update the learning rate by a factor of $0.1$ every $5$ epochs after epoch 30.
@@
 The results of ``FLIM+MLP'' and ``FLIM-FT+MLP'' show that FLIM  may dismiss fine tuning by directly providing a more effective solution.
-By comparing ``FLIM+MLP'' with VGG, one can observe that FLIM reduces the number of required training images.
-Finally, the comparison between ``FLIM+MLP'' and VGG-FT demonstrates that FLIM can provide competitive results using a considerably more simplified architecture than  VGG-FT.
+\textcolor{red}{By comparing ``FLIM+MLP'' with VGG, one can observe that FLIM reduces the number of required training images.} Finally, the comparison between \textcolor{red}{``FLIM+MLP'' and VGG-FT} demonstrates that FLIM can provide competitive results using a considerably more simplified architecture than  VGG-FT.
 The number of parameters of our feature extractor represents only $0.07\%$ of the number of parameter of VGG-FE.
@@
     }
-    \caption{Markers used for training in image regions with (first column) and without (second column) a coconut tree.}
+    \caption{\textcolor{red}{Markers used for training in image regions with (first column) and without (second column) a coconut tree.}}
     \label{fig:markers}
@@
     /pgf/number format/set thousands separator = {},
-    xlabel = Train set size,
+    xlabel = Training set size,
     ylabel = f-score,

### 2020-07-23 main.tex
change projection image in pipeline figure
pipeline add markers

@@
 \usetikzlibrary{backgrounds}
+\usetikzlibrary{decorations.pathreplacing}
 
@@
 
+    \tikzstyle{steps} = [rectangle, rounded corners, minimum width=1.3cm, minimum height=0.5cm,text centered, draw=pink, thick, fill=heliotrope!30, execute at begin node={\begin{varwidth}{2.5cm}},
+    execute at end node={\end{varwidth}}]
+    \tikzstyle{image} = [minimum width=1.3cm, minimum height=0.7cm,text centered, draw=white,fill=white!30, inner sep=0pt]
+
+    \tikzstyle{myarrow} = [navy, thick]
+
+    \tikzstyle{arrow} = [thick,->,>=stealth]
+
+    \begin{tikzpicture}[node distance=1.5cm]
+      %images----------------------------
+     \node (img1) [image] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-coco-789.png}};
+     \node (img2) [image, below left=-5em of img1] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-coco-789.png}};
+     \node (img3) [image, below left=-5em of img2] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-coco-789.png}};
+
+     \node (imgs) [fit={($(img1.north west)+(-0.45,0.1)$) ($(img3.south east)+(0.45,-0.1)$)}, inner sep=0pt, draw=none, rounded corners, very thick, label = {[label distance = 0.1em]below:{Training set}}] {};
+     %----------------------------------
+
+     \node (tsne) [steps, right=1em of imgs] {t-SNE};
+
+     \node (vis1) [image, right=1.5em of tsne, outer sep=0pt, label = {[label distance = 0.1em]below:{Projection}}] {\includegraphics[width=.12\textwidth]{figures/input_200.png}};
+
+     \node (select) [steps, right=1.35em of vis1] {Select\\ Images};
+
+    %images----------------------------
+    \node (simg1) [image, above right=-2.5em and 3em of select] {\includegraphics[width=.08\textwidth]{figures/img1.png}};
+    \node (simg2) [image, below left=-5em of simg1] {\includegraphics[width=.08\textwidth]{figures/img2.png}};
+    \node (simg3) [image, below left=-5em of simg2] {\includegraphics[width=.08\textwidth]{figures/img3.png}};
+    \node (simg4) [image, below left=-5em of simg3] {\includegraphics[width=.08\textwidth]{figures/img4.png}};
+
+    \node (simgs) [fit={($(simg1.north west)+(-0.64,0.2)$) ($(simg4.south east)+(0.64,-0.2)$)}, inner sep=0pt, draw=none, label = {[label distance = 0.0em]above:{Selected images}}] {};
+    %----------------------------------
+
+     \node (put) [steps, right=1em of simgs] {Select\\Markers};
+
+     %images----------------------------
+    \node (mimg1) [image, above right=-2em and 2.8em of put] {\includegraphics[width=.08\textwidth]{figures/markers1.png}};
+    \node (mimg2) [image, below left=-5em of mimg1] {\includegraphics[width=.08\textwidth]{figures/markers2.png}};
+    \node (mimg3) [image, below left=-5em of mimg2] {\includegraphics[width=.08\textwidth]{figures/markers3.png}};
+    \node (mimg4) [image, below left=-5em of mimg3] {\includegraphics[width=.08\textwidth]{figures/markers4.png}};
+
+    \node (mimgs) [fit={($(mimg1.north west)+(-0.64,0.2)$) ($(mimg4.south east)+(0.64,-0.2)$)}, inner sep=0pt, draw=none, label = {[label distance = 0.0em]above:{Markers}}] {};
+
+    \node (flim) [steps, right=1em of mimgs] {FLIM};
+
+    \node (ft) [steps, below=5em of flim] {Feature\\Extractor};
+
+    \node (user) [image, below=2em of simgs, outer sep = -.5em] {\includegraphics[width=.07\textwidth]{figures/user.png}};
+
+     %--------------------------------------------------------------arrows
+     \draw [arrow, myarrow] (imgs) -- (tsne);
+     \draw [arrow, myarrow] (tsne) -- (vis1);
+     \draw [arrow, myarrow] (vis1) -- (select);
+     \draw [arrow, myarrow] (select) -- (simgs);
+     \draw [arrow, myarrow] (simgs) -- (put);
+     \draw [arrow, myarrow] (put) -- (mimgs);
+     \draw [arrow, myarrow] (mimgs) -- (flim);
+     \draw [arrow, myarrow] (flim) -- (ft);
+     \draw [arrow, myarrow] (user) -- (select);
+     \draw [arrow, myarrow] (user) -- (put);
+
+     %\draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
+%(img3.west) -- (img1.north) node [black,midway,xshift=0.8cm, label = {[label distance = 0.1em]above:{Images}}] {};
+
+    \end{tikzpicture}
+  \end{center}
+  \caption{Pipeline.}
+  \label{fig:pipeline}
+\end{figure*}
+
+\begin{comment}
+
+\begin{figure*}
+  \begin{center}
+
     \tikzstyle{steps} = [rectangle, rounded corners, minimum width=1.3cm, minimum height=0.7cm,text centered, draw=black,fill=violet!30, execute at begin node={\begin{varwidth}{2.5cm}},
@@
 
-     \node (vis1) [image, right=1.5em of tsne, outer sep=-15pt] {\includegraphics[width=.15\textwidth]{figures/input.png}};
+     \node (vis1) [image, right=1.5em of tsne, outer sep=0pt] {\includegraphics[width=.12\textwidth]{figures/input_200.png}};
 
@@
 \end{figure*}
+\end{comment}
 

### 2020-07-23 main.tex
Updates from Overleaf

--- 2020-07-23 main.tex
@@
 They all lead to the importance of involving human experts during the machine learning process.
+To mitigate such problems,
+
+Explainable artificial intelligence (XAI) has appeared to address the problem and avoid wrong interpretations of the results~\cite{arrietaInfoFusion2020}.
+In methods of XAI, however, the importance of human participation during the model's training process has called little attention yet~\cite{AngelovNN2020}.
+A network designer must choose a suitable architecture and the training hyperparameters with limited control over the quality of the model's estimated weights.
+
 In this paper, we present significant advances related to the above questions.
@@
 
-In order to reduce the need for large annotated training sets by exploiting the user knowledge, we present a method, called FLIM (\emph{Feature Learning from Image Markers}), that needs a minimal set of images to learn the filters of each convolutional layer.
-The user selects the number of convolutional layers, their filter sizes, a few training images, and draws markers in regions of the input images that best discriminate the classes.
-Classes may appear in distinct clusters of a local feature space defined by patches extracted from marker pixels at each convolutional layer's input.
-FLIM defines the weights of the filters as the centroids of the clusters.
-Such filters are expected to activate discriminant regions in the given convolutional layer's output, and the process repeats in a layer-by-layer fashion.
-The architecture of the CNN could also be optimized by observing its results on a small validation set or based on data visualization techniques~\cite{rauber2016visualizing}.
+To reduce the need for large annotated training sets by exploiting the user knowledge, we present a method, called \textcolor{red}{\emph{Feature Learning from Image Markers} (FLIM)}, that needs a minimal set of images to learn the filters of each convolutional layer.
+The user selects the number of convolutional layers, their filter sizes, \textcolor{red}{a few images from a training set}, and draws markers in image regions that best discriminate the classes.
+\textcolor{red}{FLIM can find filters that enhance those regions at the output of the convolutional layers.
+It is based on clustering of patches extracted from marker pixels at the input of each convolutional layer.} The CNN architecture could also be optimized by observing its results on a small validation set or based on data visualization techniques~\cite{rauber2016visualizing}.
 
@@
 Third, by eliminating the optimization of the convolutional layers by backpropagation, FLIM can reduce the number of annotated images to train the fully connected layers.
-Fourth, FLIM is application-independent, and so it might be useful in other image classification problems.
+Fourth, FLIM is application-independent, and so it might be useful for other image classification problems.
 
@@
 
-\textcolor{red}{A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation by eliminating negative activations, and aggregate the resulting activations locally (pooling) and globally (flattening) into a highly dimensional and sparse feature space, in which a linear classifier should separate the classes with reasonable success.} Nevertheless, it is common to assume that the classes are piecewise linearly separable and apply fully connected layers -- a Multi-Layer Perceptron (MLP) classifier -- for feature space reduction and classification.
+\textcolor{red}{A CNN model should enhance image regions that best discriminate classes through a sequence of convolutional layers.
+This section shows how the identification of such regions by a simple user action can make that sequence an effective feature extractor.
+The method is named \emph{Feature Learning from Image Markers} (FLIM).
+In the feature space that results from convolutional layers,} it is common to assume that the classes are piecewise linearly separable and apply fully connected layers -- a Multi-Layer Perceptron (MLP) classifier -- for feature space reduction and classification.
 The training of a CNN usually relies on weight optimization using backpropagation, which requires a higher number of annotated images as deep as the model is.
@@
 The Humanitarian OpenStreetMap community annotated the regions.
-The dataset consists of 13587 regions with $90 \times 90$ pixels \textcolor{red}{with resolution under $10$\si{\square\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
-We converted the region images to the \textcolor{red}{LAB color space}, with each channel normalized within $[0,1]$.
+The dataset consists of 13587 regions with $90 \times 90$ pixels \textcolor{red}{with spatial resolution $8$\si{\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
+\textcolor{red}{The region images contain three channels: red, green, and blue.
+We converted them to the CIE L*a*b* color space and then normalized the channels within $[0,1]$}.
 
-To assess the impact of FLIM in reducing the need for annotated samples, we randomly selected 100 regions from each class for training and left \textcolor{red}{the remaining 11387 regions} for testing.
+\textcolor{red}{
+To assess the impact of FLIM in reducing the need for annotated samples, we
+divided the experiments in three parts.
+First, we randomly selected 200 regions (100 from each class) for training, 1000 regions (500 from each class) for the choice of architecture, and left the remaining 12387 regions for testing.
+The architecture of the network was chosen based on empirical tests on the validation set.
+Second, we used the validation set to increase the size of the training set from 200 to 1200 images.
+
 \textcolor{red}{We repeated this procedure three times to compute the mean and the standard deviation of the results in precision, recall, and f-score.} By vectorizing and projecting the 200 training images of each split on 2D with t-SNE~\cite{maaten2008visualizing}, \textcolor{red}{we identified cohesive groups of images from each class and selected two representative images from each class to draw markers.}  The selected images from one split can be seen in Figure \ref{fig:markers}.
@@
 Since it is the single and last convolutional layer, we applied a stride of $4$ in max-pooling and batch normalization to create an input as close as possible to the one of the \textcolor{red}{MLP} classifier used in the popular VGG-16~\cite{simonyan2014very}.
-\textcolor{red}{We use Pytorch to implement our feature extractor and the MLP classifier.} We trained the MLP classifier for $40$ epochs and the whole network for $60$ epochs with a batch size of $64$.
+\textcolor{red}{We use Pytorch to implement our feature extractor and the MLP classifier.The experiments used an Intel(R) Xeon(R) CPU E5-2620 v4 @ 2.10GHz CPU and a Titan X with 12 GB of memory.} We trained the MLP classifier for $40$ epochs and the whole network for $60$ epochs with a batch size of $64$.
 We use a learning rate of $0.0001$, a weight decay of $0.001$, and we update the learning rate by a factor of $0.1$ every $5$ epochs after epoch 30.

### 2020-07-23 main.tex
pipeline to train classifier

--- 2020-07-23 main.tex
@@
   \end{center}
-  \caption{Pipeline.}
-  \label{fig:pipeline}
+  \caption{Pipeline to trian a feature extractor.}
+  \label{fig:pipeline-xtractor}
 \end{figure*}
-
-\begin{comment}
 
@@
 
-    \tikzstyle{steps} = [rectangle, rounded corners, minimum width=1.3cm, minimum height=0.7cm,text centered, draw=black,fill=violet!30, execute at begin node={\begin{varwidth}{2.5cm}},
+    \tikzstyle{steps} = [rectangle, rounded corners, minimum width=1.3cm, minimum height=0.5cm,text centered, draw=pink, thick, fill=heliotrope!30, execute at begin node={\begin{varwidth}{2.5cm}},
       execute at end node={\end{varwidth}}]
@@
     \begin{tikzpicture}[node distance=1.5cm]
+
       %images----------------------------
@@
 
-     \node (imgs) [fit={($(img1.north west)+(-0.45,0.2)$) ($(img3.south east)+(0.45,-0.2)$)}, inner sep=0pt, draw=none, rounded corners, very thick, label = {[label distance = 0.3em]below:{Images}}] {};
+     \node (imgs) [fit={($(img1.north west)+(-0.45,0.1)$) ($(img3.south east)+(0.45,-0.1)$)}, inner sep=0pt, draw=none, rounded corners, very thick, label = {[label distance = 0.1em]below:{Training set}}] {};
      %----------------------------------
 
-     \node (tsne) [steps, right=1em of imgs] {t-SNE};
-
-     \node (vis1) [image, right=1.5em of tsne, outer sep=0pt] {\includegraphics[width=.12\textwidth]{figures/input_200.png}};
-
-     \node (select) [steps, right=1.35em of vis1] {Select Images};
+     \node (aft) [steps, right=2em of imgs] {Apply Featute\\Extractor};
 
       %images----------------------------
-    \node (simg1) [image, above right=-2em and 3em of select] {\includegraphics[width=.08\textwidth]{figures/img1.png}};
-    \node (simg2) [image, below left=-5em of simg1] {\includegraphics[width=.08\textwidth]{figures/img2.png}};
-    \node (simg3) [image, below left=-5em of simg2] {\includegraphics[width=.08\textwidth]{figures/img3.png}};
-    \node (simg4) [image, below left=-5em of simg3] {\includegraphics[width=.08\textwidth]{figures/img4.png}};
+    \node (f1) [image, above right=-3em and 3em of aft] {\includegraphics[width=.08\textwidth]{figures/features/f1.png}};
+    \node (f2) [image, below left=-5.5em of f1] {\includegraphics[width=.08\textwidth]{figures/features/f2.png}};
+    \node (f3) [image, below left=-5.5em of f2] {\includegraphics[width=.08\textwidth]{figures/features/f3.png}};
+    \node (f4) [image, below left=-5.5em of f3] {\includegraphics[width=.08\textwidth]{figures/features/f4.png}};
 
-    \node (simgs) [fit={($(simg1.north west)+(-0.64,0.2)$) ($(simg4.south east)+(0.64,-0.2)$)}, inner sep=0pt, draw=none] {};
+    \node (features) [fit={($(f1.north west)+(-0.3,0.2)$) ($(f4.south east)+(0.3,-0.2)$)}, draw=none, inner sep=0pt, label = {[label distance = 0.1em]below:{Features}}] {};
     %----------------------------------
 
-     \node (put) [steps, right=1em of simgs] {Put Markers};
+     \node (sup) [steps, right=2em of features] {Supervised\\Learning};
 
-     %images----------------------------
-    \node (mimg1) [image, above right=-2em and 3em of put] {\includegraphics[width=.08\textwidth]{figures/markers1.png}};
-    \node (mimg2) [image, below left=-5em of mimg1] {\includegraphics[width=.08\textwidth]{figures/markers2.png}};
-    \node (mimg3) [image, below left=-5em of mimg2] {\includegraphics[width=.08\textwidth]{figures/markers3.png}};
-    \node (mimg4) [image, below left=-5em of mimg3] {\includegraphics[width=.08\textwidth]{figures/markers4.png}};
-
-    \node (mimgs) [fit={($(mimg1.north west)+(-0.64,0.2)$) ($(mimg4.south east)+(0.64,-0.2)$)}, inner sep=0pt, draw=none] {};
-    %----------------------------------
-
-     \node (flim) [steps, below left=5em and -3em of mimgs] {FLIM};
-
-     \node (ft) [steps, left=2em of flim] {Feature Extractor};
-
-     \node (aft) [steps, left=2em of ft] {Apply Featute \\ Extractor};
-
-      %images----------------------------
-    \node (f1) [image, above left=-2.8em and 2em of aft] {\includegraphics[width=.07\textwidth]{figures/features/f1.png}};
-    \node (f2) [image, below left=-5em of f1] {\includegraphics[width=.07\textwidth]{figures/features/f2.png}};
-    \node (f3) [image, below left=-5em of f2] {\includegraphics[width=.07\textwidth]{figures/features/f3.png}};
-    \node (f4) [image, below left=-5em of f3] {\includegraphics[width=.07\textwidth]{figures/features/f4.png}};
-
-    \node (features) [fit={($(f1.north west)+(-0.17,0.2)$) ($(f4.south east)+(0.17,-0.2)$)}, draw=none, inner sep=0pt] {};
-    %----------------------------------
-
-     \node (clf) [steps, left=2em of features] {Train classifier};
+     \node (clf) [steps, right=2em of sup] {Classifier};
 
      %--------------------------------------------------------------arrows
-     \draw [arrow, myarrow] (imgs) -- (tsne);
-     \draw [arrow, myarrow] (tsne) -- (vis1);
-     \draw [arrow, myarrow] (vis1) -- (select);
-     \draw [arrow, myarrow] (select) -- (simgs);
-     \draw [arrow, myarrow] (simgs) -- (put);
-     \draw [arrow, myarrow] (put) -- (mimgs);
-     \draw [arrow, myarrow] (mimgs) -- (flim);
-     \draw [arrow, myarrow] (flim) -- (ft);
-     \draw [arrow, myarrow] (ft) -- (aft);
+
+     \draw [arrow, myarrow] (imgs) -- (aft);
      \draw [arrow, myarrow] (aft) -- (features);
-     \draw [arrow, myarrow] (features) -- (clf);
-     \draw [arrow, myarrow] (imgs) -- (aft);
+     \draw [arrow, myarrow] (features) -- (sup);
+     \draw [arrow, myarrow] (sup) -- (clf);
 
@@
   \end{center}
-  \caption{Pipeline.}
-  \label{fig:pipeline}
+  \caption{Pipeline to train a classifier.}
+  \label{fig:pipeline-classifier}
 \end{figure*}
-\end{comment}
 

### 2020-07-23 IEEE-GRSL-Response-Letter.tex

@@
 
-Dear Editor, please find in attach the revised version of our
-manuscript GRSL-00821-2020. We would like to thank you and the
-reviewers for the opportunity and comments to improve the
-presentation of this work.
-
-We have addressed all raised issues, highlighted in the text in red, and explained as follows.
+Dear Editor, please find attached the revised version of our manuscript, GRSL-00821-2020. We thank the Editor and
+reviewers for the opportunity and comments to improve the presentation of our work.
+We have addressed all raised issues and highlighted the changes in red.
+Our point-by-point response to the reviewers' comments follows next.
 
@@
 
+\ans
+
+Thanks for the comments.
+We have included more experiments to demonstrate the robustness of the proposed feature learning approach.
+The parts marked and commented by reviewer 1 in the manuscript are reproduced and addressed as follows.
+
+\eans
+
 \rwr{1} A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation by eliminating negative activations, and aggregate the resulting activations locally (pooling) and globally (flattening) into a highly dimensional and sparse feature space, in which a linear classifier should separate the classes with reasonable success.
 
-\ans   \eans
+\ans  Thanks. The sentence was long and unclear. We have simplified it and improved it in first paragraph of Section II.
+\eans
 
@@
 
-\ans CIE L*a*b* color space \eans
+\ans Thanks. These are minor corrections in different parts of the text that have been solved now. (Section III). \eans
 

### 2020-07-23 main.tex
Updates from Overleaf

--- 2020-07-23 main.tex
@@
 They all lead to the importance of involving human experts during the machine learning process.
-To mitigate such problems,
+\textcolor{red}{Explainable artificial intelligence (XAI) has appeared to address such problems~\cite{rauber2016visualizing,arrietaInfoFusion2020}.
+However, the importance of human participation during the training process has called little attention~\cite{AngelovNN2020}.
+Exceptions rely on visual analytics techniques~\cite{GarciaIJCNN2019,SpinnerTVCG2020}, and we present here a first work in which the user has direct influence over the quality of the filters.}
 
-Explainable artificial intelligence (XAI) has appeared to address the problem and avoid wrong interpretations of the results~\cite{arrietaInfoFusion2020}.
-In methods of XAI, however, the importance of human participation during the model's training process has called little attention yet~\cite{AngelovNN2020}.
-A network designer must choose a suitable architecture and the training hyperparameters with limited control over the quality of the model's estimated weights.
-
-In this paper, we present significant advances related to the above questions.
 First, we conceptually divide a CNN into convolutional layers for feature extraction and fully connected layers for feature space reduction and classification.
@@
 The user selects the number of convolutional layers, their filter sizes, \textcolor{red}{a few images from a training set}, and draws markers in image regions that best discriminate the classes.
-\textcolor{red}{FLIM can find filters that enhance those regions at the output of the convolutional layers.
-It is based on clustering of patches extracted from marker pixels at the input of each convolutional layer.} The CNN architecture could also be optimized by observing its results on a small validation set or based on data visualization techniques~\cite{rauber2016visualizing}.
+\textcolor{red}{By clustering patches extracted from marker pixels at the input of each convolutional layer, FLIM finds filters that enhance those regions at the output of the convolutional layers.} The CNN architecture could also be optimized based on visual analytics~\cite{SpinnerTVCG2020}.
 
@@
 
-We used a dataset from~\cite{8899005}, which contains regions of aerial images with and without coconut trees from the Kingdom of Tonga, as acquired in October 2017~\footnote{The images are available in \url{blog.werobotics.org/2018/01/11/open-ai-challenge-2}.}.
-Each region should be classified as containing one or none coconut tree, but part of coconut trees from adjacent regions might appears near the region's border.
+We use a dataset from~\cite{8899005}, which contains regions of aerial images with and without coconut trees from the Kingdom of Tonga, as acquired in October 2017~\footnote{The images are available in \url{blog.werobotics.org/2018/01/11/open-ai-challenge-2}.}.
+Each region is classified as containing one or none coconut tree, but part of coconut trees from adjacent regions might appear near the region's border.
 The Humanitarian OpenStreetMap community annotated the regions.
-The dataset consists of 13587 regions with $90 \times 90$ pixels \textcolor{red}{with spatial resolution $8$\si{\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
+The dataset consists of 13587 regions with $90 \times 90$ pixels \textcolor{red}{with spatial resolution of $8$\si{\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
 \textcolor{red}{The region images contain three channels: red, green, and blue.
-We converted them to the CIE L*a*b* color space and then normalized the channels within $[0,1]$}.
+We convert them to the CIE L*a*b* color space and normalize each channel within $[0,1]$}.
 
-\textcolor{red}{
+\textcolor{red}{The experiments have been designed to demonstrate the following hinvolved a dataset with X coconut-tree images from two classes, coconut-tree (Y images) and non-coconut-tree (Z images).
+All experiments were repeated T times, using random splits into training and testing sets, for statistical analysis with p-value 0.05 in Wilcoxon signed-rank test.
+The results were found statistically significant.
+The experiments have been designed to assess the following hypothesis about FLIM:
+
+It is possible to learn competitive image feature extractors from user-drawn markers in a few training images per class.
+
+We selected only 4 images from a training set with 100 images per class for the user to draw markers.
+An image feature extractor was trained by FLIM and the training set was used to design SVM and MLP classifiers for comparison with a well-established CNN, VGG-16.
+VGG was trained from scratch and also pre-trained on the ImageNet dataset and fine-tuned on the 200 training images.
+The classification results on the test set with the remaining W images indicate that the image feature extractor produced by FLIM can provide higher f-score than the feature extractor of VGG, with and without fine-tuning, and for both classifiers, SVM and MLP.
+These results are shown in Table I when we compare FLIM+SVM x VGG-FE+SVM, FLIM+MLP x VGG, and FLIM+MLP x VGG-FT.
+
+For a given network architecture, the image features obtained by FLIM can produce similar or better classification results than those obtained by backpropagation on a same network architecture.
+
+The comparison in Table I between FLIM+MLP and FLIM-FT+MLP shows that fine-tuning cannot improve the results already obtained by FLIM+MLP.
+Similarly, the comparison between FLIM+MLP x CL+MLP in Table I shows that a network with the same architecture of FLIM+MLP, trained from scratch, cannot surpass its results.
+
+The competitive classification results obtained by FLIM remain competitive independently of the number of samples used to train the classifier.
+
+We then repeated the comparisons between FLIM+MLP, FLIM-FT+MLP, and VGG-FT, the most competitive approaches, for increasing training-set sizes from 200 to 1200, and FLIM+MLP remained superior.
+The f-score curves of the methods in Figure X confirms the quality of the image feature extractor obtained by FLIM.
+
+By increasing the number of relevant images selected to draw markers, one can improve the image feature extractor produced by FLIM.
+
+            We finally compared FLIM+SVM, for the sake of simplicity, with an increasing number of images for marker selection (x1, x2, x3 images).
+As shown in Table II for X training images used to train the SVM classifier, this hypothesis is true.
+It is important to note that the results are promising but the selection of relevant images for marker selection is still a topic that deserves further investigation.
+
 To assess the impact of FLIM in reducing the need for annotated samples, we

### 2020-07-23 main.tex
Updates from Overleaf

@@
 
-\textcolor{red}{The experiments have been designed to demonstrate the following hinvolved a dataset with X coconut-tree images from two classes, coconut-tree (Y images) and non-coconut-tree (Z images).
-All experiments were repeated T times, using random splits into training and testing sets, for statistical analysis with p-value 0.05 in Wilcoxon signed-rank test.
-The results were found statistically significant.
-The experiments have been designed to assess the following hypothesis about FLIM:
-
-It is possible to learn competitive image feature extractors from user-drawn markers in a few training images per class.
-
-We selected only 4 images from a training set with 100 images per class for the user to draw markers.
-An image feature extractor was trained by FLIM and the training set was used to design SVM and MLP classifiers for comparison with a well-established CNN, VGG-16.
-VGG was trained from scratch and also pre-trained on the ImageNet dataset and fine-tuned on the 200 training images.
-The classification results on the test set with the remaining W images indicate that the image feature extractor produced by FLIM can provide higher f-score than the feature extractor of VGG, with and without fine-tuning, and for both classifiers, SVM and MLP.
-These results are shown in Table I when we compare FLIM+SVM x VGG-FE+SVM, FLIM+MLP x VGG, and FLIM+MLP x VGG-FT.
-
-For a given network architecture, the image features obtained by FLIM can produce similar or better classification results than those obtained by backpropagation on a same network architecture.
-
-The comparison in Table I between FLIM+MLP and FLIM-FT+MLP shows that fine-tuning cannot improve the results already obtained by FLIM+MLP.
-Similarly, the comparison between FLIM+MLP x CL+MLP in Table I shows that a network with the same architecture of FLIM+MLP, trained from scratch, cannot surpass its results.
-
-The competitive classification results obtained by FLIM remain competitive independently of the number of samples used to train the classifier.
-
-We then repeated the comparisons between FLIM+MLP, FLIM-FT+MLP, and VGG-FT, the most competitive approaches, for increasing training-set sizes from 200 to 1200, and FLIM+MLP remained superior.
-The f-score curves of the methods in Figure X confirms the quality of the image feature extractor obtained by FLIM.
-
-By increasing the number of relevant images selected to draw markers, one can improve the image feature extractor produced by FLIM.
-
-            We finally compared FLIM+SVM, for the sake of simplicity, with an increasing number of images for marker selection (x1, x2, x3 images).
-As shown in Table II for X training images used to train the SVM classifier, this hypothesis is true.
-It is important to note that the results are promising but the selection of relevant images for marker selection is still a topic that deserves further investigation.
-
 To assess the impact of FLIM in reducing the need for annotated samples, we

### 2020-07-23 main.tex
update pipeline figures

@@
 
-    \node (ft) [steps, below=5em of flim] {Feature\\Extractor};
+    \node (ft) [steps, below=5em of flim, fill=white, draw=none] {Feature\\Extractor};
 
@@
 
-     \node (aft) [steps, right=2em of imgs] {Apply Featute\\Extractor};
+     \node (aft) [steps, right=2em of imgs] {Feature\\Extraction};
 
@@
 
-     \node (clf) [steps, right=2em of sup] {Classifier};
+     \node (clf) [steps, right=2em of sup, fill=none, draw=none] {Classifier};
 

### 2020-07-23 main.tex
Updates from Overleaf

--- 2020-07-23 main.tex
@@
 
+\textcolor{red}{
+The experiments have been designed to evaluate the following hypotheses about FLIM.
+\begin{enumerate}
+ \item[(H1)]  FLIM can produce competitive image feature extractors from user-drawn markers in a few images selected per class from a training set. 
+\item[(H2)]  For a given network architecture, FLIM can provide better classification results than those obtained by training that network with backpropagation.
+\item[(H3)] Its feature extractor remains competitive independently of the training set size. 
+\item[(H4)] User-drawn markers on a higher number of selected images from the training set can improve image feature extraction.
+\end{enumerate}
+}
+
 To assess the impact of FLIM in reducing the need for annotated samples, we

### 2020-07-23 main.tex
Updates from Overleaf

@@
 
-     \node (imgs) [fit={($(img1.north west)+(-0.45,0.1)$) ($(img3.south east)+(0.45,-0.1)$)}, inner sep=0pt, draw=none, rounded corners, very thick, label = {[label distance = 0.1em]below:{Training set}}] {};
+     \node (imgs) [fit={($(img1.north west)+(-0.45,0.1)$) ($(img3.south east)+(0.45,-0.1)$)}, inner sep=0pt, draw=none, rounded corners, very thick, label = {[label distance = 0.1em]below:{Testing set}}] {};
      %----------------------------------
@@
 
-     \node (sup) [steps, right=2em of features] {Supervised\\Learning};
+     \node (sup) [steps, right=2em of features] {Classification};
 
@@
  \item[(H1)]  FLIM can produce competitive image feature extractors from user-drawn markers in a few images selected per class from a training set. 
-\item[(H2)]  For a given network architecture, FLIM can provide better classification results than those obtained by training that network with backpropagation.
+\item[(H2)]  For a given network architecture, FLIM can provide better classification results than those obtained by training the same architecture with backpropagation.
 \item[(H3)] Its feature extractor remains competitive independently of the training set size. 
-\item[(H4)] User-drawn markers on a higher number of selected images from the training set can improve image feature extraction.
+\item[(H4)] User-drawn markers on a higher number of selected images from the training set can improve feature extraction.
 \end{enumerate}

### 2020-07-24 main.tex
Updates from Overleaf

--- 2020-07-23 main.tex
@@
   \end{center}
-  \caption{Pipeline to trian a feature extractor.}
+  \caption{Pipeline for training a feature extractor.}
   \label{fig:pipeline-xtractor}
-\end{figure*}
-
-\begin{figure*}
-  \begin{center}
-
-    \tikzstyle{steps} = [rectangle, rounded corners, minimum width=1.3cm, minimum height=0.5cm,text centered, draw=pink, thick, fill=heliotrope!30, execute at begin node={\begin{varwidth}{2.5cm}},
-      execute at end node={\end{varwidth}}]
-      \tikzstyle{image} = [minimum width=1.3cm, minimum height=0.7cm,text centered, draw=white,fill=white!30, inner sep=0pt]
-
-    \tikzstyle{myarrow} = [navy, thick]
-
-    \tikzstyle{arrow} = [thick,->,>=stealth]
-
-    \begin{tikzpicture}[node distance=1.5cm]
-
-      %images----------------------------
-     \node (img1) [image] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-coco-789.png}};
-     \node (img2) [image, below left=-5em of img1] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-coco-789.png}};
-     \node (img3) [image, below left=-5em of img2] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-coco-789.png}};
-
-     \node (imgs) [fit={($(img1.north west)+(-0.45,0.1)$) ($(img3.south east)+(0.45,-0.1)$)}, inner sep=0pt, draw=none, rounded corners, very thick, label = {[label distance = 0.1em]below:{Training set}}] {};
-     %----------------------------------
-
-     \node (aft) [steps, right=2em of imgs] {Feature\\Extraction};
-
-      %images----------------------------
-    \node (f1) [image, above right=-3em and 3em of aft] {\includegraphics[width=.08\textwidth]{figures/features/f1.png}};
-    \node (f2) [image, below left=-5.5em of f1] {\includegraphics[width=.08\textwidth]{figures/features/f2.png}};
-    \node (f3) [image, below left=-5.5em of f2] {\includegraphics[width=.08\textwidth]{figures/features/f3.png}};
-    \node (f4) [image, below left=-5.5em of f3] {\includegraphics[width=.08\textwidth]{figures/features/f4.png}};
-
-    \node (features) [fit={($(f1.north west)+(-0.3,0.2)$) ($(f4.south east)+(0.3,-0.2)$)}, draw=none, inner sep=0pt, label = {[label distance = 0.1em]below:{Features}}] {};
-    %----------------------------------
-
-     \node (sup) [steps, right=2em of features] {Supervised\\Learning};
-
-     \node (clf) [steps, right=2em of sup, fill=none, draw=none] {Classifier};
-
-     %--------------------------------------------------------------arrows
-
-     \draw [arrow, myarrow] (imgs) -- (aft);
-     \draw [arrow, myarrow] (aft) -- (features);
-     \draw [arrow, myarrow] (features) -- (sup);
-     \draw [arrow, myarrow] (sup) -- (clf);
-
-    \end{tikzpicture}
-  \end{center}
-  \caption{Pipeline to train a classifier.}
-  \label{fig:pipeline-classifier}
 \end{figure*}
@@
 
-\textcolor{red}{
-The experiments have been designed to evaluate the following hypotheses about FLIM.
+\textcolor{red}{The experiments used an Intel(R) Xeon(R) CPU E5-2620 v4 @ 2.10GHz CPU and a Titan X with 12 GB of memory. They have been designed to evaluate the following hypotheses about FLIM.
 \begin{enumerate}
- \item[(H1)]  FLIM can produce competitive image feature extractors from user-drawn markers in a few images selected per class from a training set. 
-\item[(H2)]  For a given network architecture, FLIM can provide better classification results than those obtained by training that network with backpropagation.
-\item[(H3)] Its feature extractor remains competitive independently of the training set size. 
-\item[(H4)] User-drawn markers on a higher number of selected images from the training set can improve image feature extraction.
+ \item[H1]  It can produce competitive feature extractors from user-drawn markers in a few images selected per class from a training set. 
+\item[H2]  For a given network architecture, FLIM can provide better classification results than those obtained by training the same architecture with backpropagation.
+\item[H3] Its feature extractor remains competitive independently of the training set size. 
+\item[H4] User-drawn markers on a higher number of selected images can improve feature extraction.
 \end{enumerate}
@@
 
-To assess the impact of FLIM in reducing the need for annotated samples, we
-divided the experiments in three parts.
-First, we randomly selected 200 regions (100 from each class) for training, 1000 regions (500 from each class) for the choice of architecture, and left the remaining 12387 regions for testing.
-The architecture of the network was chosen based on empirical tests on the validation set.
-Second, we used the validation set to increase the size of the training set from 200 to 1200 images.
+\textcolor{red}{For these experiments, we have selected the following models that combine feature extraction and classification methods.
+ \begin{itemize}
+\item FLIM+SVM: Image classification is performed by support vector machines (SVM) on FLIM-based features.  
+\item FLIM+MLP: Image classification is performed by multi-layer perceptrons (MLP) on FLIM-based features.  
+\item FLIM-FT+MLP: The network in FLIM+MLP is fine-tuned by backpropagation. 
+\item CL+MLP: The same architecture used in FLIM+MLP is trained from scratch by backpropagation.
+\item VGG: Vgg-16~\cite{simonyan2014very} is trained from scratch.
+\item VGG-FT: Vgg-16 is pre-trained on ImageNet and fine-tuned on the coconut-tree training set.
+\item VGG-FE+SVM: Image features are obtained from the last convolutional layer of VGG-FT.
+\end{itemize}
+}
 
-\textcolor{red}{We repeated this procedure three times to compute the mean and the standard deviation of the results in precision, recall, and f-score.} By vectorizing and projecting the 200 training images of each split on 2D with t-SNE~\cite{maaten2008visualizing}, \textcolor{red}{we identified cohesive groups of images from each class and selected two representative images from each class to draw markers.}  The selected images from one split can be seen in Figure \ref{fig:markers}.
-We placed markers on these images in regions that we thought to better distinguish between coconut and non-coconut classes.
-
-\textcolor{red}{We have not investigated the impact of placing markers on a higher number of images and network architecture optimization procedures yet.} We just trained the single-convolutional-layer network of Figure \ref{fig:arch} from markers drawn in two images per class.
+\textcolor{red}{First, we evaluated H1 and H2 by randomly selecting 200 images for the training set, 1000 images for a validation set, and the remaining 12387 images were left in the testing set.
+The validation set was used to choose empirically the architecture of the network shown in Figure~\ref{fig:arch}.
 The convolutional layer has $60$ filters with dimension $7 \times 7$ and 3 color channels.
@@
 Since it is the single and last convolutional layer, we applied a stride of $4$ in max-pooling and batch normalization to create an input as close as possible to the one of the \textcolor{red}{MLP} classifier used in the popular VGG-16~\cite{simonyan2014very}.
-\textcolor{red}{We use Pytorch to implement our feature extractor and the MLP classifier.The experiments used an Intel(R) Xeon(R) CPU E5-2620 v4 @ 2.10GHz CPU and a Titan X with 12 GB of memory.} We trained the MLP classifier for $40$ epochs and the whole network for $60$ epochs with a batch size of $64$.
-We use a learning rate of $0.0001$, a weight decay of $0.001$, and we update the learning rate by a factor of $0.1$ every $5$ epochs after epoch 30.
+We then vectorized and projected the training images by t-SNE~\cite{maaten2008visualizing} (Figure~\ref{fig:pipeline-xtractor}).
+Four representative images, only two per class, were chosen by the user to draw markers (Figure~\ref{fig:markers}).
+These images must come from regions in the projection populated by samples from the same class (either red or green points).
+Those regions indicate groups of images from the same class in the input feature space.
+The markers must be placed in regions that best distinguish the coconut and non-coconut classes.}
 
-We selected VGG-16 for comparative analysis since it usually requires many more samples than 200 to be trained from scratch.
-It is also a very competitive CNN when pretrained in the ImageNet dataset, working as a counterpart solution for small training sets.
-Therefore, as baselines, we used three solutions based on VGG-16: (1) VGG is trained from scratch, (2) VGG-FT is pretrained in the ImageNet dataset and fine-tuned in the coconut-tree training set, and (3) VGG-FE is the feature extractor of VGG-16 (the output of its last convolutional layer) pretrained in the ImageNet dataset.
-In this case, we used for classification the  linear SVM classifier with $C=0.01$ from sklearn~\cite{fan2008liblinear, scikit-learn}.
-We use Pytorch's implementation of VGG.
-In (1) and (2), we trained for $95$ epochs with a batch size of $64$, learning rate of $0.001$, weight decay of $0.001$, and we update learning rate by a factor of $0.1$ every $5$ epochs after the model achieves an accuracy of $0.95$.
+\textcolor{red}{Pytorch was chosen to implement FLIM and MLP, being SVM the linear model available in sklearn with $C=0.01$~\cite{fan2008liblinear, scikit-learn}.
+We trained MLP for $40$ epochs and the whole network for $60$ epochs with a batch size of $64$.
+We used a learning rate of $0.0001$, a weight decay of $0.001$, and we updated the learning rate by a factor of $0.1$ at every $5$ epochs after epoch 30.
+VGG and VGG-FT were trained for $95$ epochs with a batch size of $64$, a learning rate of $0.001$, weight decay of $0.001$, and we updated the learning rate by a factor of $0.1$ at every $5$ epochs after the model achieved accuracy $0.95$.}
 
-Table~\ref{tab:results} shows the results of three FLIM-based solutions using the SVM and MLP classifiers; ``FLIM+SVM'', ``FLIM+MLP'', and ``FLIM-FT+MLP'' (i.e., FLIM with fine-tuning); as compared with the VGG-based solutions above.
-The first two rows indicate that FLIM  can provide a more effective feature extractor using a single convolutional layer than VGG-FE, which contains 13 convolutional layers pretrained on the ImageNet.
-The results of ``FLIM+MLP'' and ``FLIM-FT+MLP'' show that FLIM  may dismiss fine tuning by directly providing a more effective solution.
-\textcolor{red}{By comparing ``FLIM+MLP'' with VGG, one can observe that FLIM reduces the number of required training images.} Finally, the comparison between \textcolor{red}{``FLIM+MLP'' and VGG-FT} demonstrates that FLIM can provide competitive results using a considerably more simplified architecture than  VGG-FT.
-The number of parameters of our feature extractor represents only $0.07\%$ of the number of parameter of VGG-FE.
+\textcolor{red}{
+Table~\ref{tab:results} shows the mean results of this first experiment for three random splits of the dataset into training, validation, and testing sets.
+We considered better (in bold) the results with non-overlapping intervals of standard deviation around the mean value.
+The comparison between FLIM+SVM and VGG-FE+SVM indicates that FLIM can produce a feature extractor with a single layer competitive to VGG-FE using 13 layers (i.e., FLIM uses only $0.07\%$ of the number of parameters in VGG-FE).
+The results of FLIM+MLP, CL+MLP, and FLIM-FT+MLP show that FLIM may dismiss training and fine-tuning based on backpropagation, by directly providing a more effective solution than CL+MLP and FLIM-FT+MLP.
+This result can also be noticed when comparing  FLIM+MLP with VGG, confirming that FLIM can reduce the number of required training images for the MLP classifier.
+The comparison between FLIM+MLP and VGG-FT shows competitive results with a top-rated solution.}
+
+\textcolor{red}{To evaluate H3, we used the validation set to increase the classifiers' number of training images.
+Note that the marker selection for FLIM was still the same from those four images.
+To our surprise, as the training set increases, the differences in f-score favor FLIM+MLP over FLIM-FT+MLP, VGG-FT, and CL+MLP.
+The difference between FLIM+MLP and CL+MLP also demonstrated that FLIM was superior to backpropagation for filter estimation.
+Finally, Table II shows the results of FLIM+SVM as the number of images for marker selection increases, and the training set size remains with 200 images.
+We can notice a slight improvement in all measures, but this is a point that deserves further investigation -- i.e., choosing relevant training images and markers for FLIM.
+}
 
@@
     }
-    \caption{\textcolor{red}{Markers used for training in image regions with (first column) and without (second column) a coconut tree.}}
+    \caption{\textcolor{red}{Markers used for training, cyan markers on coconut trees and orange markers on other regions.}}
     \label{fig:markers}
@@
     FLIM-FT+MLP & $0.845 \pm 0.003$ & $0.825\pm 0.006$ & $0.832 \pm 0.005$ \\
-
+    CL+MLP & $0.822 \pm 0.011$ & $0.805 \pm 0.013$ & $0.811 \pm 0.012 $ \\
     VGG & $0.827 \pm 0.003$ & $0.770 \pm 0.016$  & $ 0.784 \pm 0.014$\\
@@
   \label{fig:vis-classifier}}
-  \caption{Projections by t-SNE of the test images at (a) the input layer, (b) the output of the convolutional layer of  ``FLIM + MLP'' after batch normalization (Figure~\ref{fig:arch}), and (c) the output of the last hidden layer of the MLP classifier in  ``FLIM + MLP''.}
+  \caption{Projections by t-SNE of the test images at (a) the input layer, (b) the output of the convolutional layer of  FLIM+MLP after batch normalization (Figure~\ref{fig:arch}), and (c) the output of the last hidden layer of the MLP classifier in  FLIM+MLP.}
 \end{figure}
@@
 As motivated by Rauber et al.
-\cite{rauber2016visualizing}, we created projections of three stages of the network ``FLIM + MLP" using t-SNE \cite{maaten2008visualizing} to understand how our CNN transforms image spaces along its layers.
-In Figure \ref{fig:vis-input}, we can see the projection of the test images in the LAB color space.
+\cite{rauber2016visualizing}, we created projections of three stages of the network FLIM+MLP using t-SNE \cite{maaten2008visualizing} to understand how our CNN transforms image spaces along its layers.
+In Figure \ref{fig:vis-input}, we can see the projection of the test images in the CIE L*a*b* color space.
 Green points are images that contain coconut trees, and red points are images that do not contain coconut trees.
@@
 The classes show more concentration in some parts of the projection and reduced overlapping between them.
-Finally, Figure \ref{fig:vis-classifier} shows the output of the last hidden layer of the MLP classifier in ``FLIM + MLP''.
+Finally, Figure \ref{fig:vis-classifier} shows the output of the last hidden layer of the MLP classifier in FLIM+MLP.
 The overlapping between classes and sample dispersion are considerably reduced, explaining our results.
-
-\begin{comment}
-  \begin{table}[!t]
-  \begin{center}
-  \begin{tabular}{|l|c|c|c|}
-  \hline
-   Method & Precision & Recall & F-score \\
-  \hline\hline
-    FLIM+MLP & $0.894 \pm 0.006$ & $0.888 \pm 0.009$ & $0.890 \pm 0.008$\\
-    FLIM-FT+MLP & $0.888 \pm 0.001$ & $0.878\pm 0.001$ & $0.881 \pm 0.001$ \\
-    VGG-FT & $0.884 \pm 0.003$ & $0.877 \pm 0.006$ & $0.879 \pm 0.005 $ \\
-  \hline
-  \end{tabular}
-  \end{center}
-  \caption{Mean and standard deviation of precision, recall, and f-score for each method: a feature extractor with a classifier with a larger train set.}
-    \label{tab:results-with-1200-images}
-  \end{table}
-\end{comment}
 
@@
 
+\begin{table}[!t]
+    \begin{center}
+        \begin{tabular}{|l|c|c|c|}
+            \hline
+            Method & Precision & Recall & F-score \\
+            \hline\hline
+            FLIM+SVM (4) & $0.856 \pm 0.011 $ & $ 0.831 \pm 0.019$ & $ 0.838 \pm 0.017$\\
+            FLIM+SVM (8) & $0.857 \pm 0.005$ & $0.838 \pm 0.010$ & $0.844 \pm 0.009$\\
+            \hline
+        \end{tabular}
+    \end{center}
+    \caption{Mean and standard deviation of precision, recall, and f-score for each method: a feature extractor trained with 8 images and a SVM classifier.}
+    \label{tab:results-more-markers}
+\end{table}
+
 \section{Conclusion}

### 2020-07-24 main.tex
results with more images with markers

@@
 \usepackage{siunitx}
+\usepackage{textcomp}
 \usepackage{hyperref}
@@
   \end{center}
-  \caption{Pipeline for training a feature extractor.}
+  \caption{\textcolor{red}{Pipeline for training a feature extractor.
+In the projection, green points are images with coconut trees, and red points are images with no coconut trees.
+Orange markers denote background regions, and cyan markers denote coconut tree regions.}}
   \label{fig:pipeline-xtractor}
@@
   \hline\hline
-    FLIM+SVM & $0.856 \pm 0.011 $ & $ 0.831 \pm 0.019$ & $ 0.838 \pm 0.017$\\ 
-    VGG-FE+SVM & $0.855 \pm 0.001$ & $0.816 \pm 0.007$ & $0.826 \pm 0.006 $ \\\hline
+    FLIM+SVM & 0.856 \textpm 0.011  &  0.831 \textpm 0.019 & 0.838 \textpm 0.017\\ 
+    VGG-FE+SVM & 0.855 \textpm 0.001 & 0.816 \textpm 0.007 & 0.826 \textpm 0.006  \\\hline
 
-    FLIM+MLP & $0.863 \pm 0.002$ & $0.849 \pm 0.005$ & $0.854 \pm 0.004$\\
-    FLIM-FT+MLP & $0.845 \pm 0.003$ & $0.825\pm 0.006$ & $0.832 \pm 0.005$ \\
-    CL+MLP & $0.822 \pm 0.011$ & $0.805 \pm 0.013$ & $0.811 \pm 0.012 $ \\
-    VGG & $0.827 \pm 0.003$ & $0.770 \pm 0.016$  & $ 0.784 \pm 0.014$\\
-    VGG-FT & $0.872 \pm 0.007$ & $0.844 \pm 0.015$ & $0.851 \pm 0.014 $ \\
+    FLIM+MLP & 0.863 \textpm 0.002 & \textbf{0.849 \textpm 0.005} & \textbf{0.854 \textpm 0.004}\\
+    FLIM-FT+MLP & 0.845 \textpm 0.003 & 0.825 \textpm 0.006 & 0.832 \textpm 0.005 \\
+    CL+MLP & 0.822 \textpm 0.011 & 0.805 \textpm 0.013 & 0.811 \textpm 0.012  \\
+    VGG & 0.827 \textpm 0.003 & 0.770 \textpm 0.016  &  0.784 \textpm 0.014\\
+    VGG-FT & \textbf{0.872 \textpm 0.007} & 0.844 \textpm 0.015 & 0.851 \textpm 0.014  \\
 
@@
             \hline\hline
-            FLIM+SVM (4) & $0.856 \pm 0.011 $ & $ 0.831 \pm 0.019$ & $ 0.838 \pm 0.017$\\
-            FLIM+SVM (8) & $0.857 \pm 0.005$ & $0.838 \pm 0.010$ & $0.844 \pm 0.009$\\
+            FLIM+SVM (4) & 0.856 \textpm 0.011 &  0.831 \textpm 0.019 &  0.838 \textpm 0.017\\
+            FLIM+SVM (8) & \textbf{0.857 \textpm 0.005} & 0.838 \textpm 0.010 & 0.844 \textpm 0.009\\
+            FLIM+SVM (12) & 0.854 \textpm 0.005 & \textbf{0.839 \textpm 0.009} & \textbf{0.844 \textpm 0.008}\\
             \hline

### 2020-07-24 main.tex
Updates from Overleaf

@@
 The third issue is related to human understanding.
-It may explore data visualization techniques to explain the model's decisions and to guide the user in the design of the model.
+It may explore visual analytics to explain the model's decisions and to guide the user in the design of the model~\cite{RauberInfoVis2018,GarciaIJCNN2019,SpinnerTVCG2020}.
 The fourth question is also essential during training, and it is related to user control over the process.
@@
 However, the importance of human participation during the training process has called little attention~\cite{AngelovNN2020}.
-Exceptions rely on visual analytics techniques~\cite{GarciaIJCNN2019,SpinnerTVCG2020}, and we present here a first work in which the user has direct influence over the quality of the filters.}
+Exceptions rely on visual analytics~\cite{RauberInfoVis2018,GarciaIJCNN2019,SpinnerTVCG2020}, and we present here a first work in which the user has direct influence over the quality of the filters.}
 
@@
 Each convolutional layer contains a filter bank, an activation function, and alternative operations (e.g., pooling, batch normalization).
-As the number of convolutional layers increases (deeper is the model), and higher is the number of annotated images required to train the model by backpropagation.
+As the number of convolutional layers increases (deeper is the model), higher is the number of annotated images required to train the model by backpropagation.
 
@@
 The user selects the number of convolutional layers, their filter sizes, \textcolor{red}{a few images from a training set}, and draws markers in image regions that best discriminate the classes.
-\textcolor{red}{By clustering patches extracted from marker pixels at the input of each convolutional layer, FLIM finds filters that enhance those regions at the output of the convolutional layers.} The CNN architecture could also be optimized based on visual analytics~\cite{SpinnerTVCG2020}.
+\textcolor{red}{By clustering patches extracted from marker pixels at the input of each convolutional layer, FLIM finds filters that enhance those regions at the output of the convolutional layers.
+The remaining training images can then be used to train the fully connected layers.} The CNN architecture could also be optimized based on visual analytics~\cite{SpinnerTVCG2020}, but we have not explored this option yet.
 
-We demonstrate the advantages of this interactive technique on the binary image classification of coconut-tree images against  VGG-16~\cite{simonyan2014very}.
+We demonstrate the advantages of this interactive technique on binary classification of coconut-tree images against  VGG-16~\cite{simonyan2014very}.
 First, the user can better understand and control the training process by observing the effectiveness of marker selection.
@@
 The method is named \emph{Feature Learning from Image Markers} (FLIM).
-In the feature space that results from convolutional layers,} it is common to assume that the classes are piecewise linearly separable and apply fully connected layers -- a Multi-Layer Perceptron (MLP) classifier -- for feature space reduction and classification.
+In the feature space that results from convolutional layers,} it is common to assume that the classes are piecewise linearly separable and apply fully connected layers -- Multi-Layer Perceptrons (MLP) -- for feature space reduction and classification.
 The training of a CNN usually relies on weight optimization using backpropagation, which requires a higher number of annotated images as deep as the model is.
-In FLIM, we eliminate the need for backpropagation to train convolutional layers by finding a set ${\cal F}$ of filters that can detect discriminant regions from all classes.
+In FLIM, we eliminate the need for backpropagation to train convolutional layers by finding a set ${\cal F}$ of filters that can \textcolor{red}{enhance} discriminant regions from all classes.
 
-The user selects a few training images to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes.
-The convolution between an image $I$ and a filter $F\in {\cal F}$ generates positive activations in regions whose local patterns are detected by $F$.
+The user selects a few \textcolor{red}{images from a training set} to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes.
+The convolution between an image $I$ and a filter $F\in {\cal F}$ \textcolor{red}{ enhances} regions whose local patterns are detected by $F$.
 We wish then to estimate the weights of $F$ such that those local patterns are characteristic of one given class.

### 2020-07-24 main.tex
Updates from Overleaf

@@
 We wish then to estimate the weights of $F$ such that those local patterns are characteristic of one given class.
-For sake of clarity, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect(P_I(p)) \cdot \vect(F)$, where $P_I(p)$ is a $k\times k$ patch with $m$ channels around $p$, $F$ is a filter with shape $k\times k \times m$, and $\vect$ is the vectorization operation.
+For sake of clarity, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect(P_I(p)) \cdot \vect(F)$, where $P_I(p)$ is a $k\times k$ patch with $m$ \textcolor{red}{bands} around $p$, $F$ is a filter with shape $k\times k \times m$, and $\vect$ is the vectorization operation.
 A filter $F$ can discriminate a class $i \in \{1,2,\ldots,c\}$ among $c$ classes when it generates positive results for local patterns $\vect(P_I(p))$ of class $i$ and negative results for patterns from other classes.
@@
 Figure~\ref{fig:filter} illustrates three groups with color-coded samples from two classes in a hypothetical 2D feature space.
-The marker-based batch normalization and convolution with filters obtained as centroids of the three groups in Figure~\ref{fig:filter}a creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters in distinct activation bands.
+The marker-based batch normalization and convolution with filters obtained as centroids of the three groups in Figure~\ref{fig:filter}a creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters in distinct activation \textcolor{red}{bands}.
 
@@
 The dataset consists of 13587 regions with $90 \times 90$ pixels \textcolor{red}{with spatial resolution of $8$\si{\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
-\textcolor{red}{The region images contain three channels: red, green, and blue.
-We convert them to the CIE L*a*b* color space and normalize each channel within $[0,1]$}.
+\textcolor{red}{The region images contain three bands: red, green, and blue.
+We convert them to the CIE L*a*b* color space and normalize each band within $[0,1]$}.
 
@@
 The validation set was used to choose empirically the architecture of the network shown in Figure~\ref{fig:arch}.
-The convolutional layer has $60$ filters with dimension $7 \times 7$ and 3 color channels.
+The convolutional layer has $60$ filters with dimension $7 \times 7$ and 3 color \textcolor{red}{bands}.
 It includes marker-based normalization before convolution, ReLu activation, and a max-pooling operation with a window of dimension $3 \times 3$.
@@
 
-\section*{Acknowledgements}
+\section*{Acknowledgments}
 

### 2020-07-24 main.tex
small revision

@@
 The user selects the number of convolutional layers, their filter sizes, \textcolor{red}{a few images from a training set}, and draws markers in image regions that best discriminate the classes.
-\textcolor{red}{By clustering patches extracted from marker pixels at the input of each convolutional layer, FLIM finds filters that enhance those regions at the output of the convolutional layers.
+\textcolor{red}{By clustering patches extracted from marker pixels at the input of each convolutional layer, FLIM finds filters that enhance those regions at the convolutional layers' output.
 The remaining training images can then be used to train the fully connected layers.} The CNN architecture could also be optimized based on visual analytics~\cite{SpinnerTVCG2020}, but we have not explored this option yet.
@@
 \textcolor{red}{A CNN model should enhance image regions that best discriminate classes through a sequence of convolutional layers.
-This section shows how the identification of such regions by a simple user action can make that sequence an effective feature extractor.
+This section shows how identifying such regions by a simple user action can make that sequence an effective feature extractor.
 The method is named \emph{Feature Learning from Image Markers} (FLIM).
@@
 
-\textcolor{red}{Pytorch was chosen to implement FLIM and MLP, being SVM the linear model available in sklearn with $C=0.01$~\cite{fan2008liblinear, scikit-learn}.
+\textcolor{red}{Pytorch was chosen to implement FLIM and MLP, being SVM the linear model available in scikit-learn with $C=0.01$~\cite{fan2008liblinear, scikit-learn}.
 We trained MLP for $40$ epochs and the whole network for $60$ epochs with a batch size of $64$.
@@
 Finally, Table II shows the results of FLIM+SVM as the number of images for marker selection increases, and the training set size remains with 200 images.
-We can notice a slight improvement in all measures, but this is a point that deserves further investigation -- i.e., choosing relevant training images and markers for FLIM.
-}
+We can notice a slight improvement in all measures, but this is a point that deserves further investigation -- i.e., choosing relevant training images and markers for FLIM.}
 

### 2020-07-24 main.tex
Updates from Overleaf

--- 2020-07-24 main.tex
@@
 
-     \node (put) [steps, right=1em of simgs] {Select\\Markers};
+     \node (put) [steps, right=1em of simgs] {Draw\\Markers};
 
@@
   \end{center}
-  \caption{\textcolor{red}{Pipeline for training a feature extractor.
+  \caption{\textcolor{red}{Pipeline for training a feature extractor using FLIM.
 In the projection, green points are images with coconut trees, and red points are images with no coconut trees.
-Orange markers denote background regions, and cyan markers denote coconut tree regions.}}
+The user selects a few images (e.g., 4) from the projection.
+The user draws orange markers on background pixels and cyan markers on coconut-tree pixels for image feature learning by FLIM.}}
   \label{fig:pipeline-xtractor}
@@
 The Humanitarian OpenStreetMap community annotated the regions.
-The dataset consists of 13587 regions with $90 \times 90$ pixels \textcolor{red}{with spatial resolution of $8$\si{\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
+The dataset consists of 13587 regions with $90 \times 90$ pixels \textcolor{red}{and spatial resolution of $8$\si{\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
 \textcolor{red}{The region images contain three bands: red, green, and blue.
@@
  \begin{itemize}
-\item FLIM+SVM: Image classification is performed by support vector machines (SVM) on FLIM-based features.  
-\item FLIM+MLP: Image classification is performed by multi-layer perceptrons (MLP) on FLIM-based features.  
-\item FLIM-FT+MLP: The network in FLIM+MLP is fine-tuned by backpropagation. 
-\item CL+MLP: The same architecture used in FLIM+MLP is trained from scratch by backpropagation.
-\item VGG: Vgg-16~\cite{simonyan2014very} is trained from scratch.
-\item VGG-FT: Vgg-16 is pre-trained on ImageNet and fine-tuned on the coconut-tree training set.
+\item FLIM+SVM: Image classification is performed by support vector machines (SVM) from FLIM-based features.  
+\item FLIM+MLP: Image classification is performed by multi-layer perceptrons (MLP) from FLIM-based features.  
+\item FLIM-FT+MLP: The architecture in FLIM+MLP is fine-tuned by backpropagation. 
+\item CL+MLP: The architecture in FLIM+MLP is trained from scratch by backpropagation.
+\item VGG: VGG-16~\cite{simonyan2014very} is trained from scratch.
+\item VGG-FT: VGG-16 is pre-trained on ImageNet and fine-tuned on the coconut-tree training set.
 \item VGG-FE+SVM: Image features are obtained from the last convolutional layer of VGG-FT.
@@
 The comparison between FLIM+SVM and VGG-FE+SVM indicates that FLIM can produce a feature extractor with a single layer competitive to VGG-FE using 13 layers (i.e., FLIM uses only $0.07\%$ of the number of parameters in VGG-FE).
-The results of FLIM+MLP, CL+MLP, and FLIM-FT+MLP show that FLIM may dismiss training and fine-tuning based on backpropagation, by directly providing a more effective solution than CL+MLP and FLIM-FT+MLP.
+The results of FLIM+MLP, CL+MLP, and FLIM-FT+MLP show that FLIM may dismiss training and fine-tuning based on backpropagation, by directly providing a more effective solution than CL+MLP and FLIM-FT+MLP (H2).
 This result can also be noticed when comparing  FLIM+MLP with VGG, confirming that FLIM can reduce the number of required training images for the MLP classifier.
-The comparison between FLIM+MLP and VGG-FT shows competitive results with a top-rated solution.}
+The comparison between FLIM+MLP and VGG-FT shows competitive results with a top-rated solution (H1).}
 
@@
 Note that the marker selection for FLIM was still the same from those four images.
-To our surprise, as the training set increases, the differences in f-score favor FLIM+MLP over FLIM-FT+MLP, VGG-FT, and CL+MLP.
-The difference between FLIM+MLP and CL+MLP also demonstrated that FLIM was superior to backpropagation for filter estimation.
-Finally, Table II shows the results of FLIM+SVM as the number of images for marker selection increases, and the training set size remains with 200 images.
-We can notice a slight improvement in all measures, but this is a point that deserves further investigation -- i.e., choosing relevant training images and markers for FLIM.
+As the training set increases, the mean differences in f-score, as obtained among the three splits, favor FLIM+MLP over FLIM-FT+MLP (H2), VGG-FT (H1 and H3), and CL+MLP (H2).
+The differences between FLIM+MLP and CL+MLP demonstrate that FLIM is superior to backpropagation in filter estimation (H2).
+Finally, Table II shows the mean results of FLIM+SVM among the three splits as the number of images for marker selection increases, and the training set size remains with 200 images (H4).
+The results are not conclusive for H4, but the choice of more markers does not deteriorate the performance of FLIM.
+This point deserves further investigation -- i.e., choosing relevant training images and markers for FLIM.
 }
@@
     }
-    \caption{\textcolor{red}{Markers used for training, cyan markers on coconut trees and orange markers on other regions.}}
+    \caption{\textcolor{red}{Markers used for training, cyan markers on coconut-tree pixels and orange markers on background pixels.}}
     \label{fig:markers}
@@
 
-    FLIM+MLP & 0.863 \textpm 0.002 & \textbf{0.849 \textpm 0.005} & \textbf{0.854 \textpm 0.004}\\
+    FLIM+MLP & \textbf{0.863 \textpm 0.002} & \textbf{0.849 \textpm 0.005} & \textbf{0.854 \textpm 0.004}\\
     FLIM-FT+MLP & 0.845 \textpm 0.003 & 0.825 \textpm 0.006 & 0.832 \textpm 0.005 \\
@@
     VGG & 0.827 \textpm 0.003 & 0.770 \textpm 0.016  &  0.784 \textpm 0.014\\
-    VGG-FT & \textbf{0.872 \textpm 0.007} & 0.844 \textpm 0.015 & 0.851 \textpm 0.014  \\
+    VGG-FT & \textbf{0.872 \textpm 0.007} & \textbf{0.844 \textpm 0.015} & \textbf{0.851 \textpm 0.014}  \\
 
@@
 \cite{rauber2016visualizing}, we created projections of three stages of the network FLIM+MLP using t-SNE \cite{maaten2008visualizing} to understand how our CNN transforms image spaces along its layers.
-In Figure \ref{fig:vis-input}, we can see the projection of the test images in the CIE L*a*b* color space.
-Green points are images that contain coconut trees, and red points are images that do not contain coconut trees.
+In Figure \ref{fig:vis-input}, we can see the projection of the test images in the \textcolor{red}{CIE L*a*b* color space.} Green points are images that contain coconut trees, and red points are images that do not contain coconut trees.
 The projection shows considerable overlapping between the two classes.
@@
             FLIM+SVM (4) & 0.856 \textpm 0.011 &  0.831 \textpm 0.019 &  0.838 \textpm 0.017\\
-            FLIM+SVM (8) & \textbf{0.857 \textpm 0.005} & 0.838 \textpm 0.010 & 0.844 \textpm 0.009\\
-            FLIM+SVM (12) & 0.854 \textpm 0.005 & \textbf{0.839 \textpm 0.009} & \textbf{0.844 \textpm 0.008}\\
+            FLIM+SVM (8) & 0.857 \textpm 0.005 & 0.838 \textpm 0.010 & 0.844 \textpm 0.009\\
+            FLIM+SVM (12) & 0.854 \textpm 0.005 & 0.839 \textpm 0.009 & 0.844 \textpm 0.008\\
             \hline
@@
 
-We introduced a first feature learning technique, named FLIM, that can estimate effective filter weights for the convolutional layers of a given network architecture from user-drawn markers in a few training images.
-The resulting feature extractor can be used with different classifiers and, when it is used with a MLP classifier, the number of required images to train the fully connected layers is reduced.
+We introduced a first feature learning technique, named FLIM, that can estimate effective filter weights for the convolutional layers of a given network architecture from user-drawn markers \textcolor{red}{in a few images of a training set.} The resulting feature extractor can be used with different classifiers and, when it is used with a MLP classifier, the number of required images to train the fully connected layers is reduced.
 We demonstrated the advantages of FLIM over solutions based on VGG-16 for the binary classification of coconut-tree images.

### 2020-07-24 main.tex

@@
     \item VGG-FE+SVM: Image features are obtained from the last convolutional layer of VGG-FT.
+
 \end{itemize}

### 2020-07-24 main.tex
:(
Updates from Overleaf

@@
 \textcolor{red}{By clustering patches extracted from marker pixels at the input of each convolutional layer, FLIM finds filters that enhance those regions at the convolutional layers' output.
-The remaining training images can then be used to train the fully connected layers.} The CNN architecture could also be optimized based on visual analytics~\cite{SpinnerTVCG2020}, but we have not explored this option yet.
+The remaining training images can then be used to train the fully connected layers or another classifier.} The CNN architecture could also be optimized based on visual analytics~\cite{SpinnerTVCG2020}, but we have not explored this option yet.
+\textcolor{red}{In interactive semantic segmentation, user-selected clicks can affect part of the network to improve segmentation, but the filters are not designed from those markers~\cite{SofiiukCVPR2020}.}
 
@@
 \item VGG-FE+SVM: Image features are obtained from the last convolutional layer of VGG-FT.
-
 \end{itemize}
@@
 Note that the marker selection for FLIM was still the same from those four images.
-As the training set increases, the mean differences in f-score, as obtained among the three splits, favor FLIM+MLP over FLIM-FT+MLP (H2), VGG-FT (H1 and H3), and CL+MLP (H2).
+To our surprise, as the training set increases, the mean differences in f-score, as obtained among the three splits, favor FLIM+MLP over FLIM-FT+MLP (H2), VGG-FT (H1 and H3), and CL+MLP (H2).
 The differences between FLIM+MLP and CL+MLP demonstrate that FLIM is superior to backpropagation in filter estimation (H2).
@@
 
-We intend to further investigate feature learning from image markers for different applications, elaborate methodologies to optimize network architecture, and extend FLIM to estimate weights in fully connected layers.
+We intend to further investigate feature learning from image markers for different applications, \textcolor{red}{image and marker selection for FLIM,} elaborate methodologies to optimize network architecture, and extend FLIM to estimate weights in fully connected layers.
 

### 2020-07-24 IEEE-GRSL-Response-Letter.tex
Updates from Overleaf

@@
 Thanks for the comments.
-We have included more experiments to demonstrate the robustness of the proposed feature learning approach.
-The parts marked and commented by reviewer 1 in the manuscript are reproduced and addressed as follows.
+We have improved the experimental section with new experiments, demonstrating the robustness of the proposed feature learning approach.
+We have also revised the text, added a figure, and clarified some parts to improve readability.
+All changes are in red.
 
@@
 
-\ans  Thanks. The sentence was long and unclear. We have simplified it and improved it in first paragraph of Section II.
+\ans  Thanks. The sentence was long and unclear. We have simplified and improved it in the first paragraph of Section II.
 \eans
@@
 
-\ans Thanks. These are minor corrections in different parts of the text that have been solved now. (Section III). \eans
+\ans Thanks. These are minor corrections in different parts of Section III that have been solved now. \eans
 
@@
 
-\ans \eans
+\ans 
+This sentence was removed.
+The dataset was randomly divided into training, validation, and testing sets.
+The validation set was used only once to choose empirically the architecture of the CNN through some preliminary tests.
+The mean results with standard deviation were computed based on those three random splits, using their respective training and testing sets.
+We hope this is clear in Section III now.
+\eans
 
@@
 
-\ans \eans
+\ans We have changed the way to explain the choice of representative images from each class to draw markers. The new Figure 3 also helps in that explanation. \eans
 
@@
 
-\ans \eans
+\ans Thanks a lot for the suggestion related to this sentence. The sentence has been removed and we have added ex\eans
 

### 2020-07-24 main.tex
Updates from Overleaf
update model figure

@@
   \begin{center}
-  \tikzset{
-    treenode/.style = {shape=rectangle, rounded corners,
-                       draw=black, align=center, line width=width, minimum width=2cm},
-    specialedge/.style = {line width=1.5pt}
-  }
+
+    \tikzstyle{steps} = [rectangle, rounded corners, minimum width=1.3cm, minimum height=0.5cm,text centered, draw=dodgerBlue, very thick, fill=dodgerBlue!25, execute at begin node={\begin{varwidth}{2.5cm}},
+      execute at end node={\end{varwidth}}]
+
+      \tikzstyle{myarrow} = [navy, very thick]
+
+      \tikzstyle{arrow} = [thick,->,>=stealth]
 
   \begin{tikzpicture}
-    [
-      grow                    = right,
-      sibling distance        = 5em,
-      level distance          = 30mm,
-      edge from parent/.style = {draw=black, -latex},
-      every node/.style       = {treenode},
-      sloped
-    ]
-    \node [treenode, fill=tealDeer] {MarkerBased \\ BatchNorm}
-    child {
-        node [treenode, fill=tealDeer] {Conv$(60 \times 7 \times 7)$}
-        child {
-          node [treenode, fill=tealDeer] {ReLU} 
-          child {
-            node [treenode, fill=tealDeer] {MaxPool$(3\times3)$}
-            child {
-              node [treenode, fill=tealDeer] {BatchNorm}
-              child {
-                node [treenode, fill=navyPurple] {Classifier}
-                edge from parent
-              }
-            }
-          } 
-        }
-      };
+
+    \node (mbn) [steps] {MarkerBased \\ BatchNorm};
+
+    \node (conv) [steps, right= of mbn] {Conv$(60 \times 7 \times 7)$};
+
+    \node (relu) [steps, right= of conv] {ReLU};
+
+    \node (pool) [steps, right=of relu] {MaxPool$(3\times3)$};
+
+    \node (bn) [steps, right=of pool] {BatchNorm};
+
+    \node (clf) [steps, fill=navyPurple!25, draw=navyPurple, right= of bn] {Classifier};
+
+    %-----------------------arrows---------------
+
+    \draw [arrow, myarrow] (mbn) -- (conv);
+    \draw [arrow, myarrow] (conv) -- (relu);
+    \draw [arrow, myarrow] (relu) -- (pool);
+    \draw [arrow, myarrow] (pool) -- (bn);
+    \draw [arrow, myarrow] (bn) -- (clf);
 
@@
 
-    \tikzstyle{steps} = [rectangle, rounded corners, minimum width=1.3cm, minimum height=0.5cm,text centered, draw=pink, thick, fill=heliotrope!30, execute at begin node={\begin{varwidth}{2.5cm}},
+    \tikzstyle{steps} = [rectangle, rounded corners, minimum width=1.3cm, minimum height=0.5cm,text centered, draw=pink, very thick, fill=heliotrope!30, execute at begin node={\begin{varwidth}{2.5cm}},
     execute at end node={\end{varwidth}}]
@@
 
-    \tikzstyle{myarrow} = [navy, thick]
+    \tikzstyle{myarrow} = [navy, very thick]
 
@@
 \textcolor{red}{First, we evaluated H1 and H2 by randomly selecting 200 images for the training set, 1000 images for a validation set, and the remaining 12387 images were left in the testing set.
-The validation set was used to choose empirically the architecture of the network shown in Figure~\ref{fig:arch}.
+The validation set was used only once to choose empirically the architecture of the network shown in Figure~\ref{fig:arch}.
 The convolutional layer has $60$ filters with dimension $7 \times 7$ and 3 color \textcolor{red}{bands}.

### 2020-07-24 main.tex
Updates from Overleaf

--- 2020-07-24 main.tex
@@
 The method is named \emph{Feature Learning from Image Markers} (FLIM).
-In the feature space that results from convolutional layers,} it is common to assume that the classes are piecewise linearly separable and apply fully connected layers -- Multi-Layer Perceptrons (MLP) -- for feature space reduction and classification.
+In the feature space that results from convolutional layers,} it is common to assume that the classes are piecewise linearly separable and apply fully connected layers -- Multi-Layer Perceptron (MLP) -- for feature space reduction and classification.
 The training of a CNN usually relies on weight optimization using backpropagation, which requires a higher number of annotated images as deep as the model is.
@@
 \item FLIM+SVM: Image classification is performed by support vector machines (SVM) from FLIM-based features.  
-\item FLIM+MLP: Image classification is performed by multi-layer perceptrons (MLP) from FLIM-based features.  
+\item FLIM+MLP: Image classification is performed by multi-layer perceptron (MLP) from FLIM-based features.  
 \item FLIM-FT+MLP: The architecture in FLIM+MLP is fine-tuned by backpropagation. 

### 2020-07-24 main.tex
change perceptron to singular

@@
 The method is named \emph{Feature Learning from Image Markers} (FLIM).
-In the feature space that results from convolutional layers,} it is common to assume that the classes are piecewise linearly separable and apply fully connected layers -- Multi-Layer Perceptron (MLP) -- for feature space reduction and classification.
+In the feature space that results from convolutional layers,} it is common to assume that the classes are piecewise linearly separable and apply fully connected layers -- Multi-Layer \textcolor{blue}{Singular} (MLP) -- for feature space reduction and classification.
 The training of a CNN usually relies on weight optimization using backpropagation, which requires a higher number of annotated images as deep as the model is.
@@
 \item FLIM+SVM: Image classification is performed by support vector machines (SVM) from FLIM-based features.  
-\item FLIM+MLP: Image classification is performed by multi-layer perceptron (MLP) from FLIM-based features.  
+\item FLIM+MLP: Image classification is performed by multi-layer \textcolor{blue}{singular} (MLP) from FLIM-based features.  
 \item FLIM-FT+MLP: The architecture in FLIM+MLP is fine-tuned by backpropagation. 

### 2020-07-24 IEEE-GRSL-Response-Letter.tex

@@
 
-\ans Thanks a lot for the suggestion related to this sentence. The sentence has been removed and we have added ex\eans
+\ans Thanks a lot for the suggestion related to this sentence.
+The sentence has been removed, and we have added experiments with an increasing number of images for marker selection.
+The results, shown in Table II, demonstrate that FLIM is not negatively affected by choice of marker pixels in a higher number of images.
+On the other hand, we consider these results inconclusive.
+We believe it is possible to assist the user in visually identifying relevant images and regions in those images for marker selection.
+However, this point requires further investigation and suitable image processing with data visualization.
+\eans
 
@@
 
-\ans \eans
+\ans MLP stands for multi-layer perceptrons \eans
 

### 2020-07-24 IEEE-GRSL-Response-Letter.tex
Updates from Overleaf

@@
 
-\ans MLP stands for multi-layer perceptrons \eans
+\ans MLP stands for multi-layer perceptron and it was defined in the first paragraph of Section II.
+We have redefined it again in Section III.
+We have also included the specifications of the machine used for the experiments in the beginning of the second paragraph in Section III.
+\eans
 

### 2020-07-24 IEEE-GRSL-Response-Letter.tex
Updates from Overleaf

--- 2020-07-24 IEEE-GRSL-Response-Letter.tex
@@
 
-\ans MLP stands for multi-layer perceptrons \eans
+\ans MLP stands for a multi-layer perceptron, and it was defined in the first paragraph of Section II.
+We have redefined it again in Section III.
+We have also added the machine used for the experiments at the beginning of the second paragraph in Section III.
+\eans
 
@@
 
-\ans \eans
+\ans Thanks, we have fixed that indicating which marker color is used to indicate each class. 
+\eans
 
@@
 
-\ans \eans
+\ans Thanks, we have removed `` '' around the names of the methods.\eans
 
@@
 
-\ans \eans
+\ans Thanks for the comments. We have improved the experiments in Section III and the presentation of the work in several parts of the text. 
+\eans
 
@@
 
-\ans \eans
+\ans Thanks. We have fixed them and revised the entire text to avoid typos. \eans
 
@@
 
-\ans \eans
+\ans Thanks to have pointed it out\eans
 

### 2020-07-24 main.tex
Updates from Overleaf

@@
 
-\textcolor{red}{Pytorch was chosen to implement FLIM and MLP, being SVM the linear model available in scikit-learn with $C=0.01$~\cite{fan2008liblinear, scikit-learn}.
+\textcolor{red}{Pytorch was chosen to implement FLIM and MLP, being SVM the linear model available in scikit-learn with~${C=0.01}$~\cite{fan2008liblinear, scikit-learn}.
 We trained MLP for $40$ epochs and the whole network for $60$ epochs with a batch size of $64$.

### 2020-07-24 IEEE-GRSL-Response-Letter.tex
Updates from Overleaf

@@
 
-\ans Thanks to have pointed it out\eans
+\ans Thanks to have pointed it out.
+We are now explaining that the dataset is randomly divided into training (200 images), validation (1000 images), and testing (remaining images) sets.
+We repeated each experiment three times, each using one of these random splits.
+However, the validation set was used only once to choose the hyperparameters of the network.
+We have not developed/used any optimization technique for this purpose.
+We have just executed tests with a few hyperparameter choices in order to select one for the experiments.
+The idea was to select the simplest and yet effective network in order to demonstrate that FLIM can create competitive solutions.
+We are clarifying this issue (Section III) and indicating it for future work (Section IV).
+\eans
 
@@
 
-\ans \eans
+\ans
+
+For three splits only and 200 training images, it was not possible to claim that this difference is statistically significant.
+However, they are undoubtedly competitive, being FLIM+MLP a much simpler (lighter) network than VGG-FT.
+Even with a new experiment in which we increase the training set up to 1200 images, the difference increases in favor of FLIM+MLP, but they are still not statistically significant (except for one case with 1100 images).
+On the other hand, the f-score curve of FLIM+MLP was consistently above one of VGG-FT.
+We have improved the experimental section to clarify this and several other issues.
+\eans
 
@@
 
-\ans \eans
+\ans Thanks a lot for raising these questions.
+We noticed that it was not clear that markers are drawn on a few images selected from the training set (we used only four images).
+The user takes a couple of minutes to select images and draw markers in those images.
+This is much less than the time taken to examine and label 200 images.
+We also added an experiment with increasing training set sizes.
+VGG-FT could never beat FLIM+MLP.
+The reason is that the image features of FLIM are better than those of VGG-FE (the one extracted from the last convolutional layer of VGG-FT), and both use a similar MLP (the only difference is its input layer).
+So, higher is the number of training samples, MLP can improve the f-score of both, being the one of FLIM+MLP always above the one o
+
+\eans
 

### 2020-07-24 main.tex
Updates from Overleaf

--- 2020-07-24 main.tex
@@
 The user selects the number of convolutional layers, their filter sizes, \textcolor{red}{a few images from a training set}, and draws markers in image regions that best discriminate the classes.
+The user takes a couple of minutes to select images and draw markers in those images.
 \textcolor{red}{By clustering patches extracted from marker pixels at the input of each convolutional layer, FLIM finds filters that enhance those regions at the convolutional layers' output.
@@
 \item FLIM+SVM: Image classification is performed by support vector machines (SVM) from FLIM-based features.  
-\item FLIM+MLP: Image classification is performed by multi-layer perceptron (MLP) from FLIM-based features.  
+\item FLIM+MLP: Image classification is performed by \textcolor{red}{Multi-Layer Perceptron} (MLP) from FLIM-based features.  
 \item FLIM-FT+MLP: The architecture in FLIM+MLP is fine-tuned by backpropagation. 
@@
 Note that the marker selection for FLIM was still the same from those four images.
-To our surprise, as the training set increases, the mean differences in f-score, as obtained among the three splits, favor FLIM+MLP over FLIM-FT+MLP (H2), VGG-FT (H1 and H3), and CL+MLP (H2).
+To our surprise, as the training set increases, the differences in mean f-score, as obtained among the three splits, favor FLIM+MLP over FLIM-FT+MLP (H2), VGG-FT (H1 and H3), and CL+MLP (H2).
 The differences between FLIM+MLP and CL+MLP demonstrate that FLIM is superior to backpropagation in filter estimation (H2).

### 2020-07-24 main.tex
indicação de tamanho na figura 3
indicação de quantidade de imagens na figura 3

--- 2020-07-24 main.tex
@@
 The user selects the number of convolutional layers, their filter sizes, \textcolor{red}{a few images from a training set}, and draws markers in image regions that best discriminate the classes.
+\textcolor{blue}{The user takes a couple of minutes to select images and draw markers in those images}.
 \textcolor{red}{By clustering patches extracted from marker pixels at the input of each convolutional layer, FLIM finds filters that enhance those regions at the convolutional layers' output.
@@
      %images----------------------------
-    \node (mimg1) [image, above right=-2em and 2.8em of put] {\includegraphics[width=.08\textwidth]{figures/markers1.png}};
+    \node (mimg1) [image, above right=-2.5em and 2.8em of put] {\includegraphics[width=.08\textwidth]{figures/markers1.png}};
     \node (mimg2) [image, below left=-5em of mimg1] {\includegraphics[width=.08\textwidth]{figures/markers2.png}};
@@
 
-     %\draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-%(img3.west) -- (img1.north) node [black,midway,xshift=0.8cm, label = {[label distance = 0.1em]above:{Images}}] {};
+     \draw [decorate,decoration={brace,amplitude=4pt,raise=5pt}]
+(img3.north west) -- (img1.north west) node [black,midway,xshift=-0.4cm, yshift=0.4cm, rotate=45] {\small 200};
+
+\draw [decorate,decoration={brace,amplitude=4pt,raise=5pt}]
+(simg4.north west) -- (simg1.north west) node [black,midway,xshift=-0.4cm, yshift=0.4cm, rotate=45] {\small 4};
+
+\draw [decorate,decoration={brace,amplitude=4pt,raise=5pt}]
+(mimg4.north west) -- (mimg1.north west) node [black,midway,xshift=-0.4cm, yshift=0.4cm, rotate=45] {\small 4};
 

### 2020-07-24 IEEE-GRSL-Response-Letter.tex

--- 2020-07-24 IEEE-GRSL-Response-Letter.tex
@@
 
-\ans \eans
+\ans
+
+For three splits only and 200 training images, it was not possible to claim that this difference is statistically significant.
+However, they are undoubtedly competitive, being FLIM+MLP a much simpler (lighter) network than VGG-FT.
+\eans
 
@@
 
-\ans \eans
+\ans Thanks a lot for raising both questions.
+We noticed that it was not clear that markers are drawn on a few images selected from the training set (we used only 4 images).
+The user takes a couple of minutes to select images and draw markers on those images.
+This time is much less than the time taken to examine and label 200 images.
+We added a sentence to clarify that in Section I.
+We also added Figure 3 to explain the user-interaction process in FLIM and modified a few parts in the text to clarify that FLIM is trained from markers in 4 images only.
+
+We added an experiment with increasing training set sizes.
+VGG-FT could never beat FLIM+MLP, even with a higher number of training images to improve its features.
+The reason is that 1200 images were not enough to improve VGG-FE (the feature extractor obtained at the last convolutional layer of VGG-FT).
+Both FLIM+MLP and VGG-FT use a similar MLP (the only difference is in the input layer due to their different number of features).
+Therefore, the higher number of training images benefit both approaches, being the mean f-score curve of FLIM+MLP consistently above the one of VGG-FT.
+
+\eans
 
@@
 \ans
-
+Thanks, we have improved the literature review in Section I.
 \eans

### 2020-07-25 main.tex
reviewing

@@
 
-\title{Learning CNN filters from user-drawn image markers for coconut-tree image classification}
+\title{Learning CNN filters from user-drawn image \textcolor{red}{markers} for coconut-tree image classification}
 
@@
 \section{Introduction}
+
 \IEEEPARstart{D}{eep} learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}.
@@
 They all lead to the importance of involving human experts during the machine learning process.
-\textcolor{red}{Explainable artificial intelligence (XAI) has appeared to address such problems~\cite{rauber2016visualizing,arrietaInfoFusion2020}.
+\textcolor{red}{Explainable artificial intelligence (XAI) has appeared to address \textcolor{blue}{(3)}~\cite{rauber2016visualizing,arrietaInfoFusion2020}.
 However, the importance of human participation during the training process has called little attention~\cite{AngelovNN2020}.

### 2020-07-27 IEEE-GRSL-Response-Letter.tex
final version

@@
 
-\ans Thanks. These are minor corrections in different parts of Section III that have been solved now. \eans
+\ans Thanks.
+These are minor corrections in different parts of Section III that have been solved now.
+We refer to the following link to a good explanation of LAB color space \url{https://en.wikipedia.org/wiki/CIELAB_color_space}.
+\eans
 
@@
 
-\rwr{1} we identified cohesive groups of images from each class and selected two representative images from each class to draw markers.}
+\rwr{1} we identified cohesive groups of images from each class and selected two representative images from each class to draw markers.
 
@@
 \ans Thanks to have pointed it out.
-We are now explaining that the dataset is randomly divided into training (200 images), validation (1000 images), and testing (remaining images) sets.
+We are now explaining that the dataset is randomly divided into training (200 images), validation (2000 images), and testing (remaining images) sets.
 We repeated each experiment three times, each using one of these random splits.

### 2020-07-27 main.tex
final version

--- 2020-07-24 main.tex
@@
 
+%\documentclass[journal,twoside,peerreview]{IEEEtran}
+
 \def\citepunt{,}
@@
 \DeclareMathOperator{\vect}{vec}
+
+\DeclareRobustCommand\changed{\textcolor{black}}
 
@@
 
+\ifCLASSOPTIONpeerreview
+  \usepackage{setspace}
+  \setstretch{1.7}
+\fi
+
 \begin{document}
 
-\title{Learning CNN filters from user-drawn image markers for coconut-tree image classification}
+\title{Learning CNN filters from user-drawn image \changed{markers} for coconut-tree image classification}
 
@@
 \section{Introduction}
-\IEEEPARstart{D}{eep} learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}.
+
+\IEEEPARstart{\changed{D}}{\changed{eep}} learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}.
 In remote sensing, the applications may involve segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}, for instance.
@@
 They all lead to the importance of involving human experts during the machine learning process.
-\textcolor{red}{Explainable artificial intelligence (XAI) has appeared to address such problems~\cite{rauber2016visualizing,arrietaInfoFusion2020}.
+\changed{Explainable artificial intelligence (XAI) has appeared to address (3)~\cite{rauber2016visualizing,arrietaInfoFusion2020}.
 However, the importance of human participation during the training process has called little attention~\cite{AngelovNN2020}.
@@
 
-To reduce the need for large annotated training sets by exploiting the user knowledge, we present a method, called \textcolor{red}{\emph{Feature Learning from Image Markers} (FLIM)}, that needs a minimal set of images to learn the filters of each convolutional layer.
-The user selects the number of convolutional layers, their filter sizes, \textcolor{red}{a few images from a training set}, and draws markers in image regions that best discriminate the classes.
+To reduce the need for large annotated training sets by exploiting the user knowledge, we present a method, called \changed{\emph{Feature Learning from Image Markers} (FLIM)}, that needs a minimal set of images to learn the filters of each convolutional layer.
+The user selects the number of convolutional layers, their filter sizes, \changed{a few images from a training set}, and draws markers in image regions that best discriminate the classes.
 The user takes a couple of minutes to select images and draw markers in those images.
-\textcolor{red}{By clustering patches extracted from marker pixels at the input of each convolutional layer, FLIM finds filters that enhance those regions at the convolutional layers' output.
-The remaining training images can then be used to train the fully connected layers or another classifier.} The CNN architecture could also be optimized based on visual analytics~\cite{SpinnerTVCG2020}, but we have not explored this option yet.
-\textcolor{red}{In interactive semantic segmentation, user-selected clicks can affect part of the network to improve segmentation, but the filters are not designed from those markers~\cite{SofiiukCVPR2020}.}
+\changed{By clustering patches extracted from marker pixels at the input of each convolutional layer, FLIM finds filters that enhance those regions at the convolutional layers' output.
+The remaining training images can then be used to train the fully connected layers or another classifier.} The CNN architecture could also be optimized \changed{with the support of} visual analytics~\cite{SpinnerTVCG2020}, but we have not explored this option yet.
+\changed{In interactive semantic segmentation, user-selected clicks can affect a part of the network to improve segmentation, but the filters are not designed from those markers~\cite{SofiiukCVPR2020}.}
 
@@
 
-\textcolor{red}{A CNN model should enhance image regions that best discriminate classes through a sequence of convolutional layers.
+\changed{A CNN model should enhance image regions that best discriminate classes through a sequence of convolutional layers.
 This section shows how identifying such regions by a simple user action can make that sequence an effective feature extractor.
@@
 The training of a CNN usually relies on weight optimization using backpropagation, which requires a higher number of annotated images as deep as the model is.
-In FLIM, we eliminate the need for backpropagation to train convolutional layers by finding a set ${\cal F}$ of filters that can \textcolor{red}{enhance} discriminant regions from all classes.
+In FLIM, we eliminate the need for backpropagation to train convolutional layers by finding a set ${\cal F}$ of filters that can \changed{enhance} discriminant regions from all classes.
 
-The user selects a few \textcolor{red}{images from a training set} to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes.
-The convolution between an image $I$ and a filter $F\in {\cal F}$ \textcolor{red}{ enhances} regions whose local patterns are detected by $F$.
+The user selects a few \changed{images from a training set} to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes.
+The convolution between an image $I$ and a filter $F\in {\cal F}$ \changed{ enhances} regions whose local patterns are detected by $F$.
 We wish then to estimate the weights of $F$ such that those local patterns are characteristic of one given class.
-For sake of clarity, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect(P_I(p)) \cdot \vect(F)$, where $P_I(p)$ is a $k\times k$ patch with $m$ \textcolor{red}{bands} around $p$, $F$ is a filter with shape $k\times k \times m$, and $\vect$ is the vectorization operation.
+For sake of clarity, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect(P_I(p)) \cdot \vect(F)$, where $P_I(p)$ is a $k\times k$ patch with $m$ \changed{bands} around $p$, $F$ is a filter with shape $k\times k \times m$, and $\vect$ is the vectorization operation.
 A filter $F$ can discriminate a class $i \in \{1,2,\ldots,c\}$ among $c$ classes when it generates positive results for local patterns $\vect(P_I(p))$ of class $i$ and negative results for patterns from other classes.
@@
 Figure~\ref{fig:filter} illustrates three groups with color-coded samples from two classes in a hypothetical 2D feature space.
-The marker-based batch normalization and convolution with filters obtained as centroids of the three groups in Figure~\ref{fig:filter}a creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters in distinct activation \textcolor{red}{bands}.
+The marker-based batch normalization and convolution with filters obtained as centroids of the three groups in Figure~\ref{fig:filter}a creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters in distinct activation \changed{bands}.
 
@@
   \end{center}
-  \caption{Example of a network with a single convolutional layer followed by a classifier, as projected by FLIM and used for the experiments in this work.}
+  \caption{Example of a network with a single convolutional layer followed by a classifier, as projected by FLIM and\changed{,} used for the experiments in this work.}
   \label{fig:arch}
@@
 
-    \begin{tikzpicture}[node distance=1.5cm]
+    \begin{tikzpicture}
     %images----------------------------
      \node (img1) [image] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-coco-789.png}};
-     \node (img2) [image, below left=-5em of img1] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-coco-789.png}};
-     \node (img3) [image, below left=-5em of img2] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-coco-789.png}};
+     \node (img2) [image, below left=-5em of img1] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-non_coco-105.png}};
+     \node (img3) [image, below left=-5em of img2] {\includegraphics[width=.08\textwidth]{figures/qualitative/vgg/lower-coco-2212.png}};
+     \node (img4) [image, below left=-5em of img3] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/upper-coco-108.png}};
 
-     \node (imgs) [fit={($(img1.north west)+(-0.45,0.1)$) ($(img3.south east)+(0.45,-0.1)$)}, inner sep=0pt, draw=none, rounded corners, very thick, label = {[label distance = 0.1em]below:{Training set}}] {};
+     \node (imgs) [fit={($(img1.north west)+(-0.64,0.0)$) ($(img4.south east)+(0.64,-0.0)$)}, inner sep=0pt, draw=none, rounded corners, very thick, label = {[label distance = 0.1em]below:{Training set}}] {};
      %----------------------------------
 
-     \node (tsne) [steps, right=1em of imgs] {t-SNE};
+     \node (tsne) [steps, right=1.4em of imgs] {t-SNE};
 
@@
 
-     \node (select) [steps, right=1.35em of vis1] {Select\\ Images};
+     \node (select) [steps, right=1.4em of vis1] {Select\\ Images};
 
     %images----------------------------
-    \node (simg1) [image, above right=-2.5em and 3em of select] {\includegraphics[width=.08\textwidth]{figures/img1.png}};
+    \node (simg1) [image, above right=-2.5em and 3.4em of select] {\includegraphics[width=.08\textwidth]{figures/img4.png}};
     \node (simg2) [image, below left=-5em of simg1] {\includegraphics[width=.08\textwidth]{figures/img2.png}};
     \node (simg3) [image, below left=-5em of simg2] {\includegraphics[width=.08\textwidth]{figures/img3.png}};
-    \node (simg4) [image, below left=-5em of simg3] {\includegraphics[width=.08\textwidth]{figures/img4.png}};
+    \node (simg4) [image, below left=-5em of simg3] {\includegraphics[width=.08\textwidth]{figures/img1.png}};
 
-    \node (simgs) [fit={($(simg1.north west)+(-0.64,0.2)$) ($(simg4.south east)+(0.64,-0.2)$)}, inner sep=0pt, draw=none, label = {[label distance = 0.0em]above:{Selected images}}] {};
+    \node (simgs) [fit={($(simg1.north west)+(-0.64,0.0)$) ($(simg4.south east)+(0.64,-0.0)$)}, inner sep=0pt, draw=none, label = {[label distance = 0.15em]above:{Selected images}}] {};
     %----------------------------------
 
-     \node (put) [steps, right=1em of simgs] {Draw\\Markers};
+     \node (put) [steps, right=1.4em of simgs] {Draw\\Markers};
 
      %images----------------------------
-    \node (mimg1) [image, above right=-2.5em and 2.8em of put] {\includegraphics[width=.08\textwidth]{figures/markers1.png}};
+    \node (mimg1) [image, above right=-2.5em and 3.4em of put] {\includegraphics[width=.08\textwidth]{figures/markers4.png}};
     \node (mimg2) [image, below left=-5em of mimg1] {\includegraphics[width=.08\textwidth]{figures/markers2.png}};
     \node (mimg3) [image, below left=-5em of mimg2] {\includegraphics[width=.08\textwidth]{figures/markers3.png}};
-    \node (mimg4) [image, below left=-5em of mimg3] {\includegraphics[width=.08\textwidth]{figures/markers4.png}};
+    \node (mimg4) [image, below left=-5em of mimg3] {\includegraphics[width=.08\textwidth]{figures/markers1.png}};
 
-    \node (mimgs) [fit={($(mimg1.north west)+(-0.64,0.2)$) ($(mimg4.south east)+(0.64,-0.2)$)}, inner sep=0pt, draw=none, label = {[label distance = 0.0em]above:{Markers}}] {};
+    \node (mimgs) [fit={($(mimg1.north west)+(-0.64,0.0)$) ($(mimg4.south east)+(0.64,-0.0)$)}, inner sep=0pt, draw=none, label = {[label distance = 0.1em]below:{Markers}}] {};
 
-    \node (flim) [steps, right=1em of mimgs] {FLIM};
+    \node (flim) [steps, right=1.4em of mimgs] {FLIM};
 
@@
 
-    \node (user) [image, below=2em of simgs, outer sep = -.5em] {\includegraphics[width=.07\textwidth]{figures/user.png}};
+    \node (user) [image, below=1.9em of simgs, outer sep = -.5em, label = {[label distance = 0.3em]below:{User}}] {\includegraphics[width=.07\textwidth]{figures/user.png}};
 
@@
      \draw [decorate,decoration={brace,amplitude=4pt,raise=5pt}]
-(img3.north west) -- (img1.north west) node [black,midway,xshift=-0.4cm, yshift=0.4cm, rotate=45] {\small 200};
+(img4.north west) -- (img1.north west) node [black,midway,xshift=-0.4cm, yshift=0.4cm, rotate=45] {\small 200};
 
@@
   \end{center}
-  \caption{\textcolor{red}{Pipeline for training a feature extractor using FLIM.
+  \caption{\changed{Pipeline for training a feature extractor using FLIM.
 In the projection, green points are images with coconut trees, and red points are images with no coconut trees.
@@
 The Humanitarian OpenStreetMap community annotated the regions.
-The dataset consists of 13587 regions with $90 \times 90$ pixels \textcolor{red}{and spatial resolution of $8$\si{\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
-\textcolor{red}{The region images contain three bands: red, green, and blue.
+The dataset consists of 13587 regions with $90 \times 90$ pixels \changed{and spatial resolution of $8$\si{\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
+\changed{The region images contain three bands: red, green, and blue.
 We convert them to the CIE L*a*b* color space and normalize each band within $[0,1]$}.
 
-\textcolor{red}{The experiments used an Intel(R) Xeon(R) CPU E5-2620 v4 @ 2.10GHz CPU and a Titan X with 12 GB of memory. They have been designed to evaluate the following hypotheses about FLIM.
+\changed{The experiments used an Intel(R) Xeon(R) CPU E5-2620 v4 @ 2.10GHz CPU and a Titan X with 12 GB of memory. They have been designed to evaluate the following hypotheses about FLIM.
 \begin{enumerate}
@@
 
-\textcolor{red}{For these experiments, we have selected the following models that combine feature extraction and classification methods.
+\changed{For these experiments, we have selected the following models that combine feature extraction and classification methods.
  \begin{itemize}
 \item FLIM+SVM: Image classification is performed by support vector machines (SVM) from FLIM-based features.  
-\item FLIM+MLP: Image classification is performed by \textcolor{red}{Multi-Layer Perceptron} (MLP) from FLIM-based features.  
+\item FLIM+MLP: Image classification is performed by \changed{Multi-Layer Perceptron} (MLP) from FLIM-based features.  
 \item FLIM-FT+MLP: The architecture in FLIM+MLP is fine-tuned by backpropagation. 
@@
 
-\textcolor{red}{First, we evaluated H1 and H2 by randomly selecting 200 images for the training set, 1000 images for a validation set, and the remaining 12387 images were left in the testing set.
-The validation set was used only once to choose empirically the architecture of the network shown in Figure~\ref{fig:arch}.
-The convolutional layer has $60$ filters with dimension $7 \times 7$ and 3 color \textcolor{red}{bands}.
+\changed{First, we evaluated H1 and H2 by randomly selecting 200 images for the training set, 2000 images for a validation set, and the remaining 11387 images were left in the testing set.
+The validation set was used only once to choose empirically the architecture of the network shown in Figure~\ref{fig:arch}, and to choose the training hyperparameters.
+The convolutional layer has $60$ filters with dimension $7 \times 7$ and 3 color \changed{bands}.
 It includes marker-based normalization before convolution, ReLu activation, and a max-pooling operation with a window of dimension $3 \times 3$.
-Since it is the single and last convolutional layer, we applied a stride of $4$ in max-pooling and batch normalization to create an input as close as possible to the one of the \textcolor{red}{MLP} classifier used in the popular VGG-16~\cite{simonyan2014very}.
+Since it is the single and last convolutional layer, we applied a stride of $4$ in max-pooling and batch normalization to create an input as close as possible to the one of the \changed{MLP} classifier used in the popular VGG-16~\cite{simonyan2014very}.
 We then vectorized and projected the training images by t-SNE~\cite{maaten2008visualizing} (Figure~\ref{fig:pipeline-xtractor}).
@@
 
-\textcolor{red}{Pytorch was chosen to implement FLIM and MLP, being SVM the linear model available in scikit-learn with~${C=0.01}$~\cite{fan2008liblinear, scikit-learn}.
+\changed{Pytorch was chosen to implement FLIM and MLP, being SVM the linear model available in scikit-learn with~${C=0.01}$~\cite{fan2008liblinear, scikit-learn}.
 We trained MLP for $40$ epochs and the whole network for $60$ epochs with a batch size of $64$.
@@
 
-\textcolor{red}{
+\changed{
 Table~\ref{tab:results} shows the mean results of this first experiment for three random splits of the dataset into training, validation, and testing sets.
@@
 
-\textcolor{red}{To evaluate H3, we used the validation set to increase the classifiers' number of training images.
+\changed{To evaluate H3, we used the validation set to increase the classifiers' number of training images.
 Note that the marker selection for FLIM was still the same from those four images.
@@
 Finally, Table II shows the mean results of FLIM+SVM among the three splits as the number of images for marker selection increases, and the training set size remains with 200 images (H4).
-The results are not conclusive for H4, but the choice of more markers does not deteriorate the performance of FLIM.
+The results are not conclusive for H4, but the choice of more markers does not deteriorate FLIM's performance.
 This point deserves further investigation -- i.e., choosing relevant training images and markers for FLIM.
@@
     }
-    \caption{\textcolor{red}{Markers used for training, cyan markers on coconut-tree pixels and orange markers on background pixels.}}
+    \caption{\changed{Markers used for training, cyan markers on coconut-tree pixels and orange markers on background pixels.}}
     \label{fig:markers}
@@
 \cite{rauber2016visualizing}, we created projections of three stages of the network FLIM+MLP using t-SNE \cite{maaten2008visualizing} to understand how our CNN transforms image spaces along its layers.
-In Figure \ref{fig:vis-input}, we can see the projection of the test images in the \textcolor{red}{CIE L*a*b* color space.} Green points are images that contain coconut trees, and red points are images that do not contain coconut trees.
+In Figure \ref{fig:vis-input}, we can see the projection of the test images in the \changed{CIE L*a*b* color space.} Green points are images that contain coconut trees, and red points are images that do not contain coconut trees.
 The projection shows considerable overlapping between the two classes.
@@
 \begin{figure}
+  \begin{center}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
+\end{center}
 \caption{Graph showing the impact of training set size on quantitative results.
@@
 
-We introduced a first feature learning technique, named FLIM, that can estimate effective filter weights for the convolutional layers of a given network architecture from user-drawn markers \textcolor{red}{in a few images of a training set.} The resulting feature extractor can be used with different classifiers and, when it is used with a MLP classifier, the number of required images to train the fully connected layers is reduced.
+We introduced a first feature learning technique, named FLIM, that can estimate effective filter weights for the convolutional layers of a given network architecture from user-drawn markers \changed{in a few images of a training set.} The resulting feature extractor can be used with different classifiers and, when it is used with \changed{an} MLP classifier, the number of required images to train the fully connected layers is reduced.
 We demonstrated the advantages of FLIM over solutions based on VGG-16 for the binary classification of coconut-tree images.
@@
 
-We intend to further investigate feature learning from image markers for different applications, \textcolor{red}{image and marker selection for FLIM,} elaborate methodologies to optimize network architecture, and extend FLIM to estimate weights in fully connected layers.
+We intend to further investigate feature learning from image markers for different applications, \changed{image and marker selection for FLIM,} elaborate methodologies to optimize network architecture, and extend FLIM to estimate weights in fully connected layers.
 

### 2020-07-27 text.tex
arxiv version

--- 2020-07-27 main.tex
@@
-%% bare_conf.tex
-%% V1.4b
-%% 2015/08/26
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
-%% conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/pkg/ieeetran
-%% and
-%% http://www.ieee.org/
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE!
-%% User assumes all risk.
-%% In no event shall the IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%*************************************************************************
-
-\documentclass[journal, twoside]{IEEEtran}
-
-%\documentclass[journal,twoside,peerreview]{IEEEtran}
-
-\def\citepunt{,}
-
-\usepackage[pdftex]{graphicx}
-\ifCLASSOPTIONcompsoc
-  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-\else
-  \usepackage[caption=false,font=footnotesize]{subfig}
-\fi
-\usepackage{comment}
-\usepackage{mathtools}
-\usepackage{amsfonts}
-\usepackage{bm}
-\usepackage{siunitx}
-\usepackage{textcomp}
-\usepackage{hyperref}
-
-\usepackage{tikz}
-\usetikzlibrary{calc}
-\usetikzlibrary{shapes.geometric, arrows, fit}
-\usepackage{varwidth}
-\usetikzlibrary{positioning}
-\usetikzlibrary{backgrounds}
-\usetikzlibrary{decorations.pathreplacing}
-
-\usepackage{pgfplots,filecontents}
-
-\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
-\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
-
-\DeclareMathOperator{\vect}{vec}
-
-\DeclareRobustCommand\changed{\textcolor{black}}
-
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\A}{\mathcal{A}}
-\newcommand{\D}{\mathcal{D}}
-\newcommand{\I}{\hat{I}}
-\newcommand{\Z}{\mathbb{Z}}
-\newcommand{\m}[1]{{\mathrm{\bf #1}}}
-\newcommand{\E}{\tilde{\m{I}}}
-\newcommand{\F}{\hat{F}}
-\newcommand{\lI}{\m{I}}
-\newcommand{\mF}{\m{F}}
-\newcommand{\bF}{\mathcal{F}}
-\newcommand{\J}{\hat{J}}
-\newcommand{\lJ}{\m{J}}
-\newcommand{\pD}{D^\prime}
-\newcommand{\eF}{\hat{\mF}}
-\newcommand{\gN}{\m{N}}
-\newcommand{\W}{\m{W}}
-\newcommand{\M}{\mathcal{M}}
-\newcommand{\Pa}{\mathcal{P}}
-\newcommand{\pDD}{\D^\prime}
-
-\definecolor{navy}{RGB}{0,0,137}
-\definecolor{tealDeer}{RGB}{148,232,180}
-\definecolor{dodgerBlue}{RGB}{18,161,255}
-\definecolor{citrine}{RGB}{230, 194, 8}
-\definecolor{violet}{RGB}{112,5,164}
-\definecolor{navyPurple}{RGB}{172,86,253}
-\definecolor{heliotrope}{RGB}{236,93,253}
-\definecolor{pink}{RGB}{250,38,160}
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\ifCLASSOPTIONpeerreview
-  \usepackage{setspace}
-  \setstretch{1.7}
-\fi
-
-\begin{document}
+%!TEX root = main.tex
 
 \title{Learning CNN filters from user-drawn image \changed{markers} for coconut-tree image classification}
-
-\author{Italos~de~Souza,~
-        and~Alexandre~ Falc\~{a}o,~\IEEEmembership{Member,~IEEE}}
-
-\markboth{IEEE GEOSCIENCE AND REMOTE SENSING LETTERS,~Vol.~17, No.~11, November~2020}%
-{de Souza \MakeLowercase{\textit{et al.}}: Learning CNN filters from user-drawn image makers for coconut-tree image classification}
 
@@
 \bibliography{bibliography}
-
-\end{document}

### 2020-08-10 text.tex
changing project structure

@@
-%!TEX root = main.tex
+%!TEX root = arxiv.tex
 
@@
   \subfloat[\label{fig:ex-groups}]
-  {\includegraphics[width=.35\linewidth]{figures/example_groups/groups_centered}}
+  {\includegraphics[width=.35\linewidth]{figs/example_groups/groups_centered}}
   ~
-  \subfloat[\label{fig:ex-groups-after}]{\includegraphics[width=.45\linewidth]{figures/example_groups/groups3d_centered}
+  \subfloat[\label{fig:ex-groups-after}]{\includegraphics[width=.45\linewidth]{figs/example_groups/groups3d_centered}
   }
@@
     %images----------------------------
-     \node (img1) [image] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-coco-789.png}};
-     \node (img2) [image, below left=-5em of img1] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-non_coco-105.png}};
-     \node (img3) [image, below left=-5em of img2] {\includegraphics[width=.08\textwidth]{figures/qualitative/vgg/lower-coco-2212.png}};
-     \node (img4) [image, below left=-5em of img3] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/upper-coco-108.png}};
+     \node (img1) [image] {\includegraphics[width=.08\textwidth]{figs/qualitative/ours/lower-coco-789.png}};
+     \node (img2) [image, below left=-5em of img1] {\includegraphics[width=.08\textwidth]{figs/qualitative/ours/lower-non_coco-105.png}};
+     \node (img3) [image, below left=-5em of img2] {\includegraphics[width=.08\textwidth]{figs/qualitative/vgg/lower-coco-2212.png}};
+     \node (img4) [image, below left=-5em of img3] {\includegraphics[width=.08\textwidth]{figs/qualitative/ours/upper-coco-108.png}};
 
@@
 
-     \node (vis1) [image, right=1.5em of tsne, outer sep=0pt, label = {[label distance = 0.1em]below:{Projection}}] {\includegraphics[width=.12\textwidth]{figures/input_200.png}};
+     \node (vis1) [image, right=1.5em of tsne, outer sep=0pt, label = {[label distance = 0.1em]below:{Projection}}] {\includegraphics[width=.12\textwidth]{figs/input_200.png}};
 
@@
     %images----------------------------
-    \node (simg1) [image, above right=-2.5em and 3.4em of select] {\includegraphics[width=.08\textwidth]{figures/img4.png}};
-    \node (simg2) [image, below left=-5em of simg1] {\includegraphics[width=.08\textwidth]{figures/img2.png}};
-    \node (simg3) [image, below left=-5em of simg2] {\includegraphics[width=.08\textwidth]{figures/img3.png}};
-    \node (simg4) [image, below left=-5em of simg3] {\includegraphics[width=.08\textwidth]{figures/img1.png}};
+    \node (simg1) [image, above right=-2.5em and 3.4em of select] {\includegraphics[width=.08\textwidth]{figs/img4.png}};
+    \node (simg2) [image, below left=-5em of simg1] {\includegraphics[width=.08\textwidth]{figs/img2.png}};
+    \node (simg3) [image, below left=-5em of simg2] {\includegraphics[width=.08\textwidth]{figs/img3.png}};
+    \node (simg4) [image, below left=-5em of simg3] {\includegraphics[width=.08\textwidth]{figs/img1.png}};
 
@@
      %images----------------------------
-    \node (mimg1) [image, above right=-2.5em and 3.4em of put] {\includegraphics[width=.08\textwidth]{figures/markers4.png}};
-    \node (mimg2) [image, below left=-5em of mimg1] {\includegraphics[width=.08\textwidth]{figures/markers2.png}};
-    \node (mimg3) [image, below left=-5em of mimg2] {\includegraphics[width=.08\textwidth]{figures/markers3.png}};
-    \node (mimg4) [image, below left=-5em of mimg3] {\includegraphics[width=.08\textwidth]{figures/markers1.png}};
+    \node (mimg1) [image, above right=-2.5em and 3.4em of put] {\includegraphics[width=.08\textwidth]{figs/markers4.png}};
+    \node (mimg2) [image, below left=-5em of mimg1] {\includegraphics[width=.08\textwidth]{figs/markers2.png}};
+    \node (mimg3) [image, below left=-5em of mimg2] {\includegraphics[width=.08\textwidth]{figs/markers3.png}};
+    \node (mimg4) [image, below left=-5em of mimg3] {\includegraphics[width=.08\textwidth]{figs/markers1.png}};
 
@@
 
-    \node (user) [image, below=1.9em of simgs, outer sep = -.5em, label = {[label distance = 0.3em]below:{User}}] {\includegraphics[width=.07\textwidth]{figures/user.png}};
+    \node (user) [image, below=1.9em of simgs, outer sep = -.5em, label = {[label distance = 0.3em]below:{User}}] {\includegraphics[width=.07\textwidth]{figs/user.png}};
 
@@
     \centering
-    \subfloat[\label{fig:markers1}]{\includegraphics[width=.25\linewidth]{figures/markers1}
+    \subfloat[\label{fig:markers1}]{\includegraphics[width=.25\linewidth]{figs/markers1}
     }
     ~
-    \subfloat[\label{fig:markers2}]{\includegraphics[width=.25\linewidth]{figures/markers2}
+    \subfloat[\label{fig:markers2}]{\includegraphics[width=.25\linewidth]{figs/markers2}
     }
     \\
-    \subfloat[\label{fig:markers3}]{\includegraphics[width=.25\linewidth]{figures/markers3}
+    \subfloat[\label{fig:markers3}]{\includegraphics[width=.25\linewidth]{figs/markers3}
     }
     ~
-    \subfloat[\label{fig:markers4}]{\includegraphics[width=.25\linewidth]{figures/markers4}
+    \subfloat[\label{fig:markers4}]{\includegraphics[width=.25\linewidth]{figs/markers4}
     }
@@
   \centering
-  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/ours/upper-coco-108.png}
+  \subfloat{\includegraphics[width=.2\linewidth]{figs/qualitative/ours/upper-coco-108.png}
   }
   ~
-  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/ours/upper-coco-2473.png}
+  \subfloat{\includegraphics[width=.2\linewidth]{figs/qualitative/ours/upper-coco-2473.png}
   }
   ~
-  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/ours/upper-non_coco-506.png}
+  \subfloat{\includegraphics[width=.2\linewidth]{figs/qualitative/ours/upper-non_coco-506.png}
   }
   \\
-  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/vgg/lower-coco-1098.png}
+  \subfloat{\includegraphics[width=.2\linewidth]{figs/qualitative/vgg/lower-coco-1098.png}
   }
   ~
-  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/vgg/lower-coco-875.png}
+  \subfloat{\includegraphics[width=.2\linewidth]{figs/qualitative/vgg/lower-coco-875.png}
   }
   ~
-  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/vgg/lower-non_coco-105.png}
+  \subfloat{\includegraphics[width=.2\linewidth]{figs/qualitative/vgg/lower-non_coco-105.png}
   }
@@
   \centering
-  \subfloat[\label{fig:vis-input}]{\includegraphics[width=.3\linewidth]{figures/input}
+  \subfloat[\label{fig:vis-input}]{\includegraphics[width=.3\linewidth]{figs/input}
   }
   ~
-  \subfloat[\label{fig:vis-feature-extractor}]{\includegraphics[width=.3\linewidth]{figures/feature_extractor}
+  \subfloat[\label{fig:vis-feature-extractor}]{\includegraphics[width=.3\linewidth]{figs/feature_extractor}
   }
   ~
-  \subfloat[]{\includegraphics[width=.3\linewidth]{figures/classifier}
+  \subfloat[]{\includegraphics[width=.3\linewidth]{figs/classifier}
   \label{fig:vis-classifier}}
@@
 \bibliographystyle{IEEEtran}
-\bibliography{bibliography}
+\bibliography{refs}

### 2020-08-13 main.tex
change authors names

@@
 
-\title{Learning CNN filters from user-drawn image \changed{markers} for coconut-tree image classification}
-
-\author{Italos~de~Souza,~
-        and~Alexandre~ Falc\~{a}o,~\IEEEmembership{Member,~IEEE}}
+\author{Italos~Estilon~de~Souza,~
+and~Alexandre~Xavier~Falc\~{a}o,~\IEEEmembership{Member,~IEEE}}
 
@@
 
-% make the title area
-\maketitle
-
-\begin{abstract}
-Identifying species of trees in aerial images is essential for land-use classification, plantation monitoring, and impact assessment of natural disasters.
-The manual identification of trees in aerial images is tedious, costly, and error-prone, so automatic classification methods are necessary.
-Convolutional Neural Network (CNN) models have well succeeded in image classification applications from different domains.
-However, CNN models usually require intensive manual annotation to create large training sets.
-One may conceptually divide a CNN into convolutional layers for feature extraction and fully connected layers for feature space reduction and classification.
-We present a method that needs a minimal set of user-selected images to train the CNN's feature extractor, reducing the number of required images to train the fully connected layers.
-The method learns the filters of each convolutional layer from user-drawn markers in image regions that discriminate classes, allowing better user control and understanding of the training process.
-It does not rely on optimization based on backpropagation, and we demonstrate its advantages on the binary classification of coconut-tree aerial images against one of the most popular CNN models.
-\end{abstract}
-
-\begin{IEEEkeywords}
-Design of convolutional neural networks, remote sensing image analysis, interactive machine learning.
-\end{IEEEkeywords}
-
-\section{Introduction}
-
-\IEEEPARstart{\changed{D}}{\changed{eep}} learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}.
-In remote sensing, the applications may involve segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}, for instance.
-In this work, we are interested in identifying species of trees from aerial images.
-The topic is important for land-use classification, plantation monitoring, and damage assessment of natural disasters.
-As the plantations can span vast areas, the manual identification of each tree is costly, tedious, and error-prone, and so automatic classification methods are necessary.
-
-Classification of tree species in aerial images has been actively investigated~\cite{fassnacht2016review}.
-In \cite{puttemans2018comparing} and~\cite{aparna2018cnn}, the authors present automatic solutions to detect coconut trees based on convolutional neural network (CNN) models.
-Despite these recent advances, CNN models usually require considerable human effort in image annotation to create large training sets.
-Vargas-Muñoz et al.~\cite{8899005} propose an active learning approach to mitigate the problem.
-The method explores data projection techniques to allow simultaneous annotation of multiple regions as having or not coconut trees.
-It then uses a CNN to identify the candidate regions with coconut trees.
-We adopt another alternative -- the design of simplified CNN models from small training sets with user interaction, being the user knowledgeable about CNN modeling and the application domain.
-
-The design of CNN models without a human as part of the training loop leaves several questions unanswered: (1) How to find a useful and simplified model for a given classification problem?
-(2) How to train that model from a minimum number of annotated images?
-(3) Can the user explain the decisions of the model?
-(4) Can the model improve from label corrections?
-The first question requires human knowledge about CNN modeling and the problem of interest.
-The second one requires to reduce the human effort to train a CNN model.
-The third issue is related to human understanding.
-It may explore visual analytics to explain the model's decisions and to guide the user in the design of the model~\cite{RauberInfoVis2018,GarciaIJCNN2019,SpinnerTVCG2020}.
-The fourth question is also essential during training, and it is related to user control over the process.
-They all lead to the importance of involving human experts during the machine learning process.
-\changed{Explainable artificial intelligence (XAI) has appeared to address (3)~\cite{rauber2016visualizing,arrietaInfoFusion2020}.
-However, the importance of human participation during the training process has called little attention~\cite{AngelovNN2020}.
-Exceptions rely on visual analytics~\cite{RauberInfoVis2018,GarciaIJCNN2019,SpinnerTVCG2020}, and we present here a first work in which the user has direct influence over the quality of the filters.}
-
-First, we conceptually divide a CNN into convolutional layers for feature extraction and fully connected layers for feature space reduction and classification.
-Each convolutional layer contains a filter bank, an activation function, and alternative operations (e.g., pooling, batch normalization).
-As the number of convolutional layers increases (deeper is the model), higher is the number of annotated images required to train the model by backpropagation.
-
-To reduce the need for large annotated training sets by exploiting the user knowledge, we present a method, called \changed{\emph{Feature Learning from Image Markers} (FLIM)}, that needs a minimal set of images to learn the filters of each convolutional layer.
-The user selects the number of convolutional layers, their filter sizes, \changed{a few images from a training set}, and draws markers in image regions that best discriminate the classes.
-The user takes a couple of minutes to select images and draw markers in those images.
-\changed{By clustering patches extracted from marker pixels at the input of each convolutional layer, FLIM finds filters that enhance those regions at the convolutional layers' output.
-The remaining training images can then be used to train the fully connected layers or another classifier.} The CNN architecture could also be optimized \changed{with the support of} visual analytics~\cite{SpinnerTVCG2020}, but we have not explored this option yet.
-\changed{In interactive semantic segmentation, user-selected clicks can affect a part of the network to improve segmentation, but the filters are not designed from those markers~\cite{SofiiukCVPR2020}.}
-
-We demonstrate the advantages of this interactive technique on binary classification of coconut-tree images against  VGG-16~\cite{simonyan2014very}.
-First, the user can better understand and control the training process by observing the effectiveness of marker selection.
-Second, very few images per class (e.g., less than five) seem to be enough.
-Third, by eliminating the optimization of the convolutional layers by backpropagation, FLIM can reduce the number of annotated images to train the fully connected layers.
-Fourth, FLIM is application-independent, and so it might be useful for other image classification problems.
-
-\section{FLIM: Feature Learning from Image Markers}
-\label{sec:method}
-
-\changed{A CNN model should enhance image regions that best discriminate classes through a sequence of convolutional layers.
-This section shows how identifying such regions by a simple user action can make that sequence an effective feature extractor.
-The method is named \emph{Feature Learning from Image Markers} (FLIM).
-In the feature space that results from convolutional layers,} it is common to assume that the classes are piecewise linearly separable and apply fully connected layers -- Multi-Layer Perceptron (MLP) -- for feature space reduction and classification.
-The training of a CNN usually relies on weight optimization using backpropagation, which requires a higher number of annotated images as deep as the model is.
-In FLIM, we eliminate the need for backpropagation to train convolutional layers by finding a set ${\cal F}$ of filters that can \changed{enhance} discriminant regions from all classes.
-
-The user selects a few \changed{images from a training set} to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes.
-The convolution between an image $I$ and a filter $F\in {\cal F}$ \changed{ enhances} regions whose local patterns are detected by $F$.
-We wish then to estimate the weights of $F$ such that those local patterns are characteristic of one given class.
-For sake of clarity, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect(P_I(p)) \cdot \vect(F)$, where $P_I(p)$ is a $k\times k$ patch with $m$ \changed{bands} around $p$, $F$ is a filter with shape $k\times k \times m$, and $\vect$ is the vectorization operation.
-A filter $F$ can discriminate a class $i \in \{1,2,\ldots,c\}$ among $c$ classes when it generates positive results for local patterns $\vect(P_I(p))$ of class $i$ and negative results for patterns from other classes.
-That is, $\vect(F)$ is the normal vector of a hyperplane in $\R^{k\times k\times m}$, which detects the patterns of class $i$ at its positive side.
-We wish to estimate ${\cal F}$ such that its filters will detect patterns from all classes in different positions $p$.
-
-For a problem with $c$ classes where $\lambda(p)=i\in \{1,2,\ldots,c\}$ is the label of a marker pixel $p$ from class $i$, let ${\cal M}_I$ be a set of marker pixels drawn in image $I\in {\cal D}$ and $P_I(p)$ be the respective patch around a pixel $p\in {\cal M}_I$.
-Let ${\cal P}_i$ be the set of all patches around marker pixels from all images $I\in {\cal D}$, with representative examples  $\vect(P_I(p))$ from class $i\in \{1,2,\ldots,c\}$.
-\begin{eqnarray}
-{\cal P}_i & = & \bigcup_{I\in {\cal D},p\in {\cal M}_I, \lambda(p) = i}{P_I(p)}.
-\label{eq:pset}
-\end{eqnarray}
-A clustering operation on each set ${\cal P}_i$, $i\in [1,c]$, guarantees groups with similar local patterns for each class.
-The groups must be shifted to the origin of $\R^{k\times k\times m}$ by subtracting the mean value of all patches, set $\mathcal{P} = \bigcup_{i \in \{1, 2, \ldots, c\}}{\mathcal{P}_i}$.
-Additionally, the standard deviation of all patches is computed and used for standardization.
-This operation allows batch normalization of image sets using the mean and standard deviation of the patches from the image markers -- i.e.,  \emph{a marker-based batch normalization}.
-The centroid of each cluster defines the weights of each filter $F\in {\cal F}$, and we force norm $\|\vect(F)\|=1$ to avoid preferences among them.
-The centralization for filter definition is paramount to eliminate activations in regions from other classes, whose patterns fall in the negative side of the hyperplane with normal $\vect(F)$.
-Figure~\ref{fig:filter} illustrates three groups with color-coded samples from two classes in a hypothetical 2D feature space.
-The marker-based batch normalization and convolution with filters obtained as centroids of the three groups in Figure~\ref{fig:filter}a creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters in distinct activation \changed{bands}.
-
-\begin{figure}[!ht]
-  \centering
-  \subfloat[\label{fig:ex-groups}]
-  {\includegraphics[width=.35\linewidth]{figures/example_groups/groups_centered}}
-  ~
-  \subfloat[\label{fig:ex-groups-after}]{\includegraphics[width=.45\linewidth]{figures/example_groups/groups3d_centered}
-  }
-  \caption{(a) A 2D feature space with color-coded samples from two classes distributed in three groups.
-(b) After marker-based batch normalization and convolution with filters obtained from each group, the classes can be more easily detected by two filters in a new 3D feature space.}
-  \label{fig:filter}
-\end{figure}
-
-Each convolutional layer is trained individually, one layer at a time, to find its filter set ${\cal F}$.
-The number of filters per layer depends on the clustering technique.
-We use $K$-means, and then the user must specify the number $K=|{\cal F}|$ of filters.
-After the convolution operation, we apply the ReLU function to eliminate negative activations and the max-pooling operation to aggregate local information.
-Note that, apart from the initial marker selection, the training process is automatic.
-We have preserved the convolutional layers' output dimensions and the marker pixels so we can find the filters of layer $L+1$ from the output of layer $L$.
-Figure~\ref{fig:arch} shows one example of a simple network projected by FLIM for the experiments of the next section.
-
-\begin{figure*}
-  \begin{center}
-
-    \tikzstyle{steps} = [rectangle, rounded corners, minimum width=1.3cm, minimum height=0.5cm,text centered, draw=dodgerBlue, very thick, fill=dodgerBlue!25, execute at begin node={\begin{varwidth}{2.5cm}},
-      execute at end node={\end{varwidth}}]
-
-      \tikzstyle{myarrow} = [navy, very thick]
-
-      \tikzstyle{arrow} = [thick,->,>=stealth]
-
-  \begin{tikzpicture}
-
-    \node (mbn) [steps] {MarkerBased \\ BatchNorm};
-
-    \node (conv) [steps, right= of mbn] {Conv$(60 \times 7 \times 7)$};
-
-    \node (relu) [steps, right= of conv] {ReLU};
-
-    \node (pool) [steps, right=of relu] {MaxPool$(3\times3)$};
-
-    \node (bn) [steps, right=of pool] {BatchNorm};
-
-    \node (clf) [steps, fill=navyPurple!25, draw=navyPurple, right= of bn] {Classifier};
-
-    %-----------------------arrows---------------
-
-    \draw [arrow, myarrow] (mbn) -- (conv);
-    \draw [arrow, myarrow] (conv) -- (relu);
-    \draw [arrow, myarrow] (relu) -- (pool);
-    \draw [arrow, myarrow] (pool) -- (bn);
-    \draw [arrow, myarrow] (bn) -- (clf);
-
-  \end{tikzpicture}
-  \end{center}
-  \caption{Example of a network with a single convolutional layer followed by a classifier, as projected by FLIM and\changed{,} used for the experiments in this work.}
-  \label{fig:arch}
-\end{figure*}
-
-\begin{figure*}
-  \begin{center}
-
-    \tikzstyle{steps} = [rectangle, rounded corners, minimum width=1.3cm, minimum height=0.5cm,text centered, draw=pink, very thick, fill=heliotrope!30, execute at begin node={\begin{varwidth}{2.5cm}},
-    execute at end node={\end{varwidth}}]
-    \tikzstyle{image} = [minimum width=1.3cm, minimum height=0.7cm,text centered, draw=white,fill=white!30, inner sep=0pt]
-
-    \tikzstyle{myarrow} = [navy, very thick]
-
-    \tikzstyle{arrow} = [thick,->,>=stealth]
-
-    \begin{tikzpicture}
-    %images----------------------------
-     \node (img1) [image] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-coco-789.png}};
-     \node (img2) [image, below left=-5em of img1] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/lower-non_coco-105.png}};
-     \node (img3) [image, below left=-5em of img2] {\includegraphics[width=.08\textwidth]{figures/qualitative/vgg/lower-coco-2212.png}};
-     \node (img4) [image, below left=-5em of img3] {\includegraphics[width=.08\textwidth]{figures/qualitative/ours/upper-coco-108.png}};
-
-     \node (imgs) [fit={($(img1.north west)+(-0.64,0.0)$) ($(img4.south east)+(0.64,-0.0)$)}, inner sep=0pt, draw=none, rounded corners, very thick, label = {[label distance = 0.1em]below:{Training set}}] {};
-     %----------------------------------
-
-     \node (tsne) [steps, right=1.4em of imgs] {t-SNE};
-
-     \node (vis1) [image, right=1.5em of tsne, outer sep=0pt, label = {[label distance = 0.1em]below:{Projection}}] {\includegraphics[width=.12\textwidth]{figures/input_200.png}};
-
-     \node (select) [steps, right=1.4em of vis1] {Select\\ Images};
-
-    %images----------------------------
-    \node (simg1) [image, above right=-2.5em and 3.4em of select] {\includegraphics[width=.08\textwidth]{figures/img4.png}};
-    \node (simg2) [image, below left=-5em of simg1] {\includegraphics[width=.08\textwidth]{figures/img2.png}};
-    \node (simg3) [image, below left=-5em of simg2] {\includegraphics[width=.08\textwidth]{figures/img3.png}};
-    \node (simg4) [image, below left=-5em of simg3] {\includegraphics[width=.08\textwidth]{figures/img1.png}};
-
-    \node (simgs) [fit={($(simg1.north west)+(-0.64,0.0)$) ($(simg4.south east)+(0.64,-0.0)$)}, inner sep=0pt, draw=none, label = {[label distance = 0.15em]above:{Selected images}}] {};
-    %----------------------------------
-
-     \node (put) [steps, right=1.4em of simgs] {Draw\\Markers};
-
-     %images----------------------------
-    \node (mimg1) [image, above right=-2.5em and 3.4em of put] {\includegraphics[width=.08\textwidth]{figures/markers4.png}};
-    \node (mimg2) [image, below left=-5em of mimg1] {\includegraphics[width=.08\textwidth]{figures/markers2.png}};
-    \node (mimg3) [image, below left=-5em of mimg2] {\includegraphics[width=.08\textwidth]{figures/markers3.png}};
-    \node (mimg4) [image, below left=-5em of mimg3] {\includegraphics[width=.08\textwidth]{figures/markers1.png}};
-
-    \node (mimgs) [fit={($(mimg1.north west)+(-0.64,0.0)$) ($(mimg4.south east)+(0.64,-0.0)$)}, inner sep=0pt, draw=none, label = {[label distance = 0.1em]below:{Markers}}] {};
-
-    \node (flim) [steps, right=1.4em of mimgs] {FLIM};
-
-    \node (ft) [steps, below=5em of flim, fill=white, draw=none] {Feature\\Extractor};
-
-    \node (user) [image, below=1.9em of simgs, outer sep = -.5em, label = {[label distance = 0.3em]below:{User}}] {\includegraphics[width=.07\textwidth]{figures/user.png}};
-
-     %--------------------------------------------------------------arrows
-     \draw [arrow, myarrow] (imgs) -- (tsne);
-     \draw [arrow, myarrow] (tsne) -- (vis1);
-     \draw [arrow, myarrow] (vis1) -- (select);
-     \draw [arrow, myarrow] (select) -- (simgs);
-     \draw [arrow, myarrow] (simgs) -- (put);
-     \draw [arrow, myarrow] (put) -- (mimgs);
-     \draw [arrow, myarrow] (mimgs) -- (flim);
-     \draw [arrow, myarrow] (flim) -- (ft);
-     \draw [arrow, myarrow] (user) -- (select);
-     \draw [arrow, myarrow] (user) -- (put);
-
-     \draw [decorate,decoration={brace,amplitude=4pt,raise=5pt}]
-(img4.north west) -- (img1.north west) node [black,midway,xshift=-0.4cm, yshift=0.4cm, rotate=45] {\small 200};
-
-\draw [decorate,decoration={brace,amplitude=4pt,raise=5pt}]
-(simg4.north west) -- (simg1.north west) node [black,midway,xshift=-0.4cm, yshift=0.4cm, rotate=45] {\small 4};
-
-\draw [decorate,decoration={brace,amplitude=4pt,raise=5pt}]
-(mimg4.north west) -- (mimg1.north west) node [black,midway,xshift=-0.4cm, yshift=0.4cm, rotate=45] {\small 4};
-
-    \end{tikzpicture}
-  \end{center}
-  \caption{\changed{Pipeline for training a feature extractor using FLIM.
-In the projection, green points are images with coconut trees, and red points are images with no coconut trees.
-The user selects a few images (e.g., 4) from the projection.
-The user draws orange markers on background pixels and cyan markers on coconut-tree pixels for image feature learning by FLIM.}}
-  \label{fig:pipeline-xtractor}
-\end{figure*}
-
-\section{Experiments}
-
-We use a dataset from~\cite{8899005}, which contains regions of aerial images with and without coconut trees from the Kingdom of Tonga, as acquired in October 2017~\footnote{The images are available in \url{blog.werobotics.org/2018/01/11/open-ai-challenge-2}.}.
-Each region is classified as containing one or none coconut tree, but part of coconut trees from adjacent regions might appear near the region's border.
-The Humanitarian OpenStreetMap community annotated the regions.
-The dataset consists of 13587 regions with $90 \times 90$ pixels \changed{and spatial resolution of $8$\si{\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively.
-\changed{The region images contain three bands: red, green, and blue.
-We convert them to the CIE L*a*b* color space and normalize each band within $[0,1]$}.
-
-\changed{The experiments used an Intel(R) Xeon(R) CPU E5-2620 v4 @ 2.10GHz CPU and a Titan X with 12 GB of memory. They have been designed to evaluate the following hypotheses about FLIM.
-\begin{enumerate}
-  \item[H1]  It can produce competitive feature extractors from user-drawn markers in a few images selected per class from a training set. 
-  \item[H2]  For a given network architecture, FLIM can provide better classification results than those obtained by training the same architecture with backpropagation.
-  \item[H3] Its feature extractor remains competitive independently of the training set size. 
-  \item[H4] User-drawn markers on a higher number of selected images can improve feature extraction.
-\end{enumerate}
-}
-
-\changed{For these experiments, we have selected the following models that combine feature extraction and classification methods.
- \begin{itemize}
-\item FLIM+SVM: Image classification is performed by support vector machines (SVM) from FLIM-based features.  
-\item FLIM+MLP: Image classification is performed by \changed{Multi-Layer Perceptron} (MLP) from FLIM-based features.  
-\item FLIM-FT+MLP: The architecture in FLIM+MLP is fine-tuned by backpropagation. 
-\item CL+MLP: The architecture in FLIM+MLP is trained from scratch by backpropagation.
-\item VGG: VGG-16~\cite{simonyan2014very} is trained from scratch.
-\item VGG-FT: VGG-16 is pre-trained on ImageNet and fine-tuned on the coconut-tree training set.
-\item VGG-FE+SVM: Image features are obtained from the last convolutional layer of VGG-FT.
-\end{itemize}
-}
-
-\changed{First, we evaluated H1 and H2 by randomly selecting 200 images for the training set, 2000 images for a validation set, and the remaining 11387 images were left in the testing set.
-The validation set was used only once to choose empirically the architecture of the network shown in Figure~\ref{fig:arch}, and to choose the training hyperparameters.
-The convolutional layer has $60$ filters with dimension $7 \times 7$ and 3 color \changed{bands}.
-It includes marker-based normalization before convolution, ReLu activation, and a max-pooling operation with a window of dimension $3 \times 3$.
-Since it is the single and last convolutional layer, we applied a stride of $4$ in max-pooling and batch normalization to create an input as close as possible to the one of the \changed{MLP} classifier used in the popular VGG-16~\cite{simonyan2014very}.
-We then vectorized and projected the training images by t-SNE~\cite{maaten2008visualizing} (Figure~\ref{fig:pipeline-xtractor}).
-Four representative images, only two per class, were chosen by the user to draw markers (Figure~\ref{fig:markers}).
-These images must come from regions in the projection populated by samples from the same class (either red or green points).
-Those regions indicate groups of images from the same class in the input feature space.
-The markers must be placed in regions that best distinguish the coconut and non-coconut classes.}
-
-\changed{Pytorch was chosen to implement FLIM and MLP, being SVM the linear model available in scikit-learn with~${C=0.01}$~\cite{fan2008liblinear, scikit-learn}.
-We trained MLP for $40$ epochs and the whole network for $60$ epochs with a batch size of $64$.
-We used a learning rate of $0.0001$, a weight decay of $0.001$, and we updated the learning rate by a factor of $0.1$ at every $5$ epochs after epoch 30.
-VGG and VGG-FT were trained for $95$ epochs with a batch size of $64$, a learning rate of $0.001$, weight decay of $0.001$, and we updated the learning rate by a factor of $0.1$ at every $5$ epochs after the model achieved accuracy $0.95$.}
-
-\changed{
-Table~\ref{tab:results} shows the mean results of this first experiment for three random splits of the dataset into training, validation, and testing sets.
-We considered better (in bold) the results with non-overlapping intervals of standard deviation around the mean value.
-The comparison between FLIM+SVM and VGG-FE+SVM indicates that FLIM can produce a feature extractor with a single layer competitive to VGG-FE using 13 layers (i.e., FLIM uses only $0.07\%$ of the number of parameters in VGG-FE).
-The results of FLIM+MLP, CL+MLP, and FLIM-FT+MLP show that FLIM may dismiss training and fine-tuning based on backpropagation, by directly providing a more effective solution than CL+MLP and FLIM-FT+MLP (H2).
-This result can also be noticed when comparing  FLIM+MLP with VGG, confirming that FLIM can reduce the number of required training images for the MLP classifier.
-The comparison between FLIM+MLP and VGG-FT shows competitive results with a top-rated solution (H1).}
-
-\changed{To evaluate H3, we used the validation set to increase the classifiers' number of training images.
-Note that the marker selection for FLIM was still the same from those four images.
-To our surprise, as the training set increases, the differences in mean f-score, as obtained among the three splits, favor FLIM+MLP over FLIM-FT+MLP (H2), VGG-FT (H1 and H3), and CL+MLP (H2).
-The differences between FLIM+MLP and CL+MLP demonstrate that FLIM is superior to backpropagation in filter estimation (H2).
-Finally, Table II shows the mean results of FLIM+SVM among the three splits as the number of images for marker selection increases, and the training set size remains with 200 images (H4).
-The results are not conclusive for H4, but the choice of more markers does not deteriorate FLIM's performance.
-This point deserves further investigation -- i.e., choosing relevant training images and markers for FLIM.
-}
-
-\begin{figure}[!t]
-    \centering
-    \subfloat[\label{fig:markers1}]{\includegraphics[width=.25\linewidth]{figures/markers1}
-    }
-    ~
-    \subfloat[\label{fig:markers2}]{\includegraphics[width=.25\linewidth]{figures/markers2}
-    }
-    \\
-    \subfloat[\label{fig:markers3}]{\includegraphics[width=.25\linewidth]{figures/markers3}
-    }
-    ~
-    \subfloat[\label{fig:markers4}]{\includegraphics[width=.25\linewidth]{figures/markers4}
-    }
-    \caption{\changed{Markers used for training, cyan markers on coconut-tree pixels and orange markers on background pixels.}}
-    \label{fig:markers}
-\end{figure}
-
-\begin{table}[!t]
-  \begin{center}
-  \begin{tabular}{|l|c|c|c|}
-  \hline
-   Method & Precision & Recall & F-score \\
-  \hline\hline
-    FLIM+SVM & 0.856 \textpm 0.011  &  0.831 \textpm 0.019 & 0.838 \textpm 0.017\\ 
-    VGG-FE+SVM & 0.855 \textpm 0.001 & 0.816 \textpm 0.007 & 0.826 \textpm 0.006  \\\hline
-
-    FLIM+MLP & \textbf{0.863 \textpm 0.002} & \textbf{0.849 \textpm 0.005} & \textbf{0.854 \textpm 0.004}\\
-    FLIM-FT+MLP & 0.845 \textpm 0.003 & 0.825 \textpm 0.006 & 0.832 \textpm 0.005 \\
-    CL+MLP & 0.822 \textpm 0.011 & 0.805 \textpm 0.013 & 0.811 \textpm 0.012  \\
-    VGG & 0.827 \textpm 0.003 & 0.770 \textpm 0.016  &  0.784 \textpm 0.014\\
-    VGG-FT & \textbf{0.872 \textpm 0.007} & \textbf{0.844 \textpm 0.015} & \textbf{0.851 \textpm 0.014}  \\
-
-  \hline
-  \end{tabular}
-  \end{center}
-
-\caption{Mean and standard deviation of precision, recall, and f-score for each method: a feature extractor with a classifier.}
-  \label{tab:results}
-\end{table}
-
-\begin{figure}
-  \centering
-  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/ours/upper-coco-108.png}
-  }
-  ~
-  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/ours/upper-coco-2473.png}
-  }
-  ~
-  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/ours/upper-non_coco-506.png}
-  }
-  \\
-  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/vgg/lower-coco-1098.png}
-  }
-  ~
-  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/vgg/lower-coco-875.png}
-  }
-  ~
-  \subfloat{\includegraphics[width=.2\linewidth]{figures/qualitative/vgg/lower-non_coco-105.png}
-  }
-  \caption{Misclassified images by ``FLIM + MLP'' (first row) and VGG-FT (second row), being the last column with images without coconut trees.}
-  \label{fig:ex-classification}
-\end{figure}
-
- In Figure \ref{fig:ex-classification}, we can see examples of misclassified images.
-In these images, coconut trees appear in different angles, sizes, and shapes, or the boundaries between the coconut tree and the background are tenuous, making it more difficult to identify them.
-In turn, images that do not contain coconut trees, contain trees that resemble their shape.
-The network designer could add markers in those images to improve the feature extractor in an active learning loop.
-
-\begin{figure}
-  \centering
-  \subfloat[\label{fig:vis-input}]{\includegraphics[width=.3\linewidth]{figures/input}
-  }
-  ~
-  \subfloat[\label{fig:vis-feature-extractor}]{\includegraphics[width=.3\linewidth]{figures/feature_extractor}
-  }
-  ~
-  \subfloat[]{\includegraphics[width=.3\linewidth]{figures/classifier}
-  \label{fig:vis-classifier}}
-  \caption{Projections by t-SNE of the test images at (a) the input layer, (b) the output of the convolutional layer of  FLIM+MLP after batch normalization (Figure~\ref{fig:arch}), and (c) the output of the last hidden layer of the MLP classifier in  FLIM+MLP.}
-\end{figure}
-
-As motivated by Rauber et al.
-\cite{rauber2016visualizing}, we created projections of three stages of the network FLIM+MLP using t-SNE \cite{maaten2008visualizing} to understand how our CNN transforms image spaces along its layers.
-In Figure \ref{fig:vis-input}, we can see the projection of the test images in the \changed{CIE L*a*b* color space.} Green points are images that contain coconut trees, and red points are images that do not contain coconut trees.
-The projection shows considerable overlapping between the two classes.
-The samples of both classes are also dispersed in the projection.
-In Figure \ref{fig:vis-feature-extractor}, the projection of the output of our feature extractor shows some reduction in sample dispersion.
-The classes show more concentration in some parts of the projection and reduced overlapping between them.
-Finally, Figure \ref{fig:vis-classifier} shows the output of the last hidden layer of the MLP classifier in FLIM+MLP.
-The overlapping between classes and sample dispersion are considerably reduced, explaining our results.
-
-\begin{figure}
-  \begin{center}
-\begin{tikzpicture}
-  \begin{axis}[
-    /pgf/number format/set thousands separator = {},
-    xlabel = Training set size,
-    ylabel = f-score,
-    xmin=300, xmax=1200,
-    grid=major,
-    grid style=dashed,
-    xtick={300,450, 600, 750, 900, 1050, 1200},
-    legend style={nodes={scale=0.65, transform shape}},
-    legend pos=south east
-    ]
-    \addplot table[x index=0,y index=1,header=false] {data/data.dat};
-    \addlegendentry{FLIM+MLP}
-
-    \addplot  table[x index=0,y index=2,header=false] {data/data.dat};
-    \addlegendentry{FLIM-FT+MLP}
-
-    \addplot  table[x index=0,y index=3,header=false] {data/data.dat};
-    \addlegendentry{VGG-FT}
-
-    \addplot table[x index=0,y index=4,header=false] {data/data.dat};
-    \addlegendentry{CL+MLP}
-
-    \legend{FLIM+MLP,FLIM-FT+MLP,VGG-FT,CL+MLP}
-
-  \end{axis}
-\end{tikzpicture}
-\end{center}
-\caption{Graph showing the impact of training set size on quantitative results.
-FLIM+MLP feature extractor was trained only with four images, and the remaining training set was used to train the MLP classifier.}
-\end{figure}
-
-\begin{table}[!t]
-    \begin{center}
-        \begin{tabular}{|l|c|c|c|}
-            \hline
-            Method & Precision & Recall & F-score \\
-            \hline\hline
-            FLIM+SVM (4) & 0.856 \textpm 0.011 &  0.831 \textpm 0.019 &  0.838 \textpm 0.017\\
-            FLIM+SVM (8) & 0.857 \textpm 0.005 & 0.838 \textpm 0.010 & 0.844 \textpm 0.009\\
-            FLIM+SVM (12) & 0.854 \textpm 0.005 & 0.839 \textpm 0.009 & 0.844 \textpm 0.008\\
-            \hline
-        \end{tabular}
-    \end{center}
-    \caption{Mean and standard deviation of precision, recall, and f-score for each method: a feature extractor trained with 8 images and a SVM classifier.}
-    \label{tab:results-more-markers}
-\end{table}
-
-\section{Conclusion}
-
-We introduced a first feature learning technique, named FLIM, that can estimate effective filter weights for the convolutional layers of a given network architecture from user-drawn markers \changed{in a few images of a training set.} The resulting feature extractor can be used with different classifiers and, when it is used with \changed{an} MLP classifier, the number of required images to train the fully connected layers is reduced.
-We demonstrated the advantages of FLIM over solutions based on VGG-16 for the binary classification of coconut-tree images.
-The experiments indicated that FLIM is an effective approach to produce considerably simplified network architectures.
-By involving the user in the training process, FLIM improves the understanding about CNN models and the user control over the training process.
-
-We intend to further investigate feature learning from image markers for different applications, \changed{image and marker selection for FLIM,} elaborate methodologies to optimize network architecture, and extend FLIM to estimate weights in fully connected layers.
-
-\section*{Acknowledgments}
-
-This research was funded by FAPESP (2014/12236-1), CNPq (303808/2018-7), Petróleo Brasileiro S.A.
-(PETROBRAS) and Agência Nacional do Petróleo, Gás Natural e Biocombustíveis (ANP), Grant Numbers 4600556376 and 4600583791.
-\bibliographystyle{IEEEtran}
-\bibliography{bibliography}
+\include{text}
 

### 2020-08-19 main.tex
fix some references

@@
 
-\documentclass[journal, twoside]{IEEEtran}
+\documentclass[journal, twoside, final]{IEEEtran}
 
@@
 
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
 \ifCLASSOPTIONpeerreview
@@
 
-\author{Italos~Estilon~de~Souza,~
-and~Alexandre~Xavier~Falc\~{a}o,~\IEEEmembership{Member,~IEEE}}
+\author{Italos~Estilon~de~Souza,~and~Alexandre~Xavier~Falc\~{a}o,~\IEEEmembership{Member,~IEEE}}
 
 \markboth{IEEE GEOSCIENCE AND REMOTE SENSING LETTERS,~Vol.~17, No.~11, November~2020}%
-{de Souza \MakeLowercase{\textit{et al.}}: Learning CNN filters from user-drawn image makers for coconut-tree image classification}
+{de Souza and Falc\~{a}o: Learning CNN filters from user-drawn image markers}
 
